### 2019-04-12 Manuscript/manuscriptPGPE.tex
Update manuscript and figures

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with 
% the REVTEX 4.1 document class and the AIP substyles.
% 
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.

%\documentclass[aip,graphicx]{revtex4-1}
%\documentclass[aip,reprint]{revtex4-1}

%\usepackage{graphicx}

%\draft % marks overfull lines with a black rule on the right
%\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}

\documentclass[aps,prl,superscriptaddress,twocolumn]{revtex4}

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\usepackage{csquotes}
\usepackage{xr}
\externaldocument{manuscriptPGPEsuppl}

\begin{document}

% Use the \preprint command to place your local institutional report number 
% on the title page in preprint mode.
% Multiple \preprint commands are allowed.
%\preprint{}

\title{NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids} %Title of paper

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
% Explanatory text should go in the []'s, 
% actual e-mail address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro for the type of information

% \affiliation command applies to all authors since the last \affiliation command. 
% The \affiliation command should follow the other information.

\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}

% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
% \noaffiliation is required (may also be used with the \author command).
%\collaboration{}
%\noaffiliation

\date{\today}

\begin{abstract}
% insert abstract here
  Primarily measured but also simulated NMR order parameters will be collected also for other than phophatidylcholine
  (these are discussed in NMRlipids I) headgroup. The information will be used to understand structural differences between 
  different lipid molecules in bilayers.
\end{abstract}

%\pacs{}% insert suggested PACS numbers in braces on next line

\maketitle %\maketitle must follow title, authors, abstract and \pacs

% Body of paper goes here. Use proper sectioning commands. 
% References should be done using the \cite, \ref, and \label commands

%\label{}
\section{Introduction}

In NMRlipids I and II project we were looking for a MD model
which would correctly reproduce headgroup and glycerol
backbone structures and cation binding for PC lipid bilayers \cite{botan15,catte16}.
Here we extend the same goal for other than PC lipids.
Currently the focus is on PE, PG and PS bilayers and their
mixtures with PC. Experimental data with different amounts of 
added salt is now collected and presented in this manuscript.

%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
%  \caption{\label{lipids}
%    Chemical structures and labels for the headgroup carbons.
%  }
%\end{figure}

Absolute values of experimental order parameters for different lipid headgroups are
collected in Fig. \ref{HGorderParameters}. Signs are measured only for PC as far as I know,
thus only absolute values are used for now.

Based on superficial reading, the conclusions in the literature are roughly \\ 
1) glycerol backbone structures are largely similar irrespectively of the headroup \cite{gally81}, \\
2) glycerol backbone and headgroup structure and behaviour are similar in model membranes and in bacteria \cite{gally81,scherer87,seelig90}, \\
3) headgroup structures are similar in PC, PE and PG lipids, while headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. \\
Extensive discussion about structural details of PE, PG or PS headgroups do not exists (as far as I know), 
In contrast to PC lipids (see \cite{botan15} and references therein).

Several simulations containing PE, PG and PS lipids have been published \cite{??}, \todo{List should be completed}
however, glycerol backbone and headgroup order parameters are not compared to
the experiments (based on superficial reading of literature).

\section{Methods}
\subsection{Experimental C--H bond order parameters}% from the natural abundance $^{13}$C\,NMR}

The headgroup and glycerol backbone C--H bond order parameter magnitudes and signs of POPE and POPG
were determined by measuring the chemical-shift resolved dipolar splittings
with a R-type Proton Detected Local Field (R-PDLF) experiment~\cite{dvinskikh04} and
%The corresponding order parameter signs were measured with a
S-DROSS experiments~\cite{gross97} using natural abundance $^{13}$C solid state NMR spectroscopy
as described previously \cite{ferreira13,ferreira16,NMRlipidsIVps}.
%The experiments were done in a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz.
%Magic angle spinning (MAS) of the sample was used at a frequency of 5.15 kHz (R-PDLF experiment) and 5 kHz (S-DROSS experiment).
%The following experimental setups were used.
\todo{The rest of the details to be written. I am not sure how much we need to repeat the NMRlipidsIVps paper.}

\section{Molecular dynamics simulations}

Molecular dynamics simulation data were collected using
the Open Collaboration method \cite{botan15}, with
the NMR\-lipids Project blog (\url{nmrlipids.blogspot.fi}) and
GitHub repository (\url{github.com/NMRlipids/NMRlipidsIVotherHGs})
as the communication platforms.
The simulated systems are listed in 
Table \ref{systems} (pure PE and PG bilayers without additional ions).
%and \ref{mixedIONsystems} (mixed PC:PS bilayers at various salt concentrations).
Further simulation details are given in the SI, and
the simulation data are indexed in a
searchable database available at \url{www.nmrlipids.fi},
and in the NMRlipids/MATCH repository (\url{github.com/NMRlipids/MATCH}).

The C--H bond order parameters were calculated directly
from the carbon and hydrogen positions using the definition
\begin{equation}
S_{\rm CH}=\frac{1}{2}\langle 3\cos^2\theta -1 \rangle,
\end{equation}
where $\theta$ is the angle between the C--H bond and the membrane normal
(taken to align with $z$, with bilayer periodicity in the $xy$-plane).
Angular brackets denote average over all sampled configurations.
The order parameters were first calculated averaging over time separately
for each lipid in the system. The average and
the standard error of the mean were then calculated over different lipids.
Python program ({\tt scripts/calcOrderParameters.py}) that uses the
MDAnalysis library \cite{agrawal11,gowers16} is available in Ref. \citenum{MATCHgit}. 
The ion number density profiles were calculated using the {\tt gmx density} tool
of the Gromacs sofware package \cite{gromacsMANUAL}.

\begin{table*}[htb]
  %\begin{sidewaystable*}[!p]
  \centering
  \caption{List of single lipid type MD simulations without additional ions.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systems}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$    & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      DPPE  & Slipids \cite{jambeck12b}    & 288 	& 9386  & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
      DPPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      \hline
      POPE  & GROMOS 43A1-S3 \cite{??}     & 128	& ?     & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & CHARMM36ua \cite{??}         & 336	& ?     & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & Slipids \cite{jambeck12b,??} & 336	& ?     & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
      POPE  & OPLS-UA vdW on H \cite{??}   & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & OPLS-UA \cite{??}            & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & Berger-based \cite{??}       & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & Berger-based2 \cite{??}      & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      POPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      \hline
      DOPE  & Berger-based \cite{??}       & 128	& ?     & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      DOPE  & Berger-based2 \cite{??}      & 128	& ?     & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      DOPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
      \hline
      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}   & 118& 4110  & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
      POPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 10664  & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
      \hline
      DPPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 11232  & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
      DPPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 11232  & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
    \end{tabular}
  \end{minipage}
  %\end{sidewaystable*} 
\end{table*}

\begin{table*}[htb]
  %\begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with lipid mixtures and additional ions.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systems}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & NaCl (mM) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0.15 \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0  \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0.15 \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0  \todoi{Concentration to be checked.}  & ? & ? & ?  &  ?  & ? & ? & \cite{??} \todoi{Data to be uploaded by J. Madsen. Details to be filled once we have the data}  \\
      \hline
      POPC             & CHARMM36 \cite{??}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
      POPC:POPE (1:1)  & CHARMM36 \cite{??}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
     \hline
      POPC             & Berger \cite{??} \todoi{This is probable not plain berger, correct force filed should be described.}  &0  & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
      POPC:POPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\
      POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.}         &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
      DOPC             & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.}         &0          & 0  & 256 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPCberger300K} \\
      DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}        &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
    \end{tabular}
  \end{minipage}
  %\end{sidewaystable*} 
\end{table*}

\section{Results and Discussion}

\subsection{Headgroup and glycerol backbone order parameters of POPE and POPG from $^{13}$C NMR}

The glycerol backbone and $\alpha$-carbon peaks in INEPT spectra of POPE were assigned based on
previously measured POPC spectra (Fig. \ref{POPEspectra}) \cite{ferreira13}. The $\beta$-carbon peak was assigned based
on $^{13}$C chemical shift table for amines available at \url{https://www.chem.wisc.edu/areas/reich/nmr/c13-data/cdata.htm}.
The order parameters for the glycerol backbone and headgroup C--H bonds were determined
from 2D-RPDLF and S-DROSS experiments (Fig. \ref{POPEspectra}), as described previously \cite{NMRlipidsIVps}.
The POPE experiments were recorded at 310~K, where the bilayer is in liquid disordered phase \cite{??}.
\todo{To be checked by Tiago}.

\todo{Figure and discussion about POPG experiments to be addded.}

The headgroup and glycerol backbone order parameters of PE lipids are similar with 
different acyl chains and also close the values for POPC, althought PE gives systematically
slightly more positive values (Fig. \ref{HGorderParameters}). These could be explained with slightly larger temperature
in PE measurements, except for the $\alpha$-carbon with the positive sign, for which the
more positive value is farther away from zero. For PG lipids, the glycerol backbone order
parameters are more positive than for other lipids. The headgroup $\alpha$-carbon gives
value close to PE, while the value of $\beta$-carbon is distinct from other lipid being
only one which has positive sign, suggesting distinct conformation of PG lipids in this region.
This was not observed in previous $^2$H NMR study, where sign was not measured and $\beta$-carbon
order parameter was apparently similar to the value for PE and PC results.

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCPSPEPG.eps}
  \caption{\label{HGorderParameters}
    (top) Chemical structure of different lipids
    (bottom) Headgroup and glycerol backbone order parameters measured from lipids
    with different headgroups in lamellar liquid disordered phase.
    The values and signs for POPE (310~K), POPG (298~K). POPS (298~K) \cite{NMRlipidsIVps} and POPC (300~K) \cite{ferreira13,ferreira16}
    are measured using $^{13}$C NMR. The literature values for
    DOPS with 0.1M of NaCl (303~K) \cite{browning80},
    POPG with 10nM PIPES (298~K) \cite{borle85},
    DPPG with 10mM PIPES and 100mM NaCl (314~K) \cite{wohlgemuth80}, 
    DPPE (341~K) \cite{seelig76},
    E.coliPE and E.coliPG (310~K) \cite{gally81}
    are measured using $^2$H NMR. The signs from $^{13}$C NMR are used also for the literature values.
  }
\end{figure}

In conclusion, the results suggests that the glycerol backbone conformations in all lipids are
relatively similar. Also, the headgroup conformations are similar for PC and PE lipids, while
PS and PG are singnificantly different. For PS lipids, the differences are discussed previously \cite{??}.

\subsection{PE headgroup and glycerol backbone}
\todo{Add results from this discussion: \url{http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1502290977005#c4964787551006026647}}
Order parameters from Slipids simulations 
and experiments for DPPE are shown in Fig. \ref{HGorderParametersPE}.
Glycerol backbone order parameters in Slipids are off from experiments,
as already observed previously for PC lipids \cite{botan15}.
Order parameter signs for PE are not experimentally measured yet.
For headgroup the signs are set to give best agreement with simulations
and for glycerol to be consistent with experimental signs for PC.
Order parameter for $\beta$ carbon shows apparent agreement
with experiments. However, the sign of beta order parameter is positive,
in contrast to PC where negative sign was measured. Thus, the the beta order
parameter agrees with experiment with the assumption that its sign is opposite
than for PC. This is yet to be confirmed by experiments.
Order parameter for $\alpha$ carbon is too close to zero, even if the sign
would be correct.

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPE.pdf}
  \caption{\label{HGorderParametersPE}
    Order parameters for DPPE headgroup and glycerol
    backbone from simulations with Slipids \cite{??} and experiments
    (DPPE from \cite{seelig76} and E.coliPE from \cite{gally81}).
    Absolute values are shown, because signs are not known experimentally.
  }
  \todo{Experimental signs of the order parameters would be highly useful.} \\
\end{figure}

\todo{Berger results are not here yet, but we should mention the ring like structures
  pointed out by T. Piggot in the blog:
  The poor performance of headgroup order parameters in Berger model can be probably explained by ring like structures seen in Fig. 6 in Ref. \citenum{mukhopadhyay04}.
  These ring like structures are a widespread feature of typical Berger based lipid force fields containing explicit hydrogen atoms in the head group \cite{zhao08,henin09,dahlberg10}.
}

\subsection{PG headgroup and glycerol backbone}
Comparison between experiments and simulations for PG lipids is shown
in Fig. \ref{HGorderParametersPOPG}. The signs are not yet measured
experimentally. They are set to give the best argeement with experiments.
This would suggest that the $\beta$ order parameter would be positive,
in contrast to PC and PS headgroups, were negative signs were measured.
Even thought the signs turned out to be correct, the tested models would
not give a very good argeement with the experiments.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPG.pdf}
  \caption{\label{HGorderParametersPOPG}
    Order parameters for PG headgroup and glycerol backbone from simulations and experiments without CaCl$_2$ 
    (POPG from \cite{borle85} contains 10mM of PIPES, DPPG from \cite{wohlgemuth80} contains 10mM PIPES and 100mM CaCl,
    E.Coli PG results from \cite{gally81}).
    Signs are not known for experimental data. They are determined to give best agreement
    with simulations. This is not reliable and should be corrected when experimental
    data becomes available.
  }
  \todo{More simulation data for lipids with different headgroups to be collected} \\
  \todo{CHARMM GUI simulation contains only counter ions as potassium.
    All experiments here contain some amount of sodium salt. The best
    ion concentrations for comparison should be figured out.} \\
  \todo{Experimental signs of the order parameters would highly useful.}
\end{figure}

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
  \caption{\label{CIdensPG}
    Counterion densities from simulations with PG lipids.
  }
\end{figure}

\subsection{Lipid headgroup interactions in PC:PE and PC:PG mixtures}
The headgroup 
order parameters increase with the addition of negatively charged PS and PG lipids,
decrease when mixed with positively charged surfactants and
are less affected by the addition of zwitterionic PE lipids or cholesterol.
In addition to the results summarized in Fig. \ref{HGorderparametersPCvsPEPSPGchol},
also mixtures of PC with negatively charged PI, CL, PA, and zwitterionic SM
follow the eletrometer concept \cite{scherer87}.
to mixtures of differently charged
lipids are collected from different experiments in Fig. \ref{HGorderparametersPCvsPEPSPGchol}.
%\section{PC lipid headgroup response to different mixtures in experiments}
As shown in Fig. \ref{HGorderparametersPCvsPEPSPGchol}, order parameters of PC
headgroup behave in various lipid mixtures as expected from the electrometer concept \cite{seelig87, scherer87},
i.e., order parameters increase when anionic lipids are mixed with PC and decrease with cationic
surfactants. The changes with the addition of neutral lipids is significantly smaller.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPCvsPEPSPGchol.eps}
  \caption{\label{HGorderparametersPCvsPEPSPGchol}
    PC headgroup order parameters from experiments of mixtures with
    PE, PS, PG and cholesterol \cite{scherer87,scherer89,ferreira13}.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
\end{figure}

The headgroup order parameters for PC lipids (POPC and DOPC)
mixed with PE, PS and PG lipids are shown in Fig. \ref{HGorderparametersPCvsPEPSPG}
from different simulation model and experiments \cite{scherer87} with different
mole fractions. As already discussed previosly, the PC lipid headgroup behaviour
follows the electrometer concept in experiments when mixed with other lipids, i.e., the order
parameters increase when mixed with negatively charged lipids (PS, PI, CL, PA and PG)
remains almost unchaged when mixed with neutral lipids (PE and SM) \cite{scherer87}.
This is not the case in simulation data shown in Fig. \ref{HGorderparametersPCvsPEPSPG}.
The addition of DOPE into a POPC and DOPC bilayers significantly decreases the PC headgroup
order parameters in simulations with OPLS compatible version of the Berger force field \cite{tieleman06}
in contrast to experiments \cite{scherer87}. On the other hand, the increase of the PC
headgroup order parameters in CHARMM36 simulations mixed with PS and PG lipids is significantly
smaller than in experiments.

\todo{This is text by P. Fuchs, copied from the blog.}
Area results in nm$^2$, the error is <= 0.003 nm$^2$
- pure POPC
CHARMM36: 0.624
Berger : 0.649
- POPC/POPE 50:50
CHARMM36 : POPC 0.609, POPE 0.557
Berger-hacked: POPC 0.637, POPE 0.632
-----
One can see that CHARMM 36 predicts a drop in the area on going from pure POPC to POPC/POPE 50:50. This means that POPC pack tightly to POPE.
In contrast, the values for Berger are not that changed.
The POPE value predicted by CHARMM 36 (in the mixture POPC/POPE 50:50) is much smaller than that predicted by Berger.
\todo{The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
  which supports the more condensed PE bilayer. In principle, this is beyond the scope of this work (lipid headgroups), but we can
consider mentioning this.}

\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPSPG.eps}
  \caption{\label{HGorderparametersPCvsPEPSPG}
    PC headgroup order parameters from mixtures with PE, PS and PG
    lipids with various mole fractions from different simulation models and experiments \cite{scherer87}.
    Signs are determined as discussed in \cite{botan15,ollila16}.
  }
  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
\end{figure*}

%\subsection{Effect of PC on PS and PG headgroups}

The headgroup order parameters for PS and PG lipid mixtures with PC
having different mole fractions from simulations and experiments \cite{borle85,roux90}
are shown in Fig. \ref{HGorderparametersPSPGvsPC}. The effect of increasing
amount of PC to PS headgroup seems to qualitatively incorrect in CHARMM36 simulations.
The $\beta$-carbon order parameter increases in experiment, but decreases
in simulations with both tested counterions (Na+ and K+). Larger
$\alpha$-carbon order parameter decreases with the addition of PC
in experiment, while the lower remains unchanged. In simulations the
larger increases and the lower decreases.
Interestingly, the $\alpha$-carbon order parameters are closer to experiments
in pure PS system with K+ counterions than with Na+. 
The changes in PG headgroup order parameters are minor in simulations, which
is in line with the only available experiment for the $\beta$-carbon.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPC.eps}
  \caption{\label{HGorderparametersPSPGvsPC}
    PS and PG lipid headgroup order parameters from mixtures with PC
    lipids with various mole fractions from different simulation models and experiments \cite{borle85,roux90}.
    Signs are not yet known experimentally for PG, thus the signs give by simulations are used.
    Signs for PS are measures as described in SI.
  }
  \todo{Some simulations contain potassium as counterions, while some sodium.
    All experiments here contain some amount of sodium salt. The best
    ion concentrations for comparison should be figured out.} \\
  \todo{Why there is difference between CHARMM36 simulation results from POPS:POPC mixture and
  pure POPS? Discussion in https://github.com/NMRLipids/NMRlipidsIVotherHGs/issues/1}
\end{figure}

\subsection{Ca$^{2+}$ binding in bilayers with negatively charged PG and PS lipids}

PC lipid headgroup order parameters can used to measure ion binding
affinity, because their magnitude is linearly proportional
to the amount of bound charge in bilayer \cite{seelig87,catte16}.
This molecular electrometer concept can be used also
for bilayers containing PC lipids mixed with charged lipids \cite{borle85,macdonald87,roux90}.
This is demonstrated in Figs \ref{OrderParametersWithCaCl},
\ref{OrderParametersWithCaClBELOW1M} and \ref{OrderParameterCHANGESWithCaClBELOW1M},
showing order parameters for PC headgroup $\alpha$ and $\beta$ carbons
as a function of CaCl$_2$ concentration in the presence of different amounts of
negatively charged PS or PG lipids.
%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/LIPIDSwithCaCl.eps}
%  \caption{\label{OrderParametersWithCaCl}
%    PC headgroup order parameters as a function of CaCl concentration from experiments containing charged lipids.
%    Pure DPPC data from \cite{akutsu81}, pure POPC data from \cite{altenbach84}, 
%    POPC:POPS mixture data from \cite{roux90}, POPC:POPG mixture data with 0.1M NaCl from \cite{macdonald87}
%    and POPC:POPG mixture data without NaCl from \cite{borle85}.
%  }
%  \todo{Check the NaCl concentrations in the samples.}
%\end{figure}
%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/LIPIDSwithCaClBELOW1M.eps}
%  \caption{\label{OrderParametersWithCaClBELOW1M}
%    Figure \ref{OrderParametersWithCaCl} zoomed to smaller concentrations.
%  }
%\end{figure}
%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/CHANGESwithCaCl.eps}
%  \caption{\label{OrderParameterCHANGESWithCaClBELOW1M}
%    The change of PC headgroup order parameters
%    in the presence of different amount of negatively charged lipids
%    respect to the values without added CaCl$_2$.
%    The original data is the same as in Fig. \ref{OrderParametersWithCaCl}. 
%  }
%\end{figure}

PC headgroup order parameters increase when negatively charged
PS or PG are added to PC bilayer in the absense of added CaCl$_2$,
as expected based on electrometer concept \cite{seelig87}
(see Fig. \ref{OrderParametersWithCaClBELOW1M}).
%In electrometer concept this is explained by the tilting of
%headgroup more parallel to membrane normal \cite{??}.
Further, the order parameters decrease with the addition
of CaCl$_2$ and the decrease is more pronounced for systems with more
negatively charged lipids (see Fig. \ref{OrderParameterCHANGESWithCaClBELOW1M}).
At CaCl$_2$ concentrations ($\sim$ 50-300mM) where order parameters reach the values for pure PC,
the Ca2+ binding presumably fully cancels the charge from negative lipids and
overcharging occurs above these concenterations.
The interpretation of this data and some other results has been that \cite{seelig90}
\begin{displayquote}
  {\it ''(i) Ca$^{2+}$ binds to neutral lipids (phosphatidylcholine, phosphatidylethanolamine) and negatively charged lipids
    (phosphatidylglycerol) with approximately the same binding constant of K = 10-20 M$^{-1}$; \\
    (ii) the free Ca$^{2+}$
    concentration at the membrane interface is distinctly enhanced if the membrane carries a negative surface
    charge, either due to protein or to lipid; \\
    (iii) increased inter-facial Ca$^{2+}$ also means increased amounts
    of bound Ca$^{2+}$ at neutral and charged lipids; \\
    (iv) the actual binding step can be described by a Langmuir
    adsorption isotherm with a 1 lipid:1 Ca$^{2+}$ stoichiometry, provided the interfacial concentration C$_M$, is
    used to describe the chemical binding equilibrium.''}
\end{displayquote}

Comparison of Ca$2+$ binding in PG between CHARMM36 simulations and experiments \cite{borle85}
is shown in Fig. \ref{changesWITHCaClPG}. The decrease of $\alpha$ order parameter
is in agreement with experiments, while decerase of $\beta$ order parameter is overestimated.
The result is very similar to the results with PC in NMRlipids II publication \cite{catte16}.
It should be, however, noted that the $\beta$-order parameters are not actually measured for PG,
but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
Anyway, the data presented in NMRlipids II project and in Fig. \ref{changesWITHCaClPG} together
suggest that Calcium binding is similarly overestimated by CHARMM36 model in pure POPC bilayers and mixtures with
POPG. The good agreement of $\alpha$ carbon would be explained by too weak dependence of its order
parameter of bound charge
\todo{The response of CHARMM36 to cationic surfactant against experiments \cite{scherer89} to be checked.
  I have already ran the simulations, analysis to be done.}.
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/CHANGESwithCaClPG.eps}
  \caption{\label{changesWITHCaClPG}
    PG order parameters as a function CaCl$_2$ concentration from experiments \cite{borle85} and CHARMM36 simulations.
    Note that beta order parameter is calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}, not actually measured. 
  }
\end{figure}

Also dependence of $\beta$-carbon of PG on CaCl$_2$ concentration is compared with
experiments \cite{borle85} in Fig. \ref{PSPGchangesWITHCaCl}. Absolute value of
the order parameter is too large without ions, but rapid decrease due to addition of
CaCl$_2$ is observed in agreement with experiments for systems with 1:1 mixture of POPC and POPG.
In addition, absolute value in systems with CaCl$_2$ is in agreement with experiments.
However, system with 4:1 mixture of POPC and POPG behaves differently, but experimental
data is not available for comparison for this mixture.

\todo{More simulation data for systems with negatively charged lipids and CaCl$_2$ to be collected}

\section{Effect of Ca$2+$ binding on PG and PE headgroup}

Also the experimental order parameters for PS and PG headgroups
as a function of CaCl$_2$ concentration are shown in Fig. \ref{PSPGchangesWITHCaCl}.
\todo{These should be compared to simulations for potential structural interpretation of the changes.}
\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{../Figs/PGwithCaCl.eps}
  \caption{\label{PSPGchangesWITHCaCl}
    PG and PS order parameters a function CaCl$_2$ concentration taken from \cite{borle85} and \cite{roux90}, respectively.
  }
  \todo{Get the small concentration data from the inserts}
\end{figure}

\todo{There are mass density profiles available from CHARMM simulations, preliminary figures should be made:
http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1501446779291#c7311317114567929503}

\section{Conclusions}

% Tables may be be put in the text as floats.
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
%
% \begin{table}
% \caption{\label{} }
% \begin{tabular}{}
% \end{tabular}
% \end{table}

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
% Put your acknowledgments here.
\end{acknowledgments}
\newpage
\appendix
\begin{center}
{\bf SUPPLEMENTARY INFORMATION}
\end{center}

%\section{Measurements of order parameter sign}

%Fig. \ref{PShgSIGNS} summarizes the experimental results on the order parameter sign
%measurement for POPS sample. The experimental protocol is the same used in Ref. \citenum{ferreira16}.
%In (a) you see the headgroup region of the INEPT spectrum where alpha and beta are
%identified. In (b) you have the R-PDLF slices for alpha and beta where you see one single
%splitting for beta (which gives an order parameter equal to 0.12), and for alpha a superposition
%of a large splitting (order parameter equal to 0.09) and a very small splitting which cannot be
%calculated. On the bottom you have the S-DROSS slices of these two carbons. The grey lines show a
%random collection of slices from noise such that it gets clear what is significant. The S-DROSS
%slice for beta clearly shows that the order parameter is negative. The slice for alpha shows that
%the higher order parameter is positive and suggests that the smaller order parameter is negative
%(from the deviation towards negative values in the longer t1 times).
%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/PShgSIGNS.pdf}
%  \caption{\label{PShgSIGNS}
%    Experimental results for sign measurement for POPS sample
%  }
%\end{figure}

%The results updated with SIMPSON simulations for the SDROSS profiles
%are shown in Fig. \ref{PShgSIGNSsimpson}. The value for the smaller
%alpha order parameter is taken from Fig 3 in Ref. \citenum{roux91},
%because resolution in 13C NMR experiments was nor high enough to determine
%numerical value for this. The plots in Fig. \ref{PShgSIGNSsimpson} (c) show
%the following. The error bars and points are the experimental SDROSS data.
%The thick lines are SIMPSON simulations. The simulations were done by using
%the order parameter for beta equal to -0.12 and for alpha one order parameter
%equal to 0.09 and the other equal to -0.02 (black) or 0.02 (grey).
%Since the black lines agree with experimental data, we conlude that
%the order parameters for $\beta$ carbon are -0.12 and for $\alpha$
%order parameters are 0.09 and -0.02.
%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/PShgSIGNSsimpson.pdf}
%  \caption{\label{PShgSIGNSsimpson}
%    Experimental results for sign measurement for POPS sample
%  }
%\end{figure}

%\section{Dihedrals}
%\begin{figure*}[]
%  \centering
%  \includegraphics[width=8.0cm]{../Figs/dihed1.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed2.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed3.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed4.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed5.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed6.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed7.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed8.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed9.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed10.png}
%  \caption{\label{dihedrals}
%    Experimental results for sign measurement for POPS sample
%  }
%\end{figure*}

% Create the reference section using BibTe
\bibliography{refs.bib}

%\newpage
%\section{APPENDIX: The NMR results reported by Tiago Ferreira}

\listoftodos

\end{document}
%
% ****** End of file aiptemplate.tex ******

### 2019-04-12 Manuscript/manuscriptPGPEsuppl.tex
Update manuscript and figures

\documentclass[journal=jpcbfk]{achemso}

\usepackage[version=3]{mhchem}
\usepackage[T1]{fontenc}
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
\newcommand{\todo}[1]{\textcolor{red}{#1}}

\usepackage{rotating}
\usepackage{upgreek}				
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xr}
\externaldocument{manuscriptPGPE}
\usepackage{soul} % for highlights with \hl{}

\author{O. H. Samuli Ollila}
\email{samuli.ollila@helsinki.fi}
%\homepage[]{Your web page}
\affiliation{Institute of Organic Chemistry and Biochemistry,
Academy of Sciences of the Czech Republic, 
Prague 6, Czech Republic}
\affiliation{Institute of Biotechnology, University of Helsinki}

\SectionNumbersOn

\renewcommand{\thetable}{S\arabic{table}}%
\renewcommand{\thefigure}{S\arabic{figure}}%
\renewcommand{\thesection}{S\arabic{section}}%
\renewcommand{\thepage}{S\arabic{page}}%

\title{ Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids }

\begin{document}

\newpage
%\tableofcontents

\section{Simulated systems}

\subsection{CHARMM36}

{\it POPC:POPE mixtures} Data is available at \cite{POPCcharmm300K,POPC1POPE1charmm36}.
300 K with v-rescale (tau=0.1 ps),
1 bar with PR semiisotropic (tau=4 ps, compressibility=4.5e-5 bar$^{-1}$),
PME order 4 and space 0.12,
rcoulomb and rvdw 1.0,
128 lipids per leaflet,
no ion \todo{Full simulation details by Fuchs et al.}

{\it POPC:POPG mixture with additional calcium} \todo{Simulation details by J. Madsen.}

\subsection{CHARMM36ua}

{\it POPE} Data is available at \cite{charmm36uaPOPEfiles}. \todo{Simulation details by  T. Piggot.}

%\subsection{MacRog}

%\subsection{Lipid17}

\subsection{Slipids}
{\it POPE} Data is available at \cite{slipidsPOPEfiles}. \todo{Simulation details by  T. Piggot.}

{\it DPPE} Data is available at \cite{slipidsDPPEfiles}. \todo{Simulation details by F. Favela.}

{\it POPG} Data is available at \cite{slipidsPOPGfiles}. \todo{Simulation details by F. Favela.}

{\it DPPG} Data in 298~K is available at \cite{slipidsDPPGfilesT298K} and in 314~K at \cite{slipidsDPPGfiles}.  \todo{Simulation details by F. Favela.}

\subsection{Berger}

{\it POPE} Data is available at \cite{bergerPOPEfiles,berger2POPEfiles}. \todo{Simulation details by  T. Piggot.}

{\it DOPE} Data is available at \cite{bergerDOPEfiles,berger2DOPEfiles}. \todo{Simulation details by  T. Piggot.}

{\it POPC:POPE, POPC:DOPE and DOPC:DOPE mixtures} Data is available at \cite{POPCberger300K,POPC1POPE1berger}. 
300 K with v-rescale (tau=0.1 ps),
1 bar with PR semiisotropic (tau=4 ps, compressibility=4.5e-5 bar$^{-1}$),
PME order 4 and space 0.12,
rcoulomb and rvdw 1.0,
128 lipids per leaflet,
no ion \todo{Simulation details by Fuchs et al.}

\subsection{GROMOS 43A1-S3}

{\it POPE} Data is available at \cite{gromos43a1s3POPEfiles}. \todo{Simulation details by  T. Piggot.}

\subsection{OPLS-UA}

{\it POPE} Data is available at \cite{OPLSuaPOPEfiles}. \todo{Simulation details by  T. Piggot.}

{\it POPE with vdW interaction in H} Data is available at \cite{OPLSuaWvdWPOPEfiles}. \todo{Simulation details by  T. Piggot.}

\subsection{GROMOS-CKP and GROMOS-CKPM}

{\it POPE} Data is available at \cite{gromosCKPpope}. \todo{Simulation details by  T. Piggot.}

{\it DOPE} Data is available at \cite{gromosCKPdope}. \todo{Simulation details by  T. Piggot.}

{\it DPPE} Data is available at \cite{gromosCKPdppe}. \todo{Simulation details by  T. Piggot.}

\clearpage
\section{R-PDLF and SDROSS experiments}

\begin{figure}[]
%  \centering
  \includegraphics[width=\textwidth]{../Figs/POPEexperiment.pdf}
  \caption{\label{POPEspectra}
    (A) Chemical structure of POPE with the labeling of headgroup and glycerol backbone carbons.
    (B) INEPT spectra from POPE sample with the headgroup and glycerol backbone peaks labeled.
    (C) 2D R-PDLF spectra
    (D) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
    (E) Experimetal S-DROSS curves giving signs of the order parameters.
  }
  \todo{A, B etc. labels to be put in the figure.}
\end{figure}

\clearpage
\bibliography{refs.bib}

\end{document}

### 2019-04-17 Manuscript/manuscriptPGPE.tex
Add PG and PE data committed by Antonio Peon https://github.com/NMRLipids/MATCH/pull/59 and others, update figures and manuscript

@@
       POPE  & CHARMM36ua \cite{??}         & 336	& ?     & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      POPE  & CHARMM36 \cite{??}           & 144	& 5760  & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
       POPE  & Slipids \cite{jambeck12b,??} & 336	& ?     & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
@@
       % some footnotes are not visible in typeset-MS (pdf)
-      lipid/counter-ions & force field for lipids / ions & NaCl (mM) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
       \hline
+      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\
+      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
       POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ?
@@
      \hline
+      POPC                   & Slipid \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCslipid150mMNaCl301K}  \\
+      POPC:POPG (7:3)        & Slipid \cite{??}        & ?         & 0  & ?     & ?     & ?   &  310  & 500 & 100 & \cite{??} \todoi{Zenodo entry unclear.}  \\
+     \hline
       POPC             & Berger \cite{??} \todoi{This is probable not plain berger, correct force filed should be described.}  &0  & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
@@
       POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.}         &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
+     \hline
       DOPC             & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.}         &0          & 0  & 256 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPCberger300K} \\
       DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}        &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
+      \hline
+      POPG             & Slipids \cite{??} \todoi{Ion parameters?}       & 0.11          & 0  & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
+      POPG             & CHARMM36 \cite{??}        & 0.11          & 0  & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
+     \hline
+      POPE             & Slipids \cite{??} \todoi{Ion parameters?}        & 0.11          & 0  & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
+      POPE             & CHARMM36 \cite{??}        & 0.11          & 0  & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
     \end{tabular}
@@
 \subsection{PE headgroup and glycerol backbone}
-\todo{Add results from this discussion: \url{http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1502290977005#c4964787551006026647}}
+%\todo{Add results from this discussion: \url{http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1502290977005#c4964787551006026647}}
 Order parameters from Slipids simulations 
@@
   }
-  \todo{Experimental signs of the order parameters would be highly useful.} \\
+  \todo{Results from united atom simulations yet to be added} \\
 \end{figure}
@@
   }
-  \todo{More simulation data for lipids with different headgroups to be collected} \\
-  \todo{CHARMM GUI simulation contains only counter ions as potassium.
-    All experiments here contain some amount of sodium salt. The best
-    ion concentrations for comparison should be figured out.} \\
-  \todo{Experimental signs of the order parameters would highly useful.}
 \end{figure}
@@
 surfactants. The changes with the addition of neutral lipids is significantly smaller.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPCvsPEPSPGchol.eps}
-  \caption{\label{HGorderparametersPCvsPEPSPGchol}
-    PC headgroup order parameters from experiments of mixtures with
-    PE, PS, PG and cholesterol \cite{scherer87,scherer89,ferreira13}.
-    Signs are determined as discussed in \cite{botan15,ollila16}.
-  }
-\end{figure}
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPCvsPEPSPGchol.eps}
+%  \caption{\label{HGorderparametersPCvsPEPSPGchol}
+%    PC headgroup order parameters from experiments of mixtures with
+%    PE, PS, PG and cholesterol \cite{scherer87,scherer89,ferreira13}.
+%    Signs are determined as discussed in \cite{botan15,ollila16}.
+%  }
+%\end{figure}
 
@@
   \centering
-  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPSPG.eps}
+  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
   \caption{\label{HGorderparametersPCvsPEPSPG}
@@
   pure POPS? Discussion in https://github.com/NMRLipids/NMRlipidsIVotherHGs/issues/1}
+\end{figure}
+
+\subsection{Ion binding to PE lipid bilayers}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
+  \caption{\label{NAdensPE}
+    Sodium (solid line) and choride ion density profiles along membrane normal
+    from different simulations with PE lipids.
+  }
 \end{figure}
@@
   I have already ran the simulations, analysis to be done.}.
-\begin{figure}[]
+\begin{figure*}[]
   \centering
-  \includegraphics[width=9.0cm]{../Figs/CHANGESwithCaClPG.eps}
+  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG.eps}
   \caption{\label{changesWITHCaClPG}
@@
   }
-\end{figure}
+\end{figure*}
 
@@
 as a function of CaCl$_2$ concentration are shown in Fig. \ref{PSPGchangesWITHCaCl}.
-\todo{These should be compared to simulations for potential structural interpretation of the changes.}
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Figs/PGwithCaCl.eps}
-  \caption{\label{PSPGchangesWITHCaCl}
-    PG and PS order parameters a function CaCl$_2$ concentration taken from \cite{borle85} and \cite{roux90}, respectively.
-  }
-  \todo{Get the small concentration data from the inserts}
-\end{figure}
+%\todo{These should be compared to simulations for potential structural interpretation of the changes.}
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=9.0cm]{../Figs/PGwithCaCl.eps}
+%  \caption{\label{PSPGchangesWITHCaCl}
+%    PG and PS order parameters a function CaCl$_2$ concentration taken from \cite{borle85} and \cite{roux90}, respectively.
+%  }
+%  \todo{Get the small concentration data from the inserts}
+%\end{figure}
 

### 2019-04-17 Manuscript/manuscriptPGPE.tex
Add calcium density profiles

@@
 
-\todo{There are mass density profiles available from CHARMM simulations, preliminary figures should be made:
-http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1501446779291#c7311317114567929503}
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/CAdensPG.eps}
+  \caption{\label{CAdensPG}
+    Calcium ion density profiles along membrane normal
+    from different simulations with PG lipids.
+  }
+\end{figure}
 

### 2019-04-18 Manuscript/manuscriptPGPE.tex
Update text

@@
 
-\section{Molecular dynamics simulations}
+\subsection{Molecular dynamics simulations}
 
@@
 The simulated systems are listed in 
-Table \ref{systems} (pure PE and PG bilayers without additional ions).
-%and \ref{mixedIONsystems} (mixed PC:PS bilayers at various salt concentrations).
+Tables \ref{systems} (pure PE and PG bilayers without additional ions).
+and \ref{systemsII} (mixtures and systems with additional ions).
 Further simulation details are given in the SI, and
@@
     % The lipid force fields named as in our previous work~\cite{botan15}.
-  }\label{systems}
+  }\label{systemsII}
   \begin{minipage}[t]{\textwidth}
@@
 
+\clearpage
 \section{Results and Discussion}
@@
 The POPE experiments were recorded at 310~K, where the bilayer is in liquid disordered phase \cite{??}.
-\todo{To be checked by Tiago}.
+\todo{Details to be checked by Tiago}.
 
@@
 
-\subsection{PE headgroup and glycerol backbone}
-%\todo{Add results from this discussion: \url{http://nmrlipids.blogspot.com/2017/03/nmrlipids-iv-headgroup-glycerol.html?showComment=1502290977005#c4964787551006026647}}
-Order parameters from Slipids simulations 
-and experiments for DPPE are shown in Fig. \ref{HGorderParametersPE}.
-Glycerol backbone order parameters in Slipids are off from experiments,
-as already observed previously for PC lipids \cite{botan15}.
-Order parameter signs for PE are not experimentally measured yet.
-For headgroup the signs are set to give best agreement with simulations
-and for glycerol to be consistent with experimental signs for PC.
-Order parameter for $\beta$ carbon shows apparent agreement
-with experiments. However, the sign of beta order parameter is positive,
-in contrast to PC where negative sign was measured. Thus, the the beta order
-parameter agrees with experiment with the assumption that its sign is opposite
-than for PC. This is yet to be confirmed by experiments.
-Order parameter for $\alpha$ carbon is too close to zero, even if the sign
-would be correct.
+\subsection{Headgroup and glycerol backbone of POPE and POPG in MD simulations}
 
+The glycerol backbone order parameters in Slipid simulations are off from experiments
+for both PE and PG lipids (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
+as already reported previously for PC lipids \cite{botan15} and PS lipids \cite{NMRlipidsIVps}.
+The Slipid simulations were able to capture the essential differences between PC and PS lipid
+headgroups \cite{NMRlipidsIVps}, but this is not the case for PE and PG lipids.
  \begin{figure}[]
@@
   \caption{\label{HGorderParametersPE}
-    Order parameters for DPPE headgroup and glycerol
-    backbone from simulations with Slipids \cite{??} and experiments
-    (DPPE from \cite{seelig76} and E.coliPE from \cite{gally81}).
-    Absolute values are shown, because signs are not known experimentally.
+    The headgroup and glycerol backbone order parameters of PE lipids
+    from experiments (POPE and signs this work, DPPE from Ref.~\citenum{seelig76}
+    and E.coliPE from Ref.~\citenum{gally81}) and simulations with different force fields.
   }
-  \todo{Results from united atom simulations yet to be added} \\
+  \todo{Results from united atom GROMOS,CHARMM36ua,OPLS-UA,Berger and GROMOS-CKP simulations yet to be analyzed and added} \\
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPG.pdf}
+  \caption{\label{HGorderParametersPOPG}
+    The headgroup and glycerol backbone order parameters of PG lipids
+    from experiments (POPG and signs from this work and from Ref.~\citenum{borle85}, %contains 10mM of PIPES,
+    DPPG with 100mM NaCl from Ref.~\citenum{wohlgemuth80},% contains 10mM PIPES and,
+    and E.Coli PG results from Ref.~\citenum{gally81}).
+    and simulations with different force fields.
+  }
 \end{figure}
@@
 
-\subsection{PG headgroup and glycerol backbone}
-Comparison between experiments and simulations for PG lipids is shown
-in Fig. \ref{HGorderParametersPOPG}. The signs are not yet measured
-experimentally. They are set to give the best argeement with experiments.
-This would suggest that the $\beta$ order parameter would be positive,
-in contrast to PC and PS headgroups, were negative signs were measured.
-Even thought the signs turned out to be correct, the tested models would
-not give a very good argeement with the experiments.
-\begin{figure}[]
+CHARMM36 simulations seems to capture all the essential differences between
+headgroup order parameters of PC PE, PG (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
+as previously observed also for PS lipids \cite{NMRlipidsIVps}. Therefore,
+the CHARMM36 simulations can be used to analyze the structural differences between
+headgroups \todo{We should do such analysis. Maybe similar to the PS part \cite{NMRlipidsIVps}?}
+
+\subsection{Lipid headgroup interactions in PC:PE and PC:PG mixtures}
+According to the electrometer concept, the headgroup order parameters increase
+with the addition of negatively charged PG lipids, but
+%decrease when mixed with positively charged surfactants and
+are not affected by the addition of zwitterionic PE lipids (Fig.~\ref{HGorderparametersPCvsPEPG}) \cite{seelig87, scherer87}. % or cholesterol.
+%In addition to the results summarized in Fig. \ref{HGorderparametersPCvsPEPSPGchol},
+%also mixtures of PC with negatively charged PI, CL, PA, and zwitterionic SM
+%follow the eletrometer concept \cite{scherer87}.
+%to mixtures of differently charged
+%lipids are collected from different experiments in Fig. \ref{HGorderparametersPCvsPEPSPGchol}.
+%%\section{PC lipid headgroup response to different mixtures in experiments}
+%As shown in Fig. \ref{HGorderparametersPCvsPEPSPGchol}, order parameters of PC
+%headgroup behave in various lipid mixtures as expected from the electrometer concept \cite{seelig87, scherer87},
+%i.e., order parameters increase when anionic lipids are mixed with PC and decrease with cationic
+%surfactants. The changes with the addition of neutral lipids is significantly smaller.
+\begin{figure*}[]
   \centering
-  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPG.pdf}
-  \caption{\label{HGorderParametersPOPG}
-    Order parameters for PG headgroup and glycerol backbone from simulations and experiments without CaCl$_2$ 
-    (POPG from \cite{borle85} contains 10mM of PIPES, DPPG from \cite{wohlgemuth80} contains 10mM PIPES and 100mM CaCl,
-    E.Coli PG results from \cite{gally81}).
-    Signs are not known for experimental data. They are determined to give best agreement
-    with simulations. This is not reliable and should be corrected when experimental
-    data becomes available.
+  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
+  \caption{\label{HGorderparametersPCvsPEPG}
+    Modulation of PC headgroup order parameters with increasing amount of PE (left) and PG (right) in bilayer
+    from experiments \cite{scherer87,macdonald87} and simulations with different force fields.
+    Signs are determined as discussed in \cite{botan15,ollila16}.
   }
-\end{figure}
-
+  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
+\end{figure*}
+%The headgroup order parameters for PC lipids (POPC and DOPC)
+%mixed with PE, PS and PG lipids are shown in Fig. \ref{HGorderparametersPCvsPEPSPG}
+%from different simulation model and experiments \cite{scherer87} with different
+%mole fractions. As already discussed previosly, the PC lipid headgroup behaviour
+%follows the electrometer concept in experiments when mixed with other lipids, i.e., the order
+%parameters increase when mixed with negatively charged lipids (PS, PI, CL, PA and PG)
+%remains almost unchaged when mixed with neutral lipids (PE and SM) \cite{scherer87}.
+This is roughly observed also in simulations here, althouht the changes in PC headgroup with
+PE are slightly overestimated and changes with PG lipid are slightly underestimated.
+The underestimated increase of PC headgroup order parameters with the addition of
+negaticely charged lipids was previously interepreted to indicate overestimated
+counterion binding affinity, which overcompensates the effect of negative charge \cite{NMRlipidsIVps}.
+Thus, the counterion binding to PG lipid bilayers (Fig. \ref{CIdensPG}) may be slightly
+overestimated also in simulations here.
 \begin{figure}[]
@@
 
-\subsection{Lipid headgroup interactions in PC:PE and PC:PG mixtures}
-The headgroup 
-order parameters increase with the addition of negatively charged PS and PG lipids,
-decrease when mixed with positively charged surfactants and
-are less affected by the addition of zwitterionic PE lipids or cholesterol.
-In addition to the results summarized in Fig. \ref{HGorderparametersPCvsPEPSPGchol},
-also mixtures of PC with negatively charged PI, CL, PA, and zwitterionic SM
-follow the eletrometer concept \cite{scherer87}.
-to mixtures of differently charged
-lipids are collected from different experiments in Fig. \ref{HGorderparametersPCvsPEPSPGchol}.
-%\section{PC lipid headgroup response to different mixtures in experiments}
-As shown in Fig. \ref{HGorderparametersPCvsPEPSPGchol}, order parameters of PC
-headgroup behave in various lipid mixtures as expected from the electrometer concept \cite{seelig87, scherer87},
-i.e., order parameters increase when anionic lipids are mixed with PC and decrease with cationic
-surfactants. The changes with the addition of neutral lipids is significantly smaller.
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPCvsPEPSPGchol.eps}
-%  \caption{\label{HGorderparametersPCvsPEPSPGchol}
-%    PC headgroup order parameters from experiments of mixtures with
-%    PE, PS, PG and cholesterol \cite{scherer87,scherer89,ferreira13}.
-%    Signs are determined as discussed in \cite{botan15,ollila16}.
-%  }
-%\end{figure}
-
-The headgroup order parameters for PC lipids (POPC and DOPC)
-mixed with PE, PS and PG lipids are shown in Fig. \ref{HGorderparametersPCvsPEPSPG}
-from different simulation model and experiments \cite{scherer87} with different
-mole fractions. As already discussed previosly, the PC lipid headgroup behaviour
-follows the electrometer concept in experiments when mixed with other lipids, i.e., the order
-parameters increase when mixed with negatively charged lipids (PS, PI, CL, PA and PG)
-remains almost unchaged when mixed with neutral lipids (PE and SM) \cite{scherer87}.
-This is not the case in simulation data shown in Fig. \ref{HGorderparametersPCvsPEPSPG}.
-The addition of DOPE into a POPC and DOPC bilayers significantly decreases the PC headgroup
-order parameters in simulations with OPLS compatible version of the Berger force field \cite{tieleman06}
-in contrast to experiments \cite{scherer87}. On the other hand, the increase of the PC
-headgroup order parameters in CHARMM36 simulations mixed with PS and PG lipids is significantly
-smaller than in experiments.
-
-\todo{This is text by P. Fuchs, copied from the blog.}
-Area results in nm$^2$, the error is <= 0.003 nm$^2$
-- pure POPC
-CHARMM36: 0.624
-Berger : 0.649
-- POPC/POPE 50:50
-CHARMM36 : POPC 0.609, POPE 0.557
-Berger-hacked: POPC 0.637, POPE 0.632
------
+\todo{This is text by P. Fuchs, copied from the blog. \\
+Area results in nm$^2$, the error is $<=$ 0.003 nm$^2$ \\
+- pure POPC \\
+CHARMM36: 0.624 \\
+Berger : 0.649 \\
+- POPC/POPE 50:50 \\
+CHARMM36 : POPC 0.609, POPE 0.557 \\
+Berger-hacked: POPC 0.637, POPE 0.632 \\
+----- \\
 One can see that CHARMM 36 predicts a drop in the area on going from pure POPC to POPC/POPE 50:50. This means that POPC pack tightly to POPE. 
 In contrast, the values for Berger are not that changed. 
-The POPE value predicted by CHARMM 36 (in the mixture POPC/POPE 50:50) is much smaller than that predicted by Berger.
-\todo{The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
+The POPE value predicted by CHARMM 36 (in the mixture POPC/POPE 50:50) is much smaller than that predicted by Berger.\\
+-------\\
+The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
   which supports the more condensed PE bilayer. In principle, this is beyond the scope of this work (lipid headgroups), but we can
@@
 
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
-  \caption{\label{HGorderparametersPCvsPEPSPG}
-    PC headgroup order parameters from mixtures with PE, PS and PG
-    lipids with various mole fractions from different simulation models and experiments \cite{scherer87}.
-    Signs are determined as discussed in \cite{botan15,ollila16}.
-  }
-  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
-\end{figure*}
-
-%\subsection{Effect of PC on PS and PG headgroups}
-
-The headgroup order parameters for PS and PG lipid mixtures with PC
-having different mole fractions from simulations and experiments \cite{borle85,roux90}
-are shown in Fig. \ref{HGorderparametersPSPGvsPC}. The effect of increasing
-amount of PC to PS headgroup seems to qualitatively incorrect in CHARMM36 simulations.
-The $\beta$-carbon order parameter increases in experiment, but decreases
-in simulations with both tested counterions (Na+ and K+). Larger
-$\alpha$-carbon order parameter decreases with the addition of PC
-in experiment, while the lower remains unchanged. In simulations the
-larger increases and the lower decreases.
-Interestingly, the $\alpha$-carbon order parameters are closer to experiments
-in pure PS system with K+ counterions than with Na+. 
-The changes in PG headgroup order parameters are minor in simulations, which
-is in line with the only available experiment for the $\beta$-carbon.
+The order parameter of headgroup $\beta$-carbon of PG slightly increases with the addition
+of PC \cite{macdonald87}, althought also smaller changes have been observed~\cite{borle85}
+(Fig. \ref{HGorderparametersPSPGvsPC}).
+The increase in PG headgroup order parameter is not observed in simulations, which
+may arise from overestimated counterion binding overcompensating the changes in
+electrostatic environment as suggested for PC headgroup above. However, the changes
+are quite small and not fully consistent in experiments.
 \begin{figure}[]
@@
   \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPC.eps}
-  \caption{\label{HGorderparametersPSPGvsPC}
-    PS and PG lipid headgroup order parameters from mixtures with PC
-    lipids with various mole fractions from different simulation models and experiments \cite{borle85,roux90}.
-    Signs are not yet known experimentally for PG, thus the signs give by simulations are used.
-    Signs for PS are measures as described in SI.
-  }
-  \todo{Some simulations contain potassium as counterions, while some sodium.
-    All experiments here contain some amount of sodium salt. The best
-    ion concentrations for comparison should be figured out.} \\
-  \todo{Why there is difference between CHARMM36 simulation results from POPS:POPC mixture and
-  pure POPS? Discussion in https://github.com/NMRLipids/NMRlipidsIVotherHGs/issues/1}
-\end{figure}
-
-\subsection{Ion binding to PE lipid bilayers}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
-  \caption{\label{NAdensPE}
-    Sodium (solid line) and choride ion density profiles along membrane normal
-    from different simulations with PE lipids.
+  \caption{\label{HGorderparametersPGvsPC}
+    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
+    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
   }
@@
 
-\subsection{Ca$^{2+}$ binding in bilayers with negatively charged PG and PS lipids}
-
-PC lipid headgroup order parameters can used to measure ion binding
-affinity, because their magnitude is linearly proportional
-to the amount of bound charge in bilayer \cite{seelig87,catte16}.
-This molecular electrometer concept can be used also
-for bilayers containing PC lipids mixed with charged lipids \cite{borle85,macdonald87,roux90}.
-This is demonstrated in Figs \ref{OrderParametersWithCaCl},
-\ref{OrderParametersWithCaClBELOW1M} and \ref{OrderParameterCHANGESWithCaClBELOW1M},
-showing order parameters for PC headgroup $\alpha$ and $\beta$ carbons
-as a function of CaCl$_2$ concentration in the presence of different amounts of
-negatively charged PS or PG lipids.
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=9.0cm]{../Figs/LIPIDSwithCaCl.eps}
-%  \caption{\label{OrderParametersWithCaCl}
-%    PC headgroup order parameters as a function of CaCl concentration from experiments containing charged lipids.
-%    Pure DPPC data from \cite{akutsu81}, pure POPC data from \cite{altenbach84}, 
-%    POPC:POPS mixture data from \cite{roux90}, POPC:POPG mixture data with 0.1M NaCl from \cite{macdonald87}
-%    and POPC:POPG mixture data without NaCl from \cite{borle85}.
-%  }
-%  \todo{Check the NaCl concentrations in the samples.}
-%\end{figure}
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=9.0cm]{../Figs/LIPIDSwithCaClBELOW1M.eps}
-%  \caption{\label{OrderParametersWithCaClBELOW1M}
-%    Figure \ref{OrderParametersWithCaCl} zoomed to smaller concentrations.
-%  }
-%\end{figure}
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=9.0cm]{../Figs/CHANGESwithCaCl.eps}
-%  \caption{\label{OrderParameterCHANGESWithCaClBELOW1M}
-%    The change of PC headgroup order parameters
-%    in the presence of different amount of negatively charged lipids
-%    respect to the values without added CaCl$_2$.
-%    The original data is the same as in Fig. \ref{OrderParametersWithCaCl}. 
-%  }
-%\end{figure}
-
-PC headgroup order parameters increase when negatively charged
-PS or PG are added to PC bilayer in the absense of added CaCl$_2$,
-as expected based on electrometer concept \cite{seelig87}
-(see Fig. \ref{OrderParametersWithCaClBELOW1M}).
-%In electrometer concept this is explained by the tilting of
-%headgroup more parallel to membrane normal \cite{??}.
-Further, the order parameters decrease with the addition
-of CaCl$_2$ and the decrease is more pronounced for systems with more
-negatively charged lipids (see Fig. \ref{OrderParameterCHANGESWithCaClBELOW1M}).
-At CaCl$_2$ concentrations ($\sim$ 50-300mM) where order parameters reach the values for pure PC,
-the Ca2+ binding presumably fully cancels the charge from negative lipids and
-overcharging occurs above these concenterations.
-The interpretation of this data and some other results has been that \cite{seelig90}
+\clearpage
+\subsection{Cation binding to PE and PG lipid bilayers}
+The headgroup order parameters of PC lipids can used to measure ion binding
+affinity to lipid bilayers, because their magnitude is linearly proportional
+to the amount of bound charge in bilayer according to the molecular
+electrometer concept \cite{seelig87,catte16}. The molecular electrometer concept
+can be used also for bilayers containing PC lipids mixed with charged
+lipids \cite{borle85,macdonald87,roux90,NMRlipidsIVps}.
+%This is demonstrated in Figs \ref{OrderParametersWithCaCl},
+%\ref{OrderParametersWithCaClBELOW1M} and \ref{OrderParameterCHANGESWithCaClBELOW1M},
+%showing order parameters for PC headgroup $\alpha$ and $\beta$ carbons
+%as a function of CaCl$_2$ concentration in the presence of different amounts of
+%negatively charged PS or PG lipids.
+%
+%PC headgroup order parameters increase when negatively charged
+%PS or PG are added to PC bilayer in the absense of added CaCl$_2$,
+%as expected based on electrometer concept \cite{seelig87}
+%(see Fig. \ref{OrderParametersWithCaClBELOW1M}).
+%Further, the order parameters decrease with the addition
+%of CaCl$_2$ and the decrease is more pronounced for systems with more
+%negatively charged lipids (see Fig. \ref{OrderParameterCHANGESWithCaClBELOW1M}).
+%At CaCl$_2$ concentrations ($\sim$ 50-300mM) where order parameters reach the values for pure PC,
+%the Ca2+ binding presumably fully cancels the charge from negative lipids and
+%overcharging occurs above these concenterations.
+Based on the electrometer concept and other data is has been suggested that \cite{seelig90}
 \begin{displayquote}
@@
 \end{displayquote}
+The electrometer concept has been very useful in evaluating ion binding affinity in simulations
+against experiments, because the headgroup order parameter changes as a function of ion concentration
+can be directly compared with experiments \cite{catte16,melcr18,NMRlipidsIVps}.
 
-Comparison of Ca$2+$ binding in PG between CHARMM36 simulations and experiments \cite{borle85}
-is shown in Fig. \ref{changesWITHCaClPG}. The decrease of $\alpha$ order parameter
-is in agreement with experiments, while decerase of $\beta$ order parameter is overestimated.
-The result is very similar to the results with PC in NMRlipids II publication \cite{catte16}.
-It should be, however, noted that the $\beta$-order parameters are not actually measured for PG,
+Sodium binding to PC lipid bilayers is significantly overestimated by most simulations models \cite{catte16}.
+Also, sodium binding to PS lipids seems to be overestimated, although the
+presence of counterions complicate the comparison for negatively charged lipids \cite{NMRlipidsIVps}.
+Sodium binding to PE lipids is weaker than to PC lipids in previous work \cite{catte16} in both CHARMM36 and Slipids simulations (Fig. \ref{NAdensPE}).
+This cannot be evaluated using electrometer concept because experimental data with PE lipids is not
+available \todo{Other experimental data should be checked and discussed}.
+The difference in Slipids may also be due to different ion model used \todo{The used ion model here should be checked}.
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
+  \caption{\label{NAdensPE}
+    Sodium (solid line) and choride ion density profiles along membrane normal
+    from different simulations with PE lipids.
+  }
+\end{figure}
+
+Calcium binding affinity to PC and PS lipid bilayers was not correctly described by
+any of the standard MD simulation forced fields \cite{catte16,NMRlipidsIVps}, while
+recently introduced force field with electronic continuum correction (ECC) performed better \cite{melcr18}.
+The decrease of $\alpha$-carbon order parameter of PC lipids in PC:PG mixtures as a function
+of calcium concentration is close to experiments CHARMM36 simulations (Fig. \ref{changesWITHCaClPG}),
+but the decrease of $\beta$-carbon order parameter seems to be overestimated.
+However, the $\beta$-carbon order parameter was not actually measured from these samples,
 but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
-Anyway, the data presented in NMRlipids II project and in Fig. \ref{changesWITHCaClPG} together
-suggest that Calcium binding is similarly overestimated by CHARMM36 model in pure POPC bilayers and mixtures with
-POPG. The good agreement of $\alpha$ carbon would be explained by too weak dependence of its order
-parameter of bound charge
-\todo{The response of CHARMM36 to cationic surfactant against experiments \cite{scherer89} to be checked.
-  I have already ran the simulations, analysis to be done.}.
+The result is similar to the $\sim$200~ns simulations with PC lipids in previous work \cite{catte16}.
+However, when simulation was continued for $\mu$s, the binding affinity substantially increased
+and interpretation was that calcium overbinds to PC lipid in CHARMM36. Therefore, the
+conclusion seems to be similar here, although the new NBfix parameters may complicate
+the situation \todo{The status of NBfix parameters in these simulations should be checked.}.
+%Anyway, the data presented in NMRlipids II project and in Fig. \ref{changesWITHCaClPG} together
+%suggest that Calcium binding is similarly overestimated by CHARMM36 model in pure POPC bilayers and mixtures with
+%POPG. The good agreement of $\alpha$ carbon would be explained by too weak dependence of its order
+%parameter of bound charge
+%\todo{The response of CHARMM36 to cationic surfactant against experiments \cite{scherer89} to be checked.
+%  I have already ran the simulations, analysis to be done.}.
 \begin{figure*}[]
@@
   \caption{\label{changesWITHCaClPG}
-    PG order parameters as a function CaCl$_2$ concentration from experiments \cite{borle85} and CHARMM36 simulations.
+    (left) The headgroup order parameters of PC from PC:PG mixtures as a function CaCl$_2$
+    concentration from experiments \cite{borle85,macdonald87} and CHARMM36 simulations.
     Note that beta order parameter is calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}, not actually measured.
+    (right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
+    concentration from experiments \cite{borle85} and CHARMM36 simulations.
   }
@@
 
-Also dependence of $\beta$-carbon of PG on CaCl$_2$ concentration is compared with
-experiments \cite{borle85} in Fig. \ref{PSPGchangesWITHCaCl}. Absolute value of
-the order parameter is too large without ions, but rapid decrease due to addition of
-CaCl$_2$ is observed in agreement with experiments for systems with 1:1 mixture of POPC and POPG.
-In addition, absolute value in systems with CaCl$_2$ is in agreement with experiments.
-However, system with 4:1 mixture of POPC and POPG behaves differently, but experimental
-data is not available for comparison for this mixture.
-
-\todo{More simulation data for systems with negatively charged lipids and CaCl$_2$ to be collected}
-
-\section{Effect of Ca$2+$ binding on PG and PE headgroup}
-
-Also the experimental order parameters for PS and PG headgroups
-as a function of CaCl$_2$ concentration are shown in Fig. \ref{PSPGchangesWITHCaCl}.
-%\todo{These should be compared to simulations for potential structural interpretation of the changes.}
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=9.0cm]{../Figs/PGwithCaCl.eps}
-%  \caption{\label{PSPGchangesWITHCaCl}
-%    PG and PS order parameters a function CaCl$_2$ concentration taken from \cite{borle85} and \cite{roux90}, respectively.
-%  }
-%  \todo{Get the small concentration data from the inserts}
-%\end{figure}
+The $\beta$-carbon order parameter of PG exhibits a rapid decrease with small CaCl$_2$ concentrations
+and a more modest decrease with larger concentrations in experiments \cite{borle85} (Fig. \ref{PSPGchangesWITHCaCl}).
+The rapid decrease with CaCl$_2$ is observed but overestimated in CHARMM36 simulation
+with POPC:POPG 1:1 mixture, but not in 4:1 mixture \todo{This is little bit weird, should be checked.}.
 
@@
 \end{figure}
+
+\todo{We need PC:PG simulations with CaCl$_2$ from different force fields to finish the discussion.}
 

### 2019-04-18 Manuscript/manuscriptPGPEsuppl.tex
Update text

@@
 
-{\it POPC:POPE mixtures} Data is available at \cite{POPCcharmm300K,POPC1POPE1charmm36}.
+\noindent {\it POPE} \todo{Simulation details by M. Javanainen}.
+
+\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.}
+
+\noindent {\it POPG} \todo{Simulation details by Ollila}.
+
+\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.}
+
+\noindent {\it POPC:POPE mixtures} Data is available at \cite{POPCcharmm300K,POPC1POPE1charmm36}.
 300 K with v-rescale (tau=0.1 ps),
@@
 
-{\it POPC:POPG mixture with additional calcium} \todo{Simulation details by J. Madsen.}
+\noindent {\it POPC:POPG mixture with additional calcium} \todo{Simulation details by J. Madsen.}
+
+\noindent {\it POPC:POPG mixture with additional NaCl} \todo{Simulation details by A. Peon.}
 
@@
 
-{\it POPE} Data is available at \cite{charmm36uaPOPEfiles}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE} Data is available at \cite{charmm36uaPOPEfiles}. \todo{Simulation details by  T. Piggot.}
 
@@
 \subsection{Slipids}
-{\it POPE} Data is available at \cite{slipidsPOPEfiles}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE} Data is available at \cite{slipidsPOPEfiles}. \todo{Simulation details by  T. Piggot.}
 
-{\it DPPE} Data is available at \cite{slipidsDPPEfiles}. \todo{Simulation details by F. Favela.}
+\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.}
 
-{\it POPG} Data is available at \cite{slipidsPOPGfiles}. \todo{Simulation details by F. Favela.}
+\noindent {\it DPPE} Data is available at \cite{slipidsDPPEfiles}. \todo{Simulation details by F. Favela.}
 
-{\it DPPG} Data in 298~K is available at \cite{slipidsDPPGfilesT298K} and in 314~K at \cite{slipidsDPPGfiles}.  \todo{Simulation details by F. Favela.}
+\noindent {\it POPG} Data is available at \cite{slipidsPOPGfiles}. \todo{Simulation details by F. Favela.}
+
+\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.}
+
+\noindent {\it DPPG} Data in 298~K is available at \cite{slipidsDPPGfilesT298K} and in 314~K at \cite{slipidsDPPGfiles}.  \todo{Simulation details by F. Favela.}
+
+\noindent {\it POPC:POPG mixture with additional NaCl} \todo{Simulation details by A. Peon.}
 
@@
 
-{\it POPE} Data is available at \cite{bergerPOPEfiles,berger2POPEfiles}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE} Data is available at \cite{bergerPOPEfiles,berger2POPEfiles}. \todo{Simulation details by  T. Piggot.}
 
-{\it DOPE} Data is available at \cite{bergerDOPEfiles,berger2DOPEfiles}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it DOPE} Data is available at \cite{bergerDOPEfiles,berger2DOPEfiles}. \todo{Simulation details by  T. Piggot.}
 
-{\it POPC:POPE, POPC:DOPE and DOPC:DOPE mixtures} Data is available at \cite{POPCberger300K,POPC1POPE1berger}. 
+\noindent {\it POPC:POPE, POPC:DOPE and DOPC:DOPE mixtures} Data is available at \cite{POPCberger300K,POPC1POPE1berger}. 
 300 K with v-rescale (tau=0.1 ps),
@@
 
-{\it POPE} Data is available at \cite{gromos43a1s3POPEfiles}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE} Data is available at \cite{gromos43a1s3POPEfiles}. \todo{Simulation details by  T. Piggot.}
 
@@
 
-{\it POPE} Data is available at \cite{OPLSuaPOPEfiles}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE} Data is available at \cite{OPLSuaPOPEfiles}. \todo{Simulation details by  T. Piggot.}
 
-{\it POPE with vdW interaction in H} Data is available at \cite{OPLSuaWvdWPOPEfiles}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE with vdW interaction in H} Data is available at \cite{OPLSuaWvdWPOPEfiles}. \todo{Simulation details by  T. Piggot.}
 
@@
 
-{\it POPE} Data is available at \cite{gromosCKPpope}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE} Data is available at \cite{gromosCKPpope}. \todo{Simulation details by  T. Piggot.}
 
-{\it DOPE} Data is available at \cite{gromosCKPdope}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it DOPE} Data is available at \cite{gromosCKPdope}. \todo{Simulation details by  T. Piggot.}
 
-{\it DPPE} Data is available at \cite{gromosCKPdppe}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it DPPE} Data is available at \cite{gromosCKPdppe}. \todo{Simulation details by  T. Piggot.}
 

### 2019-08-27 Manuscript/manuscriptPGPE.tex
Add LIPID17 POPE anad POPG data by A. Peon https://github.com/NMRLipids/MATCH/pull/61
Add data by Antonio Peon from https://github.com/NMRLipids/MATCH/pull/77/files, except the data with some unclarities: https://github.com/NMRLipids/MATCH/issues/80, https://github.com/NMRLipids/MATCH/issues/79 and https://github.com/NMRLipids/MATCH/issues/78

@@
   \centering
-  \caption{List of single lipid type MD simulations without additional ions.
+  \caption{List of MD simulations with PE lipids.
     % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
@@
       % some footnotes are not visible in typeset-MS (pdf)
-      lipid/counter-ions & force field for lipids / ions & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$    & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      lipid/counter-ions & force field for lipids / ions  & NaCl (M) & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
       \hline
-      DPPE  & Slipids \cite{jambeck12b}    & 288 	& 9386  & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
-      DPPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      POPE  & CHARMM36 \cite{??}           &0       & 144	& 5760  &0    & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
+      POPE  & CHARMM36 \cite{??}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
+      POPE  & CHARMM36 \cite{??}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
+      POPE  & CHARMM36ua \cite{??}         &0       & 336	& ?
+&0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
       \hline
-      POPE  & GROMOS 43A1-S3 \cite{??}     & 128	& ?     & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      POPE  & CHARMM36ua \cite{??}         & 336	& ?     & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      POPE  & CHARMM36 \cite{??}           & 144	& 5760  & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
-      POPE  & Slipids \cite{jambeck12b,??} & 336	& ?     & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
-      POPE  & OPLS-UA vdW on H \cite{??}   & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      POPE  & OPLS-UA \cite{??}            & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      POPE  & Berger-based \cite{??}       & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      POPE  & Berger-based2 \cite{??}      & 128	& ?     & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      POPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      DPPE  & Slipids \cite{jambeck12b}    &0    & 288 	& 9386  &0    & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
+      POPE  & Slipids \cite{jambeck12b,??} &0    & 336	& ?     &0    & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
+      POPE  & Slipids \cite{??} \todoi{Ion parameters?}        & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
+      POPE  & Slipids \cite{??} \todoi{Ion parameters?}        & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
       \hline
-      DOPE  & Berger-based \cite{??}       & 128	& ?     & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      DOPE  & Berger-based2 \cite{??}      & 128	& ?     & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      DOPE  & GROMOS-CKP    \cite{??}      & 128	& ?     & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      DPPE  & GROMOS-CKP    \cite{??}      &0    & 128	& ?     &0    & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& ?     &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
+      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& ?     &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
       \hline
-      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}   & 118& 4110  & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
-      POPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 10664  & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
+      POPE  & GROMOS 43A1-S3 \cite{??}     &0    & 128	& ?
+&0    & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
       \hline
-      DPPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 11232  & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
-      DPPG/Na$^+$  & Slipids \cite{jambeck13}   & 288 	& 11232  & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
+      POPE  & OPLS-UA vdW on H \cite{??}   &0    & 128	& ?
+&0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      POPE  & OPLS-UA \cite{??}            &0    & 128	& ?
+&0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      \hline
+      POPE  & Berger-based \cite{??}       &0    & 128	& ?
+&0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      POPE  & Berger-based2 \cite{??}      &0    & 128	& ?
+&0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      DOPE  & Berger-based \cite{??}       &0    & 128	& ?
+&0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      DOPE  & Berger-based2 \cite{??}      &0    & 128	& ?
+&0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      \hline
+%      POPE             & LIPID17 \cite{??} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
+      POPE             & LIPID17 \cite{??} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
     \end{tabular}
@@
   \centering
-  \caption{List of MD simulations with lipid mixtures and additional ions.
+  \caption{List of MD simulations with PG lipids.
+    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
+    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
+    % The lipid force fields named as in our previous work~\cite{botan15}.
+  }\label{systemsII}
+  \begin{minipage}[t]{\textwidth}
+    \begin{tabular}{l c c r r r r r r c c}
+      %\hline
+      % some footnotes are not visible in typeset-MS (pdf)
+      lipid/counter-ions & force field for lipids / ions & NaCl (M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      \hline
+      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}    &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
+      POPG             & CHARMM36 \cite{??}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
+      POPG             & CHARMM36 \cite{??}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
+      \hline
+      POPG/Na$^+$  & Slipids \cite{jambeck13}    &0         & 288 	& 10664   &0     & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
+      DPPG/Na$^+$  & Slipids \cite{jambeck13}    &0         & 288 	& 11232  &0     & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
+      DPPG/Na$^+$  & Slipids \cite{jambeck13}    &0         & 288 	& 11232   &0     & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
+      POPG         & Slipids \cite{??} \todoi{Ion parameters?}       & 0    & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGslipids} \\
+      POPG         & Slipids \cite{??} \todoi{Ion parameters?}       & 0.11    & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
+      \hline
+      POPG             & LIPID17 \cite{??}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
+      POPG             & LIPID17 \cite{??}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
+      \hline
+      POPG             & GROMOS-CKP \cite{??}         & 0              & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGgromosCKP} \\
+      POPG             & GROMOS-CKP \cite{??}         & 0.11           & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGgromosCKP150mMNaCl} \\
+    \end{tabular}
+  \end{minipage}
+  %\end{sidewaystable*} 
+\end{table*}
+
+\begin{table*}[htb]
+  %\begin{sidewaystable*}[!p]
+  \centering
+  \caption{List of MD simulations with PE and PG lipids mixed with PC.
     % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
@@
       DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}        &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
-      \hline
-      POPG             & Slipids \cite{??} \todoi{Ion parameters?}       & 0.11          & 0  & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
-      POPG             & CHARMM36 \cite{??}        & 0.11          & 0  & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
-     \hline
-      POPE             & Slipids \cite{??} \todoi{Ion parameters?}        & 0.11          & 0  & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
-      POPE             & CHARMM36 \cite{??}        & 0.11          & 0  & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
     \end{tabular}

### 2019-08-30 Manuscript/manuscriptPGPE.tex
Add results from united atom models calculated with the code by Fuchs et al.

@@
 
-Absolute values of experimental order parameters for different lipid headgroups are
-collected in Fig. \ref{HGorderParameters}. Signs are measured only for PC as far as I know,
-thus only absolute values are used for now.
+Experimental order parameters for different lipid headgroups are
+collected in Fig. \ref{HGorderParameters}.
 
@@
 the standard error of the mean were then calculated over different lipids.
-Python program ({\tt scripts/calcOrderParameters.py}) that uses the
-MDAnalysis library \cite{agrawal11,gowers16} is available in Ref. \citenum{MATCHgit}. 
+Python programs that use the MDAnalysis library \cite{agrawal11,gowers16}
+used for all atom simulations is available in Ref. \citenum{MATCHgit}
+({\tt scripts/calcOrderParameters.py}). For united atom simulations, the trajectories
+with hydrogens having ideal geometry were constructed first using either buildH program~\cite{buildH}
+or ({\tt scratch/opAAUA\_prod.py}) in  Ref. \citenum{MATCHgit}, and the order parameters were
+then calculated from these trajectories. This approach has been tested against trajectories
+with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot18}.
 The ion number density profiles were calculated using the {\tt gmx density} tool
@@
       POPE  & CHARMM36 \cite{??}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
-      POPE  & CHARMM36ua \cite{??}         &0       & 336	& ?
-&0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      POPE  & CHARMM36ua \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
       \hline
@@
       POPE  & Slipids \cite{jambeck12b,??} &0    & 336	& ?     &0    & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
-      POPE  & Slipids \cite{??} \todoi{Ion parameters?}        & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
+      POPE  & Slipids \cite{??}            & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
       POPE  & Slipids \cite{??} \todoi{Ion parameters?}        & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
       \hline
-      DPPE  & GROMOS-CKP    \cite{??}      &0    & 128	& ?     &0    & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& ?     &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      DPPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3655  &0    & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \\
+      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
       POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
-      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& ?     &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
       \hline
-      POPE  & GROMOS 43A1-S3 \cite{??}     &0    & 128	& ?
-&0    & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      POPE  & GROMOS 43A1-S3 \cite{??}     &0    & 128	& 3552     &0    & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles}  \\
       \hline
-      POPE  & OPLS-UA vdW on H \cite{??}   &0    & 128	& ?
-&0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      POPE  & OPLS-UA \cite{??}            &0    & 128	& ?
-&0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      POPE  & OPLS-UA vdW on H \cite{??}   &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \\
+      POPE  & OPLS-UA \cite{??}            &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \\
       \hline
-      POPE  & Berger-based \cite{??}       &0    & 128	& ?
-&0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      POPE  & Berger-based2 \cite{??}      &0    & 128	& ?
-&0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      DOPE  & Berger-based \cite{??}       &0    & 128	& ?
-&0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
-      DOPE  & Berger-based2 \cite{??}      &0    & 128	& ?
-&0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \todoi{Not analyzed yet, waiting for the code for UA simulations.}  \\
+      POPE  & Berger-Vries \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
+      POPE  & Berger-largeH \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
+      DOPE  & Berger-Vries \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
+      DOPE  & Berger-largeH \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
       \hline
-%      POPE             & LIPID17 \cite{??} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
+      POPE             & LIPID17 \cite{??} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
       POPE             & LIPID17 \cite{??} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
@@
       POPG             & GROMOS-CKP \cite{??}         & 0              & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGgromosCKP} \\
-      POPG             & GROMOS-CKP \cite{??}         & 0.11           & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGgromosCKP150mMNaCl} \\
+      POPG             & GROMOS-CKP \cite{??}         & 0.11           & 500 & 25000 & 49 &  310  & 500 & 100 & \cite{POPGgromosCKP150mMNaCl} \\
     \end{tabular}
@@
   }
-  \todo{Results from united atom GROMOS,CHARMM36ua,OPLS-UA,Berger and GROMOS-CKP simulations yet to be analyzed and added} \\
+  \todo{This should be clarified as in NMRlipidsI and error bars should be added.
+    Probably larger error bars for united atom models based on the report by Fuchs et al.
+  }
 \end{figure}
@@
 the CHARMM36 simulations can be used to analyze the structural differences between
-headgroups \todo{We should do such analysis. Maybe similar to the PS part \cite{NMRlipidsIVps}?}
+headgroups (Fig.~\ref{structures}).
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=9.0cm]{../Figs/struct.png}
+  \caption{\label{structures}
+    Overlayed snapshots from CHARMM36 simulations of different lipids which give the best agreement with experiments.
+  }
+\end{figure}
 

### 2019-10-04 Manuscript/manuscriptPGPE.tex
Add todos
Update figures and text

@@
       POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
+      POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
       DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
@@
   %\end{sidewaystable*}
+  \todo{Data for POPC:POPG mixtures by listed by Antonio Peon is missing from this table}
 \end{table*}
@@
 \subsection{Headgroup and glycerol backbone of POPE and POPG in MD simulations}
+As reported previously for PC \cite{botan15} and PS \cite{NMRlipidsIVps} lipids,
+the headgroup and glycerol backbone order parameters show wide variation between different force fields
+for both PE and PG lipids (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
+and none of the force fields reproduce all values within experimental error bars.
+%The Slipid simulations were able to capture the essential differences between PC and PS lipid
+%headgroups \cite{NMRlipidsIVps}, but this is not the case for PE and PG lipids.
+%\todo{Berger results are not here yet, but we should mention the ring like structures
+%  pointed out by T. Piggot in the blog:
+The poor performance of headgroup order parameters in Berger model can be probably explained by ring like structures seen in Fig. 6 in Ref. \citenum{mukhopadhyay04},
+which is a typical feature for Berger based lipid force fields containing explicit hydrogen atoms in the head group \cite{zhao08,henin09,dahlberg10}.
+%}
 
-The glycerol backbone order parameters in Slipid simulations are off from experiments
-for both PE and PG lipids (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
-as already reported previously for PC lipids \cite{botan15} and PS lipids \cite{NMRlipidsIVps}.
-The Slipid simulations were able to capture the essential differences between PC and PS lipid
-headgroups \cite{NMRlipidsIVps}, but this is not the case for PE and PG lipids.
 \begin{figure}[]
@@
 \end{figure}
-
-\todo{Berger results are not here yet, but we should mention the ring like structures
-  pointed out by T. Piggot in the blog:
-  The poor performance of headgroup order parameters in Berger model can be probably explained by ring like structures seen in Fig. 6 in Ref. \citenum{mukhopadhyay04}.
-  These ring like structures are a widespread feature of typical Berger based lipid force fields containing explicit hydrogen atoms in the head group \cite{zhao08,henin09,dahlberg10}.
-}
-
-CHARMM36 simulations seems to capture all the essential differences between
-headgroup order parameters of PC PE, PG (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
-as previously observed also for PS lipids \cite{NMRlipidsIVps}. Therefore,
-the CHARMM36 simulations can be used to analyze the structural differences between
-headgroups (Fig.~\ref{structures}).
-\begin{figure}[!h]
+\begin{figure*}[!h]
   \centering
-  \includegraphics[width=9.0cm]{../Figs/struct.png}
+  \includegraphics[width=18.0cm]{../Figs/struct.png}
   \caption{\label{structures}
@@
   }
-\end{figure}
+\end{figure*}
+Also in line with previous studies for PC \cite{botan15} and PS \cite{NMRlipidsIVps} lipids,
+CHARMM36 simulations are in best agreement with experiments for headgroup and glycerol backbone
+order parameters, and seem to capture all the essential differences between different headroups
+(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
+%headgroup order parameters of PC PE, PG (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
+%as previously observed also for PS lipids \cite{NMRlipidsIVps}.
+In previous study, CHARMM36 predicted the more negative $\beta$-carbon order parameter and larger forking of
+the $\alpha$-carbon is PS headgroup than in PC \cite{NMRlipidsIVps}.
+In this work, the CHARMM36 simulations reproduce also the other essential differences in experimental
+headgroup order parameters between different headgroups (Fig. \ref{HGorderParameters}) despite the inaccuracies
+in individual segments:
+The PE headgroup order parameters in CHARMM36 simulations are similar to the PC \cite{botan15}
+and $\beta$-carbon order parameters is positive in PG headgroup, in contrast to the negative values observed in
+other lipids. Therefore, we use CHARMM36 simulations to analyze the structural differences between
+headgroups (Fig.~\ref{structures}). While rotation around N-C$_\alpha$-C$_\beta$-O$_\alpha$ dihedral was
+significantly different in PS and PC headgroups in previous work \cite{NMRlipidsIVps}, essential differences
+between PC, PE and PG are not observed here (Fig.~\ref{structures}) \todo{We need also the dihedral distributions to finish this discussion.}. \\
+\todo{Why is the $\beta$-carbon order parameter of PG different to PC and PE then?}
 
@@
   }
-  \todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
+  \todo{Data for CHARMM, Lipid17, Slipids, and also GROMOS also without ions is now available.
+    This data should be updated into this plot}
+  %\todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
 \end{figure*}
@@
   \caption{\label{CIdensPG}
-    Counterion densities from simulations with PG lipids.
+    Counterion densities and area per lipids from simulations with PG lipids.
+    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
   }
+  \todo{Data for GROMOS is also available without ions, but I could not analyze it, probably because tpr file was not correct.}
 \end{figure}

### 2019-10-07 Manuscript/manuscriptPGPE.tex
Update PCvsPG results
Update text and figures

@@
 
-\subsection{Lipid headgroup interactions in PC:PE and PC:PG mixtures}
-According to the electrometer concept, the headgroup order parameters increase
-with the addition of negatively charged PG lipids, but
-%decrease when mixed with positively charged surfactants and
-are not affected by the addition of zwitterionic PE lipids (Fig.~\ref{HGorderparametersPCvsPEPG}) \cite{seelig87, scherer87}. % or cholesterol.
-%In addition to the results summarized in Fig. \ref{HGorderparametersPCvsPEPSPGchol},
-%also mixtures of PC with negatively charged PI, CL, PA, and zwitterionic SM
-%follow the eletrometer concept \cite{scherer87}.
-%to mixtures of differently charged
-%lipids are collected from different experiments in Fig. \ref{HGorderparametersPCvsPEPSPGchol}.
-%%\section{PC lipid headgroup response to different mixtures in experiments}
-%As shown in Fig. \ref{HGorderparametersPCvsPEPSPGchol}, order parameters of PC
-%headgroup behave in various lipid mixtures as expected from the electrometer concept \cite{seelig87, scherer87},
-%i.e., order parameters increase when anionic lipids are mixed with PC and decrease with cationic
-%surfactants. The changes with the addition of neutral lipids is significantly smaller.
+\subsection{Lipid headgroup interactions in PC:PE and PC:PG mixtures and counterion binding}
 \begin{figure*}[]
@@
   \caption{\label{HGorderparametersPCvsPEPG}
-    Modulation of PC headgroup order parameters with increasing amount of PE (left) and PG (right) in bilayer
+    Modulation of POPC headgroup order parameters with increasing amount of POPE (left) and POPG (right) in bilayer
     from experiments \cite{scherer87,macdonald87} and simulations with different force fields.
@@
   }
-  \todo{Data for CHARMM, Lipid17, Slipids, and also GROMOS also without ions is now available.
-    This data should be updated into this plot}
+  \todo{Data for CHARMM without ions to be added once this issue is solved: https://github.com/NMRLipids/MATCH/issues/83}
   %\todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
 \end{figure*}
-%The headgroup order parameters for PC lipids (POPC and DOPC)
-%mixed with PE, PS and PG lipids are shown in Fig. \ref{HGorderparametersPCvsPEPSPG}
-%from different simulation model and experiments \cite{scherer87} with different
-%mole fractions. As already discussed previosly, the PC lipid headgroup behaviour
-%follows the electrometer concept in experiments when mixed with other lipids, i.e., the order
-%parameters increase when mixed with negatively charged lipids (PS, PI, CL, PA and PG)
-%remains almost unchaged when mixed with neutral lipids (PE and SM) \cite{scherer87}.
-This is roughly observed also in simulations here, althouht the changes in PC headgroup with
-PE are slightly overestimated and changes with PG lipid are slightly underestimated.
-The underestimated increase of PC headgroup order parameters with the addition of
-negaticely charged lipids was previously interepreted to indicate overestimated
-counterion binding affinity, which overcompensates the effect of negative charge \cite{NMRlipidsIVps}.
+According to the electrometer concept, the headgroup order parameters increase
+with the addition of negatively charged PG lipids, but
+are not affected by the addition of zwitterionic PE and SM lipids
+or cholesterol (Fig.~\ref{HGorderparametersPCvsPEPG}) \cite{seelig87, scherer87,antila18}. 
+The different responses of PC headgroup order parameters to PE and PG are roughly
+reproduced in CHARMM36 simulations, although the changes are slightly
+overestimated upon addition of PE and underestimated upon addition of PG.
+Interestingly, the response of POPC headgroup order parameters to POPE is
+close to experiments also in Berger-OPLS simulations, even though the response
+of Berger force field to cholesterol was significantly overestimated \cite{botan15}.
+In all force fields except Slipids, the $\alpha$-carbon order parameters of different
+hydrogens are responsing differently when mixed with PE or PG lipids \todo{Maybe we should figure out what is the reason for this}.
+\todo{Maybe we should analyze the P-N vector angle from different simulations}.
+
+In our previous study about PS lipids \cite{antila18}, we related the underestimated increase of PC
+headgroup order parameters upon addition of negaticely charged lipids
+%was previously interepreted
+to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
+Similar correlation is observed here in other than GROMOS-CKP simulation:
+Lipid 17 simulation exhibits the stronger counter-ion binding affinitiy to the pure POPG bilayer (Fig. \ref{CIdensPG})
+and larger decrease in POPC $\beta$ carbon order parameter upon addition of POPG than other simulations.
+This, together with the low area per molecule in Lipid17 (59.5 {\AA} vs. 66.1 {\AA} in experiments) suggests
+that the counterion binding this model is too strong.
+However, $\beta$-carbon order parameter of POPC in GROMOS-CKP simulations is in line with experiments and
+increases more than in CHARMM36 and Slipids simulations even though the binding affinities in these
+simulations are simialar. On the other hand, the different responses of different hydrogens in $\alpha$-carbons
+complicate the interpretation. 
 Thus, the counterion binding to PG lipid bilayers (Fig. \ref{CIdensPG}) may be slightly

### 2019-10-08 Manuscript/manuscriptPGPE.tex
Update text and figures

@@
 
-\subsection{Lipid headgroup interactions in PC:PE and PC:PG mixtures and counterion binding}
+\subsection{PC headgroup interactions with PE and PG}
 \begin{figure*}[]
@@
 \end{figure*}
-According to the electrometer concept, the headgroup order parameters increase
-with the addition of negatively charged PG lipids, but
+According to the electrometer concept, the PC headgroup order parameters increase
+with the addition of negatively charged PG or PS lipids, but
 are not affected by the addition of zwitterionic PE and SM lipids
@@
 reproduced in CHARMM36 simulations, although the changes are slightly
-overestimated upon addition of PE and underestimated upon addition of PG.
+overestimated upon addition of PE and underestimated upon addition of PG (Fig.~\ref{HGorderparametersPCvsPEPG}).
 Interestingly, the response of POPC headgroup order parameters to POPE is
 close to experiments also in Berger-OPLS simulations, even though the response
-of Berger force field to cholesterol was significantly overestimated \cite{botan15}.
+of Berger force field to cholesterol was significantly overestimated in our previous work \cite{botan15}.
 In all force fields except Slipids, the $\alpha$-carbon order parameters of different
@@
 
-In our previous study about PS lipids \cite{antila18}, we related the underestimated increase of PC
-headgroup order parameters upon addition of negaticely charged lipids
-%was previously interepreted
-to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
-Similar correlation is observed here in other than GROMOS-CKP simulation:
-Lipid 17 simulation exhibits the stronger counter-ion binding affinitiy to the pure POPG bilayer (Fig. \ref{CIdensPG})
-and larger decrease in POPC $\beta$ carbon order parameter upon addition of POPG than other simulations.
-This, together with the low area per molecule in Lipid17 (59.5 {\AA} vs. 66.1 {\AA} in experiments) suggests
-that the counterion binding this model is too strong.
-However, $\beta$-carbon order parameter of POPC in GROMOS-CKP simulations is in line with experiments and
-increases more than in CHARMM36 and Slipids simulations even though the binding affinities in these
-simulations are simialar. On the other hand, the different responses of different hydrogens in $\alpha$-carbons
-complicate the interpretation. 
-Thus, the counterion binding to PG lipid bilayers (Fig. \ref{CIdensPG}) may be slightly
-overestimated also in simulations here.
+For $\beta$-carbon order parameter in PG headgroup, experiments report
+mild increase \cite{macdonald87} or no change~\cite{borle85} upon addition 
+of PC lipids (Fig. \ref{HGorderparametersPGvsPCchange}). 
+Simulations with all the tested force fields give only very small changes also for
+the $\alpha$-carbon order parameter (Figs. \ref{HGorderparametersPGvsPC} and \ref{HGorderparametersPGvsPCchange}). 
+Therefore, the simulations are generally in line with experimetns, suggesting that the
+interactions with PC do not essentially effect the PG headgroup structure.
+This is in contrast with previous results for PS headgroup \cite{antila18}, where
+all the force fieds significantly overestimated the structural response of PS headgroup
+to the interactions with PC lipids.
 \begin{figure}[]
   \centering
-  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
-  \caption{\label{CIdensPG}
-    Counterion densities and area per lipids from simulations with PG lipids.
-    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
+  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPCchange.eps}
+  \caption{\label{HGorderparametersPGvsPCchange}
+    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
+    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
   }
-  \todo{Data for GROMOS is also available without ions, but I could not analyze it, probably because tpr file was not correct.}
 \end{figure}
@@
 The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
-  which supports the more condensed PE bilayer. In principle, this is beyond the scope of this work (lipid headgroups), but we can
-consider mentioning this.}
+which supports the more condensed PE bilayer.
+This is interesting, but to avoid the overexapansion of the manuscript, it is probably better to keep the focus in headgroups and ion binding also in this manuscript.}
 
-The order parameter of headgroup $\beta$-carbon of PG slightly increases with the addition
-of PC \cite{macdonald87}, althought also smaller changes have been observed~\cite{borle85}
-(Fig. \ref{HGorderparametersPSPGvsPC}).
-The increase in PG headgroup order parameter is not observed in simulations, which
-may arise from overestimated counterion binding overcompensating the changes in
-electrostatic environment as suggested for PC headgroup above. However, the changes
-are quite small and not fully consistent in experiments.
+\subsection{Counterion binding to PE and PG lipid bilayers}
+In our previous study about PS lipids \cite{antila18}, the underestimated increase of PC
+headgroup order parameters upon addition of negaticely charged lipids
+was related
+to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
+Similar correlation is observed here in %other than GROMOS-CKP simulation:
+Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
+and larger decrease in POPC $\beta$ carbon order parameter upon addition of POPG than other simulations.
+Together with the lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$),
+the results suggest that the counterions bind too strongly and shield the electrostatic repulsion between PG headgroups
+in bilayers simulated with Lipid17 parameters.
+In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of POPC
+increases upon addition of POPG more than in CHARMM36 and Slipids simulations even though the binding affinities in these
+simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
+are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
+and other decreases \todo{Can we figure out the reason for this?}.
+In conclusion, the counter-ion binding seems to be overestimated in Lipid17 with the strongest affinity among the
+tested models, while with this data we cannot fully evaluate binding affinity in the other models.
 \begin{figure}[]
   \centering
-  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPC.eps}
-  \caption{\label{HGorderparametersPGvsPC}
-    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
-    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
+  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
+  \caption{\label{CIdensPG}
+    Counterion densities and area per lipids from simulations with PG lipids.
+    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
+  }
+  \todo{Data for GROMOS is also available without ions, but I could not analyze it, probably because tpr file was not correct.}
+\end{figure}
+
+%Sodium binding to PC lipid bilayers is significantly overestimated by most simulations models \cite{catte16}.
+%Also, sodium binding to PS lipids seems to be overestimated, although the
+%presence of counterions complicate the comparison for negatively charged lipids \cite{NMRlipidsIVps}.
+Similarly to POPG, the Lipid17 exhibits stronger sodium binding affinity
+to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
+Experimental data for sodium binding to PE lipids is not available, but
+large deviations from PC would be surprising.
+However, in Slipids and GROMOS-CKP simulations, the binding affinity to POPE is much weaker than to POPC
+(our previous work \cite{catte16} and Fig. ??). On the other hand, 
+the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
+previous Amber lipid version, Lipid14.
+\todo{The potential differences in the used ion models should be checked}.
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
+  \caption{\label{NAdensPE}
+    Sodium (solid line) and choride ion density profiles along membrane normal
+    from different simulations with PE lipids.
   }
@@
 can be directly compared with experiments \cite{catte16,melcr18,NMRlipidsIVps}.
-
-Sodium binding to PC lipid bilayers is significantly overestimated by most simulations models \cite{catte16}.
-Also, sodium binding to PS lipids seems to be overestimated, although the
-presence of counterions complicate the comparison for negatively charged lipids \cite{NMRlipidsIVps}.
-Sodium binding to PE lipids is weaker than to PC lipids in previous work \cite{catte16} in both CHARMM36 and Slipids simulations (Fig. \ref{NAdensPE}).
-This cannot be evaluated using electrometer concept because experimental data with PE lipids is not
-available \todo{Other experimental data should be checked and discussed}.
-The difference in Slipids may also be due to different ion model used \todo{The used ion model here should be checked}.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
-  \caption{\label{NAdensPE}
-    Sodium (solid line) and choride ion density profiles along membrane normal
-    from different simulations with PE lipids.
-  }
-\end{figure}
 

### 2019-10-09 Manuscript/manuscriptPGPE.tex
Update text and figures

@@
 
-\subsection{Counterion binding to PE and PG lipid bilayers}
+\subsection{Sodium binding to PE and PG lipid bilayers}
 In our previous study about PS lipids \cite{antila18}, the underestimated increase of PC
@@
 to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
-Experimental data for sodium binding to PE lipids is not available, but
-large deviations from PC would be surprising.
-However, in Slipids and GROMOS-CKP simulations, the binding affinity to POPE is much weaker than to POPC
-(our previous work \cite{catte16} and Fig. ??). On the other hand, 
-the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
-previous Amber lipid version, Lipid14.
+In Lipid17 the sodium binding affinity to POPE is approximately similar as to POPC,
+while in other models exhibit weaker binding to POPE (Fig. \ref{NAdensPC}).
+The difference is small in Slipids and CHARMM36, while GROMOS-CKP predicts
+subtantially stronger binding to POPC.
+Experimental data for sodium binding to PE lipids is not available, while
+simulations in agreement with electrometer data from NMR suggests that
+sodium binding to POPC is weaker than in any of the simulations
+here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}. Assuming that the binding
+to POPE would be similar than to POPC, the sodium binding affinity to POPE
+is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
+but substantially overestimated in Lipid17 simulation.
+%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
+%previous Amber lipid version, Lipid14.
 \todo{The potential differences in the used ion models should be checked}.

### 2019-10-09 Manuscript/manuscriptPGPEsuppl.tex
Update text and figures

@@
 \clearpage
+
+\section{Changes of PG headroup order parameters upon addition of PC}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPC.eps}
+  \caption{\label{HGorderparametersPGvsPC}
+    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
+    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
+  }
+\end{figure}
+
+\clearpage
+
+\section{Sodium binding to POPC simulations}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/NAdensPC.eps}
+  \caption{\label{NAdensPC}
+    Sodium (solid line) and choride ion density profiles along membrane normal
+    from different simulations with PC lipids.
+  }
+  \todo{Discussion about differences to the NMRlipids II to be discussed once we have the details on ions models.}
+\end{figure}
+
 \bibliography{refs.bib}

### 2019-10-22 Manuscript/manuscriptPGPE.tex
UPdate text
Add density profile from GROMOS-CKP POPG simulatio with counterions only
Update text

@@
 \section{Introduction}
+
+Zwitterionic PE is the second most abundant glycerophospholipid in eukaryotic cells
+and has been related to the diseases \cite{vance15,calzada16,patel17}.
+PE and PG affect membrane protein functionality \cite{hariharan18} and bind to various proteins \cite{yeagle14}.
 
@@
       POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
-      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ?
-& ?
-& ?
-&  ?
-& ?
-& ?
-& \cite{??} \todoi{Data to be uploaded by J.
-Madsen.
-Details to be filled once we have the data}  \\
-      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0.15 \todoi{Concentration to be checked.}  & ?
-& ?
-& ?
-&  ?
-& ?
-& ?
-& \cite{??} \todoi{Data to be uploaded by J.
-Madsen.
-Details to be filled once we have the data}  \\
-      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0  \todoi{Concentration to be checked.}  & ?
-& ?
-& ?
-&  ?
-& ?
-& ?
-& \cite{??} \todoi{Data to be uploaded by J.
-Madsen.
-Details to be filled once we have the data}  \\
-      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0      & ?
-& ?
-& ?
-&  ?
-& ?
-& ?
-& \cite{??} \todoi{Data to be uploaded by J.
-Madsen.
-Details to be filled once we have the data}  \\
-      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0.15 \todoi{Concentration to be checked.}  & ?
-& ?
-& ?
-&  ?
-& ?
-& ?
-& \cite{??} \todoi{Data to be uploaded by J.
-Madsen.
-Details to be filled once we have the data}  \\
-      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0  \todoi{Concentration to be checked.}  & ?
-& ?
-& ?
-&  ?
-& ?
-& ?
-& \cite{??} \todoi{Data to be uploaded by J.
-Madsen.
-Details to be filled once we have the data}  \\
+      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0  & 250:250 & 18158 & 0   &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG5050} \\ 
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.34 \todoi{Concentration calculated based in total amount of calcium ions.
+This may not be reasonable due to the lack of counterions.}  & 250:250 & 20798 & 128 &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG5050150mMCaCl} \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.36  \todoi{Concentration calculated based in total amount of calcium ions.
+This may not be reasonable due to the lack of counterions.} & 250:250 & 18114 & 445  &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
+      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0  & 400:100 & 18664 & 0  &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG4010} \\
+      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0 \todoi{Concentration calculated based in total amount of calcium ions.
+This may not be reasonable due to the lack of counterions.} & 400:100 & 18647 & 419  &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
       \hline
@@
   }
-  \todo{Data for GROMOS is also available without ions, but I could not analyze it, probably because tpr file was not correct.}
 \end{figure}

### 2019-10-22 Manuscript/manuscriptPGPE.tex
Update text

@@
 
+PE and PG lipids are most common lipids in bacteria \cite{sohlenkamp16}.
 Zwitterionic PE is the second most abundant glycerophospholipid in eukaryotic cells
 and has been related to the diseases \cite{vance15,calzada16,patel17}.
+Anionic PG lipids are less abundant, but is also proposed to be fundamental for terrestrial life \cite{furse17}.
 PE and PG affect membrane protein functionality \cite{hariharan18} and bind to various proteins \cite{yeagle14}.
+PE headgroup is also prone for negative membrane curvature and causes membrane fusion \cite{??}.
+Therefore, the PE and PG headgroup structures play probably essential roles in 
+many biological processes.
 
-In NMRlipids I and II project we were looking for a MD model
-which would correctly reproduce headgroup and glycerol
-backbone structures and cation binding for PC lipid bilayers \cite{botan15,catte16}.
-Here we extend the same goal for other than PC lipids.
-Currently the focus is on PE, PG and PS bilayers and their
-mixtures with PC. Experimental data with different amounts of 
-added salt is now collected and presented in this manuscript.
+Structural details of lipid headgroups are mainly studied using NMR experiments, which
+suggest that the glycerol backbone structures are largely similar irrespectively of the headroup \cite{gally81}, 
+glycerol backbone and headgroup structure and behaviour are similar in model membranes and in bacteria \cite{gally81,scherer87,seelig90},
+and the headgroup structures are similar in PC, PE and PG lipids, while headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. 
+%Extensive discussion about structural details of PE, PG or PS headgroups do not exists (as far as I know), 
+%In contrast to PC lipids (see \cite{botan15} and references therein).
+Some attempts to resolve conformational ensembles from NMR for PC lipids have been made,
+but lesser extend for other lipids \cite{??}.
+Classical molecular dynamics simulations could potentially give such ensembles and therefore enable
+the detailed studies of lipid headgroup behaviour in complex biomolecular systems, but current
+force fields are not accurate enough to reproduce the correct conformational ensembles for PC and PS headgroups \cite{botan15,antila19}.
+Several MD simulations of PE and PG lipids have been published especially in the context of modeling
+inner membrane of Gram-negative bacteria \cite{??}, but glycerol backbone and headgroup structures have not
+been evaluated against experiments.
+
+Besides the structure, also ion binding may regulate biophysical activity of especially negatively
+charged lipid headgroups \cite{??}. Monovalent cation (except Lithium) binding to zwitterionic PC headgroups is
+very weak, while multivalent ion binding is stronger but still weak \cite{??}. The ion binding affinity
+data for PE is more scarce \cite{??}, but large differences to PC would be surprising.
+Negatively charged lipids are suggested to bear same cation binding constants than zwitterionic lipids, but
+the amount of bound ions to negatively charged membranes would still be larger because
+the concentration of cations in the vicinity of membranes would be higher  \cite{seelig90}.
+%is stronger
+%Based on the electrometer concept and other data is has been suggested that \cite{seelig90}
+%\begin{displayquote}
+%  {\it ''(i) Ca$^{2+}$ binds to neutral lipids (phosphatidylcholine, phosphatidylethanolamine) and negatively charged lipids
+%    (phosphatidylglycerol) with approximately the same binding constant of K = 10-20 M$^{-1}$; \\
+%    (ii) the free Ca$^{2+}$
+%    concentration at the membrane interface is distinctly enhanced if the membrane carries a negative surface
+%    charge, either due to protein or to lipid; \\
+%    (iii) increased inter-facial Ca$^{2+}$ also means increased amounts
+%    of bound Ca$^{2+}$ at neutral and charged lipids; \\
+%    (iv) the actual binding step can be described by a Langmuir
+%    adsorption isotherm with a 1 lipid:1 Ca$^{2+}$ stoichiometry, provided the interfacial concentration C$_M$, is
+%    used to describe the chemical binding equilibrium.''}
+%\end{displayquote}
+On the other hand, anionic PS lipids are proposed chelate with calcium ions \cite{??}.
+In simulations, the cation binding affinity to PC and PS membranes is typically overestimated \cite{catte16,antila19},
+which can be improved by applying the ECC to the partial charges of the force fields \cite{melcr18,melcr19}.
+
+Here, we use open collaboration and order parameters of glycerol backbone and headgroup
+to evaluate the accuracy of PE and PG heagroup structures, and the cation binding affinity
+to anionic membranes containing PG lipids in the current MD simulation force fields.
+The force field giving the best description for glycerol backbone and headgroup structures
+of PC, PS, PG and PE headgroups (CHARMM36) reproduces the essential differences in order parameters
+between these headgroups, and therefore enables the analysis of structural differences between the
+headgroups.
 
@@
 
-Experimental order parameters for different lipid headgroups are
-collected in Fig. \ref{HGorderParameters}.
-
-Based on superficial reading, the conclusions in the literature are roughly \\ 
-1) glycerol backbone structures are largely similar irrespectively of the headroup \cite{gally81}, \\
-2) glycerol backbone and headgroup structure and behaviour are similar in model membranes and in bacteria \cite{gally81,scherer87,seelig90}, \\
-3) headgroup structures are similar in PC, PE and PG lipids, while headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. \\
-Extensive discussion about structural details of PE, PG or PS headgroups do not exists (as far as I know), 
-In contrast to PC lipids (see \cite{botan15} and references therein).
-
-Several simulations containing PE, PG and PS lipids have been published \cite{??}, \todo{List should be completed}
-however, glycerol backbone and headgroup order parameters are not compared to
-the experiments (based on superficial reading of literature).
-
 \section{Methods}
@@
 S-DROSS experiments~\cite{gross97} using natural abundance $^{13}$C solid state NMR spectroscopy
-as described previously \cite{ferreira13,ferreira16,NMRlipidsIVps}.
+as described previously \cite{ferreira13,ferreira16,antila19}.
 %The experiments were done in a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz.
@@
 then calculated from these trajectories. This approach has been tested against trajectories
-with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot18}.
+with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot17}.
 The ion number density profiles were calculated using the {\tt gmx density} tool
@@
 The order parameters for the glycerol backbone and headgroup C--H bonds were determined
-from 2D-RPDLF and S-DROSS experiments (Fig. \ref{POPEspectra}), as described previously \cite{NMRlipidsIVps}.
+from 2D-RPDLF and S-DROSS experiments (Fig. \ref{POPEspectra}), as described previously \cite{antila19}.
 The POPE experiments were recorded at 310~K, where the bilayer is in liquid disordered phase \cite{??}.
@@
     with different headgroups in lamellar liquid disordered phase.
-    The values and signs for POPE (310~K), POPG (298~K). POPS (298~K) \cite{NMRlipidsIVps} and POPC (300~K) \cite{ferreira13,ferreira16}
+    The values and signs for POPE (310~K), POPG (298~K). POPS (298~K) \cite{antila19} and POPC (300~K) \cite{ferreira13,ferreira16}
     are measured using $^{13}$C NMR. The literature values for
@@
 \subsection{Headgroup and glycerol backbone of POPE and POPG in MD simulations}
-As reported previously for PC \cite{botan15} and PS \cite{NMRlipidsIVps} lipids,
+As reported previously for PC \cite{botan15} and PS \cite{antila19} lipids,
 the headgroup and glycerol backbone order parameters show wide variation between different force fields
@@
 %The Slipid simulations were able to capture the essential differences between PC and PS lipid
-%headgroups \cite{NMRlipidsIVps}, but this is not the case for PE and PG lipids.
+%headgroups \cite{antila19}, but this is not the case for PE and PG lipids.
 %\todo{Berger results are not here yet, but we should mention the ring like structures
@@
 \end{figure*}
-Also in line with previous studies for PC \cite{botan15} and PS \cite{NMRlipidsIVps} lipids,
+Also in line with previous studies for PC \cite{botan15} and PS \cite{antila19} lipids,
 CHARMM36 simulations are in best agreement with experiments for headgroup and glycerol backbone
@@
 %headgroup order parameters of PC PE, PG (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
-%as previously observed also for PS lipids \cite{NMRlipidsIVps}.
+%as previously observed also for PS lipids \cite{antila19}.
 In previous study, CHARMM36 predicted the more negative $\beta$-carbon order parameter and larger forking of
-the $\alpha$-carbon is PS headgroup than in PC \cite{NMRlipidsIVps}.
+the $\alpha$-carbon is PS headgroup than in PC \cite{antila19}.
 In this work, the CHARMM36 simulations reproduce also the other essential differences in experimental
@@
 headgroups (Fig.~\ref{structures}). While rotation around N-C$_\alpha$-C$_\beta$-O$_\alpha$ dihedral was
-significantly different in PS and PC headgroups in previous work \cite{NMRlipidsIVps}, essential differences
+significantly different in PS and PC headgroups in previous work \cite{antila19}, essential differences
 between PC, PE and PG are not observed here (Fig.~\ref{structures}) \todo{We need also the dihedral distributions to finish this discussion.}. \\
@@
 %Also, sodium binding to PS lipids seems to be overestimated, although the
-%presence of counterions complicate the comparison for negatively charged lipids \cite{NMRlipidsIVps}.
+%presence of counterions complicate the comparison for negatively charged lipids \cite{antila19}.
 Similarly to POPG, the Lipid17 exhibits stronger sodium binding affinity
@@
 can be used also for bilayers containing PC lipids mixed with charged
-lipids \cite{borle85,macdonald87,roux90,NMRlipidsIVps}.
+lipids \cite{borle85,macdonald87,roux90,antila19}.
 %This is demonstrated in Figs \ref{OrderParametersWithCaCl},
@@
 %overcharging occurs above these concenterations.
-Based on the electrometer concept and other data is has been suggested that \cite{seelig90}
-\begin{displayquote}
-  {\it ''(i) Ca$^{2+}$ binds to neutral lipids (phosphatidylcholine, phosphatidylethanolamine) and negatively charged lipids
-    (phosphatidylglycerol) with approximately the same binding constant of K = 10-20 M$^{-1}$; \\
-    (ii) the free Ca$^{2+}$
-    concentration at the membrane interface is distinctly enhanced if the membrane carries a negative surface
-    charge, either due to protein or to lipid; \\
-    (iii) increased inter-facial Ca$^{2+}$ also means increased amounts
-    of bound Ca$^{2+}$ at neutral and charged lipids; \\
-    (iv) the actual binding step can be described by a Langmuir
-    adsorption isotherm with a 1 lipid:1 Ca$^{2+}$ stoichiometry, provided the interfacial concentration C$_M$, is
-    used to describe the chemical binding equilibrium.''}
-\end{displayquote}
 The electrometer concept has been very useful in evaluating ion binding affinity in simulations
 against experiments, because the headgroup order parameter changes as a function of ion concentration
-can be directly compared with experiments \cite{catte16,melcr18,NMRlipidsIVps}.
+can be directly compared with experiments \cite{catte16,melcr18,antila19}.
 
 Calcium binding affinity to PC and PS lipid bilayers was not correctly described by
-any of the standard MD simulation forced fields \cite{catte16,NMRlipidsIVps}, while
+any of the standard MD simulation forced fields \cite{catte16,antila19}, while
 recently introduced force field with electronic continuum correction (ECC) performed better \cite{melcr18}.

### 2019-10-23 Manuscript/manuscriptPGPE.tex
Update text and figures and add POPG data from Tiago Ferreira send by email

@@
 \section{Methods}
-\subsection{Experimental C--H bond order parameters}% from the natural abundance $^{13}$C\,NMR}
-
+\subsection{Experimental C--H bond order parameters}
 The headgroup and glycerol backbone C--H bond order parameter magnitudes and signs of POPE and POPG
@@
 S-DROSS experiments~\cite{gross97} using natural abundance $^{13}$C solid state NMR spectroscopy
-as described previously \cite{ferreira13,ferreira16,antila19}.
+as described previously \cite{ferreira13,ferreira16}.
 %The experiments were done in a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz.
@@
 %The following experimental setups were used.
-\todo{The rest of the details to be written. I am not sure how much we need to repeat the NMRlipidsIVps paper.}
+POPE and POPG powder were purchased from Avanti polar lipids.
+The NMR experiments were identical to our previous work~\cite{antila19}.
+\todo{Is this enough and correct, or should we repeat some methods from the NMRlipidsIVps paper?}
 
@@
 as the communication platforms.
-The simulated systems are listed in 
-Tables \ref{systems} (pure PE and PG bilayers without additional ions).
-and \ref{systemsII} (mixtures and systems with additional ions).
+The simulated systems of pure PE and PG bilayers without additional ions
+are listed in Tables \ref{systemsPE} and \ref{systemsPG},
+and lipid mixtures with additional ions in Table~\ref{systemsMIX}.
 Further simulation details are given in the SI, and
@@
 then calculated from these trajectories. This approach has been tested against trajectories
-with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot17}.
+with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot17}.\\
+\todo{BuildH program is now cited with a direct link to the GitHub repo. I think that a release to Zenodo would be nice in the final publication.}\\
+\todo{Maybe we should also shortly discuss here about the reasons for slight dependence of order parameter values on the method used to reconstruct hydrogens?}\\
 The ion number density profiles were calculated using the {\tt gmx density} tool
@@
     % The lipid force fields named as in our previous work~\cite{botan15}.
-  }\label{systems}
+  }\label{systemsPE}
   \begin{minipage}[t]{\textwidth}
@@
   %\end{sidewaystable*} 
+  \todo{We need citations for the force fields.}
 \end{table*}
@@
     % The lipid force fields named as in our previous work~\cite{botan15}.
-  }\label{systemsII}
+  }\label{systemsPG}
   \begin{minipage}[t]{\textwidth}
@@
   \end{minipage}
-  %\end{sidewaystable*} 
+  % \end{sidewaystable*}
+    \todo{We need citations for the force fields.}
 \end{table*}
@@
     % The lipid force fields named as in our previous work~\cite{botan15}.
-  }\label{systemsII}
+  }\label{systemsMIX}
   \begin{minipage}[t]{\textwidth}
@@
   %\end{sidewaystable*}
-  \todo{Data for POPC:POPG mixtures by listed by Antonio Peon is missing from this table}
+  \todo{Data for POPC:POPG mixtures by listed by Antonio Peon is missing from this table} \\
+    \todo{We need citations for the force fields.}
 \end{table*}
@@
 \subsection{Headgroup and glycerol backbone order parameters of POPE and POPG from $^{13}$C NMR}
-
-The glycerol backbone and $\alpha$-carbon peaks in INEPT spectra of POPE were assigned based on
-previously measured POPC spectra (Fig. \ref{POPEspectra}) \cite{ferreira13}. The $\beta$-carbon peak was assigned based
-on $^{13}$C chemical shift table for amines available at \url{https://www.chem.wisc.edu/areas/reich/nmr/c13-data/cdata.htm}.
-The order parameters for the glycerol backbone and headgroup C--H bonds were determined
-from 2D-RPDLF and S-DROSS experiments (Fig. \ref{POPEspectra}), as described previously \cite{antila19}.
-The POPE experiments were recorded at 310~K, where the bilayer is in liquid disordered phase \cite{??}.
+Absolute values of the headgroup and glycerol backbone order parameters from PE and PG lipids are measured
+previously using $^2$H NMR~\cite{seelig76,gally81,wohlgemuth80,borle85}. Because also the order parameter
+signs bear essential information about the lipid structures \cite{botan15,ollila16}, we measured the
+magnitudes and signs of POPE and POPG C--H bond headgroup and glycerol backbone order parameter in liquid phase
+using the 
+%order parameters The order parameters for the glycerol backbone and headgroup s were determined
+2D-RPDLF and S-DROSS experiments, as described previously \cite{ferreira13,ferreira16,antila19}.
+For POPE, the glycerol backbone and $\alpha$-carbon peaks in INEPT spectra were assigned based on
+previously measured POPC spectra~\cite{ferreira13} and
+the $\beta$-carbon peak was assigned based on $^{13}$C chemical shift table for amines available
+at \url{https://www.chem.wisc.edu/areas/reich/nmr/c13-data/cdata.htm} (Fig. \ref{POPEspectra}).
+For POPG, the glycerol backbone peaks in INEPT spectra were assigned based on
+previously measured POPC spectra~\cite{ferreira13}, while $\alpha$ and  $\gamma$-carbon peaks
+\todo{How were these assigned?} (Fig. \ref{POPGspectra}). The numerical value of the $\beta$-carbon
+order parameter could not be determined, because its peak overlapped with the g$_2$ peak from glycerol backbone in POPG.
+However, the order parameter of $\beta$-carbon is expected to be clearly smaller than for g$_2$
+based on previous $^2$H NMR measurements \cite{wohlgemuth80,gally81,borle85}.
+Therefore, the beginning of the S-DROSS curve gives the sign for g$_2$ order parameter and end for $\beta$ (Fig. \ref{POPGspectra} (E)).
+This is confirmed with SIMPSON calculations using negative value for g$_2$ and positive value for $\beta$ order parameter (Fig. \ref{POPGsimpson}).
+The POPE experiments were recorded at 310~K and POPG experiments at 298~K, where the bilayers are in the liquid disordered phase \cite{marsh13}.
 \todo{Details to be checked by Tiago}.
 
-\todo{Figure and discussion about POPG experiments to be addded.}
+As discussed previously for PC and PS headgroups \cite{ollila16,antila19}, also 
+the headgroup and glycerol backbone order parameters of PE and PG are essentially independent of 
+acyl chain composition, and therefore manifest mainly headgroup chemistry (Fig. \ref{HGorderParameters}).
+The glycerol backbone order parameters are similar for all the lipids, although they move slightly toward
+positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG. 
+Also the headgroup order parameters of PC lipids are close PE, althought the latter gives systematically
+slightly more positive values (Fig. \ref{HGorderParameters}).
+% These could be explained with slightly larger temperature in PE measurements, except for the $\alpha$-carbon with the positive sign, for which the
+% more positive value is farther away from zero.
+The $\alpha$-carbon order parameter of PG is similar to PE and PG, while the positive value of $\beta$-carbon
+is distinct from the other lipids. Notably, this difference was not observed in previous $^2$H NMR experiments,
+because absolute value of $\beta$-carbon order parameter is similar in PG, PE and PC lipids and the order parameter
+signs were not measured~\cite{wohlgemuth80,gally81,borle85}.
 
-The headgroup and glycerol backbone order parameters of PE lipids are similar with 
-different acyl chains and also close the values for POPC, althought PE gives systematically
-slightly more positive values (Fig. \ref{HGorderParameters}). These could be explained with slightly larger temperature
-in PE measurements, except for the $\alpha$-carbon with the positive sign, for which the
-more positive value is farther away from zero. For PG lipids, the glycerol backbone order
-parameters are more positive than for other lipids. The headgroup $\alpha$-carbon gives
-value close to PE, while the value of $\beta$-carbon is distinct from other lipid being
-only one which has positive sign, suggesting distinct conformation of PG lipids in this region.
-This was not observed in previous $^2$H NMR study, where sign was not measured and $\beta$-carbon
-order parameter was apparently similar to the value for PE and PC results.
+In conclusion, the order parameter experiments suggest that the glycerol backbone conformations in all lipids
+and the headgroup conformations in PC and PE lipids are relatively similar, while 
+PS and PG headgropus exhibit distinct conformational sampling.
+The details of sampled conformation are difficult to deduce from order parameters only,
+but the distinct headgroup order parameters
+of PS lipids are previously related to the more rigid structure of the headgroup~\cite{browning80,buldt81,antila19}.
 
@@
   \caption{\label{HGorderParameters}
-    (top) Chemical structure of different lipids
-    (bottom) Headgroup and glycerol backbone order parameters measured from lipids
-    with different headgroups in lamellar liquid disordered phase.
-    The values and signs for POPE (310~K), POPG (298~K). POPS (298~K) \cite{antila19} and POPC (300~K) \cite{ferreira13,ferreira16}
-    are measured using $^{13}$C NMR. The literature values for
+    (top) Chemical structure of different lipids.
+    (bottom) Headgroup and glycerol backbone order parameters
+    from different experiments in lamellar liquid disordered phase.
+    The values and signs for POPE (310~K) and POPG (298~K)
+    measured in this work, and for POPS (298~K) \cite{antila19} and POPC (300~K) \cite{ferreira13,ferreira16}
+    previously using $^{13}$C NMR. The literature values for
     DOPS with 0.1M of NaCl (303~K) \cite{browning80},
@@
   }
+  \todo{The bottom figure could be clarified as Fig. 2 in the NMRlipids IVps paper.}
 \end{figure}
 
-In conclusion, the results suggests that the glycerol backbone conformations in all lipids are
-relatively similar. Also, the headgroup conformations are similar for PC and PE lipids, while
-PS and PG are singnificantly different. For PS lipids, the differences are discussed previously \cite{??}.
-
 \subsection{Headgroup and glycerol backbone of POPE and POPG in MD simulations}
-As reported previously for PC \cite{botan15} and PS \cite{antila19} lipids,
-the headgroup and glycerol backbone order parameters show wide variation between different force fields
-for both PE and PG lipids (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
+Headgroup and glycerol backbone order parameters of PE and PG lipids
+show wide variation between different force fields
 and none of the force fields reproduce all values within experimental error bars.
+(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
+as observed previously also for PC and PS lipids \cite{botan15,antila19}.
 %The Slipid simulations were able to capture the essential differences between PC and PS lipid
 %headgroups \cite{antila19}, but this is not the case for PE and PG lipids.
-%\todo{Berger results are not here yet, but we should mention the ring like structures
-%  pointed out by T. Piggot in the blog:
-The poor performance of headgroup order parameters in Berger model can be probably explained by ring like structures seen in Fig. 6 in Ref. \citenum{mukhopadhyay04},
-which is a typical feature for Berger based lipid force fields containing explicit hydrogen atoms in the head group \cite{zhao08,henin09,dahlberg10}.
-%}
+The poor performance of headgroup order parameters in Berger model can be probably explained by
+ring like structures seen in Fig. 6 in Ref. \citenum{mukhopadhyay04}, which is a typical feature
+for Berger based lipid force fields containing explicit hydrogen atoms in the head group \cite{zhao08,henin09,dahlberg10}.
+The poor performance of glycerol backbone of Slipids simulations is systemically observed also
+for other lipids in previous studies \cite{botan15,antila19}.
+\todo{Should we comment more the relative quality of different force fields and/or make the subjective force field ranking figures?}
+
+Without further discussion about poorly performing force fields, we focus on more detailed analysis
+of CHARMM36 simulations, which give the best overall agreement with experiments for
+the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
+studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
+not within the experimental error bars, it seems to capture the essential differences between different headroups.
+%(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
+In previous study, CHARMM36 predicted the more negative $\beta$-carbon order parameter and larger forking of
+the $\alpha$-carbon in PS headgroup than in PC \cite{antila19}.
+In this work,
+%the CHARMM36 simulations reproduce also the other essential differences in experimental
+%headgroup order parameters between different headgroups (Fig. \ref{HGorderParameters}) despite the inaccuracies
+%in individual segments:
+the PE headgroup order parameters are similar to the PC \cite{botan15}
+and $\beta$-carbon order parameters is positive in PG headgroup, while negative values observed in
+other lipids. Therefore, we use CHARMM36 simulations to analyze the structural differences between
+headgroups (Fig.~\ref{structures}).
+\todo{The structures are already in Fig. \ref{structures}, but we need also the dihedral distributions to finish the discussion}.
+%While rotation around N-C$_\alpha$-C$_\beta$-O$_\alpha$ dihedral was
+%significantly different in PS and PC headgroups in previous work \cite{antila19}, essential differences
+%between PC, PE and PG are not observed here (Fig.~\ref{structures}) \todo{We need also the dihedral distributions to finish this discussion.}. \\
+%\todo{Why is the $\beta$-carbon order parameter of PG different to PC and PE then?}
+
+\todo{Maybe we should analyze also some other structural features, at least P-N angles.}
 
@@
 \end{figure*}
-Also in line with previous studies for PC \cite{botan15} and PS \cite{antila19} lipids,
-CHARMM36 simulations are in best agreement with experiments for headgroup and glycerol backbone
-order parameters, and seem to capture all the essential differences between different headroups
-(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
-%headgroup order parameters of PC PE, PG (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
-%as previously observed also for PS lipids \cite{antila19}.
-In previous study, CHARMM36 predicted the more negative $\beta$-carbon order parameter and larger forking of
-the $\alpha$-carbon is PS headgroup than in PC \cite{antila19}.
-In this work, the CHARMM36 simulations reproduce also the other essential differences in experimental
-headgroup order parameters between different headgroups (Fig. \ref{HGorderParameters}) despite the inaccuracies
-in individual segments:
-The PE headgroup order parameters in CHARMM36 simulations are similar to the PC \cite{botan15}
-and $\beta$-carbon order parameters is positive in PG headgroup, in contrast to the negative values observed in
-other lipids. Therefore, we use CHARMM36 simulations to analyze the structural differences between
-headgroups (Fig.~\ref{structures}). While rotation around N-C$_\alpha$-C$_\beta$-O$_\alpha$ dihedral was
-significantly different in PS and PC headgroups in previous work \cite{antila19}, essential differences
-between PC, PE and PG are not observed here (Fig.~\ref{structures}) \todo{We need also the dihedral distributions to finish this discussion.}. \\
-\todo{Why is the $\beta$-carbon order parameter of PG different to PC and PE then?}
 
@@
 \end{figure*}
-According to the electrometer concept, the PC headgroup order parameters increase
+
+In experiments, the PC headgroup order parameters increase
 with the addition of negatively charged PG or PS lipids, but
 are not affected by the addition of zwitterionic PE and SM lipids
-or cholesterol (Fig.~\ref{HGorderparametersPCvsPEPG}) \cite{seelig87, scherer87,antila18}. 
+or cholesterol (Fig.~\ref{HGorderparametersPCvsPEPG}).
+This can be explained by the electrometer concept, which suggests that the 
+headgroup dipole tilts more parallel to the membrane plane upon addition of
+negative charge to the membrane~\cite{seelig87, scherer87,antila18}.
 The different responses of PC headgroup order parameters to PE and PG are roughly

### 2019-10-23 Manuscript/manuscriptPGPEsuppl.tex
Update text and figures and add POPG data from Tiago Ferreira send by email

@@
 %  \centering
-  \includegraphics[width=\textwidth]{../Figs/POPEexperiment.pdf}
+  \includegraphics[width=0.8\textwidth]{../Figs/POPEexperiment.pdf}
   \caption{\label{POPEspectra}
@@
   \todo{A, B etc. labels to be put in the figure.}
+\end{figure}
+
+\begin{figure}[]
+%  \centering
+  \includegraphics[width=0.8\textwidth]{../Figs/PGexpRPDLF.eps}
+  \caption{\label{POPGspectra}
+    (A) Chemical structure of POPG with the labeling of headgroup and glycerol backbone carbons.
+    (B) INEPT spectra from POPG sample with the headgroup and glycerol backbone peaks labeled.
+    (C) 2D R-PDLF spectra
+    (D) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
+    (E) Experimetal S-DROSS curves giving signs of the order parameters.
+  }
+\end{figure}
+
+\begin{figure}[]
+%  \centering
+  \includegraphics[width=0.8\textwidth]{../Figs/SIMPSON.pdf}
+  \caption{\label{POPGsimpson}
+    Simpson simulaton of S-DROSS curve of $\beta$-carbon of POPG.
+  }
 \end{figure}

### 2019-10-29 Manuscript/manuscriptPGPE.tex
Add CHARMM PCvsPG results without ions
Update texty

@@
 
+\clearpage
 \subsection{PC headgroup interactions with PE and PG}
@@
   }
-  \todo{Data for CHARMM without ions to be added once this issue is solved: https://github.com/NMRLipids/MATCH/issues/83}
   %\todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
@@
 negative charge to the membrane~\cite{seelig87, scherer87,antila18}.
-The different responses of PC headgroup order parameters to PE and PG are roughly
-reproduced in CHARMM36 simulations, although the changes are slightly
-overestimated upon addition of PE and underestimated upon addition of PG (Fig.~\ref{HGorderparametersPCvsPEPG}).
-Interestingly, the response of POPC headgroup order parameters to POPE is
-close to experiments also in Berger-OPLS simulations, even though the response
-of Berger force field to cholesterol was significantly overestimated in our previous work \cite{botan15}.
-In all force fields except Slipids, the $\alpha$-carbon order parameters of different
-hydrogens are responsing differently when mixed with PE or PG lipids \todo{Maybe we should figure out what is the reason for this}.
-\todo{Maybe we should analyze the P-N vector angle from different simulations}.
+The response of PC headgroup order parameters to PE by the tested CHARMM36 and
+Berger-OPLS force fields, although CHARMM36 slightly overestimates the changes (Fig.~\ref{HGorderparametersPCvsPEPG}).
+The good performance of Berger-OPLS simulations is notable because the response of headgroup order parameters
+to cholesterol was significantly overestimated by the Berger/Höltje force field in our previous work \cite{botan15}.
+None of the force fields fully reproduces the PC headgroup order parameter response to the increasing amount of
+PG, which may be related to the counterion binding affinity (see also the next section) \cite{antila19}.
+In all force fields except Slipids, the order parameters of different
+hydrogens attached to the $\alpha$-carbon are responsing differently when mixed with PE or PG lipids
+\todo{Maybe we should figure out what is the reason for this?}.\\
+\todo{Maybe we should analyze the P-N vector angle from different simulations?}.
 
@@
 the $\alpha$-carbon order parameter (Figs. \ref{HGorderparametersPGvsPC} and \ref{HGorderparametersPGvsPCchange}). 
-Therefore, the simulations are generally in line with experimetns, suggesting that the
+Therefore, the simulations are generally in line with experiments, suggesting that the
 interactions with PC do not essentially effect the PG headgroup structure.
-This is in contrast with previous results for PS headgroup \cite{antila18}, where
+This suggests that the interactions between PG and PC headgroups are captured better in
+simulations than for PS headgroup, where
 all the force fieds significantly overestimated the structural response of PS headgroup
-to the interactions with PC lipids.
+to the interactions with PC lipids \cite{antila18}.
 \begin{figure}[]
@@
 
+\clearpage
 \subsection{Sodium binding to PE and PG lipid bilayers}
-In our previous study about PS lipids \cite{antila18}, the underestimated increase of PC
-headgroup order parameters upon addition of negaticely charged lipids
-was related
-to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
-Similar correlation is observed here in %other than GROMOS-CKP simulation:
-Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
-and larger decrease in POPC $\beta$ carbon order parameter upon addition of POPG than other simulations.
-Together with the lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$),
-the results suggest that the counterions bind too strongly and shield the electrostatic repulsion between PG headgroups
-in bilayers simulated with Lipid17 parameters.
-In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of POPC
-increases upon addition of POPG more than in CHARMM36 and Slipids simulations even though the binding affinities in these
-simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
-are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
-and other decreases \todo{Can we figure out the reason for this?}.
-In conclusion, the counter-ion binding seems to be overestimated in Lipid17 with the strongest affinity among the
-tested models, while with this data we cannot fully evaluate binding affinity in the other models.
+Sodium binding affinity to PE lipids has not been measured experimentally,
+but large differences to PC would be surprising. 
+The sodium binding affinity to POPE depends on the used force field (Fig. \ref{NAdensPE}),
+but lesser extend than reported previously for PC~\cite{catte16}.
+%Similarly to POPG,
+Because some simulation and ion parameters are not identical with the previous work~\cite{catte16}
+\todo{This should be finished once we have the simulation details},
+we compare POPE results to the POPC simulations ran with identical parameters (Fig. \ref{NAdensPC}).
+In Lipid17 with the strongest sodium binding affinity to POPE,
+the binding affinity is approximately similar to POPC.
+Slipids and CHARMM36 exhibit slightly, and
+GROMOS-CKP subtantially weaker binding to POPE than to POPC.
+%The Lipid17 exhibits stronger sodium binding affinity
+%to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
+%Experimental data for sodium binding to PE lipids is not available, while
+%simulations in agreement with electrometer data from NMR suggests that
+%sodium binding to POPC is weaker than in any of the simulations
+%here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}.
+Assuming that the binding
+to POPE would be similar than to POPC, the sodium binding affinity to POPE
+is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
+but substantially overestimated in Lipid17 simulation.
+%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
+%previous Amber lipid version, Lipid14.
+%\todo{The potential differences in the used ion models should be checked}.
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
+  \caption{\label{NAdensPE}
+    Sodium (solid line) and choride ion density profiles along membrane normal
+    from different simulations with PE lipids.
+  }
+\end{figure}
+
+Simulations with PG lipids give similar dependence on force field as observed in POPE simulations:
+Lipid 17 simulations exhibits stronger counter-ion binding affinity to pure POPG bilayer
+than CHARMM36, Slipids, and GROMOS-CKP simulations, which are roughly similar (Fig. \ref{CIdensPG}).
+Lipid17 also exhibits less increase in POPC headgroup
+%$\beta$ carbon
+order parameters upon addition of POPG than other simulations (Fig. \ref{HGorderparametersPCvsPEPG}),
+and lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$).
+In our previous study about PS lipids \cite{antila18}, such behaviour was related to 
+%the results suggest that
+the overestimated counterions binding and shielding the electrostatic repulsion between PG headgroups
+in bilayers.% simulated with Lipid17 parameters.
+%the underestimated increase of PC
+%headgroup order parameters upon addition of negaticely charged lipids
+%was related
+%to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
+%Similar correlation is observed here in %other than GROMOS-CKP simulation:
+%%Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
+%
+Even though the area per lipid in CHARMM36, Slipids, and GROMOS-CKP simulations is in good agreement
+with experiments (Fig. \ref{CIdensPG}), the experimental increase in 
+%In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of
+POPC headgroup order parameters upon addition of POPG are not fully reproduced (Fig. \ref{HGorderparametersPCvsPEPG}).
+%more than in CHARMM36 and Slipids simulations even though the binding affinities in these
+%simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
+%are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
+%and other decreases \todo{Can we figure out the reason for this?}.
+Therefore we conclude that the counter-ion binding affinity is overestimated in Lipid17 simulations,
+% with the strongest affinity among the tested models,
+while the other simulations are more realistic, but slight overbinding cannot be excluded.
 \begin{figure}[]
@@
     Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
-  }
-\end{figure}
-
-%Sodium binding to PC lipid bilayers is significantly overestimated by most simulations models \cite{catte16}.
-%Also, sodium binding to PS lipids seems to be overestimated, although the
-%presence of counterions complicate the comparison for negatively charged lipids \cite{antila19}.
-Similarly to POPG, the Lipid17 exhibits stronger sodium binding affinity
-to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
-In Lipid17 the sodium binding affinity to POPE is approximately similar as to POPC,
-while in other models exhibit weaker binding to POPE (Fig. \ref{NAdensPC}).
-The difference is small in Slipids and CHARMM36, while GROMOS-CKP predicts
-subtantially stronger binding to POPC.
-Experimental data for sodium binding to PE lipids is not available, while
-simulations in agreement with electrometer data from NMR suggests that
-sodium binding to POPC is weaker than in any of the simulations
-here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}. Assuming that the binding
-to POPE would be similar than to POPC, the sodium binding affinity to POPE
-is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
-but substantially overestimated in Lipid17 simulation.
-%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
-%previous Amber lipid version, Lipid14.
-\todo{The potential differences in the used ion models should be checked}.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
-  \caption{\label{NAdensPE}
-    Sodium (solid line) and choride ion density profiles along membrane normal
-    from different simulations with PE lipids.
   }
@@
 
+\clearpage
 \section{Conclusions}

### 2019-10-31 Manuscript/manuscriptPGPE.tex
Update text

@@
 % insert abstract here
-  Primarily measured but also simulated NMR order parameters will be collected also for other than phophatidylcholine
-  (these are discussed in NMRlipids I) headgroup. The information will be used to understand structural differences between 
-  different lipid molecules in bilayers.
+Abstract
 \end{abstract}
@@
 PE and PG affect membrane protein functionality \cite{hariharan18} and bind to various proteins \cite{yeagle14}.
-PE headgroup is also prone for negative membrane curvature and causes membrane fusion \cite{??}.
+PE headgroup is also prone for negative membrane curvature and causes membrane fusion \cite{Chernomordik08,calzada16}.
 Therefore, the PE and PG headgroup structures play probably essential roles in 
@@
 %In contrast to PC lipids (see \cite{botan15} and references therein).
-Some attempts to resolve conformational ensembles from NMR for PC lipids have been made,
-but lesser extend for other lipids \cite{??}.
+Some attempts to resolve conformational ensembles from NMR for PC and PE lipids have been made,
+but lesser extend for PG or PS lipids \cite{seelig77c,davis83,Semchyschyn04}.
 Classical molecular dynamics simulations could potentially give such ensembles and therefore enable
@@
 Several MD simulations of PE and PG lipids have been published especially in the context of modeling
-inner membrane of Gram-negative bacteria \cite{??}, but glycerol backbone and headgroup structures have not
+inner membrane of Gram-negative bacteria \cite{devries04,murzyn05,pedersen06,zhao07,gurtovenko08,zhao08,kukol09,tsai12,dickson12,venable13,dickson14,berglund15}
+\todo{There may be some relevant publication missing from here},
+but glycerol backbone and headgroup structures have not
 been evaluated against experiments.
@@
 Besides the structure, also ion binding may regulate biophysical activity of especially negatively
-charged lipid headgroups \cite{??}. Monovalent cation (except Lithium) binding to zwitterionic PC headgroups is
-very weak, while multivalent ion binding is stronger but still weak \cite{??}. The ion binding affinity
-data for PE is more scarce \cite{??}, but large differences to PC would be surprising.
+charged lipid headgroups \cite{seelig90}. Monovalent cation (except Lithium) binding to zwitterionic PC and anionic PS headgroups is
+very weak, while multivalent ion binding is stronger but still weak \cite{cevc90,tocanne90,roux90,catte16,antila19}. The ion binding affinity
+data for PE is more scarce \cite{marra85}, but large differences to PC would be surprising.
 Negatively charged lipids are suggested to bear same cation binding constants than zwitterionic lipids, but
@@
 %\end{displayquote}
-On the other hand, anionic PS lipids are proposed chelate with calcium ions \cite{??}.
+On the other hand, anionic PS lipids are proposed chelate with calcium ions \cite{hauser85,feigenson86,roux91}.
 In simulations, the cation binding affinity to PC and PS membranes is typically overestimated \cite{catte16,antila19},

### 2019-11-01 Manuscript/manuscriptPGPE.tex
Update text

@@
 Several MD simulations of PE and PG lipids have been published especially in the context of modeling
-inner membrane of Gram-negative bacteria \cite{devries04,murzyn05,pedersen06,zhao07,gurtovenko08,zhao08,kukol09,tsai12,dickson12,venable13,dickson14,berglund15}
+inner membrane of Gram-negative
+bacteria \cite{devries04,murzyn05,pedersen06,zhao07,gurtovenko08,zhao08,henin09,kukol09,tsai12,dickson12,venable13,dickson14,berglund15}
 \todo{There may be some relevant publication missing from here},
-but glycerol backbone and headgroup structures have not
-been evaluated against experiments.
+but evaluation of glycerol backbone and headgroup structures against experiments is rare \cite{henin09}.
 
@@
 to cholesterol was significantly overestimated by the Berger/Höltje force field in our previous work \cite{botan15}.
+\todo{This is text by P. Fuchs, copied from the blog. \\
+Area results in nm$^2$, the error is $<=$ 0.003 nm$^2$ \\
+- pure POPC \\
+CHARMM36: 0.624 \\
+Berger : 0.649 \\
+- POPC/POPE 50:50 \\
+CHARMM36 : POPC 0.609, POPE 0.557 \\
+Berger-hacked: POPC 0.637, POPE 0.632 \\
+----- \\
+One can see that CHARMM 36 predicts a drop in the area on going from pure POPC to POPC/POPE 50:50. This means that POPC pack tightly to POPE. 
+In contrast, the values for Berger are not that changed. 
+The POPE value predicted by CHARMM 36 (in the mixture POPC/POPE 50:50) is much smaller than that predicted by Berger.\\
+-------\\
+The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
+which supports the more condensed PE bilayer.
+This is interesting, but not exactly the core message of the manuscript.
+Maybe we should mention this very briefly? For example, we could just report the areas per lipid (without distinguishing PC and PE)
+and mention the difference between CHARMM36 and Berger. I have opened an issue for this: https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/7
+}
+
 None of the force fields fully reproduces the PC headgroup order parameter response to the increasing amount of
@@
 \end{figure}
-
-\todo{This is text by P. Fuchs, copied from the blog. \\
-Area results in nm$^2$, the error is $<=$ 0.003 nm$^2$ \\
-- pure POPC \\
-CHARMM36: 0.624 \\
-Berger : 0.649 \\
-- POPC/POPE 50:50 \\
-CHARMM36 : POPC 0.609, POPE 0.557 \\
-Berger-hacked: POPC 0.637, POPE 0.632 \\
------ \\
-One can see that CHARMM 36 predicts a drop in the area on going from pure POPC to POPC/POPE 50:50. This means that POPC pack tightly to POPE. 
-In contrast, the values for Berger are not that changed. 
-The POPE value predicted by CHARMM 36 (in the mixture POPC/POPE 50:50) is much smaller than that predicted by Berger.\\
--------\\
-The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
-which supports the more condensed PE bilayer.
-This is interesting, but to avoid the overexapansion of the manuscript, it is probably better to keep the focus in headgroups and ion binding also in this manuscript.}
 

### 2019-11-13 Manuscript/manuscriptPGPE.tex

@@
 for other lipids in previous studies \cite{botan15,antila19}.
-\todo{Should we comment more the relative quality of different force fields and/or make the subjective force field ranking figures?}
+\todo{Should we comment more the relative quality of different force fields and/or make the subjective force field ranking figures?
+https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/8}
 
@@
 headgroups (Fig.~\ref{structures}).
-\todo{The structures are already in Fig. \ref{structures}, but we need also the dihedral distributions to finish the discussion}.
+\todo{The structures are already in Fig. \ref{structures}, but we need also the dihedral distributions to finish the discussion.
+Maybe we should analyze also some other structural features, at least P-N angles. More discussion in issue
+https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9}.
 %While rotation around N-C$_\alpha$-C$_\beta$-O$_\alpha$ dihedral was
@@
 %\todo{Why is the $\beta$-carbon order parameter of PG different to PC and PE then?}
-
-\todo{Maybe we should analyze also some other structural features, at least P-N angles.}
 
@@
 hydrogens attached to the $\alpha$-carbon are responsing differently when mixed with PE or PG lipids
-\todo{Maybe we should figure out what is the reason for this?}.\\
-\todo{Maybe we should analyze the P-N vector angle from different simulations?}.
+\todo{Maybe we should figure out what is the reason for this? \\
+  Maybe we should analyze the P-N vector angle from different simulations? \\
+  https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/10}.
 

### 2019-11-18 Manuscript/manuscriptPGPE.tex
Update calcium binding section

@@
 %\homepage[]{Your web page}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
+%\affiliation{Institute of Organic Chemistry and Biochemistry,
+%Academy of Sciences of the Czech Republic, 
+%Prague 6, Czech Republic}
 \affiliation{Institute of Biotechnology, University of Helsinki}
@@
 \clearpage
-\subsection{Cation binding to PE and PG lipid bilayers}
+\subsection{Calcium binding to PE and PG lipid bilayers}
 The headgroup order parameters of PC lipids can used to measure ion binding
@@
 lipids \cite{borle85,macdonald87,roux90,antila19}.
-%This is demonstrated in Figs \ref{OrderParametersWithCaCl},
-%\ref{OrderParametersWithCaClBELOW1M} and \ref{OrderParameterCHANGESWithCaClBELOW1M},
-%showing order parameters for PC headgroup $\alpha$ and $\beta$ carbons
-%as a function of CaCl$_2$ concentration in the presence of different amounts of
-%negatively charged PS or PG lipids.
-%
-%PC headgroup order parameters increase when negatively charged
-%PS or PG are added to PC bilayer in the absense of added CaCl$_2$,
-%as expected based on electrometer concept \cite{seelig87}
-%(see Fig. \ref{OrderParametersWithCaClBELOW1M}).
-%Further, the order parameters decrease with the addition
-%of CaCl$_2$ and the decrease is more pronounced for systems with more
-%negatively charged lipids (see Fig. \ref{OrderParameterCHANGESWithCaClBELOW1M}).
-%At CaCl$_2$ concentrations ($\sim$ 50-300mM) where order parameters reach the values for pure PC,
-%the Ca2+ binding presumably fully cancels the charge from negative lipids and
-%overcharging occurs above these concenterations.
 The electrometer concept has been very useful in evaluating ion binding affinity in simulations
@@
 but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
-The result is similar to the $\sim$200~ns simulations with PC lipids in previous work \cite{catte16}.
-However, when simulation was continued for $\mu$s, the binding affinity substantially increased
-and interpretation was that calcium overbinds to PC lipid in CHARMM36. Therefore, the
-conclusion seems to be similar here, although the new NBfix parameters may complicate
-the situation \todo{The status of NBfix parameters in these simulations should be checked.}.
-%Anyway, the data presented in NMRlipids II project and in Fig. \ref{changesWITHCaClPG} together
-%suggest that Calcium binding is similarly overestimated by CHARMM36 model in pure POPC bilayers and mixtures with
-%POPG. The good agreement of $\alpha$ carbon would be explained by too weak dependence of its order
-%parameter of bound charge
-%\todo{The response of CHARMM36 to cationic surfactant against experiments \cite{scherer89} to be checked.
-%  I have already ran the simulations, analysis to be done.}.
+On the other hand, the simulations are not converged.
+\todo{We still need more data to finish the discussion. More detailed discussion is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}
+% The result is similar to the $\sim$200~ns simulations with PC lipids in previous work \cite{catte16}.
+%However, when simulation was continued for $\mu$s, the binding affinity substantially increased
+%and interpretation was that calcium overbinds to PC lipid in CHARMM36. Therefore, the
+%conclusion seems to be similar here, although the new NBfix parameters may complicate
+%the situation \todo{The status of NBfix parameters in these simulations should be checked.}.
 \begin{figure*}[]
@@
 
-The $\beta$-carbon order parameter of PG exhibits a rapid decrease with small CaCl$_2$ concentrations
-and a more modest decrease with larger concentrations in experiments \cite{borle85} (Fig. \ref{PSPGchangesWITHCaCl}).
-The rapid decrease with CaCl$_2$ is observed but overestimated in CHARMM36 simulation
-with POPC:POPG 1:1 mixture, but not in 4:1 mixture \todo{This is little bit weird, should be checked.}.
+%The $\beta$-carbon order parameter of PG exhibits a rapid decrease with small CaCl$_2$ concentrations
+%and a more modest decrease with larger concentrations in experiments \cite{borle85} (Fig. \ref{PSPGchangesWITHCaCl}).
+%The rapid decrease with CaCl$_2$ is observed but overestimated in CHARMM36 simulation
+%with POPC:POPG 1:1 mixture, but not in 4:1 mixture \todo{This is little bit weird, should be checked.}.
 
@@
 
-\todo{We need PC:PG simulations with CaCl$_2$ from different force fields to finish the discussion.}
+%\todo{We need PC:PG simulations with CaCl$_2$ from different force fields to finish the discussion.}
 

### 2020-01-03 Manuscript/manuscriptPGPE.tex
Add authors

@@
 
+\author{Pavel Buslaev}
+\affiliation{University of Jyv{\"a}skyl{\"a}}
+
+\author{Fernando Favela-Rosales}
+\affiliation{Departamento de Investigaci\'{o}n, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}
+
+\author{Patrick Fuchs}
+\affiliation{Paris, France}
+
+\author{Matti Javanainen}
+\affiliation{Institute of Organic Chemistry and Biochemistry of the 
+Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}
+
+\author{Jesper J. Madsen}
+\affiliation{Department of Chemistry, The University of Chicago, Chicago, Illinois, United States of America}
+\affiliation{Department of Global Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}
+
+\author{Josef Melcr}
+\affiliation{Institute of Organic Chemistry and Biochemistry of the 
+Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}
+\affiliation{Groningen Biomolecular Sciences and Biotechnology Institute 
+and The Zernike Institute for Advanced Materials, 
+University of Groningen, 9747 AG Groningen, The Netherlands}
+
+\author{Markus S. Miettinen}
+% \affiliation[Max Planck Institute of Colloids and Interfaces]{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
+\affiliation{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
+
 \author{O. H. Samuli Ollila}
 \email[]{samuli.ollila@helsinki.fi}
-%\homepage[]{Your web page}
-%\affiliation{Institute of Organic Chemistry and Biochemistry,
-%Academy of Sciences of the Czech Republic, 
-%Prague 6, Czech Republic}
 \affiliation{Institute of Biotechnology, University of Helsinki}
+
+\author{Chris G. Papadopoulos}
+\affiliation{I2BC - University Paris Sud}
+
+\author{Antonio Pe{\'o}n}
+\affiliation{Spain}
+
+\author{Thomas J. Piggot}
+\affiliation{Chemistry, University of Southampton, Highfield, Southampton SO17 1BJ, United Kingdom}
+
+\author{Pierre Poulain}
+\affiliation{Paris, France}
 
@@
 \begin{acknowledgments}
+AP is grateful to the Centro de
+Supercomputación de Galicia (CESGA) for use of the Finis
+Terrae computer
     %     Put your acknowledgments here.
 \end{acknowledgments}
-\newpage
-\appendix
-\begin{center}
-{\bf SUPPLEMENTARY INFORMATION}
-\end{center}
+%\newpage
+%\appendix
+%\begin{center}
+%{\bf SUPPLEMENTARY INFORMATION}
+%\end{center}
 

### 2020-01-03 Manuscript/manuscriptPGPEsuppl.tex
Add authors

@@
 
+\author{Pavel Buslaev}
+\affiliation{University of Jyv{\"a}skyl{\"a}}
+
+\author{Fernando Favela-Rosales}
+\affiliation[Tecnol\'{o}gico Nacional de M\'{e}xico]{Departamento de Investigaci\'{o}n, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}
+
+\author{Patrick Fuchs}
+\affiliation{Paris, France}
+
+\author{Matti Javanainen}
+\affiliation[Czech Academy of Sciences]{Institute of Organic Chemistry and Biochemistry of the 
+Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}
+
+\author{Jesper J. Madsen}
+\affiliation[University of Chicago]{Department of Chemistry, The University of Chicago, Chicago, Illinois, United States of America}
+\alsoaffiliation[University of South Florida]{Department of Global Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}
+
+\author{Josef Melcr}
+\affiliation[Czech Academy of Sciences]{Institute of Organic Chemistry and Biochemistry of the 
+Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}
+\alsoaffiliation{Groningen Biomolecular Sciences and Biotechnology Institute 
+and The Zernike Institute for Advanced Materials, 
+University of Groningen, 9747 AG Groningen, The Netherlands}
+
+\author{Markus S. Miettinen}
+% \affiliation[Max Planck Institute of Colloids and Interfaces]{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
+\affiliation{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
+
 \author{O. H. Samuli Ollila}
 \email{samuli.ollila@helsinki.fi}
-%\homepage[]{Your web page}
-\affiliation{Institute of Organic Chemistry and Biochemistry,
-Academy of Sciences of the Czech Republic, 
-Prague 6, Czech Republic}
 \affiliation{Institute of Biotechnology, University of Helsinki}
+
+\author{Chris G. Papadopoulos}
+\affiliation[]{I2BC - University Paris Sud}
+
+\author{Antonio Pe{\'o}n}
+\affiliation[]{Spain}
+
+\author{Thomas J. Piggot}
+\affiliation[University of Southampton]{Chemistry, University of Southampton, Highfield, Southampton SO17 1BJ, United Kingdom}
+
+\author{Pierre Poulain}
+\affiliation{Paris, France}
+
+%\author{O. H. Samuli Ollila}
+%\email{samuli.ollila@helsinki.fi}
+%%\homepage[]{Your web page}
+%\affiliation{Institute of Organic Chemistry and Biochemistry,
+%Academy of Sciences of the Czech Republic, 
+%Prague 6, Czech Republic}
+%\affiliation{Institute of Biotechnology, University of Helsinki}
 

### 2020-01-03 Manuscript/manuscriptPGPE.tex
Add figure from issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9 into the manuscript

@@
 headgroups (Fig.~\ref{structures}).
-\todo{The structures are already in Fig. \ref{structures}, but we need also the dihedral distributions to finish the discussion.
+\todo{The structures and dihedrals are in Fig. \ref{structures}, but we need to finish the discussion.
 Maybe we should analyze also some other structural features, at least P-N angles. More discussion in issue
@@
   \centering
-  \includegraphics[width=18.0cm]{../Figs/struct.png}
+  \includegraphics[width=18.0cm]{../Figs/PCPEPGPSstructures.png}
+  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals1.png}
+  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals2.png}
   \caption{\label{structures}
-    Overlayed snapshots from CHARMM36 simulations of different lipids which give the best agreement with experiments.
+    Overlayed snapshots and dihedral angle distributions from CHARMM36 simulations of different lipids which give the best agreement with experiments.
   }

### 2020-02-10 Manuscript/manuscriptPGPE.tex
Add results from headgroup tilt by P. Buslaev to the manuscript from https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9
Update PG, PE and PC mixtures data

@@
       \hline
+      POPE  & OPLS-MacRog \cite{rog16}        &0    & 144	& 5760     &0    & 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
+      \hline
       POPE  & Berger-Vries \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
@@
       POPC                   & Slipid \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCslipid150mMNaCl301K}  \\
-      POPC:POPG (7:3)        & Slipid \cite{??}        & ?         & 0  & ?     & ?     & ?   &  310  & 500 & 100 & \cite{??} \todoi{Zenodo entry unclear.}  \\
+      POPC:POPG (7:3)        & Slipid \cite{??}        & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
+      POPC:POPG (1:1)        & Slipid \cite{??}        & 0         & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid \cite{??}        & 0         & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid \cite{??}        & 0         & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid \cite{??}        & 0         & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid \cite{??}        & 0         & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+
+      \hline
+      POPC:POPG (4:1)        & Lipid17 \cite{??}       &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
+      POPC:POPG (4:1)        & Lipid17 \cite{??}       &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 150 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
+      POPC:POPG (4:1)        & Lipid17 \cite{??}       &0          & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17 \cite{??}       &0          & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
+      POPC:POPG (1:1)        & Lipid17 \cite{??}       &0          & 0.1& 150:150 & 31401 & 57 &  298  & 720 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17 \cite{??}       &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
       \hline
@@
   \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals2.png}
+  \includegraphics[width=8.0cm]{../Figs/HGtilt.png}
   \caption{\label{structures}

### 2020-05-06 scripts/AddData.py
Add order parameters from the CHARMM simulations that are used to analyze structures between different lipids in issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9. The results are analyzed with the new jupyter notebook that will be used to build the databank.

@@ -1,0 +1,636 @@
+#!/usr/bin/env python
+# coding: utf-8
+
+# # Database test
+
+# ## Simulation description file
+
+# In[1]:
+
+
+DOI="10.5281/zenodo.1011096"
+def_file  = "/home/samuli/work/NMRlipids/MATCH/scripts/orderParm_defs/order_parameter_definitions_MODEL_CHARMM36_POPG.def"
+
+user_information = """
+POPC multiple trajectories test
+#NMRLIPIDS BEGIN
+
+@SIM
+@SYSTEM=POPG_T298K
+@SOFTWARE=gromacs
+@FF=CHARMM36
+@FF_SOURCE=CHARMM-GUI
+@FF_DATE=??
+@TRJ=run2.xtc
+@TPR=run2.tpr
+
+#NMRLIPIDS END
+
+"""
+
+
+# Working directory
+dir_wrk  = "/home/samuli/work/temp/DATAbankTST/POPCcharmm/"
+
+
+# ## General Imports
+
+# In[2]:
+
+
+# Working with files and directories
+import os
+
+#For quering webs
+import urllib.request
+from urllib.error import URLError,HTTPError
+
+# From time monitoring
+from tqdm import tqdm
+
+# Python program to find SHA256 hash string of a file
+import hashlib
+
+# For dealing with excel and cvs 
+import pandas as pd
+pd.set_option('display.max_rows', 500)
+pd.set_option('display.max_columns', 500)
+pd.set_option('display.width', 1000)
+pd.set_option('display.max_colwidth', 1000)
+
+#To make real independent copies of lists
+from copy import deepcopy
+
+
+# ## Directories
+
+# In[3]:
+
+
+dir_wrk = os.path.normpath(dir_wrk)
+print(dir_wrk)
+def_file = os.path.normpath(def_file)
+print(def_file)
+
+
+# ## Check that the DOI link is valid
+
+# In[4]:
+
+
+DOI_url = 'https://doi.org/' + DOI
+print(DOI_url)
+
+try:
+    response = urllib.request.urlopen(DOI_url)
+    print("Status of the DOI link: {0}".format(response.msg))
+except HTTPError as e:
+    print(DOI_url)
+    print('The server couldn\'t fulfill the request.')
+    print('Error code: ', e.code)
+    user_information = ""
+    print('The code will not proceed, please fix DOI')
+except URLError as e:
+    print(DOI_url)
+    print('We failed to reach a server.')
+    print('Reason: ', e.reason)
+    user_information = ""
[…]

### 2020-05-12 Manuscript/manuscriptPGPE.tex
Add order parameter results figure from CHARMM into the manuscript
Update discussion related to differences in conformations

@@
 
-Without further discussion about poorly performing force fields, we focus on more detailed analysis
-of CHARMM36 simulations, which give the best overall agreement with experiments for
-the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
-studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
-not within the experimental error bars, it seems to capture the essential differences between different headroups.
-%(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
-In previous study, CHARMM36 predicted the more negative $\beta$-carbon order parameter and larger forking of
-the $\alpha$-carbon in PS headgroup than in PC \cite{antila19}.
-In this work,
-%the CHARMM36 simulations reproduce also the other essential differences in experimental
-%headgroup order parameters between different headgroups (Fig. \ref{HGorderParameters}) despite the inaccuracies
-%in individual segments:
-the PE headgroup order parameters are similar to the PC \cite{botan15}
-and $\beta$-carbon order parameters is positive in PG headgroup, while negative values observed in
-other lipids. Therefore, we use CHARMM36 simulations to analyze the structural differences between
-headgroups (Fig.~\ref{structures}).
-\todo{The structures and dihedrals are in Fig. \ref{structures}, but we need to finish the discussion.
-Maybe we should analyze also some other structural features, at least P-N angles. More discussion in issue
-https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9}.
-%While rotation around N-C$_\alpha$-C$_\beta$-O$_\alpha$ dihedral was
-%significantly different in PS and PC headgroups in previous work \cite{antila19}, essential differences
-%between PC, PE and PG are not observed here (Fig.~\ref{structures}) \todo{We need also the dihedral distributions to finish this discussion.}. \\
-%\todo{Why is the $\beta$-carbon order parameter of PG different to PC and PE then?}
-
 \begin{figure}[]
@@
 \end{figure}
+
+Without further discussion about poorly performing force fields, we focus on more detailed analysis
+of CHARMM36 simulations, which give the best overall agreement with experiments for
+the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
+studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
+not within the experimental error bars, it captures the essential differences between different headroups
+with the exception of $\beta$-carbon order parameter of PC (Fig.~\ref{structures}),
+which is too large when compared with PS or PE order parameter, or with experiments \cite{botan15}.
+As already pointed out in our previous study \cite{antila19}, N-C$_\beta$-C$_\alpha$-O$_\alpha$
+dihedral exhibits a more asymmetric angle distribution for PS, having largest probability being
+in gauche 270$^\degree$ conformation, than for the PC headgroup.
+Also C$_\beta$-C$_\alpha$-O$_\alpha$-P in PS has asymmetric tendency to be in this conformation.
+In PG, O$_\beta$-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral (corresponding N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral in other lipids)
+is mostly in trans conformation, and C$_\beta$-C$_\alpha$-O$_\alpha$-P  has asymmetric tendency to be in
+gauche 60$^\degree$ conformation.
+
 \begin{figure*}[!h]
@@
   \includegraphics[width=18.0cm]{../Figs/PCPEPGPSstructures.png}
+  \includegraphics[width=8.0cm]{../Figs/CHARMMfromLIPIDS.eps}
   \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals1.png}
@@
   \caption{\label{structures}
-    Overlayed snapshots and dihedral angle distributions from CHARMM36 simulations of different lipids which give the best agreement with experiments.
+    Overlayed snapshots and dihedral angle distributions from CHARMM36 simulations of different lipids
+    which give the best agreement with experiments.
   }
+  \todo{The differences observed in dihedral distributions are not visible in the snapshot figures.} \\
+  \todo{Why integral of the P-N angle distributions are not the same?} \\
+  \todo{More detailed discussion of this figure is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9} \\
 \end{figure*}

### 2020-05-15 Manuscript/manuscriptPGPE.tex
Update discussion about headgroup structures

@@
 Without further discussion about poorly performing force fields, we focus on more detailed analysis
-of CHARMM36 simulations, which give the best overall agreement with experiments for
-the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
-studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
-not within the experimental error bars, it captures the essential differences between different headroups
-with the exception of $\beta$-carbon order parameter of PC (Fig.~\ref{structures}),
-which is too large when compared with PS or PE order parameter, or with experiments \cite{botan15}.
-As already pointed out in our previous study \cite{antila19}, N-C$_\beta$-C$_\alpha$-O$_\alpha$
-dihedral exhibits a more asymmetric angle distribution for PS, having largest probability being
-in gauche 270$^\degree$ conformation, than for the PC headgroup.
-Also C$_\beta$-C$_\alpha$-O$_\alpha$-P in PS has asymmetric tendency to be in this conformation.
-In PG, O$_\beta$-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral (corresponding N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral in other lipids)
+of CHARMM36 simulations, which captures the essential differences between
+PC, PS, PG and PE headroup order parameters  (Fig.~\ref{structures}) with the exception of $\beta$-carbon order parameter of PC
+%give the best overall agreement with experiments for
+%the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
+%studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
+%not within the experimental error bars, it 
+which is too negative when compared with PS or PE order parameter, or with experiments \cite{botan15}.
+%As already pointed out in our previous study \cite{antila19},
+Characteristic dihedral conformations in PS headgroup are asymmetric conformations
+preferring gauche 270$^\circ$ conformations in N-C$_\beta$-C$_\alpha$-O$_\alpha$ and
+%dihedral exhibits a more asymmetric angle distribution for PS, having largest probability being
+%in , than for the PC headgroup.
+C$_\beta$-C$_\alpha$-O$_\alpha$-P dihedrals. 
+In PG headgroup, the O$_\beta$-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral (corresponding N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral in other lipids)
 is mostly in trans conformation, and C$_\beta$-C$_\alpha$-O$_\alpha$-P  has asymmetric tendency to be in
-gauche 60$^\degree$ conformation.
+gauche 60$^\circ$ conformation. 
+Main difference between PC and PE is the lower probability of trans state in C$_\beta$-C$_\alpha$-O$_\alpha$-P PC dihedral,
+which could be a potential reason for the too negative $\beta$ headgroup order parameter in PC.
 

### 2020-05-19 Manuscript/manuscriptPGPE.tex
Add first POPG-ECC results and update figure and text

@@
 \todo{Is this enough and correct, or should we repeat some methods from the NMRlipidsIVps paper?}
+The POPE experiments were recorded at 310~K and POPG experiments at 298~K, where the bilayers are in the liquid disordered phase \cite{marsh13}.
 
@@
 This is confirmed with SIMPSON calculations using negative value for g$_2$ and positive value for $\beta$ order parameter (Fig. \ref{POPGsimpson}).
-The POPE experiments were recorded at 310~K and POPG experiments at 298~K, where the bilayers are in the liquid disordered phase \cite{marsh13}.
 \todo{Details to be checked by Tiago}.
@@
 \subsection{Calcium binding to PE and PG lipid bilayers}
-The headgroup order parameters of PC lipids can used to measure ion binding
-affinity to lipid bilayers, because their magnitude is linearly proportional
-to the amount of bound charge in bilayer according to the molecular
-electrometer concept \cite{seelig87,catte16}. The molecular electrometer concept
-can be used also for bilayers containing PC lipids mixed with charged
-lipids \cite{borle85,macdonald87,roux90,antila19}.
-The electrometer concept has been very useful in evaluating ion binding affinity in simulations
-against experiments, because the headgroup order parameter changes as a function of ion concentration
-can be directly compared with experiments \cite{catte16,melcr18,antila19}.
+\begin{figure*}[b]
+  \centering
+  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC1.eps}
+  \caption{\label{changesWITHCaClPG}
+    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
+    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
+    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
+    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
+    The changes with respect to the systems without CaCl$_2$ are shown for other data than
+    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
+    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
+    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
+  }
+    \todo{CHARMM36 simulations should be longer and with Na couterions.}
+\end{figure*}
 
-Calcium binding affinity to PC and PS lipid bilayers was not correctly described by
-any of the standard MD simulation forced fields \cite{catte16,antila19}, while
-recently introduced force field with electronic continuum correction (ECC) performed better \cite{melcr18}.
-The decrease of $\alpha$-carbon order parameter of PC lipids in PC:PG mixtures as a function
-of calcium concentration is close to experiments CHARMM36 simulations (Fig. \ref{changesWITHCaClPG}),
-but the decrease of $\beta$-carbon order parameter seems to be overestimated.
-However, the $\beta$-carbon order parameter was not actually measured from these samples,
-but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
-On the other hand, the simulations are not converged.
-\todo{We still need more data to finish the discussion. More detailed discussion is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}
+To evaluate calcium binding to PG containing lipid bilayers in simulations,
+we calculated the changes in headgroup order parameters upon addition of CaCl$_2$
+to POPC:POPG (1:1) and (4:1) mixtures for which experimental data is available \cite{borle85,macdonald87}.
+The headgroup order parameters of PC lipids can be used to measure the ion binding
+affinity to lipid bilayers and evaluate MD simulations against experimental data,
+because their magnitude is linearly proportional to the amount of bound charge in bilayer
+and order parameters can be directly calculated from MD simulations \cite{seelig87,catte16}.
+This molecular electrometer concept can be used also for bilayers containing PC lipids mixed with charged lipids \cite{borle85,macdonald87,roux90,antila19}.
+%The electrometer concept has been very useful in evaluating ion binding affinity in simulations
+%against experiments, because the headgroup order parameter changes as a function of ion concentration
+%can be directly compared with experiments \cite{catte16,melcr18,antila19}.
+%
+
+%Calcium binding affinity to PC and PS lipid bilayers was not correctly described by
+%any of the standard MD simulation forced fields \cite{catte16,antila19}, while
+%recently introduced force field with electronic continuum correction (ECC) performed better \cite{melcr18}.
+As previously observed for PC and PS lipids, the decrease in PC headgroup order parameters with
+increasing CaCl$_2$ concentration is overestimated in Slipids and Lipid17 simulations (Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4})
+indicating too strong binding affinity of calcium into the bilayers.
+\todo{CHARMM results to be mentioned once we have the new simulations.}
+%The decrease of $\alpha$-carbon order parameter of PC lipids in PC:PG mixtures as a function
+%of calcium concentration is close to experiments CHARMM36 simulations (Fig. \ref{changesWITHCaClPG}),
+%but the decrease of $\beta$-carbon order parameter seems to be overestimated.
+%However, the $\beta$-carbon order parameter was not actually measured from these samples,
+%but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
+%On the other hand, the simulations are not converged.
+The behaviour of $\beta$-carbon order parameter of POPG is similar to POPC upon addition of calcium in experiments,
+exhibiting a rapid decrease with CaCl$_2$ concentrations below 10 mM,
+and then a more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
+The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.
+Lipid17 and Slipids force fields correctly capture the PG $\beta$-carbon order parameter response to CaCl$_2$
+even thought the binding affinity was too large based on the comparison of PC headgroup order parameter changes with experiments.
+\todo{To be finished once we have the new CHARMM simulations.} For the response of PG $\alpha$-carbon order parameters to CaCl$_2$,
+different force fields give different predictions, but experimental data to evaluate these predictions is not available. 
 % The result is similar to the $\sim$200~ns simulations with PC lipids in previous work \cite{catte16}.
@@
 %the situation \todo{The status of NBfix parameters in these simulations should be checked.}.
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG.eps}
-  \caption{\label{changesWITHCaClPG}
-    (left) The headgroup order parameters of PC from PC:PG mixtures as a function CaCl$_2$
-    concentration from experiments \cite{borle85,macdonald87} and CHARMM36 simulations.
-    Note that beta order parameter is calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}, not actually measured.
-    (right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
-    concentration from experiments \cite{borle85} and CHARMM36 simulations.
-  }
-\end{figure*}
 
@@
 
+\todo{We still need more data to finish the discussion. More detailed discussion is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}
+
 \begin{figure}[]
   \centering
-  \includegraphics[width=8.0cm]{../Figs/CAdensPG.eps}
+  \includegraphics[width=8.0cm]{../Figs/CAdensPG1PC1.eps}
   \caption{\label{CAdensPG}
-    Calcium ion density profiles along membrane normal
-    from different simulations with PG lipids.
+    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (1:1) mixtures with different force fields.
   }

### 2020-05-19 Manuscript/manuscriptPGPEsuppl.tex
Add first POPG-ECC results and update figure and text

@@
 
+\clearpage
+
+\section{Calcium binding to POPC:POPG (4:1) mixtures}
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC4.eps}
+  \caption{\label{changesWITHCaClPG1PC4}
+    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (4:1)
+    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{macdonald87} and simulations.
+    The changes with respect to the systems without CaCl$_2$ are shown for other data than
+    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
+  }
+  \todo{CHARMM36 simulations should be longer and with Na couterions.}
+\end{figure*}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/CAdensPG1PC4.eps}
+  \caption{\label{CAdensPG1PC4}
+    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (4:1) mixtures with different force fields.
+  }
+\end{figure}
+
+\clearpage
 \bibliography{refs.bib}

### 2020-05-19 Manuscript/manuscriptPGPE.tex
Update references

@@
      \hline
-      POPC                   & Slipid \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCslipid150mMNaCl301K}  \\
-      POPC:POPG (7:3)        & Slipid \cite{??}        & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
-      POPC:POPG (1:1)        & Slipid \cite{??}        & 0         & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid \cite{??}        & 0         & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid \cite{??}        & 0         & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid \cite{??}        & 0         & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid \cite{??}        & 0         & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC                   & Slipid / ?? \cite{jambeck12b,??}  & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCslipid150mMNaCl301K}  \\
+      POPC:POPG (7:3)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06}        & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06}        & 0         & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06}   & 0         & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06}   & 0         & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06}   & 0         & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06}   & 0         & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
 
       \hline
-      POPC:POPG (4:1)        & Lipid17 \cite{??}       &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
-      POPC:POPG (4:1)        & Lipid17 \cite{??}       &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 150 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
-      POPC:POPG (4:1)        & Lipid17 \cite{??}       &0          & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
-      POPC:POPG (1:1)        & Lipid17 \cite{??}       &0          & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
-      POPC:POPG (1:1)        & Lipid17 \cite{??}       &0          & 0.1& 150:150 & 31401 & 57 &  298  & 720 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
-      POPC:POPG (1:1)        & Lipid17 \cite{??}       &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}   &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 150 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
+      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}     &0          & 0.1& 150:150 & 31401 & 57 &  298  & 720 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
       \hline

### 2020-05-20 Manuscript/manuscriptPGPE.tex
Update discussion on calcium binding

@@
 \subsection{Calcium binding to PE and PG lipid bilayers}
-\begin{figure*}[b]
+\begin{figure*}[bt]
   \centering
@@
 
-To evaluate calcium binding to PG containing lipid bilayers in simulations,
-we calculated the changes in headgroup order parameters upon addition of CaCl$_2$
-to POPC:POPG (1:1) and (4:1) mixtures for which experimental data is available \cite{borle85,macdonald87}.
+To evaluate the calcium binding in simulations of lipid bilayers containing PG lipids,
+we calculated the changes in headgroup order parameters of POPC:POPG (1:1) and (4:1) mixtures
+upon addition of CaCl$_2$, and compared these with the available experimental data \cite{borle85,macdonald87}.
 The headgroup order parameters of PC lipids can be used to measure the ion binding
-affinity to lipid bilayers and evaluate MD simulations against experimental data,
-because their magnitude is linearly proportional to the amount of bound charge in bilayer
-and order parameters can be directly calculated from MD simulations \cite{seelig87,catte16}.
+affinity to lipid bilayers because their magnitude is linearly proportional to the amount of bound charge in bilayer \cite{seelig87,catte16}.
 This molecular electrometer concept can be used also for bilayers containing PC lipids mixed with charged lipids \cite{borle85,macdonald87,roux90,antila19}.
-%The electrometer concept has been very useful in evaluating ion binding affinity in simulations
-%against experiments, because the headgroup order parameter changes as a function of ion concentration
-%can be directly compared with experiments \cite{catte16,melcr18,antila19}.
-%
+The headgroup order parameters can be used to evaluate MD simulations against experimental data,
+because they can be directly calculated from MD simulations \cite{catte16}.
 
-%Calcium binding affinity to PC and PS lipid bilayers was not correctly described by
-%any of the standard MD simulation forced fields \cite{catte16,antila19}, while
-%recently introduced force field with electronic continuum correction (ECC) performed better \cite{melcr18}.
-As previously observed for PC and PS lipids, the decrease in PC headgroup order parameters with
-increasing CaCl$_2$ concentration is overestimated in Slipids and Lipid17 simulations (Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4})
-indicating too strong binding affinity of calcium into the bilayers.
+The decrease of POPC headgroup order parameters in mixtures with POPG lipids with
+increasing CaCl$_2$ concentration is overestimated in Slipids and Lipid17
+simulations (Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4})
+indicating too strong binding affinity of calcium into the bilayers as previously
+observed for pure PC lipid bilayers and mixtures with PS lipids \cite{catte16,antila19}.
 \todo{CHARMM results to be mentioned once we have the new simulations.}
@@
 %On the other hand, the simulations are not converged.
-The behaviour of $\beta$-carbon order parameter of POPG is similar to POPC upon addition of calcium in experiments,
-exhibiting a rapid decrease with CaCl$_2$ concentrations below 10 mM,
-and then a more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
-The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.
+The calcium binding affinity to lipid bilayers with PC and PS lipids
+was recently improved by applying the electronic continuum correction (ECC) to Amber Lipid14/17 force fields \cite{melcr18,melcr20}.
+In this approach, the electronic polarizability is implicitly included in the classical force fields
+by scaling the charges with constant factors \cite{leontyev11}. Here, we make a ECC-POPG force field
+by applying the scaling factors originally used for POPS also to POPG, i.e., we multiply
+charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions with 
+$f_q$=0.75 and $f_\sigma$=0.89, respectively \cite{melcr20}.
+ECC-POPG model gives a weaker calcium binding affinity (Fig.\ref{CAdensPG}) and better agreement with the experimental
+PC headgroup order parameter changes (Fig. \ref{changesWITHCaClPG}) for POPC:POPG mixtures than the original Lipid17 model
+\todo{to be finished when we have all the data}, indicating that the ECC improves the simulation predictions of
+calcium binding affinity as previously observed for PC and PS lipids \cite{melcr18,melcr20}.
+
+Experimental data for the $\beta$-carbon order parameter of POPG
+shows a rapid decrease with increasing CaCl$_2$ concentrations up to 10 mM
+and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
+This behaviour is similar to that of $\beta$-carbon order parameters of POPC,
+but essentially different than observed for POPS, where $\beta$-carbon order parameters increases with
+addition of calcium \cite{antila19}. Experimentally measured changes of PG $\alpha$-carbon order parameters upon addition of
+calcium are not available.
+%upon addition of calcium in experiments,
+%exhibiting 
+%The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.
 Lipid17 and Slipids force fields correctly capture the PG $\beta$-carbon order parameter response to CaCl$_2$
 even thought the binding affinity was too large based on the comparison of PC headgroup order parameter changes with experiments.
-\todo{To be finished once we have the new CHARMM simulations.} For the response of PG $\alpha$-carbon order parameters to CaCl$_2$,
-different force fields give different predictions, but experimental data to evaluate these predictions is not available. 
+While applying ECC to Lipid17 improved the PC headgroup order parameter response and binding affinity,
+the response of PG $\beta$-carbon order parameter to calcium is too weak in this model.
+The response of PG $\alpha$-carbon order parameters to CaCl$_2$ differs between force fields,
+but experimental data to evaluate these predictions is not available. 
+\todo{To be finished once we have the new CHARMM simulations and conformational changes of PG analyzed.}
 % The result is similar to the $\sim$200~ns simulations with PC lipids in previous work \cite{catte16}.
@@
 \end{figure}
-
-%\todo{We need PC:PG simulations with CaCl$_2$ from different force fields to finish the discussion.}
 

### 2020-05-20 Manuscript/manuscriptPGPE.tex
Update results with corrected dihedral for Lipid17 POPG from https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12

@@
       DPPE  & Slipids \cite{jambeck12b}    &0    & 288 	& 9386  &0    & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
-      POPE  & Slipids \cite{jambeck12b,??} &0    & 336	& ?     &0    & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
-      POPE  & Slipids \cite{??}            & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
-      POPE  & Slipids \cite{??} \todoi{Ion parameters?}        & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
+      POPE  & Slipids \cite{jambeck12b} &0    & 336	& ?     &0    & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
+      POPE  & Slipids \cite{jambeck12b}            & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
+      POPE  & Slipids / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
       \hline
@@
       \hline
-      POPG/Na$^+$  & Slipids \cite{jambeck13}    &0         & 288 	& 10664   &0     & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
-      DPPG/Na$^+$  & Slipids \cite{jambeck13}    &0         & 288 	& 11232  &0     & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
-      DPPG/Na$^+$  & Slipids \cite{jambeck13}    &0         & 288 	& 11232   &0     & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
-      POPG         & Slipids \cite{??} \todoi{Ion parameters?}       & 0    & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGslipids} \\
-      POPG         & Slipids \cite{??} \todoi{Ion parameters?}       & 0.11    & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
+      POPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 10664   &0     & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
+      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232  &0     & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
+      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232   &0     & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
+      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0    & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGslipids} \\
+      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0.11    & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
       \hline
-      POPG             & LIPID17 \cite{??}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
+      POPG             & LIPID17 / Dang \cite{gould18,smith94}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
       POPG             & LIPID17 \cite{??}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
@@
      \hline
-      POPC                   & Slipid / ?? \cite{jambeck12b,??}  & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCslipid150mMNaCl301K}  \\
-      POPC:POPG (7:3)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06}        & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
+      POPC                   & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
+      POPC:POPG (7:3)        & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
       POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\

### 2020-05-25 Manuscript/manuscriptPGPE.tex
Add data from MacRog POPE

@@
       POPE  & OPLS-MacRog \cite{rog16}     &0    & 144	& 5760     &0    & 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
+      POPE  & OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
       \hline
@@
       \hline
-      POPE             & LIPID17 \cite{??} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
-      POPE             & LIPID17 \cite{??} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
+      POPE             & LIPID17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
+      POPE             & LIPID17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
     \end{tabular}
@@
   %\end{sidewaystable*} 
-  \todo{We need citations for the force fields.}
+  \todo{Citation for CHARMM36 PE?} \\
+  \todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
+  \todo{Citation for GROMOS-CKP?} \\
+  \todo{Citation for GROMOS 43A1-S3?} \\
+  \todo{Citation for OPLS-UA models?} \\
+  \todo{Citations for Berger-* simulations?} \\
+  \todo{LIPID17 simulations with correct dihedrals still coming}
 \end{table*}
@@
       POPC:POPE (1:1)  & CHARMM36 \cite{??}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
+      \hline
+      POPC             & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPCmacrog300K} \\
+      POPC:POPE (1:1)  & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1macrogT300K} \\
+      \hline
+      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  300  & 170 & 100 & \cite{POPCslipid300K} \\
+      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\
      \hline
@@
     Probably larger error bars for united atom models based on the report by Fuchs et al.
-  }
+  } \\
 \end{figure}
@@
 but large differences to PC would be surprising. 
-The sodium binding affinity to POPE depends on the used force field (Fig. \ref{NAdensPE}),
+In simulations, the sodium binding affinity to POPE depends on the used force field (Fig. \ref{NAdensPE}),
 but lesser extend than reported previously for PC~\cite{catte16}.
-%Similarly to POPG,
-Because some simulation and ion parameters are not identical with the previous work~\cite{catte16}
-\todo{This should be finished once we have the simulation details},
+\todo{This will be finished once we have all the simulation details and Lipid17 simulations with correct dihedrals from issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12},
+Because some simulation and ion parameters are not identical with the previous work~\cite{catte16},
 we compare POPE results to the POPC simulations ran with identical parameters (Fig. \ref{NAdensPC}).
@@
 Simulations with PG lipids give similar dependence on force field as observed in POPE simulations:
-Lipid 17 simulations exhibits stronger counter-ion binding affinity to pure POPG bilayer
+Lipid 17 simulations with Dang ion parameters exhibits stronger counter-ion binding affinity to pure POPG bilayer
 than CHARMM36, Slipids, and GROMOS-CKP simulations, which are roughly similar (Fig. \ref{CIdensPG}).

### 2020-05-26 Manuscript/manuscriptPGPE.tex
Update ECC-lipid results

@@
   % \end{sidewaystable*}
-    \todo{We need citations for the force fields.}
+  \todo{Citations and ion model for CHARMM36?} \\
+  \todo{Lipid17 simulation with correct dihedral potentials still coming.} \\
+  \todo{Citation and ion model for GROMOS-CKP?}
 \end{table*}
@@
      \hline
+      POPC                   & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCgromosCKPT310K}  \\
+      POPC:POPG (7:3)        & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG3gromosCKPT310K} \\
+     \hline
       POPC                   & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
@@
   %\end{sidewaystable*}
-  \todo{Data for POPC:POPG mixtures by listed by Antonio Peon is missing from this table} \\
-    \todo{We need citations for the force fields.}
+  \todo{New CHARMM simulations are still to be ran.} \\
+  \todo{Citation and ion model for GROMOS-CKP?} \\
+  \todo{Citation and description for ''Berger'' model?}
 \end{table*}

### 2020-05-27 Manuscript/manuscriptPGPE.tex
Update ECC-lipid data

@@
       POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
-      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 150 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 250 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
       POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
@@
       POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
+      \hline
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \cite{Lipid17eccPOPCPOPG5050} \\
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0.1& 150:150 & 29865 & 54 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17eccPOPCPOPG50501000mMCaCl} \\
       \hline
@@
   \todo{Citation and ion model for GROMOS-CKP?} \\
-  \todo{Citation and description for ''Berger'' model?}
+  \todo{Citation and description for ''Berger'' model?} \\
+  \todo{Missing Zenodo entries of Lipid17 should be updated to the ones with corrected dihedrals.}\\
+  \todo{Lipid17ecc simulations to be added.}
 \end{table*}

### 2020-05-27 Manuscript/manuscriptPGPEsuppl.tex
Update results with corrected dihedral for Lipid17 POPG from https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12
Update ECC-lipid data

@@
 
-\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.}
+\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.
+  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.
+}
 
@@
 
-\noindent {\it POPG} Data is available at \cite{slipidsPOPGfiles}. \todo{Simulation details by F. Favela.}
+\noindent {\it POPG} Data is available at \cite{slipidsPOPGfiles}. \todo{Simulation details by F. Favela.
+  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
 
-\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.}
+\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.
+  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
 
-\noindent {\it DPPG} Data in 298~K is available at \cite{slipidsDPPGfilesT298K} and in 314~K at \cite{slipidsDPPGfiles}.  \todo{Simulation details by F. Favela.}
+\noindent {\it DPPG} Data in 298~K is available at \cite{slipidsDPPGfilesT298K} and in 314~K at \cite{slipidsDPPGfiles}.
+\todo{Simulation details by F. Favela.
+  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
 
-\noindent {\it POPC:POPG mixture with additional NaCl} \todo{Simulation details by A. Peon.}
+\noindent {\it POPC:POPG mixture with additional NaCl} \todo{Simulation details by A. Peon.
+  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
 
@@
 \noindent {\it DPPE} Data is available at \cite{gromosCKPdppe}. \todo{Simulation details by  T. Piggot.}
+
+\subsection{Lipid17}
+
+\subsection{ECC-LIPID POPG}
+In ECC-lipid models, electronic continuum correction (ECC) is applied to 
+implicitly include the missing electronic polarizability into the force field description \cite{melcr18,melcr19}.
+In practise, this is implemented by scaling the charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions
+of Amber Lipid14/17 models are scaled by constant factors.
+Here, we follow the approach that previously improved ion binding to bilayers containing negatively charged PS lipids \cite{melcr19}:
+ECC-POPC parameters (scaling factors $f_q$=0.8 and $f_\sigma$=0.89 applied to Lipid14 POPC parameters) \cite{melcr18} were used for POPC
+and scaling factors of $f_q$=0.75 and $f_\sigma$=0.89 were applied to the charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions
+of Amber Lipid17 POPG parameters. The Lipid17 parameters (described above) and initial configurations
+were taken from Ref. \cite{POPGlipid17} with the correct dihedral type,
+and the resulting parameters are available from Ref. \citenum{??}. ECC-ion parameters with the scaled charges \cite{pluharova14,kohagen16,martinek18},
+downloaded from \url{bitbucket.org/hseara/ions/src/master/}, were used in these simulations.
 

### 2020-06-04 Manuscript/manuscriptPGPE.tex
Update links to the data

--- 2020-05-26 Manuscript/manuscriptPGPE.tex
@@
       POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
-      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 150 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 250 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
       POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
       POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
-      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}     &0          & 0.1& 150:150 & 31401 & 57 &  298  & 720 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 150:150 & 31401 & 57 &  298  & 718 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
       POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
+      \hline
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \cite{Lipid17eccPOPCPOPG5050} \\
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0.1& 150:150 & 29865 & 54 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 1.0& 150:150 & 29865 & 569 &  298  & 600 & 400 & \cite{Lipid17eccPOPCPOPG50501000mMCaCl} \\
       \hline
@@
   \todo{Citation and ion model for GROMOS-CKP?} \\
-  \todo{Citation and description for ''Berger'' model?}
+  \todo{Citation and description for ''Berger'' model?} \\
+  \todo{Lipid17 POPC and POPC:POPG mixtures (https://doi.org/10.5281/zenodo.3241242 and https://doi.org/10.5281/zenodo.3237656)
+    should be added after simulated with corrected dihedrals.}\\
+  \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
 \end{table*}

### 2020-08-17 scripts/OrderParameter.py
Choose all heavy atoms for dihedral calculation

@@ -355,3 +355,97 @@ def read_trj_PN_angles(molname,atoms, top_fname, traj_fname, gro_fname):
 
     return angles, resAverageAngles, totalAverage, totalSTDerror
 
+#####Dihedrals#####
+#Anne: Functions needed for dihedral analysis
+
+class Dihedrals:
+
+    def __init__(self, resname, atom_1_name, atom_2_name, atom_3_name, atom_4_name, *args):
+        self.resname = resname       # name of residue atoms are in
+        self.at1name = atom_1_name
+        self.at2name = atom_2_name
+        self.at3name = atom_3_name
+        self.at4name = atom_4_name
+        self.name = atom_1_name + " " + atom_2_name + " " + atom_3_name + " " + atom_4_name  # generic name of atom A and atom B
+
+        self.traj = []
+        self.selection = []
+
+#def calcDihedrals(atom1, atom2, atom3, atom4):
+    
+#    return
+
+def parseDihedralInput(mapping_file):
+#reads all heavy atoms from mapping file and forms groups of four consecutive atoms for dihedral calculation
+#not ready yet!!!!
+    G1 = []
+    G2 = []
+    G3 = []
+    resname = ""
+    dihGroups = []
+
+    regexp_H = re.compile(r'M_[A-Z0-9]*H[0-9]*_M')
+    regexp_1 = re.compile(r'M_G[0-9]*[A-Z][0-9]*_M')
+#    try:
+    with open(mapping_file,"r") as f:
+        for ind, line in enumerate(f,1):
+            if line.startswith("#Whole molecules"):
+#                print(getline(f.name, ind + 1).split())
+                resname = getline(f.name, ind + 1).split()[1]
+                break
+    with open(mapping_file, "r") as f:
+        for line in f.readlines():
+            if not line.startswith("#") and not regexp_H.search(line):
+                    if line.startswith("M_G1"):
+                        atom = line.split()
+                        G1.append(atom[0])
+                    elif line.startswith("M_G2"):
+                        atom = line.split()
+                        G2.append(atom[0])
+                    elif line.startswith("M_G3"):
+                        atom = line.split()
+                        G3.append(atom[0])
+
+    G1.reverse()
+    G2.reverse()
+    print(G1)
+    print(G2)
+    print(G3)
+    G1G3 = G1 + G3
+    G2G3 = G2 + G3
+
+    dihAtoms = []
+
+    for atom in G1G3:
[…]

### 2020-09-01 scripts/DataBankINFO/AddData.py
Add results with ions and preliminary plotting script

@@ -882,24 +882,18 @@ for sim in sims_working_links:
     
     print("Calculating order parameters")
     
-    outfile=open(str(dir_tmp) + '/' + str(ID)+'/'+'OrderParameters.dat','w')
-    line1="Atom     Average OP     OP stem"+'\n'
-    outfile.write(line1)
-    
-    data = {}
-    outfile2=str(dir_tmp) + '/' + str(ID)+'/'+'OrderParameters.json' 
 
     for key in sim['MAPPING_DICT']:    
         mapping_file = sim['MAPPING_DICT'][key]
-    
-     #   try:
         OrdParam=find_OP('../mapping_files/'+mapping_file,tpr,xtcwhole)
-        
-      #  except OSError:
-            #OrdParam=find_OP(mapping_file,tpr,xtcwhole)
-        
+
+        outfile=open(str(dir_tmp) + '/' + str(ID)+'/' + key + 'OrderParameters.dat','w')
+        line1="Atom     Average OP     OP stem"+'\n'
+        outfile.write(line1)
     
-    
+        data = {}
+        outfile2=str(dir_tmp) + '/' + str(ID)+'/' + key + 'OrderParameters.json' 
+        
         for i,op in enumerate(OrdParam):
             resops =op.get_op_res
             (op.avg, op.std, op.stem) =op.get_avg_std_stem_OP
@@ -912,9 +906,10 @@ for sim in sims_working_links:
         with open(outfile2, 'w') as f:
             json.dump(data,f)
 
-    outfile.close()
-    get_ipython().system("cp {str(dir_tmp)}'/'{str(ID)}'/OrderParameters.dat' {data_directory.get(str(ID))}    ")
-    get_ipython().system("cp {str(dir_tmp)}'/'{str(ID)}'/OrderParameters.json' {data_directory.get(str(ID))}  ")
+        outfile.close()
+
+        get_ipython().system("cp {str(dir_tmp)}'/'{str(ID)}'/'{key}'OrderParameters.dat' {data_directory.get(str(ID))}    ")
+        get_ipython().system("cp {str(dir_tmp)}'/'{str(ID)}'/'{key}'OrderParameters.json' {data_directory.get(str(ID))}  ")
     
     print("Done calculating order parameters.")
 

### 2020-09-01 Manuscript/manuscriptPGPE.tex
Update P-N vector angle and order parameters results to the manuscript
Update dihedral and ion binding results in the manuscript

@@
       POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
-      POPC:POPG (1:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0  & 250:250 & 18158 & 0   &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG5050} \\ 
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
       POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.34 \todoi{Concentration calculated based in total amount of calcium ions.
 This may not be reasonable due to the lack of counterions.}  & 250:250 & 20798 & 128 &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG5050150mMCaCl} \\
-      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.36  \todoi{Concentration calculated based in total amount of calcium ions.
-This may not be reasonable due to the lack of counterions.} & 250:250 & 18114 & 445  &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
-      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 0  & 400:100 & 18664 & 0  &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG4010} \\
-      POPC:POPG (4:1)/K$^+$  & CHARMM36 \cite{??}        &0          & 1.0 \todoi{Concentration calculated based in total amount of calcium ions.
-This may not be reasonable due to the lack of counterions.} & 400:100 & 18647 & 419  &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
       \hline
@@
   \includegraphics[width=8.0cm]{../Figs/CHARMMfromLIPIDS.eps}
-  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals1.png}
-  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals2.png}
-  \includegraphics[width=8.0cm]{../Figs/HGtilt.png}
+%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals1.png}
+%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals2.png}
+  \includegraphics[width=8.0cm]{../Figs/PNangleNORM.eps}
+  \includegraphics[width=8.0cm]{../Figs/DIHEDRALS.eps}
   \caption{\label{structures}
@@
   \todo{The differences observed in dihedral distributions are not visible in the snapshot figures.} \\
-  \todo{Why integral of the P-N angle distributions are not the same?} \\
   \todo{More detailed discussion of this figure is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9} \\
@@
   }
-    \todo{CHARMM36 simulations should be longer and with Na couterions.}
+  \todo{Data for CHARMM with 100mM is still coming.}
 \end{figure*}
@@
   }
+  \todo{Density of CHARMM simulation still to be updated from the new data.}
 \end{figure}

### 2020-09-16 Manuscript/manuscriptPGPE.tex

@@
       \hline
-      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\
-      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
+%      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\
+%      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
+      POPC                   & CHARMM36 \cite{??}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
+      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
       POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
@@
       \hline
-      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  300  & 170 & 100 & \cite{POPCslipid300K} \\
-      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\
+      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  298  & 170 & 100 & \cite{POPCslipid298K} \\
+      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120  & 0  &  298  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\
      \hline
@@
      \hline
-      POPC                   & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
+      POPC                   & Slipid \cite{jambeck12b}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
       POPC:POPG (7:3)        & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
@@
   }
-  %\todo{Simulation of CHARMM36 at 298K should be maybe rerun with Gromacs 5.}
+  \todo{P-N vector angles from Berger-OPLS, Lipid17, and GROMOS-CKP simulations are yet to be analyzed.}
 \end{figure*}
@@
   }
+  \todo{P-N angles from Lipid17, Slipids, and GROMOS-CKP yet to be calculated.}
 \end{figure}
@@
   }
-  \todo{Data for CHARMM with 100mM is still coming.}
+  \todo{Data for CHARMM with 100mM is still coming.} \\
+  \todo{P-N vector angles from Lipid17 and Lipid17ecc models yet to be analyzed.}
 \end{figure*}

### 2020-10-14 scripts/DataBankINFO/AddData.py
some small errors fixed

@@ -630,15 +630,6 @@ for sim in sims_working_links :
 
 # ## Read molecule numbers into dictionary
 
-# In[21]:
-
-for sim in sims_working_links :
-    print(sim['MAPPING_DICT'])
-    mapping_filesTST = []
-    for value in sim['MAPPING_DICT'].values():
-        mapping_filesTST.append(value)
-    print(mapping_filesTST)
-
 # In[22]:
 
 
@@ -679,7 +670,7 @@ for sim in sims_working_links :
         selection = ""
         if key_mol in sim['MAPPING_DICT'].keys():
             m_file = sim['MAPPING_DICT'][key_mol]
-            with open('mapping_files/'+m_file,"r") as f:
+            with open('../mapping_files/'+m_file,"r") as f:
                     for line in f:
                         if len(line.split()) > 2 and "Individual atoms" not in line:
                             selection = selection + "(resname " + line.split()[2] + " and name " + line.split()[1] + ") and "
@@ -710,7 +701,7 @@ for sim in sims_working_links :
         selection = ""
         if key_mol in sim['MAPPING_DICT'].keys():
             m_file = sim['MAPPING_DICT'][key_mol]
-            with open('mapping_files/'+m_file,"r") as f:
+            with open('../mapping_files/'+m_file,"r") as f:
                     for line in f:
                         if len(line.split()) > 2 and "Individual atoms" not in line:
                             selection = selection + "resname " + line.split()[2] + " and name " + line.split()[1] + " or "
@@ -899,7 +890,7 @@ for sim in sims_working_links:
 
     for key in sim['MAPPING_DICT']:    
         mapping_file = sim['MAPPING_DICT'][key]
-        OrdParam=find_OP('../mapping_files/'+mapping_file,tpr,xtcwhole)
+        OrdParam=find_OP('../mapping_files/'+mapping_file,tpr,xtcwhole,key)
 
         outfile=open(str(dir_tmp) + '/' + str(ID)+'/' + key + 'OrderParameters.dat','w')
         line1="Atom     Average OP     OP stem"+'\n'

### 2020-10-14 scripts/DataBankINFO/AddData.py
Update AddData.py

@@ -637,6 +637,7 @@ for sim in sims_working_links :
 
 # ## Read molecule numbers into dictionary
 
+
 # In[22]:
 
 
@@ -651,9 +652,9 @@ from MDAnalysis import Universe
     
 for sim in sims_working_links :
     ID = sim.get('ID')
-    tpr = str(dir_wrk)+ '/tmp/' + str(ID) + '/' + str(sim.get('TPR')).translate({ord(c): None for c in "']["})
-    trj = str(dir_wrk)+ '/tmp/' + str(ID) + '/' + str(sim.get('TRJ')).translate({ord(c): None for c in "']["})
-    gro = str(dir_wrk)+ '/tmp/' + str(ID) + '/conf.gro'
+    tpr = str(dir_tmp)+ '/' + str(ID) + '/' + str(sim.get('TPR')).translate({ord(c): None for c in "']["})
+    trj = str(dir_tmp)+ '/' + str(ID) + '/' + str(sim.get('TRJ')).translate({ord(c): None for c in "']["})
+    gro = str(dir_tmp)+ '/' + str(ID) + '/conf.gro'
     print(gro)
     
  #   if str(sim.get('INI')).translate({ord(c): None for c in "']["}) != '':
@@ -662,7 +663,7 @@ for sim in sims_working_links :
 
   #  else:
     get_ipython().system('echo System | gmx trjconv -f {trj} -s {tpr} -dump 0 -o {gro}')
-    gro_path = str(dir_wrk) + '/tmp/' + str(ID) + '/' + 'conf.gro'
+    gro_path = str(dir_tmp) + '/' + str(ID) + '/' + 'conf.gro'
     
     
     

### 2020-10-14 scripts/DataBankINFO/POPCPOPG4:1lipid17eccINFO.py

@@ -20,11 +20,11 @@ POPC:POPG (4:1) lipid17ecc
 @POPS=POPS
 @POPE=POPE
 
-@POT=M_K_M
-@SOD=M_NA_M
-@CLA=M_CL_M
-@CAL=M_CA_M
-@SOL=M_SOL_M
+@POT=K
+@SOD=NA
+@CLA=CL
+@CAL=CA
+@SOL=SOL
 
 @NPOPC=[0,0]
 @NPOPG=[0,0]
@@ -43,4 +43,4 @@ POPC:POPG (4:1) lipid17ecc
 #NMRLIPIDS END
 
 """
-dir_wrk = "/media/osollila/Data/tmp/DATABANK/"
+dir_wrk = "/media/osollila/Data1/tmp/DATABANK/"

### 2020-10-14 scripts/DataBankINFO/POPCPOPGcharmmINFO.py

@@ -20,11 +20,11 @@ POPC:POPG (1:1) charmm-gui
 @POPS=POPS
 @POPE=POPE
 
-@POT=M_K_M
-@SOD=M_NA_M
-@CLA=M_CL_M
-@CAL=M_CA_M
-@SOL=M_SOL_M
+@POT=K
+@SOD=SOD
+@CLA=CL
+@CAL=CA
+@SOL=SOL
 
 @NPOPC=[0,0]
 @NPOPG=[0,0]
@@ -43,4 +43,4 @@ POPC:POPG (1:1) charmm-gui
 #NMRLIPIDS END
 
 """
-dir_wrk = "/media/osollila/Data/tmp/DATABANK/"
+dir_wrk = "/media/osollila/Data1/tmp/DATABANK/"

### 2020-10-15 scripts/DataBankINFO/AddData.py
fixed AddData.py, OrderParameter.py

@@ -654,27 +654,26 @@ for sim in sims_working_links :
     ID = sim.get('ID')
     tpr = str(dir_tmp)+ '/' + str(ID) + '/' + str(sim.get('TPR')).translate({ord(c): None for c in "']["})
     trj = str(dir_tmp)+ '/' + str(ID) + '/' + str(sim.get('TRJ')).translate({ord(c): None for c in "']["})
-    gro = str(dir_tmp)+ '/' + str(ID) + '/conf.gro'
-    print(gro)
-    
+    gro = str(dir_tmp) + '/' + str(ID) + '/' + 'conf.gro'
  #   if str(sim.get('INI')).translate({ord(c): None for c in "']["}) != '':
  #       gro = str(sim.get('INI')).translate({ord(c): None for c in "']["})
  #       gro_path = str(dir_wrk) + '/tmp/' + str(ID) + '/' + gro
 
   #  else:
     get_ipython().system('echo System | gmx trjconv -f {trj} -s {tpr} -dump 0 -o {gro}')
-    gro_path = str(dir_tmp) + '/' + str(ID) + '/' + 'conf.gro'
+  #  gro_path = str(dir_tmp) + '/' + str(ID) + '/' + 'conf.gro'
     
     
     
     leaflet1 = 0 #total number of lipids in upper leaflet
     leaflet2 = 0 #total number of lipids in lower leaflet
     
-    u = Universe(gro_path)
+    u = Universe(gro)
     lipids = []
 
 # select lipids 
     for key_mol in lipids_dict:
+        print(key_mol)
         selection = ""
         if key_mol in sim['MAPPING_DICT'].keys():
             m_file = sim['MAPPING_DICT'][key_mol]
@@ -686,19 +685,23 @@ for sim in sims_working_links :
                             continue
                         else:
                             selection = "resname " + sim.get(key_mol)
+                            print(selection)
                             break
         selection = selection.rstrip(' and ')
-  #      print(selection)
+        print(selection)
         molecules = u.select_atoms(selection)
        # print(molecules)
         if molecules.n_residues > 0:
-            lipids.append(selection)
+            lipids.append(u.select_atoms(selection))
 # join all the selected the lipids together to make a selection of the entire membrane and calculate the
 # z component of the centre of mass of the membrane
-    membrane = " and ".join(str(x) for x in lipids)
- #   print(membrane)
-    R_membrane_z = u.select_atoms(membrane).center_of_mass()[2]
[…]

### 2020-10-15 scripts/DataBankINFO/OrderParameter.py

@@ -182,9 +182,9 @@ def read_trajs_calc_OPs(ordPars, top, trajs):
         for res in selection:
             # check if we have only 2 atoms (A & B) selected
             if res.n_atoms != 2:
-                #print(res.resnames, res.resids)
+                print(res.resnames, res.resids)
                 for atom in res.atoms:
-#                   print(atom.name, atom.id)
+                   print(atom.name, atom.id)
                    atA=op.atAname
                    atB=op.atBname
                    nat=res.n_atoms

### 2020-10-15 scripts/DataBankINFO/AddData.py
Fix for proper ouput of lipids with zero quantity

@@ -733,7 +733,7 @@ for sim in sims_working_links :
                 elif R[2] - R_membrane_z < 0:
                     leaflet2 = leaflet2 +1
                        # print('layer2  ' + str(leaflet2))
-            sim[x] = [leaflet1, leaflet2] 
+        sim[x] = [leaflet1, leaflet2] 
 
         
 #print("upper leaflet: " + str(leaflet1))

### 2020-10-21 scripts/DataBankINFO/AddData.py
Update data adding to databank, lipid17 entries do not work yet.

@@ -690,7 +690,6 @@ for sim in sims_working_links :
         selection = selection.rstrip(' and ')
         print(selection)
         molecules = u.select_atoms(selection)
-       # print(molecules)
         if molecules.n_residues > 0:
             lipids.append(u.select_atoms(selection))
 # join all the selected the lipids together to make a selection of the entire membrane and calculate the

### 2020-10-21 scripts/DataBankINFO/POPCPOPG100mMCaClcharmmINFO.py

@@ -1,10 +1,10 @@
-DOI="10.5281/zenodo.3997135"
+DOI="10.5281/zenodo.4005515"
 user_information = """
-POPC:POPG (1:1) +1M CaCl charmm-gui
+POPC:POPG (1:1) +100 mM CaCl charmm-gui
 #NMRLIPIDS BEGIN
 
 @SIM
-@SYSTEM=POPC:POPG(1:1)+1MCaCl_T298K
+@SYSTEM=POPC:POPG(1:1)+100mMCaCl_T298K
 @MAPPING=POPC,mappingPOPCcharmm.txt,POPG,mappingPOPGcharmm.txt
 @SOFTWARE=gromacs
 @FF=CHARMM36

### 2020-10-21 scripts/DataBankINFO/POPCPOPG1:1_1000mMCaCl_lipid17eccINFO.py

@@ -1,18 +1,18 @@
-DOI="10.5281/zenodo.3997135"
+DOI="10.5281/zenodo.3862036"
 user_information = """
-POPC:POPG (1:1) +1M CaCl charmm-gui
+POPC:POPG (1:1) lipid17ecc 1000mM CaCl
 #NMRLIPIDS BEGIN
 
 @SIM
-@SYSTEM=POPC:POPG(1:1)+1MCaCl_T298K
-@MAPPING=POPC,mappingPOPCcharmm.txt,POPG,mappingPOPGcharmm.txt
+@SYSTEM=POPC:POPG(1:1)_T298K_1000mM_CaCl
+@MAPPING=POPC,mappingPOPClipid17ecc.txt,POPG,mappingPOPGlipid17ecc.txt
 @SOFTWARE=gromacs
-@FF=CHARMM36
-@FF_SOURCE=CHARMM-GUI
+@FF=lipid17ecc
+@FF_SOURCE=NMRlipidsIV
 @FF_DATE=?/?/2020
-@TRJ=100-400ns.xtc
-@TPR=run_400ns.tpr
-@PREEQTIME=100
+@TRJ=traj.xtc
+@TPR=topol.tpr
+@PREEQTIME=200
 @TIMELEFTOUT=0
 
 @POPC=POPC
@@ -21,10 +21,10 @@ POPC:POPG (1:1) +1M CaCl charmm-gui
 @POPE=POPE
 
 @POT=K
-@SOD=SOD
-@CLA=CLA
-@CAL=CAL
-@SOL=TIP3
+@SOD=NA
+@CLA=CL
+@CAL=CA
+@SOL=SOL
 
 @NPOPC=[0,0]
 @NPOPG=[0,0]

### 2020-10-21 scripts/DataBankINFO/POPCPOPG1:1_100mMCaCl_lipid17eccINFO.py

@@ -1,17 +1,17 @@
-DOI="10.5281/zenodo.3997135"
+DOI="10.5281/zenodo.3855729"
 user_information = """
-POPC:POPG (1:1) +1M CaCl charmm-gui
+POPC:POPG (1:1) lipid17ecc 100mM CaCl
 #NMRLIPIDS BEGIN
 
 @SIM
-@SYSTEM=POPC:POPG(1:1)+1MCaCl_T298K
-@MAPPING=POPC,mappingPOPCcharmm.txt,POPG,mappingPOPGcharmm.txt
+@SYSTEM=POPC:POPG(1:1)_T298K_100mM_CaCl
+@MAPPING=POPC,mappingPOPClipid17ecc.txt,POPG,mappingPOPGlipid17ecc.txt
 @SOFTWARE=gromacs
-@FF=CHARMM36
-@FF_SOURCE=CHARMM-GUI
+@FF=lipid17ecc
+@FF_SOURCE=NMRlipidsIV
 @FF_DATE=?/?/2020
-@TRJ=100-400ns.xtc
-@TPR=run_400ns.tpr
+@TRJ=traj.xtc
+@TPR=topol.tpr
 @PREEQTIME=100
 @TIMELEFTOUT=0
 
@@ -21,10 +21,10 @@ POPC:POPG (1:1) +1M CaCl charmm-gui
 @POPE=POPE
 
 @POT=K
-@SOD=SOD
-@CLA=CLA
-@CAL=CAL
-@SOL=TIP3
+@SOD=NA
+@CLA=CL
+@CAL=CA
+@SOL=SOL
 
 @NPOPC=[0,0]
 @NPOPG=[0,0]

### 2020-10-21 scripts/DataBankINFO/POPCPOPG1:1lipid17eccINFO.py

@@ -1,19 +1,19 @@
-DOI="10.5281/zenodo.3997135"
+DOI="10.5281/zenodo.3859339"
 user_information = """
-POPC:POPG (1:1) +1M CaCl charmm-gui
+POPC:POPG (1:1) lipid17ecc
 #NMRLIPIDS BEGIN
 
 @SIM
-@SYSTEM=POPC:POPG(1:1)+1MCaCl_T298K
-@MAPPING=POPC,mappingPOPCcharmm.txt,POPG,mappingPOPGcharmm.txt
+@SYSTEM=POPC:POPG(1:1)_T298K
+@MAPPING=POPC,mappingPOPClipid17ecc.txt,POPG,mappingPOPGlipid17ecc.txt
 @SOFTWARE=gromacs
-@FF=CHARMM36
-@FF_SOURCE=CHARMM-GUI
+@FF=lipid17ecc
+@FF_SOURCE=NMRlipidsIV
 @FF_DATE=?/?/2020
-@TRJ=100-400ns.xtc
-@TPR=run_400ns.tpr
-@PREEQTIME=100
-@TIMELEFTOUT=0
+@TRJ=traj.xtc
+@TPR=topol.tpr
+@PREEQTIME=0
+@TIMELEFTOUT=15
 
 @POPC=POPC
 @POPG=POPG
@@ -21,10 +21,10 @@ POPC:POPG (1:1) +1M CaCl charmm-gui
 @POPE=POPE
 
 @POT=K
-@SOD=SOD
-@CLA=CLA
-@CAL=CAL
-@SOL=TIP3
+@SOD=NA
+@CLA=CL
+@CAL=CA
+@SOL=SOL
 
 @NPOPC=[0,0]
 @NPOPG=[0,0]

### 2020-10-21 scripts/DataBankINFO/POPCPOPG4101MCaCllipid17eccINFO.py

@@ -1,14 +1,14 @@
-DOI="10.5281/zenodo.3997135"
+DOI="10.5281/zenodo.3997176"
 user_information = """
-POPC:POPG (1:1) +1M CaCl charmm-gui
+POPC:POPG (4:1) lipid17ecc + 0.1 M CaCl
 #NMRLIPIDS BEGIN
 
 @SIM
-@SYSTEM=POPC:POPG(1:1)+1MCaCl_T298K
-@MAPPING=POPC,mappingPOPCcharmm.txt,POPG,mappingPOPGcharmm.txt
+@SYSTEM=POPC:POPG(4:1)_T298K_0.1MCaCl
+@MAPPING=POPC,mappingPOPClipid17ecc.txt,POPG,mappingPOPGlipid17ecc.txt
 @SOFTWARE=gromacs
-@FF=CHARMM36
-@FF_SOURCE=CHARMM-GUI
+@FF=lipid17ecc
+@FF_SOURCE=NMRlipidsIV
 @FF_DATE=?/?/2020
 @TRJ=100-400ns.xtc
 @TPR=run_400ns.tpr
@@ -21,10 +21,10 @@ POPC:POPG (1:1) +1M CaCl charmm-gui
 @POPE=POPE
 
 @POT=K
-@SOD=SOD
-@CLA=CLA
-@CAL=CAL
-@SOL=TIP3
+@SOD=NA
+@CLA=CL
+@CAL=CA
+@SOL=SOL
 
 @NPOPC=[0,0]
 @NPOPG=[0,0]

### 2020-10-21 scripts/DataBankINFO/POPCPOPG4:11MCaCllipid17eccINFO.py

@@ -1,14 +1,14 @@
-DOI="10.5281/zenodo.3997135"
+DOI="10.5281/zenodo.3997184"
 user_information = """
-POPC:POPG (1:1) +1M CaCl charmm-gui
+POPC:POPG (4:1) lipid17ecc + 1 M CaCl
 #NMRLIPIDS BEGIN
 
 @SIM
-@SYSTEM=POPC:POPG(1:1)+1MCaCl_T298K
-@MAPPING=POPC,mappingPOPCcharmm.txt,POPG,mappingPOPGcharmm.txt
+@SYSTEM=POPC:POPG(4:1)_T298K_1MCaCl
+@MAPPING=POPC,mappingPOPClipid17ecc.txt,POPG,mappingPOPGlipid17ecc.txt
 @SOFTWARE=gromacs
-@FF=CHARMM36
-@FF_SOURCE=CHARMM-GUI
+@FF=lipid17ecc
+@FF_SOURCE=NMRlipidsIV
 @FF_DATE=?/?/2020
 @TRJ=100-400ns.xtc
 @TPR=run_400ns.tpr
@@ -21,10 +21,10 @@ POPC:POPG (1:1) +1M CaCl charmm-gui
 @POPE=POPE
 
 @POT=K
-@SOD=SOD
-@CLA=CLA
-@CAL=CAL
-@SOL=TIP3
+@SOD=NA
+@CLA=CL
+@CAL=CA
+@SOL=SOL
 
 @NPOPC=[0,0]
 @NPOPG=[0,0]

### 2020-10-21 scripts/DataBankINFO/POPCPOPG7:3lipid17T310KINFO.py

@@ -1,18 +1,18 @@
-DOI="10.5281/zenodo.3693681"
+DOI="10.5281/zenodo.3241243"
 user_information = """
-POPC:POPG (4:1) lipid17
+POPC:POPG (7:3) lipid17
 #NMRLIPIDS BEGIN
 
 @SIM
-@SYSTEM=POPC:POPG(4:1)_T298K
+@SYSTEM=POPC:POPG(7:3)_T310K
 @MAPPING=POPC,mappingPOPClipid17.txt,POPG,mappingPOPGlipid17.txt
 @SOFTWARE=gromacs
 @FF=lipid17
-@FF_SOURCE=10.5281/zenodo.2585523
+@FF_SOURCE=??
 @FF_DATE=??
-@TRJ=trajSKIP.xtc
-@TPR=topol.tpr
-@PREEQTIME=50
+@TRJ=total_4_500.xtc
+@TPR=md_0.tpr
+@PREEQTIME=400
 @TIMELEFTOUT=0
 
 @POPC=system1
@@ -21,7 +21,7 @@ POPC:POPG (4:1) lipid17
 @POPE=POPE
 
 @POT=K
-@SOD=NA
+@SOD=NA+
 @CLA=CL
 @CAL=CA
 @SOL=SOL

### 2020-10-21 scripts/DataBankINFO/POPClipid17T310KINFO.py

@@ -1,18 +1,18 @@
-DOI="10.5281/zenodo.3693681"
+DOI="10.5281/zenodo.3237657"
 user_information = """
-POPC:POPG (4:1) lipid17
+POPC lipid17
 #NMRLIPIDS BEGIN
 
 @SIM
-@SYSTEM=POPC:POPG(4:1)_T298K
-@MAPPING=POPC,mappingPOPClipid17.txt,POPG,mappingPOPGlipid17.txt
+@SYSTEM=POPC_T310K
+@MAPPING=POPC,mappingPOPClipid17.txt
 @SOFTWARE=gromacs
 @FF=lipid17
-@FF_SOURCE=10.5281/zenodo.2585523
+@FF_SOURCE=??
 @FF_DATE=??
-@TRJ=trajSKIP.xtc
-@TPR=topol.tpr
-@PREEQTIME=50
+@TRJ=total_4_500.xtc
+@TPR=md_0.tpr
+@PREEQTIME=400
 @TIMELEFTOUT=0
 
 @POPC=system1
@@ -21,7 +21,7 @@ POPC:POPG (4:1) lipid17
 @POPE=POPE
 
 @POT=K
-@SOD=NA
+@SOD=NA+
 @CLA=CL
 @CAL=CA
 @SOL=SOL

### 2020-10-21 scripts/DataBankINFO/POPGlipid17T310KINFO.py

@@ -1,18 +1,18 @@
-DOI="10.5281/zenodo.3693681"
+DOI="10.5281/zenodo.3832219"
 user_information = """
-POPC:POPG (4:1) lipid17
+POPG lipid17
 #NMRLIPIDS BEGIN
 
 @SIM
-@SYSTEM=POPC:POPG(4:1)_T298K
-@MAPPING=POPC,mappingPOPClipid17.txt,POPG,mappingPOPGlipid17.txt
+@SYSTEM=POPG_T310K
+@MAPPING=POPG,mappingPOPGlipid17.txt
 @SOFTWARE=gromacs
 @FF=lipid17
-@FF_SOURCE=10.5281/zenodo.2585523
+@FF_SOURCE=??
 @FF_DATE=??
-@TRJ=trajSKIP.xtc
-@TPR=topol.tpr
-@PREEQTIME=50
+@TRJ=total_4_500_last.xtc
+@TPR=md.tpr
+@PREEQTIME=400
 @TIMELEFTOUT=0
 
 @POPC=system1
@@ -21,10 +21,10 @@ POPC:POPG (4:1) lipid17
 @POPE=POPE
 
 @POT=K
-@SOD=NA
+@SOD=NA_d
 @CLA=CL
 @CAL=CA
-@SOL=SOL
+@SOL=WAT
 
 @NPOPC=[0,0]
 @NPOPG=[0,0]

### 2020-10-21 Manuscript/manuscriptPGPE.tex
Update manuscript and data

@@
       POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
-      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.34 \todoi{Concentration calculated based in total amount of calcium ions.
-This may not be reasonable due to the lack of counterions.}  & 250:250 & 20798 & 128 &  298  & 200 & 200 & \cite{CHARMM36POPCPOPG5050150mMCaCl} \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
       POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
@@
   %\end{sidewaystable*}
-  \todo{New CHARMM simulations are still to be ran.} \\
   \todo{Citation and ion model for GROMOS-CKP?} \\
@@
   \todo{More detailed discussion of this figure is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9} \\
+\end{figure*}
+
+\begin{figure*}[!h]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPCfromPDB.eps}
+  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPEfromPDB.eps}
+  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPGfromPDB.eps}
+  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPSfromPDB.eps}
+  \caption{\label{structures}
+    Dihedral distributions from simulations and lipid structures in PDB.
+  }
 \end{figure*}
@@
   }
-  \todo{Data for CHARMM with 100mM is still coming.} \\
-  \todo{P-N vector angles from Lipid17 and Lipid17ecc models yet to be analyzed.}
+  \todo{P-N vector angles from Lipid17 model yet to be analyzed.}
 \end{figure*}

### 2020-11-03 Manuscript/manuscriptPGPE.tex
Update PNangle results

@@
   }
-  \todo{P-N vector angles from Berger-OPLS, Lipid17, and GROMOS-CKP simulations are yet to be analyzed.}
+  \todo{P-N vector angles from Berger-OPLS and GROMOS-CKP simulations are yet to be analyzed.}
 \end{figure*}
@@
   }
-  \todo{P-N angles from Lipid17, Slipids, and GROMOS-CKP yet to be calculated.}
+  \todo{P-N angles from Slipids and GROMOS-CKP yet to be calculated.}
 \end{figure}
@@
   }
-  \todo{P-N vector angles from Lipid17 model yet to be analyzed.}
 \end{figure*}
@@
   }
-  \todo{Density of CHARMM simulation still to be updated from the new data.}
 \end{figure}

### 2020-11-03 scripts/DataBankINFO/POPCPOPG7:3lipid17T310KINFO.py

@@ -43,4 +43,4 @@ POPC:POPG (7:3) lipid17
 #NMRLIPIDS END
 
 """
-dir_wrk = "/media/osollila/Data1/tmp/DATABANK/"
+dir_wrk = "/media/osollila/Data/tmp/DATABANK/"

### 2020-11-03 scripts/DataBankINFO/POPClipid17T310KINFO.py

@@ -43,4 +43,4 @@ POPC lipid17
 #NMRLIPIDS END
 
 """
-dir_wrk = "/media/osollila/Data1/tmp/DATABANK/"
+dir_wrk = "/media/osollila/Data/tmp/DATABANK/"

### 2020-11-03 scripts/DataBankINFO/POPGlipid17T310KINFO.py

@@ -43,4 +43,4 @@ POPG lipid17
 #NMRLIPIDS END
 
 """
-dir_wrk = "/media/osollila/Data1/tmp/DATABANK/"
+dir_wrk = "/media/osollila/Data/tmp/DATABANK/"

### 2020-11-03 scripts/calcDIHEDRALSgromacs.py

@@ -165,7 +165,7 @@ def calcDihedrals(lipids,DIHatoms):
                                 ndxfile.write('[ '+ atom1 + '_' + atom2 + '_' + atom3 + '_' + atom4 + ' ] \n')
                                 for item in selection:
                                     ndxfile.write(str(item) + " ")
-                                ndxfile.write(' \n')
+                                    ndxfile.write(' \n')
                             
                             #get_ipython().system('echo "{res}_&_{atom1}_{atom2}_{atom3}_{atom4} \n" | gmx angle -f {xtcwhole} -od {outfilename} -type dihedral -n {ndx_name} -xvg none ')
                             get_ipython().system('gmx angle -f {xtcwhole} -od {outfilename} -type dihedral -n {ndx_name} -xvg none ')               

### 2020-11-03 scripts/calcPNvectors.py

@@ -186,7 +186,7 @@ for subdir, dirs, files in os.walk(r'../Data/Simulations/'):
 # In[45]:
 
 
-print(totalAverage,totalAverageError)
+#print(totalAverage,totalAverageError)
 
 
 # In[ ]:

### 2020-11-03 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 
+Absolute values of the headgroup and glycerol backbone order parameters from PE and PG lipids are measured
+previously using $^2$H NMR~\cite{seelig76,gally81,wohlgemuth80,borle85}. Because also the order parameter
+signs bear essential information about the lipid structures \cite{botan15,ollila16}, we measured the
+magnitudes and signs of POPE and POPG C--H bond headgroup and glycerol backbone order parameter in liquid phase
+using the 2D-RPDLF and S-DROSS experiments, as described previously \cite{ferreira13,ferreira16,antila19}.
+For POPE, the glycerol backbone and $\alpha$-carbon peaks in INEPT spectra were assigned based on
+previously measured POPC spectra~\cite{ferreira13} and
+the $\beta$-carbon peak was assigned based on $^{13}$C chemical shift table for amines available
+at \url{https://www.chem.wisc.edu/areas/reich/nmr/c13-data/cdata.htm} (Fig. \ref{POPEspectra}).
+For POPG, the glycerol backbone peaks in INEPT spectra were assigned based on
+previously measured POPC spectra~\cite{ferreira13}, while $\alpha$ and  $\gamma$-carbon peaks
+\todo{How were these assigned?} (Fig. \ref{POPGspectra}). The numerical value of the $\beta$-carbon
+order parameter could not be determined, because its peak overlapped with the g$_2$ peak from glycerol backbone in POPG.
+However, the order parameter of $\beta$-carbon is expected to be clearly smaller than for g$_2$
+based on previous $^2$H NMR measurements \cite{wohlgemuth80,gally81,borle85}.
+Therefore, the beginning of the S-DROSS curve gives the sign for g$_2$ order parameter and end for $\beta$ (Fig. \ref{POPGspectra} (E)).
+This is confirmed with SIMPSON calculations using negative value for g$_2$ and positive value for $\beta$ order parameter (Fig. \ref{POPGsimpson}).
+\todo{Details to be checked by Tiago}.
+
 \subsection{Molecular dynamics simulations}
@@
 
-\begin{table*}[htb]
-  %\begin{sidewaystable*}[!p]
-  \centering
-  \caption{List of MD simulations with PE lipids.
-    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
-    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
-    % The lipid force fields named as in our previous work~\cite{botan15}.
-  }\label{systemsPE}
-  \begin{minipage}[t]{\textwidth}
-    \begin{tabular}{l c c r r r r r r c c}
-      %\hline
-      % some footnotes are not visible in typeset-MS (pdf)
-      lipid/counter-ions & force field for lipids / ions  & NaCl (M) & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
-      \hline
-      POPE  & CHARMM36 \cite{??}           &0       & 144	& 5760  &0    & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
-      POPE  & CHARMM36 \cite{??}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
-      POPE  & CHARMM36 \cite{??}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
-      POPE  & CHARMM36ua \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
-      \hline
-      DPPE  & Slipids \cite{jambeck12b}    &0    & 288 	& 9386  &0    & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
-      POPE  & Slipids \cite{jambeck12b} &0    & 336	& ?     &0    & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
-      POPE  & Slipids \cite{jambeck12b}            & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
-      POPE  & Slipids / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
-      \hline
-      DPPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3655  &0    & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \\
-      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
-      POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
-      POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
-      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
-      \hline
-      POPE  & GROMOS 43A1-S3 \cite{??}     &0    & 128	& 3552     &0    & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles}  \\
-      \hline
-      POPE  & OPLS-UA vdW on H \cite{??}   &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \\
-      POPE  & OPLS-UA \cite{??}            &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \\
-      \hline
-      POPE  & OPLS-MacRog \cite{rog16}     &0    & 144	& 5760     &0    & 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
-      POPE  & OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
-      \hline
-      POPE  & Berger-Vries \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
-      POPE  & Berger-largeH \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
-      DOPE  & Berger-Vries \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
-      DOPE  & Berger-largeH \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
-      \hline
-      POPE             & LIPID17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
-      POPE             & LIPID17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
-    \end{tabular}
-  \end{minipage}
-  %\end{sidewaystable*} 
-  \todo{Citation for CHARMM36 PE?} \\
-  \todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
-  \todo{Citation for GROMOS-CKP?} \\
-  \todo{Citation for GROMOS 43A1-S3?} \\
-  \todo{Citation for OPLS-UA models?} \\
-  \todo{Citations for Berger-* simulations?} \\
-  \todo{LIPID17 simulations with correct dihedrals still coming}
-\end{table*}
+\subsection{Analysis of molecular dynamics simulation data}
+The big data set of MD simulations was analysed in the NMRlipids databank manner.
+Unique naming convention for lipid atoms in each force field was defined using the mapping files
+and analysis for all simulations indexed in NMRlipids databank manner were performed
+using python codes.
 
-      \begin{table*}[htb]
-  %\begin{sidewaystable*}[!p]
-  \centering
-  \caption{List of MD simulations with PG lipids.
-    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
-    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
-    % The lipid force fields named as in our previous work~\cite{botan15}.
-  }\label{systemsPG}
-  \begin{minipage}[t]{\textwidth}
-    \begin{tabular}{l c c r r r r r r c c}
-      %\hline
-      % some footnotes are not visible in typeset-MS (pdf)
-      lipid/counter-ions & force field for lipids / ions & NaCl (M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
-      \hline
-      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}    &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
-      POPG             & CHARMM36 \cite{??}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
-      POPG             & CHARMM36 \cite{??}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
-      \hline
-      POPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 10664   &0     & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
-      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232  &0     & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
-      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232   &0     & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
-      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0    & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGslipids} \\
-      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0.11    & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
-      \hline
-      POPG             & LIPID17 / Dang \cite{gould18,smith94}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
-      POPG             & LIPID17 \cite{??}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
-      \hline
-      POPG             & GROMOS-CKP \cite{??}         & 0              & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGgromosCKP} \\
-      POPG             & GROMOS-CKP \cite{??}         & 0.11           & 500 & 25000 & 49 &  310  & 500 & 100 & \cite{POPGgromosCKP150mMNaCl} \\
-    \end{tabular}
-  \end{minipage}
-  % \end{sidewaystable*}
-  \todo{Citations and ion model for CHARMM36?} \\
-  \todo{Lipid17 simulation with correct dihedral potentials still coming.} \\
-  \todo{Citation and ion model for GROMOS-CKP?}
-\end{table*}
-
-\begin{table*}[htb]
-  %\begin{sidewaystable*}[!p]
-  \centering
-  \caption{List of MD simulations with PE and PG lipids mixed with PC.
-    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
-    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
-    % The lipid force fields named as in our previous work~\cite{botan15}.
-  }\label{systemsMIX}
-  \begin{minipage}[t]{\textwidth}
-    \begin{tabular}{l c c r r r r r r c c}
-      %\hline
-      % some footnotes are not visible in typeset-MS (pdf)
-      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
-      \hline
-%      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\
-%      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
-      POPC                   & CHARMM36 \cite{??}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
-      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
-      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
-      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
-      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
-      \hline
-      POPC             & CHARMM36 \cite{??}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
-      POPC:POPE (1:1)  & CHARMM36 \cite{??}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
-      \hline
-      POPC             & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPCmacrog300K} \\
-      POPC:POPE (1:1)  & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1macrogT300K} \\
-      \hline
-      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  298  & 170 & 100 & \cite{POPCslipid298K} \\
-      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120  & 0  &  298  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\
-     \hline
-      POPC                   & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCgromosCKPT310K}  \\
-      POPC:POPG (7:3)        & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG3gromosCKPT310K} \\
-     \hline
-      POPC                   & Slipid \cite{jambeck12b}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
-      POPC:POPG (7:3)        & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
-      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
-
-      \hline
-      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
-      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 250 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
-      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
-      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
-      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 150:150 & 31401 & 57 &  298  & 718 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
-      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
-      \hline
-      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \cite{Lipid17eccPOPCPOPG5050} \\
-      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0.1& 150:150 & 29865 & 54 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG5050100mMCaCl} \\
-      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 1.0& 150:150 & 29865 & 569 &  298  & 600 & 400 & \cite{Lipid17eccPOPCPOPG50501000mMCaCl} \\
-      \hline
-      POPC             & Berger \cite{??} \todoi{This is probable not plain berger, correct force filed should be described.}  &0  & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
-      POPC:POPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\
-      POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
-     \hline
-      DOPC             & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 256 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPCberger300K} \\
-      DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}  &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
-    \end{tabular}
-  \end{minipage}
-  %\end{sidewaystable*}
-  \todo{Citation and ion model for GROMOS-CKP?} \\
-  \todo{Citation and description for ''Berger'' model?} \\
-  \todo{Lipid17 POPC and POPC:POPG mixtures (https://doi.org/10.5281/zenodo.3241242 and https://doi.org/10.5281/zenodo.3237656)
-    should be added after simulated with corrected dihedrals.}\\
-  \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
-\end{table*}
+\subsection{Analysis of lipid conformations bound to proteins}
+Dihedral angles of all available conformations in the PDB databank were calculated using the
+API access to the databank.
 
@@
 
-\subsection{Headgroup and glycerol backbone order parameters of POPE and POPG from $^{13}$C NMR}
-Absolute values of the headgroup and glycerol backbone order parameters from PE and PG lipids are measured
-previously using $^2$H NMR~\cite{seelig76,gally81,wohlgemuth80,borle85}. Because also the order parameter
-signs bear essential information about the lipid structures \cite{botan15,ollila16}, we measured the
-magnitudes and signs of POPE and POPG C--H bond headgroup and glycerol backbone order parameter in liquid phase
-using the 
-%order parameters The order parameters for the glycerol backbone and headgroup s were determined
-2D-RPDLF and S-DROSS experiments, as described previously \cite{ferreira13,ferreira16,antila19}.
-For POPE, the glycerol backbone and $\alpha$-carbon peaks in INEPT spectra were assigned based on
-previously measured POPC spectra~\cite{ferreira13} and
-the $\beta$-carbon peak was assigned based on $^{13}$C chemical shift table for amines available
-at \url{https://www.chem.wisc.edu/areas/reich/nmr/c13-data/cdata.htm} (Fig. \ref{POPEspectra}).
-For POPG, the glycerol backbone peaks in INEPT spectra were assigned based on
-previously measured POPC spectra~\cite{ferreira13}, while $\alpha$ and  $\gamma$-carbon peaks
-\todo{How were these assigned?} (Fig. \ref{POPGspectra}). The numerical value of the $\beta$-carbon
-order parameter could not be determined, because its peak overlapped with the g$_2$ peak from glycerol backbone in POPG.
-However, the order parameter of $\beta$-carbon is expected to be clearly smaller than for g$_2$
-based on previous $^2$H NMR measurements \cite{wohlgemuth80,gally81,borle85}.
-Therefore, the beginning of the S-DROSS curve gives the sign for g$_2$ order parameter and end for $\beta$ (Fig. \ref{POPGspectra} (E)).
-This is confirmed with SIMPSON calculations using negative value for g$_2$ and positive value for $\beta$ order parameter (Fig. \ref{POPGsimpson}).
-\todo{Details to be checked by Tiago}.
+%\subsection{Headgroup and glycerol backbone order parameters of POPE and POPG from $^{13}$C NMR}
+\subsection{Conformational ensembles of different lipid headgroups in bulk bilayer}
 
-As discussed previously for PC and PS headgroups \cite{ollila16,antila19}, also 
-the headgroup and glycerol backbone order parameters of PE and PG are essentially independent of 
-acyl chain composition, and therefore manifest mainly headgroup chemistry (Fig. \ref{HGorderParameters}).
+The conformational ensembles of different lipid headgroups and their order parameters have been
+discussed previously \cite{??}, but these studies have been inconclusive because order parameter
+signs have been unknown and tools to interpret the conformational ensembles have not been available.
+Our experimental order parameters witht the information of the sign from different lipid headgroups
+combined with the literature data are shown in Fig. \ref{HGorderParameters}.
 The glycerol backbone order parameters are similar for all the lipids, although they move slightly toward
 positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG. 
-Also the headgroup order parameters of PC lipids are close PE, althought the latter gives systematically
-slightly more positive values (Fig. \ref{HGorderParameters}).
+%Also the headgroup order parameters of PC lipids are close PE, althought the latter gives systematically
+%slightly more positive values (Fig. \ref{HGorderParameters}).
 % These could be explained with slightly larger temperature in PE measurements, except for the $\alpha$-carbon with the positive sign, for which the
 % more positive value is farther away from zero.
+While headgroup order parameters of PC and PE lipids are rather similar,
+PG and PS lipids exhibit distinct values compared with other lipids.
+In PS lipids, the $\alpha$-carbon order parameter exhibits significant forking and
+the $\beta$-carbon order parameter is more negative than others.
 The $\alpha$-carbon order parameter of PG is similar to PE and PG, while the positive value of $\beta$-carbon
-is distinct from the other lipids. Notably, this difference was not observed in previous $^2$H NMR experiments,
+is distinct from the other lipids.
+Notably, this difference was not observed in previous $^2$H NMR experiments,
 because absolute value of $\beta$-carbon order parameter is similar in PG, PE and PC lipids and the order parameter
@@
 
-In conclusion, the order parameter experiments suggest that the glycerol backbone conformations in all lipids
-and the headgroup conformations in PC and PE lipids are relatively similar, while 
-PS and PG headgropus exhibit distinct conformational sampling.
-The details of sampled conformation are difficult to deduce from order parameters only,
-but the distinct headgroup order parameters
-of PS lipids are previously related to the more rigid structure of the headgroup~\cite{browning80,buldt81,antila19}.
+%As discussed previously for PC and PS headgroups \cite{ollila16,antila19}, also 
+%the headgroup and glycerol backbone order parameters of PE and PG are essentially independent of 
+%acyl chain composition, and therefore manifest mainly headgroup chemistry (Fig. \ref{HGorderParameters}).
 
@@
 
-\subsection{Headgroup and glycerol backbone of POPE and POPG in MD simulations}
+In conclusion, the order parameter experiments suggest that the glycerol backbone conformations in all lipids
+and the headgroup conformations in PC and PE lipids are relatively similar, while 
+PS and PG headgropus exhibit distinct conformational sampling.
+The details of sampled conformation are difficult to deduce from order parameters only,
+but the distinct headgroup order parameters
+of PS lipids are previously related to the more rigid structure of the headgroup~\cite{browning80,buldt81,antila19}.
+
+%\subsection{Headgroup and glycerol backbone of POPE and POPG in MD simulations}
 Headgroup and glycerol backbone order parameters of PE and PG lipids
@@
 https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/8}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPE.pdf}
-  \caption{\label{HGorderParametersPE}
-    The headgroup and glycerol backbone order parameters of PE lipids
-    from experiments (POPE and signs this work, DPPE from Ref.~\citenum{seelig76}
-    and E.coliPE from Ref.~\citenum{gally81}) and simulations with different force fields.
-  }
-  \todo{This should be clarified as in NMRlipidsI and error bars should be added.
-    Probably larger error bars for united atom models based on the report by Fuchs et al.
-  } \\
-\end{figure}
-
-\begin{figure}[!h]
-  \centering
-  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPG.pdf}
-  \caption{\label{HGorderParametersPOPG}
-    The headgroup and glycerol backbone order parameters of PG lipids
-    from experiments (POPG and signs from this work and from Ref.~\citenum{borle85}, %contains 10mM of PIPES,
-    DPPG with 100mM NaCl from Ref.~\citenum{wohlgemuth80},% contains 10mM PIPES and,
-    and E.Coli PG results from Ref.~\citenum{gally81}).
-    and simulations with different force fields.
-  }
-\end{figure}
 
@@
 
-\begin{figure*}[!h]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPCfromPDB.eps}
-  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPEfromPDB.eps}
-  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPGfromPDB.eps}
-  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPSfromPDB.eps}
-  \caption{\label{structures}
-    Dihedral distributions from simulations and lipid structures in PDB.
-  }
-\end{figure*}
-
 \clearpage
-\subsection{PC headgroup interactions with PE and PG}
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
-  \caption{\label{HGorderparametersPCvsPEPG}
-    Modulation of POPC headgroup order parameters with increasing amount of POPE (left) and POPG (right) in bilayer
-    from experiments \cite{scherer87,macdonald87} and simulations with different force fields.
-    Signs are determined as discussed in \cite{botan15,ollila16}.
-  }
-  \todo{P-N vector angles from Berger-OPLS and GROMOS-CKP simulations are yet to be analyzed.}
-\end{figure*}
-
-In experiments, the PC headgroup order parameters increase
-with the addition of negatively charged PG or PS lipids, but
-are not affected by the addition of zwitterionic PE and SM lipids
-or cholesterol (Fig.~\ref{HGorderparametersPCvsPEPG}).
-This can be explained by the electrometer concept, which suggests that the 
-headgroup dipole tilts more parallel to the membrane plane upon addition of
-negative charge to the membrane~\cite{seelig87, scherer87,antila18}.
-The response of PC headgroup order parameters to PE by the tested CHARMM36 and
-Berger-OPLS force fields, although CHARMM36 slightly overestimates the changes (Fig.~\ref{HGorderparametersPCvsPEPG}).
-The good performance of Berger-OPLS simulations is notable because the response of headgroup order parameters
-to cholesterol was significantly overestimated by the Berger/Höltje force field in our previous work \cite{botan15}.
-\todo{This is text by P. Fuchs, copied from the blog. \\
-Area results in nm$^2$, the error is $<=$ 0.003 nm$^2$ \\
-- pure POPC \\
-CHARMM36: 0.624 \\
-Berger : 0.649 \\
-- POPC/POPE 50:50 \\
-CHARMM36 : POPC 0.609, POPE 0.557 \\
-Berger-hacked: POPC 0.637, POPE 0.632 \\
------ \\
-One can see that CHARMM 36 predicts a drop in the area on going from pure POPC to POPC/POPE 50:50. This means that POPC pack tightly to POPE. 
-In contrast, the values for Berger are not that changed. 
-The POPE value predicted by CHARMM 36 (in the mixture POPC/POPE 50:50) is much smaller than that predicted by Berger.\\
--------\\
-The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
-which supports the more condensed PE bilayer.
-This is interesting, but not exactly the core message of the manuscript.
-Maybe we should mention this very briefly? For example, we could just report the areas per lipid (without distinguishing PC and PE)
-and mention the difference between CHARMM36 and Berger. I have opened an issue for this: https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/7
-}
-
-None of the force fields fully reproduces the PC headgroup order parameter response to the increasing amount of
-PG, which may be related to the counterion binding affinity (see also the next section) \cite{antila19}.
-In all force fields except Slipids, the order parameters of different
-hydrogens attached to the $\alpha$-carbon are responsing differently when mixed with PE or PG lipids
-\todo{Maybe we should figure out what is the reason for this? \\
-  Maybe we should analyze the P-N vector angle from different simulations? \\
-  https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/10}.
-
-For $\beta$-carbon order parameter in PG headgroup, experiments report
-mild increase \cite{macdonald87} or no change~\cite{borle85} upon addition 
-of PC lipids (Fig. \ref{HGorderparametersPGvsPCchange}). 
-Simulations with all the tested force fields give only very small changes also for
-the $\alpha$-carbon order parameter (Figs. \ref{HGorderparametersPGvsPC} and \ref{HGorderparametersPGvsPCchange}). 
-Therefore, the simulations are generally in line with experiments, suggesting that the
-interactions with PC do not essentially effect the PG headgroup structure.
-This suggests that the interactions between PG and PC headgroups are captured better in
-simulations than for PS headgroup, where
-all the force fieds significantly overestimated the structural response of PS headgroup
-to the interactions with PC lipids \cite{antila18}.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPCchange.eps}
-  \caption{\label{HGorderparametersPGvsPCchange}
-    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
-    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
-  }
-  \todo{P-N angles from Slipids and GROMOS-CKP yet to be calculated.}
-\end{figure}
-
-\clearpage
-\subsection{Sodium binding to PE and PG lipid bilayers}
-Sodium binding affinity to PE lipids has not been measured experimentally,
-but large differences to PC would be surprising. 
-In simulations, the sodium binding affinity to POPE depends on the used force field (Fig. \ref{NAdensPE}),
-but lesser extend than reported previously for PC~\cite{catte16}.
-\todo{This will be finished once we have all the simulation details and Lipid17 simulations with correct dihedrals from issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12},
-Because some simulation and ion parameters are not identical with the previous work~\cite{catte16},
-we compare POPE results to the POPC simulations ran with identical parameters (Fig. \ref{NAdensPC}).
-In Lipid17 with the strongest sodium binding affinity to POPE,
-the binding affinity is approximately similar to POPC.
-Slipids and CHARMM36 exhibit slightly, and
-GROMOS-CKP subtantially weaker binding to POPE than to POPC.
-%The Lipid17 exhibits stronger sodium binding affinity
-%to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
-%Experimental data for sodium binding to PE lipids is not available, while
-%simulations in agreement with electrometer data from NMR suggests that
-%sodium binding to POPC is weaker than in any of the simulations
-%here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}.
-Assuming that the binding
-to POPE would be similar than to POPC, the sodium binding affinity to POPE
-is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
-but substantially overestimated in Lipid17 simulation.
-%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
-%previous Amber lipid version, Lipid14.
-%\todo{The potential differences in the used ion models should be checked}.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
-  \caption{\label{NAdensPE}
-    Sodium (solid line) and choride ion density profiles along membrane normal
-    from different simulations with PE lipids.
-  }
-\end{figure}
-
-Simulations with PG lipids give similar dependence on force field as observed in POPE simulations:
-Lipid 17 simulations with Dang ion parameters exhibits stronger counter-ion binding affinity to pure POPG bilayer
-than CHARMM36, Slipids, and GROMOS-CKP simulations, which are roughly similar (Fig. \ref{CIdensPG}).
-Lipid17 also exhibits less increase in POPC headgroup
-%$\beta$ carbon
-order parameters upon addition of POPG than other simulations (Fig. \ref{HGorderparametersPCvsPEPG}),
-and lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$).
-In our previous study about PS lipids \cite{antila18}, such behaviour was related to 
-%the results suggest that
-the overestimated counterions binding and shielding the electrostatic repulsion between PG headgroups
-in bilayers.% simulated with Lipid17 parameters.
-%the underestimated increase of PC
-%headgroup order parameters upon addition of negaticely charged lipids
-%was related
-%to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
-%Similar correlation is observed here in %other than GROMOS-CKP simulation:
-%%Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
-%
-Even though the area per lipid in CHARMM36, Slipids, and GROMOS-CKP simulations is in good agreement
-with experiments (Fig. \ref{CIdensPG}), the experimental increase in 
-%In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of
-POPC headgroup order parameters upon addition of POPG are not fully reproduced (Fig. \ref{HGorderparametersPCvsPEPG}).
-%more than in CHARMM36 and Slipids simulations even though the binding affinities in these
-%simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
-%are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
-%and other decreases \todo{Can we figure out the reason for this?}.
-Therefore we conclude that the counter-ion binding affinity is overestimated in Lipid17 simulations,
-% with the strongest affinity among the tested models,
-while the other simulations are more realistic, but slight overbinding cannot be excluded.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
-  \caption{\label{CIdensPG}
-    Counterion densities and area per lipids from simulations with PG lipids.
-    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
-  }
-\end{figure}
-
-\clearpage
-\subsection{Calcium binding to PE and PG lipid bilayers}
+\subsection{Lipid conformational ensembles in lipid bilayers with bound ions}
 \begin{figure*}[bt]
@@
 \end{figure}
+
+\subsection{Protein bound lipid conformations }
+\begin{figure*}[!h]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPCfromPDB.eps}
+  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPEfromPDB.eps}
+  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPGfromPDB.eps}
+  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPSfromPDB.eps}
+  \caption{\label{structures}
+    Dihedral distributions from simulations and lipid structures in PDB.
+  }
+\end{figure*}
 

### 2020-11-03 Manuscript/manuscriptPGPEsuppl.tex
Update dihedral and ion binding results in the manuscript
Update manuscript

@@
 \newcommand*\mycommand[1]{\texttt{\emph{#1}}}
-\newcommand{\todo}[1]{\textcolor{red}{#1}}
-
+%\newcommand{\todo}[1]{\textcolor{red}{#1}}
+\usepackage[obeyFinal]{easy-todo}
 \usepackage{rotating}
@@
 
+\section{Comparison of headgroup order parameters from different force fields to experiments}
+\begin{figure}[]
+  \centering
+  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPE.pdf}
+  \caption{\label{HGorderParametersPE}
+    The headgroup and glycerol backbone order parameters of PE lipids
+    from experiments (POPE and signs this work, DPPE from Ref.~\citenum{seelig76}
+    and E.coliPE from Ref.~\citenum{gally81}) and simulations with different force fields.
+  }
+  \todo{This should be clarified as in NMRlipidsI and error bars should be added.
+    Probably larger error bars for united atom models based on the report by Fuchs et al.
+  } \\
+\end{figure}
+
+\begin{figure}[!h]
+  \centering
+  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPG.pdf}
+  \caption{\label{HGorderParametersPOPG}
+    The headgroup and glycerol backbone order parameters of PG lipids
+    from experiments (POPG and signs from this work and from Ref.~\citenum{borle85}, %contains 10mM of PIPES,
+    DPPG with 100mM NaCl from Ref.~\citenum{wohlgemuth80},% contains 10mM PIPES and,
+    and E.Coli PG results from Ref.~\citenum{gally81}).
+    and simulations with different force fields.
+  }
+\end{figure}
+
+\pagebreak
+
+\subsection{PC headgroup interactions with PE and PG}
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=16.0cm]{../Figs/HGorderparametersPCvsPEPG.eps}
+  \caption{\label{HGorderparametersPCvsPEPG}
+    Modulation of POPC headgroup order parameters with increasing amount of POPE (left) and POPG (right) in bilayer
+    from experiments \cite{scherer87,macdonald87} and simulations with different force fields.
+    Signs are determined as discussed in \cite{botan15,ollila16}.
+  }
+  \todo{P-N vector angles from Berger-OPLS and GROMOS-CKP simulations are yet to be analyzed.}
+\end{figure*}
+
+In experiments, the PC headgroup order parameters increase
+with the addition of negatively charged PG or PS lipids, but
+are not affected by the addition of zwitterionic PE and SM lipids
+or cholesterol (Fig.~\ref{HGorderparametersPCvsPEPG}).
+This can be explained by the electrometer concept, which suggests that the 
+headgroup dipole tilts more parallel to the membrane plane upon addition of
+negative charge to the membrane~\cite{seelig87, scherer87,antila18}.
+The response of PC headgroup order parameters to PE by the tested CHARMM36 and
+Berger-OPLS force fields, although CHARMM36 slightly overestimates the changes (Fig.~\ref{HGorderparametersPCvsPEPG}).
+The good performance of Berger-OPLS simulations is notable because the response of headgroup order parameters
+to cholesterol was significantly overestimated by the Berger/Höltje force field in our previous work \cite{botan15}.
+\todo{This is text by P. Fuchs, copied from the blog. \\
+Area results in nm$^2$, the error is $<=$ 0.003 nm$^2$ \\
+- pure POPC \\
+CHARMM36: 0.624 \\
+Berger : 0.649 \\
+- POPC/POPE 50:50 \\
+CHARMM36 : POPC 0.609, POPE 0.557 \\
+Berger-hacked: POPC 0.637, POPE 0.632 \\
+----- \\
+One can see that CHARMM 36 predicts a drop in the area on going from pure POPC to POPC/POPE 50:50. This means that POPC pack tightly to POPE. 
+In contrast, the values for Berger are not that changed. 
+The POPE value predicted by CHARMM 36 (in the mixture POPC/POPE 50:50) is much smaller than that predicted by Berger.\\
+-------\\
+The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
+which supports the more condensed PE bilayer.
+This is interesting, but not exactly the core message of the manuscript.
+Maybe we should mention this very briefly? For example, we could just report the areas per lipid (without distinguishing PC and PE)
+and mention the difference between CHARMM36 and Berger. I have opened an issue for this: https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/7
+}
+
+None of the force fields fully reproduces the PC headgroup order parameter response to the increasing amount of
+PG, which may be related to the counterion binding affinity (see also the next section) \cite{antila19}.
+In all force fields except Slipids, the order parameters of different
+hydrogens attached to the $\alpha$-carbon are responsing differently when mixed with PE or PG lipids
+\todo{Maybe we should figure out what is the reason for this? \\
+  Maybe we should analyze the P-N vector angle from different simulations? \\
+  https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/10}.
+
+For $\beta$-carbon order parameter in PG headgroup, experiments report
+mild increase \cite{macdonald87} or no change~\cite{borle85} upon addition 
+of PC lipids (Fig. \ref{HGorderparametersPGvsPCchange}). 
+Simulations with all the tested force fields give only very small changes also for
+the $\alpha$-carbon order parameter (Figs. \ref{HGorderparametersPGvsPC} and \ref{HGorderparametersPGvsPCchange}). 
+Therefore, the simulations are generally in line with experiments, suggesting that the
+interactions with PC do not essentially effect the PG headgroup structure.
+This suggests that the interactions between PG and PC headgroups are captured better in
+simulations than for PS headgroup, where
+all the force fieds significantly overestimated the structural response of PS headgroup
+to the interactions with PC lipids \cite{antila18}.
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPCchange.eps}
+  \caption{\label{HGorderparametersPGvsPCchange}
+    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
+    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
+  }
+  \todo{P-N angles from Slipids and GROMOS-CKP yet to be calculated.}
+\end{figure}
+
+\clearpage
+\subsection{Sodium binding to PE and PG lipid bilayers}
+Sodium binding affinity to PE lipids has not been measured experimentally,
+but large differences to PC would be surprising. 
+In simulations, the sodium binding affinity to POPE depends on the used force field (Fig. \ref{NAdensPE}),
+but lesser extend than reported previously for PC~\cite{catte16}.
+\todo{This will be finished once we have all the simulation details and Lipid17 simulations with correct dihedrals from issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12},
+Because some simulation and ion parameters are not identical with the previous work~\cite{catte16},
+we compare POPE results to the POPC simulations ran with identical parameters (Fig. \ref{NAdensPC}).
+In Lipid17 with the strongest sodium binding affinity to POPE,
+the binding affinity is approximately similar to POPC.
+Slipids and CHARMM36 exhibit slightly, and
+GROMOS-CKP subtantially weaker binding to POPE than to POPC.
+%The Lipid17 exhibits stronger sodium binding affinity
+%to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
+%Experimental data for sodium binding to PE lipids is not available, while
+%simulations in agreement with electrometer data from NMR suggests that
+%sodium binding to POPC is weaker than in any of the simulations
+%here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}.
+Assuming that the binding
+to POPE would be similar than to POPC, the sodium binding affinity to POPE
+is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
+but substantially overestimated in Lipid17 simulation.
+%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
+%previous Amber lipid version, Lipid14.
+%\todo{The potential differences in the used ion models should be checked}.
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
+  \caption{\label{NAdensPE}
+    Sodium (solid line) and choride ion density profiles along membrane normal
+    from different simulations with PE lipids.
+  }
+\end{figure}
+
+Simulations with PG lipids give similar dependence on force field as observed in POPE simulations:
+Lipid 17 simulations with Dang ion parameters exhibits stronger counter-ion binding affinity to pure POPG bilayer
+than CHARMM36, Slipids, and GROMOS-CKP simulations, which are roughly similar (Fig. \ref{CIdensPG}).
+Lipid17 also exhibits less increase in POPC headgroup
+%$\beta$ carbon
+order parameters upon addition of POPG than other simulations (Fig. \ref{HGorderparametersPCvsPEPG}),
+and lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$).
+In our previous study about PS lipids \cite{antila18}, such behaviour was related to 
+%the results suggest that
+the overestimated counterions binding and shielding the electrostatic repulsion between PG headgroups
+in bilayers.% simulated with Lipid17 parameters.
+%the underestimated increase of PC
+%headgroup order parameters upon addition of negaticely charged lipids
+%was related
+%to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
+%Similar correlation is observed here in %other than GROMOS-CKP simulation:
+%%Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
+%
+Even though the area per lipid in CHARMM36, Slipids, and GROMOS-CKP simulations is in good agreement
+with experiments (Fig. \ref{CIdensPG}), the experimental increase in 
+%In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of
+POPC headgroup order parameters upon addition of POPG are not fully reproduced (Fig. \ref{HGorderparametersPCvsPEPG}).
+%more than in CHARMM36 and Slipids simulations even though the binding affinities in these
+%simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
+%are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
+%and other decreases \todo{Can we figure out the reason for this?}.
+Therefore we conclude that the counter-ion binding affinity is overestimated in Lipid17 simulations,
+% with the strongest affinity among the tested models,
+while the other simulations are more realistic, but slight overbinding cannot be excluded.
+\begin{figure}[]
+  \centering
+  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
+  \caption{\label{CIdensPG}
+    Counterion densities and area per lipids from simulations with PG lipids.
+    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
+  }
+\end{figure}
+
 \section{Simulated systems}
+
+%\begin{table*}[htb]
+%
+\begin{sidewaystable*}[!p]
+  \centering
+  \caption{List of MD simulations with PE lipids.
+    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
+    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
+    % The lipid force fields named as in our previous work~\cite{botan15}.
+  }\label{systemsPE}
+  \begin{minipage}[t]{\textwidth}
+    \begin{tabular}{l c c r r r r r r c c}
+      %\hline
+      % some footnotes are not visible in typeset-MS (pdf)
+      lipid/counter-ions & force field for lipids / ions  & NaCl (M) & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      \hline
+      POPE  & CHARMM36 \cite{??}           &0       & 144	& 5760  &0    & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
+      POPE  & CHARMM36 \cite{??}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
+      POPE  & CHARMM36 \cite{??}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
+      POPE  & CHARMM36ua \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
+      \hline
+      DPPE  & Slipids \cite{jambeck12b}    &0    & 288 	& 9386  &0    & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
+      POPE  & Slipids \cite{jambeck12b} &0    & 336	& ?     &0    & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
+      POPE  & Slipids \cite{jambeck12b}            & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
+      POPE  & Slipids / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
+      \hline
+      DPPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3655  &0    & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \\
+      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
+      POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
+      POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
+      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
+      \hline
+      POPE  & GROMOS 43A1-S3 \cite{??}     &0    & 128	& 3552     &0    & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles}  \\
+      \hline
+      POPE  & OPLS-UA vdW on H \cite{??}   &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \\
+      POPE  & OPLS-UA \cite{??}            &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \\
+      \hline
+      POPE  & OPLS-MacRog \cite{rog16}     &0    & 144	& 5760     &0    & 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
+      POPE  & OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
+      \hline
+      POPE  & Berger-Vries \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
+      POPE  & Berger-largeH \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
+      DOPE  & Berger-Vries \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
+      DOPE  & Berger-largeH \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
+      \hline
+      POPE             & LIPID17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
+      POPE             & LIPID17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
+    \end{tabular}
+  \end{minipage}
+  \todo{Citation for CHARMM36 PE?} \\
+  \todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
+  \todo{Citation for GROMOS-CKP?} \\
+  \todo{Citation for GROMOS 43A1-S3?} \\
+  \todo{Citation for OPLS-UA models?} \\
+  \todo{Citations for Berger-* simulations?} \\
+  \todo{LIPID17 simulations with correct dihedrals still coming}
+%\end{table*}
+  \end{sidewaystable*}
+
+  %    \begin{table*}[htb]
+  \begin{sidewaystable*}[!p]
+  \centering
+  \caption{List of MD simulations with PG lipids.
+    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
+    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
+    % The lipid force fields named as in our previous work~\cite{botan15}.
+  }\label{systemsPG}
+  \begin{minipage}[t]{\textwidth}
+    \begin{tabular}{l c c r r r r r r c c}
+      %\hline
+      % some footnotes are not visible in typeset-MS (pdf)
+      lipid/counter-ions & force field for lipids / ions & NaCl (M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      \hline
+      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}    &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
+      POPG             & CHARMM36 \cite{??}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
+      POPG             & CHARMM36 \cite{??}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
+      \hline
+      POPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 10664   &0     & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
+      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232  &0     & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
+      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232   &0     & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
+      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0    & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGslipids} \\
+      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0.11    & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
+      \hline
+      POPG             & LIPID17 / Dang \cite{gould18,smith94}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
+      POPG             & LIPID17 \cite{??}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
+      \hline
+      POPG             & GROMOS-CKP \cite{??}         & 0              & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGgromosCKP} \\
+      POPG             & GROMOS-CKP \cite{??}         & 0.11           & 500 & 25000 & 49 &  310  & 500 & 100 & \cite{POPGgromosCKP150mMNaCl} \\
+    \end{tabular}
+  \end{minipage}
+    \todo{Citations and ion model for CHARMM36?} \\
+  \todo{Lipid17 simulation with correct dihedral potentials still coming.} \\
+  \todo{Citation and ion model for GROMOS-CKP?}
+%\end{table*}
+ \end{sidewaystable*}
+
+%\begin{table*}[htb]
+  \begin{sidewaystable*}[!p]
+  \centering
+  \caption{List of MD simulations with PE and PG lipids mixed with PC.
+    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
+    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
+    % The lipid force fields named as in our previous work~\cite{botan15}.
+  }\label{systemsMIX}
+  \begin{minipage}[t]{\textwidth}
+    \begin{tabular}{l c c r r r r r r c c}
+      %\hline
+      % some footnotes are not visible in typeset-MS (pdf)
+      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      \hline
+%      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\
+%      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
+      POPC                   & CHARMM36 \cite{??}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
+      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
+      \hline
+      POPC             & CHARMM36 \cite{??}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
+      POPC:POPE (1:1)  & CHARMM36 \cite{??}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
+      \hline
+      POPC             & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPCmacrog300K} \\
+      POPC:POPE (1:1)  & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1macrogT300K} \\
+      \hline
+      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  298  & 170 & 100 & \cite{POPCslipid298K} \\
+      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120  & 0  &  298  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\
+     \hline
+      POPC                   & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCgromosCKPT310K}  \\
+      POPC:POPG (7:3)        & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG3gromosCKPT310K} \\
+     \hline
+      POPC                   & Slipid \cite{jambeck12b}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
+      POPC:POPG (7:3)        & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+    \end{tabular}
+  \end{minipage}
+    \todo{Citation and ion model for GROMOS-CKP?} \\
+  \todo{Citation and description for ''Berger'' model?} \\
+  \todo{Lipid17 POPC and POPC:POPG mixtures (https://doi.org/10.5281/zenodo.3241242 and https://doi.org/10.5281/zenodo.3237656)
+    should be added after simulated with corrected dihedrals.}\\
+  \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
+%\end{table*}
+\end{sidewaystable*}
+
+  \begin{sidewaystable*}[!p]
+  \centering
+  \caption{List of MD simulations with PE and PG lipids mixed with PC.
+    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
+    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
+    % The lipid force fields named as in our previous work~\cite{botan15}.
+  }\label{systemsMIX}
+  \begin{minipage}[t]{\textwidth}
+    \begin{tabular}{l c c r r r r r r c c}
+      %\hline
+      % some footnotes are not visible in typeset-MS (pdf)
+      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      \hline
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 250 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
+      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 150:150 & 31401 & 57 &  298  & 718 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
+      \hline
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \cite{Lipid17eccPOPCPOPG5050} \\
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0.1& 150:150 & 29865 & 54 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 1.0& 150:150 & 29865 & 569 &  298  & 600 & 400 & \cite{Lipid17eccPOPCPOPG50501000mMCaCl} \\
+      \hline
+      POPC             & Berger \cite{??} \todoi{This is probable not plain berger, correct force filed should be described.}  &0  & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
+      POPC:POPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\
+      POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
+     \hline
+      DOPC             & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 256 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPCberger300K} \\
+      DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}  &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
+    \end{tabular}
+  \end{minipage}
+    \todo{Citation and ion model for GROMOS-CKP?} \\
+  \todo{Citation and description for ''Berger'' model?} \\
+  \todo{Lipid17 POPC and POPC:POPG mixtures (https://doi.org/10.5281/zenodo.3241242 and https://doi.org/10.5281/zenodo.3237656)
+    should be added after simulated with corrected dihedrals.}\\
+  \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
+%\end{table*}
+\end{sidewaystable*}
 
@@
   }
-  \todo{CHARMM36 simulations should be longer and with Na couterions.}
+  \todo{Lipid17ecc data to be analyzed and added.}
 \end{figure*}

### 2020-11-04 Manuscript/manuscriptPGPE.tex
Update manuscript

--- 2020-11-03 Manuscript/manuscriptPGPE.tex
@@
 
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in CHARMM36 dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
 \clearpage

### 2020-11-04 Manuscript/manuscriptPGPEsuppl.tex
Update manuscript

--- 2020-05-27 Manuscript/manuscriptPGPEsuppl.tex
@@
 
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in Slipids dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in lipid17ecc dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in CHARMM36 dihedrals with increasing amount of cationic surfactant.
+  }
+\end{figure}
+
 \clearpage

### 2020-11-05 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 
-The conformational ensembles of different lipid headgroups and their order parameters have been
-discussed previously \cite{??}, but these studies have been inconclusive because order parameter
-signs have been unknown and tools to interpret the conformational ensembles have not been available.
-Our experimental order parameters witht the information of the sign from different lipid headgroups
+Derivation of different lipid headgroup conformational ensembles in liquid state has been
+%discussed previously \cite{??}, but these studies have been
+inconclusive due to lack of suitable experimental data and tools to interpret the conformational ensembles.
+Our experimental order parameters, including the signs, from different lipid headgroups
 combined with the literature data are shown in Fig. \ref{HGorderParameters}.
@@
 % more positive value is farther away from zero.
-While headgroup order parameters of PC and PE lipids are rather similar,
-PG and PS lipids exhibit distinct values compared with other lipids.
-In PS lipids, the $\alpha$-carbon order parameter exhibits significant forking and
-the $\beta$-carbon order parameter is more negative than others.
-The $\alpha$-carbon order parameter of PG is similar to PE and PG, while the positive value of $\beta$-carbon
-is distinct from the other lipids.
+While the headgroup order parameters are similar for PC and PE lipids,
+PG and PS lipids exhibit distinct values in comparison with other lipids.
+In PS headgroup, the $\alpha$-carbon order parameter exhibits significant forking and
+the $\beta$-carbon has more negative value than in other lipids.
+In PG headgroup, the $\alpha$-carbon order parameter is similar to PE and PG,
+while the positive value of $\beta$-carbon is distinct from all the other lipids.
 Notably, this difference was not observed in previous $^2$H NMR experiments,
@@
 
-In conclusion, the order parameter experiments suggest that the glycerol backbone conformations in all lipids
-and the headgroup conformations in PC and PE lipids are relatively similar, while 
-PS and PG headgropus exhibit distinct conformational sampling.
-The details of sampled conformation are difficult to deduce from order parameters only,
-but the distinct headgroup order parameters
-of PS lipids are previously related to the more rigid structure of the headgroup~\cite{browning80,buldt81,antila19}.
-
-%\subsection{Headgroup and glycerol backbone of POPE and POPG in MD simulations}
-Headgroup and glycerol backbone order parameters of PE and PG lipids
-show wide variation between different force fields
-and none of the force fields reproduce all values within experimental error bars.
-(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
-as observed previously also for PC and PS lipids \cite{botan15,antila19}.
+As in previous NMRlipids project results for PC and PS lipids \cite{botan15,antila19},
+none of the MD simulation force fields correctly captures all the
+headgroup and glycerol backbone order parameters of PE and PG lipids
+(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG})
+that would enable a straightforward interpretation of conformational ensembles.
+%show wide variation between different force fields
+%and none of the force fields reproduce all values within experimental error bars.
+%(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
+%as observed previously also for PC and PS lipids \cite{botan15,antila19}.
 %The Slipid simulations were able to capture the essential differences between PC and PS lipid
 %headgroups \cite{antila19}, but this is not the case for PE and PG lipids.
-The poor performance of headgroup order parameters in Berger model can be probably explained by
-ring like structures seen in Fig. 6 in Ref. \citenum{mukhopadhyay04}, which is a typical feature
-for Berger based lipid force fields containing explicit hydrogen atoms in the head group \cite{zhao08,henin09,dahlberg10}.
-The poor performance of glycerol backbone of Slipids simulations is systemically observed also
-for other lipids in previous studies \cite{botan15,antila19}.
-\todo{Should we comment more the relative quality of different force fields and/or make the subjective force field ranking figures?
-https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/8}
-
-Without further discussion about poorly performing force fields, we focus on more detailed analysis
-of CHARMM36 simulations, which captures the essential differences between
+%Without further discussion about poorly performing force fields, we focus on more detailed analysis
+%of
+Nevertheless, CHARMM36 force field, which gives the results closest to the experiments for all lipids,
+captures the essential differences between
 PC, PS, PG and PE headroup order parameters  (Fig.~\ref{structures}) with the exception of $\beta$-carbon order parameter of PC
@@
 %not within the experimental error bars, it 
-which is too negative when compared with PS or PE order parameter, or with experiments \cite{botan15}.
-%As already pointed out in our previous study \cite{antila19},
-Characteristic dihedral conformations in PS headgroup are asymmetric conformations
-preferring gauche 270$^\circ$ conformations in N-C$_\beta$-C$_\alpha$-O$_\alpha$ and
-%dihedral exhibits a more asymmetric angle distribution for PS, having largest probability being
-%in , than for the PC headgroup.
-C$_\beta$-C$_\alpha$-O$_\alpha$-P dihedrals. 
-In PG headgroup, the O$_\beta$-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral (corresponding N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral in other lipids)
-is mostly in trans conformation, and C$_\beta$-C$_\alpha$-O$_\alpha$-P  has asymmetric tendency to be in
-gauche 60$^\circ$ conformation. 
-Main difference between PC and PE is the lower probability of trans state in C$_\beta$-C$_\alpha$-O$_\alpha$-P PC dihedral,
-which could be a potential reason for the too negative $\beta$ headgroup order parameter in PC.
+which is too negative when compared with other lipids or experiments \cite{botan15}.
+% As already pointed out in our previous study \cite{antila19},
+
+The conformational ensembles characterized by the distributions of heavy atom dihedral angles
+show major differences between lipids only for O$_\alpha$-C$_\alpha$ and  C$_\alpha$-C$_\beta$ bonds
+near the end of the headgroup (Fig. \ref{structures}).
+These probably explain the slight differences in the
+angle between headgroup dipole and membrane normal, which decreases 
+in the order PG $>$ PE  $>$ PC  $>$ PS, although difference between PC and PE may an artefact as
+the $\beta$-carbon order parameter in PC is poorly reproduced by the CHARMM36 force field.
+Dihedrals of both O-P bonds can freely rotate in all lipids as all possible angles
+are observed in the distributions. For PS, however, some angles are quite unlikely,
+possible explaining the more rigid headgroup structures proposed for PS lipids~\cite{browning80,buldt81}.
+The eclipsed (0 or 360 degrees) are not present in any other dihedrals, and anti-eclipsed (120 or 240 degrees)
+are not present in C$_\alpha$-C$_\beta$ and g$_1$-g$_2$ bonds.
+
+%Characteristic dihedral conformations in PS headgroup are asymmetric conformations
+%preferring gauche 270$^\circ$ conformations in N-C$_\beta$-C$_\alpha$-O$_\alpha$ and
+%%dihedral exhibits a more asymmetric angle distribution for PS, having largest probability being
+%%in , than for the PC headgroup.
+%C$_\beta$-C$_\alpha$-O$_\alpha$-P dihedrals. 
+%In PG headgroup, the O$_\beta$-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral (corresponding N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral in other% lipids)
+%is mostly in trans conformation, and C$_\beta$-C$_\alpha$-O$_\alpha$-P  has asymmetric tendency to be in
+%gauche 60$^\circ$ conformation. 
+%Main difference between PC and PE is the lower probability of trans state in C$_\beta$-C$_\alpha$-O$_\alpha$-P PC dihedral,
+%which could be a potential reason for the too negative $\beta$ headgroup order parameter in PC.%
+
+In conclusion, the results suggest that free rotations around O-P bonds decouple the
+headgroup and glycerol backbone structure and dynamics in similarly in all lipids,
+althought this rotation is slower in PS lipids where crossing certain barriers are slower.
+Despite significant differences in dihedral distributions near the headgroup end,
+all lipids have dihedral angles within approximately the same ranges.
+This suggests that all lipids can quite freely arrange to multiple headgroup conformations
+when interaction with proteins, ions or other biomolecules.
+%the order parameter experiments suggest that the glycerol backbone conformations in all lipids
+%and the headgroup conformations in PC and PE lipids are relatively similar, while 
+%PS and PG headgropus exhibit distinct conformational sampling.
+%The details of sampled conformation are difficult to deduce from order parameters only,
+%but the distinct headgroup order parameters
+%of PS lipids are previously related to the more rigid structure of the headgroup~\cite{browning80,buldt81,antila19}.
 
@@
 \subsection{Lipid conformational ensembles in lipid bilayers with bound ions}
+
+Headgroup order parameters of PC lipids decrease or increase upon addition of positive
+or negative charge into membrane, respectively, while they remain unchanged upon addition
+of neutral molecules such as cholesterol or sphingomyelin \cite{??}.
+This has been explained by the tilting of the headgroup dipole due to electrostatic
+interactions with bound charges, and has been used to measure the binding affinity
+of ions to membranes \cite{??}.
+%The headgroup order parameters of PC lipids can be used to measure the ion binding
+%affinity to lipid bilayers because their magnitude is linearly proportional to the amount of bound charge in bilayer \cite{seelig87,catte16}.
+%This molecular electrometer concept can be used also for bilayers containing PC lipids mixed with charged lipids \cite{borle85,macdonald87,roux90,antila19}.
+%The headgroup order parameters can be used to evaluate MD simulations against experimental data,
+%because they can be directly calculated from MD simulations \cite{catte16}.
+
+%To evaluate the calcium binding in simulations of lipid bilayers containing PG lipids,
+%we calculated
+The changes of PC headgroup order parameters in POPC:POPG (1:1) and (4:1) mixtures
+upon addition of CaCl$_2$ from simulations and experiments \cite{borle85,macdonald87}
+%, and compared these with the available experimental data \cite{borle85,macdonald87}.
+in Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4} are in line with
+conclusion from our previous studies \cite{??}:  
+%The decrease of POPC headgroup order parameters in mixtures with POPG lipids with
+%increasing CaCl$_2$ concentration is overestimated in Slipids and Lipid17
+order parameters decrease in the response to bound cations in qualitative agreement with
+experiments, but calcium binding affinity to membranes is overestimated in 
+simulations except in CHARMM36 with the NBfix correction which underestimates the binding affinity.
+On the other hand, implicit inclusion of electronic polarizability to Amber based lipid models
+improves the ion binding affinity leading to a good agreement with experiments.
+%(Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4})
+%
+%indicating too strong binding affinity of calcium into the bilayers as previously
+%observed for pure PC lipid bilayers and mixtures with PS lipids \cite{catte16,antila19}.
+%
+%The decrease of $\alpha$-carbon order parameter of PC lipids in PC:PG mixtures as a function
+%of calcium concentration is close to experiments CHARMM36 simulations (Fig. \ref{changesWITHCaClPG}),
+%but the decrease of $\beta$-carbon order parameter seems to be overestimated.
+%However, the $\beta$-carbon order parameter was not actually measured from these samples,
+%but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
+%On the other hand, the simulations are not converged.
+%
+%The calcium binding affinity to lipid bilayers with PC and PS lipids
+%was recently improved by applying the electronic continuum correction (ECC) to Amber Lipid14/17 force fields \cite{melcr18,melcr20}.
+%In this approach, the electronic polarizability is implicitly included in the classical force fields
+%by scaling the charges with constant factors \cite{leontyev11}. Here, we make a ECC-POPG force field
+%by applying the scaling factors originally used for POPS also to POPG, i.e., we multiply
+%charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions with 
+%$f_q$=0.75 and $f_\sigma$=0.89, respectively \cite{melcr20}.
+%ECC-POPG model gives a weaker calcium binding affinity (Fig.\ref{CAdensPG}) and better agreement with the experimental
+%PC headgroup order parameter changes (Fig. \ref{changesWITHCaClPG}) for POPC:POPG mixtures than the original Lipid17 model
+%\todo{to be finished when we have all the data}, indicating that the ECC improves the simulation predictions of
+%calcium binding affinity as previously observed for PC and PS lipids \cite{melcr18,melcr20}.
+
 \begin{figure*}[bt]
@@
 
-To evaluate the calcium binding in simulations of lipid bilayers containing PG lipids,
-we calculated the changes in headgroup order parameters of POPC:POPG (1:1) and (4:1) mixtures
-upon addition of CaCl$_2$, and compared these with the available experimental data \cite{borle85,macdonald87}.
-The headgroup order parameters of PC lipids can be used to measure the ion binding
-affinity to lipid bilayers because their magnitude is linearly proportional to the amount of bound charge in bilayer \cite{seelig87,catte16}.
-This molecular electrometer concept can be used also for bilayers containing PC lipids mixed with charged lipids \cite{borle85,macdonald87,roux90,antila19}.
-The headgroup order parameters can be used to evaluate MD simulations against experimental data,
-because they can be directly calculated from MD simulations \cite{catte16}.
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/CAdensPG1PC1.eps}
+  \caption{\label{CAdensPG}
+    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (1:1) mixtures with different force fields.
+  }
+\end{figure}
 
-The decrease of POPC headgroup order parameters in mixtures with POPG lipids with
-increasing CaCl$_2$ concentration is overestimated in Slipids and Lipid17
-simulations (Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4})
-indicating too strong binding affinity of calcium into the bilayers as previously
-observed for pure PC lipid bilayers and mixtures with PS lipids \cite{catte16,antila19}.
-\todo{CHARMM results to be mentioned once we have the new simulations.}
-%The decrease of $\alpha$-carbon order parameter of PC lipids in PC:PG mixtures as a function
-%of calcium concentration is close to experiments CHARMM36 simulations (Fig. \ref{changesWITHCaClPG}),
-%but the decrease of $\beta$-carbon order parameter seems to be overestimated.
-%However, the $\beta$-carbon order parameter was not actually measured from these samples,
-%but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
-%On the other hand, the simulations are not converged.
-The calcium binding affinity to lipid bilayers with PC and PS lipids
-was recently improved by applying the electronic continuum correction (ECC) to Amber Lipid14/17 force fields \cite{melcr18,melcr20}.
-In this approach, the electronic polarizability is implicitly included in the classical force fields
-by scaling the charges with constant factors \cite{leontyev11}. Here, we make a ECC-POPG force field
-by applying the scaling factors originally used for POPS also to POPG, i.e., we multiply
-charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions with 
-$f_q$=0.75 and $f_\sigma$=0.89, respectively \cite{melcr20}.
-ECC-POPG model gives a weaker calcium binding affinity (Fig.\ref{CAdensPG}) and better agreement with the experimental
-PC headgroup order parameter changes (Fig. \ref{changesWITHCaClPG}) for POPC:POPG mixtures than the original Lipid17 model
-\todo{to be finished when we have all the data}, indicating that the ECC improves the simulation predictions of
-calcium binding affinity as previously observed for PC and PS lipids \cite{melcr18,melcr20}.
+Changes in lipid conformational ensembles characterized by distributions of heavy atom dihedral angles
+are very small upon binding of calcium or addition of cationic surfactants to membranes (Figs. \ref{??}).
+Yet, such changes are sufficient to tilt the headgroup dipole angle and reproduce the experimentally observed
+order parameter changes in $\alpha$ and $\beta$ carbons.
 
-Experimental data for the $\beta$-carbon order parameter of POPG
-shows a rapid decrease with increasing CaCl$_2$ concentrations up to 10 mM
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in CHARMM36 dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+Simulations also qualitatively reproduce the experimentally observed decrease of POPG
+%Experimental data for the
+$\beta$-carbon order parameter
+%of POPG shows a rapid decrease with increasing
+upon addition of CaCl$_2$,
+%concentrations up to 10 mM
+suggesting that the PG headgroup response is qualitatively captured by simulations, in contrast
+to PS in our previous work \cite{antila19}.
 and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
@@
 
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/CAdensPG1PC1.eps}
-  \caption{\label{CAdensPG}
-    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (1:1) mixtures with different force fields.
-  }
-\end{figure}
-
 \subsection{Protein bound lipid conformations }
@@
 \end{figure*}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in CHARMM36 dihedrals with increasing amount of CaCl$_2$.
-  }
-\end{figure}
 

### 2020-11-05 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 to evaluate the accuracy of PE and PG heagroup structures, and the cation binding affinity
+
 to anionic membranes containing PG lipids in the current MD simulation force fields.
@@
 
-Simulations also qualitatively reproduce the experimentally observed decrease of POPG
-%Experimental data for the
-$\beta$-carbon order parameter
-%of POPG shows a rapid decrease with increasing
-upon addition of CaCl$_2$,
-%concentrations up to 10 mM
-suggesting that the PG headgroup response is qualitatively captured by simulations, in contrast
-to PS in our previous work \cite{antila19}.
-and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
-This behaviour is similar to that of $\beta$-carbon order parameters of POPC,
-but essentially different than observed for POPS, where $\beta$-carbon order parameters increases with
-addition of calcium \cite{antila19}. Experimentally measured changes of PG $\alpha$-carbon order parameters upon addition of
-calcium are not available.
-%upon addition of calcium in experiments,
-%exhibiting 
-%The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.
 Lipid17 and Slipids force fields correctly capture the PG $\beta$-carbon order parameter response to CaCl$_2$
 even thought the binding affinity was too large based on the comparison of PC headgroup order parameter changes with experiments.
-While applying ECC to Lipid17 improved the PC headgroup order parameter response and binding affinity,
-the response of PG $\beta$-carbon order parameter to calcium is too weak in this model.
-The response of PG $\alpha$-carbon order parameters to CaCl$_2$ differs between force fields,
-but experimental data to evaluate these predictions is not available. 
-\todo{To be finished once we have the new CHARMM simulations and conformational changes of PG analyzed.}
+%While applying ECC to Lipid17 improved the PC headgroup order parameter response and binding affinity,
+The response of PG $\beta$-carbon order parameter to calcium is too weak in CHARMM36 and ECC-lipid17 force fields,
+but decrease of the order parameter is associated with decreased P-N vector angle in all
+force fields. Even thought the change in headgroup orientation upon addition of calcium is smaller
+in PG lipids than in PC, the conformational changes in PG are more significant because also
+dihedral angle distribution around O$_\alpha$-C$_\alpha$ bond is affected by calcium binding (Fig. ??).
+However, because none of the force fields simulatneously capture both the PG conformational ensemble and calcium binding,
+more accurate force fields are required for more detailed studies of the effect of ions to PG conformational enembles.
+%and PG
+%The response of PG $\alpha$-carbon order parameters to CaCl$_2$ differs between force fields,
+%but experimental data to evaluate these predictions is not available. 
+%Simulations also qualitatively reproduce the experimentally observed decrease of POPG
+%%Experimental data for the
+%$\beta$-carbon order parameter
+%%of POPG shows a rapid decrease with increasing
+%upon addition of CaCl$_2$,
+%%concentrations up to 10 mM
+%suggesting that the PG headgroup response is qualitatively captured by simulations, in contrast
+%to PS in our previous work \cite{antila19}.
+%and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
+%This behaviour is similar to that of $\beta$-carbon order parameters of POPC,
+%but essentially different than observed for POPS, where $\beta$-carbon order parameters increases with
+%addition of calcium \cite{antila19}. Experimentally measured changes of PG $\alpha$-carbon order parameters upon addition of
+%calcium are not available.
+%%upon addition of calcium in experiments,
+%%exhibiting 
+%%The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.
+
 % The result is similar to the $\sim$200~ns simulations with PC lipids in previous work \cite{catte16}.
@@
 
-\todo{We still need more data to finish the discussion. More detailed discussion is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}
+%\todo{We still need more data to finish the discussion. More detailed discussion is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}
 

### 2020-11-05 Manuscript/manuscriptPGPEsuppl.tex

@@
 \section{Comparison of headgroup order parameters from different force fields to experiments}
+The poor performance of headgroup order parameters in Berger model can be probably explained by
+ring like structures seen in Fig. 6 in Ref. \citenum{mukhopadhyay04}, which is a typical feature
+for Berger based lipid force fields containing explicit hydrogen atoms in the head group \cite{zhao08,henin09,dahlberg10}.
+The poor performance of glycerol backbone of Slipids simulations is systemically observed also
+for other lipids in previous studies \cite{botan15,antila19}.
+\todo{Should we comment more the relative quality of different force fields and/or make the subjective force field ranking figures?
+https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/8}
 \begin{figure}[]

### 2020-11-17 Manuscript/manuscriptPGPE.tex
Update dihedral results into the manuscript

--- 2020-11-05 Manuscript/manuscriptPGPE.tex
@@
   \caption{\label{DIHSwithCA}
-    Changes in CHARMM36 dihedrals with increasing amount of CaCl$_2$.
+    Changes in POPC CHARMM36 dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in POPG CHARMM36 dihedrals with increasing amount of CaCl$_2$.
   }

### 2020-11-17 Manuscript/manuscriptPGPEsuppl.tex
Update dihedral results into the manuscript

@@
 
+\section{Changes in headgroup conformations upon addition of CaCl$_2$}
+
 \begin{figure}[]
@@
   \caption{\label{DIHSwithCA}
-    Changes in Slipids dihedrals with increasing amount of CaCl$_2$.
+    Changes in POPC Slipids dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in POPG Slipids dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in POPC lipid17 dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in POPG lipid17 dihedrals with increasing amount of CaCl$_2$.
   }
@@
   \caption{\label{DIHSwithCA}
-    Changes in lipid17ecc dihedrals with increasing amount of CaCl$_2$.
+    Changes in POPC lipid17ecc dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in POPG lipid17ecc dihedrals with increasing amount of CaCl$_2$.
   }

### 2020-11-17 Manuscript/manuscriptPGPE.tex

--- 2020-11-05 Manuscript/manuscriptPGPE.tex
@@
   \caption{\label{DIHSwithCA}
-    Changes in CHARMM36 dihedrals with increasing amount of CaCl$_2$.
+    Changes in POPC CHARMM36 dihedrals with increasing amount of CaCl$_2$.
   }
 \end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in POPG CHARMM36 dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+%Simulations also qualitatively reproduce the experimentally observed decrease of POPG
+%%Experimental data for the
+%$\beta$-carbon order parameter
+%%of POPG shows a rapid decrease with increasing
+%upon addition of CaCl$_2$,
+%%concentrations up to 10 mM
+%suggesting that the PG headgroup response is qualitatively captured by simulations, in contrast
+%to PS in our previous work \cite{antila19}.
+%and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
+%This behaviour is similar to that of $\beta$-carbon order parameters of POPC,
+%but essentially different than observed for POPS, where $\beta$-carbon order parameters increases with
+%addition of calcium \cite{antila19}. Experimentally measured changes of PG $\alpha$-carbon order parameters upon addition of
+%calcium are not available.
+%%upon addition of calcium in experiments,
+%%exhibiting 
+%%The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.
 

### 2020-11-17 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 \subsection{Lipid conformational ensembles in lipid bilayers with bound ions}
-
-Headgroup order parameters of PC lipids decrease or increase upon addition of positive
-or negative charge into membrane, respectively, while they remain unchanged upon addition
-of neutral molecules such as cholesterol or sphingomyelin \cite{??}.
-This has been explained by the tilting of the headgroup dipole due to electrostatic
-interactions with bound charges, and has been used to measure the binding affinity
-of ions to membranes \cite{??}.
+Phosphatidylcholine headgroup conformational ensemble and orientation are unchanged
+upon addition of neutral molecules such as cholesterol or sphingomyelin into a bilayer,
+while incorporation of charges affect the headgroup dipole vector tilt which can
+be detected by measuring C-H bond order parameters of $\alpha$ and $\beta$ carbons \cite{??}.
+Such changes have been observed upon addition of charged lipids, proteins, surfactants,
+drugs, or ions \cite{??}, but how charges affect the lipid conformational ensemble remains unknown.
+%Headgroup order parameters of PC lipids decrease or increase upon addition of positive
+%or negative charge into membrane, respectively, while they remain 
+%\cite{??}.
+%This has been explained by the tilting of the headgroup dipole due to electrostatic
+%interactions with bound charges, and has been used to measure the binding affinity
+%of ions to membranes \cite{??}.
 %The headgroup order parameters of PC lipids can be used to measure the ion binding
@@
 
-%To evaluate the calcium binding in simulations of lipid bilayers containing PG lipids,
-%we calculated
-The changes of PC headgroup order parameters in POPC:POPG (1:1) and (4:1) mixtures
-upon addition of CaCl$_2$ from simulations and experiments \cite{borle85,macdonald87}
-%, and compared these with the available experimental data \cite{borle85,macdonald87}.
+\begin{figure*}[bt]
+  \centering
+  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCvsSURFchange.eps}
+  \caption{\label{changesWITHsurf}
+    Modulation of PC headgroup order parameters and P-N vector angle upon addition of cationic surfactant.
+  }
+\end{figure*}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+  \caption{\label{DIHSwithSURF}
+    Changes in CHARMM36 dihedrals with increasing amount of cationic surfactant.
+  }
+\end{figure}
+
+The response of PC headgroup order parameters to the addition of cationic surfactants into a
+membrane in CHARMM36 simulations are in very good agreement with the experimental data (Fig. \ref{changesWITHsurf})
+suggesting that these simulations can be used to interpret how lipid headgroups respond to the
+inclusion of charges into a membrane. Analysis of lipid dihedral bond distributions suggest that
+the lipid conformational changes upon addition of charge occur in phosphate-oxygen and g$_3$-phosphate-oxygen bonds.
+
+To resolve the lipid headgroup conformational ensembles in mixed membranes we analyzed
+headgroup order parameters from mixture of PC and PE or PG and compared the changes
+upon mixing to the experimental data \cite{??} (Figs. \ref{??}). However, the currently
+available force fields failed to capture the headgroup conformational ensembles in mixed membranes
+as we previously observed also for mixtures with PS lipids \cite{antila19}. The best performing force field for
+single componen membrane, CHARMM36, overestimated the influcence of PE to PC conformations and
+undersestimated response to anionic PG headgroup. The latter may arise from overbinding of sodium
+counterions \cite{antila19}.
+
+Also the changes of headgroup order parameters in POPC:POPG (1:1) and (4:1) mixtures
+upon addition of CaCl$_2$
+%from simulations and experiments \cite{borle85,macdonald87}
 in Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4} are in line with
-conclusion from our previous studies \cite{??}:  
+conclusions from our previous studies \cite{catte16,melcr18,antila19,melcr20}:  
 %The decrease of POPC headgroup order parameters in mixtures with POPG lipids with
 %increasing CaCl$_2$ concentration is overestimated in Slipids and Lipid17
-order parameters decrease in the response to bound cations in qualitative agreement with
-experiments, but calcium binding affinity to membranes is overestimated in 
-simulations except in CHARMM36 with the NBfix correction which underestimates the binding affinity.
-On the other hand, implicit inclusion of electronic polarizability to Amber based lipid models
-improves the ion binding affinity leading to a good agreement with experiments.
+%order parameters decrease in the response to bound cations in qualitative agreement with
+%experiments, but
+calcium binding affinity to membranes is typically overestimated in simulations, but
+CHARMM36 with the NBfix correction underestimates the binding affinity, and that the
+%On the other hand,
+results can be improved by implicit inclusion of electronic polarizability.
+Among the available simulations, there are two models that correctly reproduce the
+PG headgroup order parameter changes upon addition of calcium, Lipid17 and Slipids,
+even though the binding affinity of calcium is overestimated in these simulations.
+The response of PC to calcium binding is correctly reproduced by Lipid17ecc model, as
+previously reported for other membranes \cite{melcr18,melcr19}.
+%to Amber based lipid models
+%improves the ion binding affinity leading to a good agreement with experiments.
 %(Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4})

### 2020-11-17 Manuscript/manuscriptPGPEsuppl.tex
Update manuscript

@@
 
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in CHARMM36 dihedrals with increasing amount of cationic surfactant.
-  }
-\end{figure}
-
 \clearpage

### 2020-11-18 scripts/DataBankINFO/AddData.py
United atom simulation handling added

@@ -268,7 +268,9 @@ gromacs_dict = {
                              "TYPE" : "string",
  #                        "EXTENSION": ("txt"),
                              },
-              
+            'UNITEDATOM' : {"REQUIRED": False,
+                            "TYPE" : "string",
+                         },
                }
 
 # Amber
@@ -633,8 +635,18 @@ for sim in sims_working_links :
     sim['MAPPING_DICT'] = mapping_dict
 
     print(sim['MAPPING_DICT'])
-    print(sim)
 
+#in case of a united atom simulation make a dictionary of united atom names 
+    if sim.get('UNITEDATOM'):
+        unitedAtoms = sim['UNITEDATOM'].split(',')
+        unitedAtomsDic = {}
+        for i in range(0, int(len(unitedAtoms)/2)):
+            lipid = unitedAtoms[2*i]
+            UAlipid = unitedAtoms[2*i+1]
+            unitedAtomsDic[lipid]=UAlipid
+        sim['UADICTIONARY'] = unitedAtomsDic
+        
+    print(sim)
 
 # ## Read molecule numbers into dictionary
 
@@ -656,13 +668,12 @@ for sim in sims_working_links :
     tpr = str(dir_tmp)+ '/' + str(ID) + '/' + str(sim.get('TPR')).translate({ord(c): None for c in "']["})
     trj = str(dir_tmp)+ '/' + str(ID) + '/' + str(sim.get('TRJ')).translate({ord(c): None for c in "']["})
     gro = str(dir_tmp) + '/' + str(ID) + '/' + 'conf.gro'
- #   if str(sim.get('INI')).translate({ord(c): None for c in "']["}) != '':
- #       gro = str(sim.get('INI')).translate({ord(c): None for c in "']["})
- #       gro_path = str(dir_wrk) + '/tmp/' + str(ID) + '/' + gro
 
-  #  else:
+    #make gro file
     get_ipython().system('echo System | gmx trjconv -f {trj} -s {tpr} -dump 0 -o {gro}')
-  #  gro_path = str(dir_tmp) + '/' + str(ID) + '/' + 'conf.gro'
+
+    # add gro into dictionary for later use
+    sim['GRO'] = gro
     
     
     
@@ -859,6 +870,10 @@ import warnings
 import subprocess
 import mdtraj
 import json
+import sys
+#sys.path.insert(1, '../')
+import buildH_calcOP_test
+
 #!cp corr_ftios_ind.sh {dir_wrk}
[…]

### 2020-11-19 scripts/DataBankINFO/AddData.py
Added opnmm dictionary and os instead of ipython as suggested in pull request by batukav

@@ -313,7 +313,73 @@ namd_dict = {
 charmm_dict = {}
 
 # OPENMM
-openmm_dict = {}
+openmm_dict = {
+               'INI' : {"REQUIRED": False,
+                        "TYPE" : "files",
+                        "EXTENSION" : ("gro", "pdb",),
+                       }, # Could be not needed in the future (tpr)
+               'MDP' : {"REQUIRED": False,
+                        "TYPE" : "file",
+                        "EXTENSION" : ("mdp",),
+                       }, # Could be not needed in the future (tpr)
+               'TRJ' : {"REQUIRED": True,
+                        "TYPE" : "files",
+                        "EXTENSION" : ("xtc","trr",),
+                       },
+               'PDB' : {"REQUIRED": True,
+                        "TYPE" : "file",
+                        "EXTENSION" : ("pdb",),
+                       },
+               'TPR' : {"REQUIRED": False,
+                        "TYPE" : "file",
+                        "EXTENSION" : ("tpr",),
+                       },
+               'CPT' : {"REQUIRED": False,
+                        "TYPE" : "file",
+                        "EXTENSION" : ("cpt",),
+                       },
+               'TOP' : {"REQUIRED": False,
+                        "TYPE" : "file",
+                        "EXTENSION" : ("top",),
+                       },
+               'ITP' : {"REQUIRED": False,
+                        "TYPE" : "files",
+                        "EXTENSION" : ("itp",),
+                       },
+               'FF'  : {"REQUIRED": False,
+                        "TYPE" : "string",
+                       },
+               'FF_SOURCE' : {"REQUIRED": False,
+                              "TYPE" : "string",
+                              },
+               'FF_DATE' : {"REQUIRED": False,
+                            "TYPE" : "date",
+                           },
+               'DOI' : {"REQUIRED": True,
+                            "TYPE" : "string",
+                           },
+
+               'SYSTEM' : {"REQUIRED": True,
+                            "TYPE" : "string",
+                           },
[…]

### 2020-11-30 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 
+Derivation of different lipid headgroup conformational ensembles in liquid state has been
+inconclusive due to lack of suitable experimental data and tools to interpret the conformational ensembles.
+
 Besides the structure, also ion binding may regulate biophysical activity of especially negatively
@@
 
-Derivation of different lipid headgroup conformational ensembles in liquid state has been
-%discussed previously \cite{??}, but these studies have been
-inconclusive due to lack of suitable experimental data and tools to interpret the conformational ensembles.
-Our experimental order parameters, including the signs, from different lipid headgroups
-combined with the literature data are shown in Fig. \ref{HGorderParameters}.
-The glycerol backbone order parameters are similar for all the lipids, although they move slightly toward
-positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG. 
+To experimentally characterize the differences in lipid headgroup conformational ensembles
+in liquid bilayer phase, we measured the C-H bond order parameters magnitudes and signs
+from POPG and POPE bilayers, and compared those to our previously published results
+for POPC \cite{ferreira13,ferreira16} and POPS \cite{antila19} in figure \ref{HGorderParameters}.
+%Our experimental order parameters, including the signs, from different lipid headgroups
+%combined with the literature data are shown in Fig. \ref{HGorderParameters}.
 %Also the headgroup order parameters of PC lipids are close PE, althought the latter gives systematically
@@
 % more positive value is farther away from zero.
-While the headgroup order parameters are similar for PC and PE lipids,
-PG and PS lipids exhibit distinct values in comparison with other lipids.
-In PS headgroup, the $\alpha$-carbon order parameter exhibits significant forking and
-the $\beta$-carbon has more negative value than in other lipids.
-In PG headgroup, the $\alpha$-carbon order parameter is similar to PE and PG,
-while the positive value of $\beta$-carbon is distinct from all the other lipids.
-Notably, this difference was not observed in previous $^2$H NMR experiments,
-because absolute value of $\beta$-carbon order parameter is similar in PG, PE and PC lipids and the order parameter
-signs were not measured~\cite{wohlgemuth80,gally81,borle85}.
+%While the headgroup order parameters are similar for PC and PE lipids,
+%PG and PS lipids exhibit distinct values in comparison with other lipids.
+The most distinct order parameters are observed for PS and PG headgroups.
+In PS, the $\alpha$-carbon order parameter exhibits significant forking
+and the $\beta$-carbon has more negative value than in other lipids.
+In PG headgroup,
+%the $\alpha$-carbon order parameter is similar to PE and PG, while
+the $\beta$-carbon order parameter has positive sign in contrast to all the other lipids.
+Notably, this has not been observed in traditional $^2$H NMR experiments,
+%because absolute value of $\beta$-carbon order parameter is similar in PG, PE and PC lipids and
+where only the absolute value of the order parameters are measured~\cite{wohlgemuth80,gally81,borle85}.
+The glycerol backbone order parameters are similar for all the lipids, although they move slightly toward
+positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG.
 
@@
 
+To characterize the differences in lipid headgroup conformational ensembles
+leading to the distinct order parameters in PS and PG lipids, we calculate the
+the distributions of heavy atom dihedral angles from CHARMM36 simulations.
+Among available simulation models, this force field has least problems in
+reproducing the experimental lipid headgroup and glycerol backbone order parameters in different lipids,
+see Refs. \cite{botan15} and \cite{antila19}, and Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}.
+Importantly, it captures the experimentally observed distinct order parameters for PS and PG headgroups (Fig. \ref{??}).
+
+show major differences between lipids only for O$_\alpha$-C$_\alpha$ and  C$_\alpha$-C$_\beta$ bonds
+near the end of the headgroup (Fig. \ref{structures}).
+These probably explain the slight differences in the
+angle between headgroup dipole and membrane normal, which decreases 
+in the order PG $>$ PE  $>$ PC  $>$ PS, although difference between PC and PE may an artefact as
+the $\beta$-carbon order parameter in PC is poorly reproduced by the CHARMM36 force field.
 As in previous NMRlipids project results for PC and PS lipids \cite{botan15,antila19},
@@
 
-The conformational ensembles characterized by the distributions of heavy atom dihedral angles
-show major differences between lipids only for O$_\alpha$-C$_\alpha$ and  C$_\alpha$-C$_\beta$ bonds
-near the end of the headgroup (Fig. \ref{structures}).
-These probably explain the slight differences in the
-angle between headgroup dipole and membrane normal, which decreases 
-in the order PG $>$ PE  $>$ PC  $>$ PS, although difference between PC and PE may an artefact as
-the $\beta$-carbon order parameter in PC is poorly reproduced by the CHARMM36 force field.
 Dihedrals of both O-P bonds can freely rotate in all lipids as all possible angles
@@
   \centering
-  \includegraphics[width=18.0cm]{../Figs/PCPEPGPSstructures.png}
+  \includegraphics[width=18.0cm]{../Figs/PCPEPGPSstructures2.png}
   \includegraphics[width=8.0cm]{../Figs/CHARMMfromLIPIDS.eps}
@@
   }
-  \todo{The differences observed in dihedral distributions are not visible in the snapshot figures.} \\
   \todo{More detailed discussion of this figure is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9} \\

### 2020-11-30 user_instructions/user_instructions.tex
AddData instructions

\documentclass[12pt,a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{makeidx}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage[a4paper, total={7.5in, 9in}]{geometry}
\usepackage{float}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{hyperref}
\date{}

\title{Instructions}

\begin{document}
\maketitle
\section*{AddData.py}
AddData.py is a script that builds a database that contains a dictionary file and analysis data of each simulation. The dictionary file contains information about the simulation. The script also calculates order parameters of all CH bonds of the lipids in the simulation. To add a simulation it must be first uploaded to Zenodo (www.zenodo.org). The trajectory and topology files of the simulation are downloaded to the working directory from Zenodo but these are not saved into the database. To add a simulation to the database the user has to give some essential information about the simulation. This is done by writing a info file (*INFO.py) which is passed to AddData.py. 
\newline \\
AddData.py requires GROMACS and MDAnalysis library to be installed.
\newline \\
Run AddData.py as follows:
\newline \\
python3 AddData.py exampleINFO.py
\newline \\
There is also a script runAddData.sh that can be used to loop over several info files to add many simulations at one go.

\section*{Simulation dictionary}
A simulation dictionary contains information about the simulation. Some of the information is provided by the user. The numbers of lipid molecules, solvent and ions are automatically read from the files and so are the simulation temperature and trajectory length. All this information is saved to a file named README.yaml.

\section*{User input}

\subsubsection*{DOI}
DOI is the DOI access number of the simulation. Use the version DOI.

\subsubsection*{SYSTEM}
SYSTEM is name of the system.

\subsubsection*{MAPPING}
MAPPING provides the name of the mapping file for a lipid. It must be given as a string of alternating lipid name and the corresponding mapping file name separated by a comma. The existing mapping files are in the directory named "mapping\_files". If a mapping file of a certain lipid does not exist the user must construct it.
\newline \\
The purpose of a mapping file is to circumvent the problem caused by different atom naming conventions used by different force fields. The first column of a mapping file contains general atom names. The second column contains the name of the atom as it is in the force field. If the lipid consists of several residues which is the case with some AMBER force fields, then a third column is needed which contains the name of the residue to which each atom belongs to.

\subsubsection*{SOFTWARE}
SOFTWARE stands for the name of the software used for running the simulation. The options are GROMACS, AMBER, NAMD, CHARMM and OPENMM. So far, only simulations run with GROMACS are accepted by the script.

\subsubsection*{FF}
FF is the name of the forcefield used in the simulation.
\subsubsection*{FF\_SOURCE}
FF\_SOURCE tells where the forcefield parameters are taken from.
\subsubsection*{FF\_DATE}
FF\_DATE is the date when the forcefield parameters were created. The format is day/month/year.

\subsubsection*{TRJ}
TRJ stands for the name of the trajectory file.
\subsubsection*{TPR}
TPR stands for the name of the topology file.
\subsubsection*{PREEQTIME}
PREEQTIME means the time used for pre-equilibrating the system. This should be in nano seconds.
\subsubsection*{TIMELEFTOUT}
TIMELEFTOUT stands for the length of the simulation in nano seconds to be left out at the beginning of the simulation.
\subsubsection*{UNITEDATOM}
The handling of united atom simulations is enabled by a separate script called buildH\_calcOP.py. In case of a united atom simulation, the user has to give the names of the lipids and the corresponding names of those lipids as they are in the dic\_lipids.py dictionary used by buildH\_calcOP.py script. In case of an all atom simulation this can be given as an empty string.

\subsubsection*{Molecule names}
The user must also provide the names of the molecules, ions and solvent that are used. These are obtained from the tpr file. If a lipid consists of more than one residue the name of the head group residue is given.
\subsubsection*{dir\_wrk}
Finally the user has to give the path of the working directory.

\end{document}

### 2020-11-30 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 Among available simulation models, this force field has least problems in
-reproducing the experimental lipid headgroup and glycerol backbone order parameters in different lipids,
-see Refs. \cite{botan15} and \cite{antila19}, and Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}.
+reproducing the experimental lipid headgroup and glycerol backbone order parameters in different
+lipids \cite{botan15,antila19} (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
 Importantly, it captures the experimentally observed distinct order parameters for PS and PG headgroups (Fig. \ref{??}).
+The dihedral distributions show significant similarity between different lipids for all the other
+bonds, except for the last two in the headgroup end, O$_\alpha$-C$_\alpha$ and  C$_\alpha$-C$_\beta$ bonds,
+in figure \ref{structures}. Therefore, we conclude that the disctinct order parameters of
+PS and PG headgroups can be explained by the differences in these two dihedrals with respenct to other lipids.
+The differences between PC and PE lipids, especially in O$_\alpha$-C$_\alpha$ bond dihedral,
+may be an artefact because the $\beta$-carbon order parameter in PC is poorly reproduced by the CHARMM36 force field \cite{botan15}.
+In larger scale, these differences are reflected, for example, to the 
+angle between headgroup dipole and membrane normal,
+which decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig. \ref{structures}).
 
-show major differences between lipids only for O$_\alpha$-C$_\alpha$ and  C$_\alpha$-C$_\beta$ bonds
-near the end of the headgroup (Fig. \ref{structures}).
-These probably explain the slight differences in the
-angle between headgroup dipole and membrane normal, which decreases 
-in the order PG $>$ PE  $>$ PC  $>$ PS, although difference between PC and PE may an artefact as
-the $\beta$-carbon order parameter in PC is poorly reproduced by the CHARMM36 force field.
-As in previous NMRlipids project results for PC and PS lipids \cite{botan15,antila19},
-none of the MD simulation force fields correctly captures all the
-headgroup and glycerol backbone order parameters of PE and PG lipids
-(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG})
-that would enable a straightforward interpretation of conformational ensembles.
+%
+%As in previous NMRlipids project results for PC and PS lipids \cite{botan15,antila19},
+%none of the MD simulation force fields correctly captures all the
+%headgroup and glycerol backbone order parameters of PE and PG lipids
+%(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG})
+%that would enable a straightforward interpretation of conformational ensembles.
+%
 %show wide variation between different force fields
@@
 %of
-Nevertheless, CHARMM36 force field, which gives the results closest to the experiments for all lipids,
-captures the essential differences between
-PC, PS, PG and PE headroup order parameters  (Fig.~\ref{structures}) with the exception of $\beta$-carbon order parameter of PC
+%
+%Nevertheless, CHARMM36 force field, which gives the results closest to the experiments for all lipids,
+%captures the essential differences between
+%PC, PS, PG and PE headroup order parameters  (Fig.~\ref{structures}) with the exception of $\beta$-carbon order parameter of PC
+%
 %give the best overall agreement with experiments for
@@
 %not within the experimental error bars, it 
-which is too negative when compared with other lipids or experiments \cite{botan15}.
+%
+%which is too negative when compared with other lipids or experiments \cite{botan15}.
+%
 % As already pointed out in our previous study \cite{antila19},

### 2020-12-01 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 
-\clearpage
+%\clearpage
+\begin{figure*}[]
+  \centering
+%  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
+  %  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCPSPEPG.eps}
+   \includegraphics[width=18.0cm]{../Figs/figure1.eps}
+   \caption{\label{HGorderParameters}
+     Chemical structure, refocused-INEPT spectrum, 2D R-PDLF spectra, and S-DROSS data (from top to bottom) of {\bf A)} POPE  and {\bf B)} POPG.
+    {\bf C)} Chemical structure of POPC and POPS.
+    {\bf D)} Headgroup and glycerol backbone order parameters from different experiments in lamellar liquid disordered phase.
+    The values and signs for POPE (310~K) and POPG (298~K)
+    measured in this work, and for POPS (298~K) \cite{antila19} and POPC (300~K) \cite{ferreira13,ferreira16}
+    previously using $^{13}$C NMR. The literature values for
+    DOPS with 0.1M of NaCl (303~K) \cite{browning80},
+    POPG with 10nM PIPES (298~K) \cite{borle85},
+    DPPG with 10mM PIPES and 100mM NaCl (314~K) \cite{wohlgemuth80}, 
+    DPPE (341~K) \cite{seelig76},
+    E.coliPE and E.coliPG (310~K) \cite{gally81}
+    are measured using $^2$H NMR. The signs from $^{13}$C NMR are used also for the literature values.
+   }
+   \todo{This is a sketch, needs a lot of polishing.} \\
+  \todo{D) could be clarified as Fig. 2 in the NMRlipids IVps paper.}
+\end{figure*}
+
 \section{Results and Discussion}
@@
 %\subsection{Headgroup and glycerol backbone order parameters of POPE and POPG from $^{13}$C NMR}
+
+\begin{figure*}[]
+  \centering
+   \includegraphics[width=18.0cm]{../Figs/figure2.eps}
+%  \includegraphics[width=18.0cm]{../Figs/PCPEPGPSstructures2.png}
+%  \includegraphics[width=8.0cm]{../Figs/CHARMMfromLIPIDS.eps}
+%%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals1.png}
+%%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals2.png}
+%  \includegraphics[width=8.0cm]{../Figs/PNangleNORM.eps}
+%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALS.eps}
+   \caption{\label{structures}
+     Results from CHARMM36 simulations demonstrating the differences in conformational ensembles betwee different lipids. 
+     {\bf A)} Snapshots with overlayed C$_\beta$, C$_\alpha$ and O$_\alpha$ atoms and occurence of different conformations.
+     {\bf B)} Headgroup and glycerol backbone region order parameters of different different lipids.
+     {\bf C)} Distributions of heavy atom dihedral angles of different lipids from CHARMM36 simulations.
+     {\bf D)} Distributions of P-N vector angle with respect to membrane normal.
+  }
+  \todo{This is a draft and requires quite a bit of polishing.
+    More detailed discussion of this figure is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9} \\
+\end{figure*}
+
 \subsection{Conformational ensembles of different lipid headgroups in bulk bilayer}
@@
 in liquid bilayer phase, we measured the C-H bond order parameters magnitudes and signs
-from POPG and POPE bilayers, and compared those to our previously published results
-for POPC \cite{ferreira13,ferreira16} and POPS \cite{antila19} in figure \ref{HGorderParameters}.
+from POPG and POPE bilayers.
+Order parameters for POPE are in good agreement with previous $^2$H NMR experiments \cite{seelig76}
+and similar to our previous results for POPC \cite{ferreira16}.
+In POPG, the $\beta$ and g$_2$ carbons have similar chemical environment and their peaks
+overlap in the NMR spectra (Fig. \ref{HGorderParameters} B).
+The signs of these order parameters were solved using SIMPSON simulations to interpre the
+S-DROSS experiments in similar fashion as we did previously for PS lipids \cite{antila19}.
+The resulting order parameters are compared with our previously published results
+for POPC \cite{ferreira13,ferreira16} and POPS \cite{antila19} in figure \ref{HGorderParameters} D.
+
 %Our experimental order parameters, including the signs, from different lipid headgroups
@@
 %PG and PS lipids exhibit distinct values in comparison with other lipids.
+
 The most distinct order parameters are observed for PS and PG headgroups.
@@
 
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
-  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCPSPEPG.eps}
-  \caption{\label{HGorderParameters}
-    (top) Chemical structure of different lipids.
-    (bottom) Headgroup and glycerol backbone order parameters
-    from different experiments in lamellar liquid disordered phase.
-    The values and signs for POPE (310~K) and POPG (298~K)
-    measured in this work, and for POPS (298~K) \cite{antila19} and POPC (300~K) \cite{ferreira13,ferreira16}
-    previously using $^{13}$C NMR. The literature values for
-    DOPS with 0.1M of NaCl (303~K) \cite{browning80},
-    POPG with 10nM PIPES (298~K) \cite{borle85},
-    DPPG with 10mM PIPES and 100mM NaCl (314~K) \cite{wohlgemuth80}, 
-    DPPE (341~K) \cite{seelig76},
-    E.coliPE and E.coliPG (310~K) \cite{gally81}
-    are measured using $^2$H NMR. The signs from $^{13}$C NMR are used also for the literature values.
-  }
-  \todo{The bottom figure could be clarified as Fig. 2 in the NMRlipids IVps paper.}
-\end{figure}
-
 To characterize the differences in lipid headgroup conformational ensembles
@@
 Importantly, it captures the experimentally observed distinct order parameters for PS and PG headgroups (Fig. \ref{??}).
-The dihedral distributions show significant similarity between different lipids for all the other
-bonds, except for the last two in the headgroup end, O$_\alpha$-C$_\alpha$ and  C$_\alpha$-C$_\beta$ bonds,
-in figure \ref{structures}. Therefore, we conclude that the disctinct order parameters of
-PS and PG headgroups can be explained by the differences in these two dihedrals with respenct to other lipids.
-The differences between PC and PE lipids, especially in O$_\alpha$-C$_\alpha$ bond dihedral,
+The dihedral distributions in figure \ref{structures} show significant similarity between different lipids for all the other
+bonds, except for the last two bonds in the headgroup end, O$_\alpha$-C$_\alpha$ and  C$_\alpha$-C$_\beta$.
+These differences explaing the disctinct order parameters of
+PS and PG headgroups with respenct to the other lipids.
+The difference between PC and PE lipids in the O$_\alpha$-C$_\alpha$ bond dihedral
 may be an artefact because the $\beta$-carbon order parameter in PC is poorly reproduced by the CHARMM36 force field \cite{botan15}.
-In larger scale, these differences are reflected, for example, to the 
+The differences between lipids are reflected also to the 
 angle between headgroup dipole and membrane normal,
@@
 
-Dihedrals of both O-P bonds can freely rotate in all lipids as all possible angles
-are observed in the distributions. For PS, however, some angles are quite unlikely,
-possible explaining the more rigid headgroup structures proposed for PS lipids~\cite{browning80,buldt81}.
-The eclipsed (0 or 360 degrees) are not present in any other dihedrals, and anti-eclipsed (120 or 240 degrees)
-are not present in C$_\alpha$-C$_\beta$ and g$_1$-g$_2$ bonds.
+The analysis of dihedral angle distributions reveal the free rotation of dihedral around
+phosphorus-oxygen bonds in all lipids as all possible angles are observed in the distributions.
+Results in figure \ref{structure} support the recently proposed models where this free rotation 
+decouples the dynamics and conformational ensembles between acyl chains and headgroup in all lipids containing
+the phosphorus group \cite{antila21}. 
+However, some dihedral angles of phosphorus-oxygen bonds are less likely for PS, which
+%suggests that the decoupling may be weaker for PS lipids
+possibly explains the more rigid headgroup structures proposed for PS lipids~\cite{browning80,buldt81}.
+
+%The eclipsed (0 or 360 degrees) are not present in any other dihedrals, and anti-eclipsed (120 or 240 degrees)
+%are not present in C$_\alpha$-C$_\beta$ and g$_1$-g$_2$ bonds.
 
@@
 
-In conclusion, the results suggest that free rotations around O-P bonds decouple the
-headgroup and glycerol backbone structure and dynamics in similarly in all lipids,
-althought this rotation is slower in PS lipids where crossing certain barriers are slower.
-Despite significant differences in dihedral distributions near the headgroup end,
-all lipids have dihedral angles within approximately the same ranges.
-This suggests that all lipids can quite freely arrange to multiple headgroup conformations
-when interaction with proteins, ions or other biomolecules.
+\begin{figure*}[bt]
+  \centering
+  \includegraphics[width=18.0cm]{../Figs/figure3.eps}
+  \caption{\label{changesWITHsurf}
+    {\bf A)} Modulation of PC headgroup order parameters and P-N vector angle upon addition of cationic surfactant
+    from CHARMM36 simulations compared with experimental data \cite{??}.
+    {\bf B)} Changes in PC headgroup conformational ensembles upon increasing amount of positive charge in bilayer,
+    characterized by the heavy atom dihedral distributions, from CHARMM36 simulations.
+  }
+\end{figure*}
+
+In conclusion, all lipid headgroups sample very broad conformational ensembles in liquid lamellar phase.
+%the results suggest that free rotations around O-P bonds decouple the
+%headgroup and glycerol backbone structure and dynamics in similarly in all lipids,
+%althought this rotation is slower in PS lipids where crossing certain barriers are slower.
+Despite the differences in dihedral distributions of two the last bonds at the headgroup end,
+leading to the distinct order parameters for PS and PG lipids,
+lipids with different headgroup sample dihedral angles within approximately same ranges.
+Altogether, our results suggest that lipid headgroups and very flexible thereby being
+able to adopt multiple conformations when interacting with proteins, ions or other biomolecules.
+Furthermore, approximately same conformations are present liquid lamellar phase for different
+headgroups studied here, althought the probability weight for the conformations are slightly different.
 %the order parameter experiments suggest that the glycerol backbone conformations in all lipids
@@
 
-\begin{figure*}[!h]
-  \centering
-  \includegraphics[width=18.0cm]{../Figs/PCPEPGPSstructures2.png}
-  \includegraphics[width=8.0cm]{../Figs/CHARMMfromLIPIDS.eps}
-%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals1.png}
-%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals2.png}
-  \includegraphics[width=8.0cm]{../Figs/PNangleNORM.eps}
-  \includegraphics[width=8.0cm]{../Figs/DIHEDRALS.eps}
-  \caption{\label{structures}
-    Overlayed snapshots and dihedral angle distributions from CHARMM36 simulations of different lipids
-    which give the best agreement with experiments.
-  }
-  \todo{More detailed discussion of this figure is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9} \\
-\end{figure*}
-
-\clearpage
 \subsection{Lipid conformational ensembles in lipid bilayers with bound ions}
-Phosphatidylcholine headgroup conformational ensemble and orientation are unchanged
-upon addition of neutral molecules such as cholesterol or sphingomyelin into a bilayer,
-while incorporation of charges affect the headgroup dipole vector tilt which can
-be detected by measuring C-H bond order parameters of $\alpha$ and $\beta$ carbons \cite{??}.
+%Phosphatidylcholine headgroup conformational ensemble and orientation are unchanged
+%upon addition of neutral molecules such as cholesterol or sphingomyelin into a bilayer,
+%while
+Order parameters of $\alpha$ and $\beta$ carbons in PC lipid headgroup are known to depend
+linearly on the incorporation of charged molecules into a lipid bilayer,
+which is explained by tilting of the headgroup dipole orientation \cite{??}.
+%are kn affect the headgroup dipole vector tilt which can
+%be detected by measuring C-H bond order parameters of $\alpha$ and $\beta$ carbons.
 Such changes have been observed upon addition of charged lipids, proteins, surfactants,
@@
 
-\begin{figure*}[bt]
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+%  \caption{\label{DIHSwithSURF}
+%
+%  }
+%\end{figure}
+
+The experimentally measured decrease of PC headgroup order parameters upon addition of cationic surfactants into a bilayer
+are well captured by CHARMM36 simulations in figure \ref{changesWITHsurf} A,
+suggesting that these simulations can be used to interpret how lipid headgroup conformational ensembles
+respond to the binding of positively charged into a membrane.
+Charactization of conformational ensembles using heavy atom dihedral angle distributions in figure \ref{changesWITHsurf} B
+reveal that choline region is essentially unchanged upon addition of carge.
+The major changes upon addition of charge are observed in dihedrals related to the phosphate oxygens,
+while small change is observed also in the g$_2$-g$_3$ bond in the glycerol backbone.
+This result is in line with the recently proposed model suggesting that the flexible phosphate enables the headgroup
+orient according to the charge accumulated into a membrane \cite{antila21}.
+
+\begin{figure*}[]
   \centering
-  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCvsSURFchange.eps}
-  \caption{\label{changesWITHsurf}
-    Modulation of PC headgroup order parameters and P-N vector angle upon addition of cationic surfactant.
+  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC1.eps}
+  \caption{\label{changesWITHCaClPG}
+    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
+    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
+    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
+    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
+    The changes with respect to the systems without CaCl$_2$ are shown for other data than
+    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
+    Calsium density distributions are shown in figure \ref{CAdensPG}.
+    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
+    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
   }
@@
 
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
-  \caption{\label{DIHSwithSURF}
-    Changes in CHARMM36 dihedrals with increasing amount of cationic surfactant.
-  }
-\end{figure}
+Headgroup order parameters of PC lipids respond similarly also to the addition of charged lipids in experiments \cite{??}.
+To resolve the conformational ensembles of lipid headgroups also in biologically relevant
+mixtures, % of zwitterionic and charged lipids,
+we compared the response of headgroup order parameters in
+PC lipids to the addition of PE or PG (Fig. \ref{??}),
+as well as in PG lipids to the addition of PC (Fig. \ref{??}),
+between simulations and experiments. However, the accuracy of currently available force fields
+is not sufficient to capture the headgroup conformational ensembles
+in mixed membranes: 
+The best performing force field for single componen membranes, CHARMM36, overestimates
+the effect of PE to PC conformations and underestimates the response PC to the addition
+of anionic PG headgroup. Previously we explained similar result for PC/PS mixtures  
+%as we previously observed also for mixtures with PS lipids \cite{antila19}.
+%The latter may arise from
+by overbinding of the sodium counterions \cite{antila19}.
 
-The response of PC headgroup order parameters to the addition of cationic surfactants into a
-membrane in CHARMM36 simulations are in very good agreement with the experimental data (Fig. \ref{changesWITHsurf})
-suggesting that these simulations can be used to interpret how lipid headgroups respond to the
-inclusion of charges into a membrane. Analysis of lipid dihedral bond distributions suggest that
-the lipid conformational changes upon addition of charge occur in phosphate-oxygen and g$_3$-phosphate-oxygen bonds.
-
-To resolve the lipid headgroup conformational ensembles in mixed membranes we analyzed
-headgroup order parameters from mixture of PC and PE or PG and compared the changes
-upon mixing to the experimental data \cite{??} (Figs. \ref{??}). However, the currently
-available force fields failed to capture the headgroup conformational ensembles in mixed membranes
-as we previously observed also for mixtures with PS lipids \cite{antila19}. The best performing force field for
-single componen membrane, CHARMM36, overestimated the influcence of PE to PC conformations and
-undersestimated response to anionic PG headgroup. The latter may arise from overbinding of sodium
-counterions \cite{antila19}.
-
-Also the changes of headgroup order parameters in POPC:POPG (1:1) and (4:1) mixtures
-upon addition of CaCl$_2$
-%from simulations and experiments \cite{borle85,macdonald87}
-in Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4} are in line with
-conclusions from our previous studies \cite{catte16,melcr18,antila19,melcr20}:  
+In order to analyze how conformational ensemble of PC and PG lipids response to the
+bound ions, we compared the changes of headgroup order parameters
+in POPC:POPG (1:1) and (4:1) mixtures upon addition of CaCl$_2$ from different simulations
+to the experimental data available in the literature \cite{borle85,macdonald87}
+in figures \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4}.
+As in our previous studies \cite{catte16,melcr18,antila19,melcr20},  
 %The decrease of POPC headgroup order parameters in mixtures with POPG lipids with
@@
 %experiments, but
-calcium binding affinity to membranes is typically overestimated in simulations, but
-CHARMM36 with the NBfix correction underestimates the binding affinity, and that the
-%On the other hand,
-results can be improved by implicit inclusion of electronic polarizability.
-Among the available simulations, there are two models that correctly reproduce the
-PG headgroup order parameter changes upon addition of calcium, Lipid17 and Slipids,
-even though the binding affinity of calcium is overestimated in these simulations.
-The response of PC to calcium binding is correctly reproduced by Lipid17ecc model, as
-previously reported for other membranes \cite{melcr18,melcr19}.
+the calcium binding affinity to membranes is typically overestimated in simulations,
+except by CHARMM36 with the NBfix correction which underestimates the binding affinity,
+and the implicit inclusion of electronic polarizability improves the results.
+Lipid17ecc model with the implicit inclusion of electronic polarizability gives
+the most realistic response of PC lipid headgroup order parameters to the binding of calcium.
+In this model, the main effect of calcium to the lipid conformational ensemble is
+%the decrease is headgroup order parameters observed in this model is
+the slight change of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral distribution from trans state
+to eclipsed conformations (Fig. \ref{DIHSwithCAlipid17eccPOPC}).
+This is in line with the changes observed in CHARMM36 simulations upon addition of charged surgactants (Fig. \ref{changesWITHsurf}),
+despite the major differences in lipid headgroup conformational ensembles between
+these models without ions (Fig. \ref{structures} vs. Fig. \ref{DIHSwithCAlipid17eccPOPC}).
+%of lipid headgroup is not 
+%Changes in lipid conformational ensembles characterized by distributions of heavy atom dihedral angles
+%are very small upon binding of calcium or addition of cationic surfactants to membranes (Figs. \ref{??}).
+%Yet, such changes are sufficient to tilt the headgroup dipole angle and reproduce the experimentally observed
+%order parameter changes in $\alpha$ and $\beta$ carbons.
+
+%Among the available simulations, there are two models that correctly reproduce the
+%PG headgroup order parameter changes upon addition of calcium, Lipid17 and Slipids,
+%even though the binding affinity of calcium is overestimated in these simulations.
+%The response of PC to calcium binding is correctly reproduced by Lipid17ecc model, as
+%previously reported for other membranes \cite{melcr18,melcr19}.
+
 %to Amber based lipid models
@@
 
-\begin{figure*}[bt]
-  \centering
-  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC1.eps}
-  \caption{\label{changesWITHCaClPG}
-    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
-    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
-    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
-    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
-    The changes with respect to the systems without CaCl$_2$ are shown for other data than
-    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
-    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
-    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
-  }
-\end{figure*}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/CAdensPG1PC1.eps}
-  \caption{\label{CAdensPG}
-    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (1:1) mixtures with different force fields.
-  }
-\end{figure}
-
-Changes in lipid conformational ensembles characterized by distributions of heavy atom dihedral angles
-are very small upon binding of calcium or addition of cationic surfactants to membranes (Figs. \ref{??}).
-Yet, such changes are sufficient to tilt the headgroup dipole angle and reproduce the experimentally observed
-order parameter changes in $\alpha$ and $\beta$ carbons.
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in POPC CHARMM36 dihedrals with increasing amount of CaCl$_2$.
-  }
-\end{figure}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in POPG CHARMM36 dihedrals with increasing amount of CaCl$_2$.
-  }
-\end{figure}
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+%  \caption{\label{DIHSwithCA}
+%    Changes in POPG CHARMM36 dihedrals with increasing amount of CaCl$_2$.
+%  }
+%\end{figure}
 
@@
 
+Decrease in headgroup order parameters of PC lipids upon addition of charges is
+usually captured qualitatitely despite inaccurate lipid conformational ensembles in the models \cite{catte16},
+but situation with PS lipids is more complicated \cite{antila19,meclr20}.
+Here we observe in figure \ref{changesWITHCaClPG} that
 Lipid17 and Slipids force fields correctly capture the PG $\beta$-carbon order parameter response to CaCl$_2$
-even thought the binding affinity was too large based on the comparison of PC headgroup order parameter changes with experiments.
+even thought the calcium binding affinity was overestimated.
+%based on the comparison of PC headgroup order parameter changes with experiments.
 %While applying ECC to Lipid17 improved the PC headgroup order parameter response and binding affinity,
-The response of PG $\beta$-carbon order parameter to calcium is too weak in CHARMM36 and ECC-lipid17 force fields,
-but decrease of the order parameter is associated with decreased P-N vector angle in all
-force fields. Even thought the change in headgroup orientation upon addition of calcium is smaller
-in PG lipids than in PC, the conformational changes in PG are more significant because also
-dihedral angle distribution around O$_\alpha$-C$_\alpha$ bond is affected by calcium binding (Fig. ??).
-However, because none of the force fields simulatneously capture both the PG conformational ensemble and calcium binding,
-more accurate force fields are required for more detailed studies of the effect of ions to PG conformational enembles.
+%The response of PG $\beta$-carbon order parameter to calcium is too weak in CHARMM36 and ECC-lipid17 force fields,
+%but decrease of the order parameter is associated with decreased P-N vector angle in all force fields.
+Analysis of the changes in PG conformational ensemble from these models suggest that
+calcium is affecting also to the headgroup glycerol conformations
+(Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}),
+in contrast to PC lipids where only conformations near phosphate were affected.
+%dihedral angle distribution around O$_\alpha$-C$_\alpha$ bond is affected by calcium binding (Fig. ??).
+On the other hand, the upward tilting of the headgroup dipole is weaker
+%Even thought the change in headgroup orientation upon addition of calcium is smaller
+in PG than in PC headgroup, possible due to the compenating effects from the changes in
+phosphate and glycerol regions. 
+However, none of the simulations captures and calcium binding affinity and conformational ensemble of PG lipids
+simulationeously, and differences between force fields are observed for the response of PG $\alpha$ carbon to
+the added calcium for which experimental data is not available.
+Therefore, more accurate force fields are required for the solid analysis of PG conformations enembles in different ionic conditions.
+
 %and PG

### 2020-12-01 Manuscript/manuscriptPGPEsuppl.tex
Update manuscript

@@
 
+\section{R-PDLF and SDROSS experiments}
+
+\begin{figure}[]
+%  \centering
+  \includegraphics[width=0.8\textwidth]{../Figs/POPEexperiment.pdf}
+  \caption{\label{POPEspectra}
+    (A) Chemical structure of POPE with the labeling of headgroup and glycerol backbone carbons.
+    (B) INEPT spectra from POPE sample with the headgroup and glycerol backbone peaks labeled.
+    (C) 2D R-PDLF spectra
+    (D) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
+    (E) Experimetal S-DROSS curves giving signs of the order parameters.
+  }
+  \todo{A, B etc. labels to be put in the figure.}
+\end{figure}
+
+\begin{figure}[]
+%  \centering
+  \includegraphics[width=0.8\textwidth]{../Figs/PGexpRPDLF.eps}
+  \caption{\label{POPGspectra}
+    (A) Chemical structure of POPG with the labeling of headgroup and glycerol backbone carbons.
+    (B) INEPT spectra from POPG sample with the headgroup and glycerol backbone peaks labeled.
+    (C) 2D R-PDLF spectra
+    (D) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
+    (E) Experimetal S-DROSS curves giving signs of the order parameters.
+  }
+\end{figure}
+
+\begin{figure}[]
+%  \centering
+  \includegraphics[width=0.8\textwidth]{../Figs/SIMPSON.pdf}
+  \caption{\label{POPGsimpson}
+    Simpson simulaton of S-DROSS curve of $\beta$-carbon of POPG.
+  }
+\end{figure}
+
+\clearpage
 \section{Comparison of headgroup order parameters from different force fields to experiments}
@@
 \end{figure}
+
+\section{Calcium density distributions}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/CAdensPG1PC1.eps}
+  \caption{\label{CAdensPG}
+    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (1:1) mixtures with different force fields.
+  }
+\end{figure}
+
+\clearpage
+
+\section{Changes of PG headroup order parameters upon addition of PC}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPC.eps}
+  \caption{\label{HGorderparametersPGvsPC}
+    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
+    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
+  }
+\end{figure}
+
+\clearpage
+
+\section{Sodium binding to POPC simulations}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/NAdensPC.eps}
+  \caption{\label{NAdensPC}
+    Sodium (solid line) and choride ion density profiles along membrane normal
+    from different simulations with PC lipids.
+  }
+  \todo{Discussion about differences to the NMRlipids II to be discussed once we have the details on ions models.}
+\end{figure}
+
+\clearpage
+
+\section{Calcium binding to POPC:POPG (4:1) mixtures}
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC4.eps}
+  \caption{\label{changesWITHCaClPG1PC4}
+    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (4:1)
+    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{macdonald87} and simulations.
+    The changes with respect to the systems without CaCl$_2$ are shown for other data than
+    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
+  }
+  \todo{Lipid17ecc data to be analyzed and added.}
+\end{figure*}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{../Figs/CAdensPG1PC4.eps}
+  \caption{\label{CAdensPG1PC4}
+    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (4:1) mixtures with different force fields.
+  }
+\end{figure}
+
+\section{Changes in headgroup conformations upon addition of CaCl$_2$}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+  \caption{\label{DIHSwithCA}
+    Changes in POPC CHARMM36 dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
+  \caption{\label{DIHSwithCAlipid17eccPOPC}
+    Changes in POPC lipid17ecc dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MSlipids.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
+%  \caption{\label{DIHSwithCA}
+%    Changes in POPC Slipids dihedrals with increasing amount of CaCl$_2$.
+%  }
+%\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
+  \caption{\label{DIHSwithCAslipidsPOPG}
+    Changes in POPG Slipids dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
+%  \caption{\label{DIHSwithCA}
+%    Changes in POPC lipid17 dihedrals with increasing amount of CaCl$_2$.
+%  }
+%\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
+  \caption{\label{DIHSwithCAlipid17POPG}
+    Changes in POPG lipid17 dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17ecc.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
+%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
+%  \caption{\label{DIHSwithCA}
+%    Changes in POPG lipid17ecc dihedrals with increasing amount of CaCl$_2$.
+%  }
+%\end{figure}
 
@@
 \clearpage
-\section{R-PDLF and SDROSS experiments}
-
-\begin{figure}[]
-%  \centering
-  \includegraphics[width=0.8\textwidth]{../Figs/POPEexperiment.pdf}
-  \caption{\label{POPEspectra}
-    (A) Chemical structure of POPE with the labeling of headgroup and glycerol backbone carbons.
-    (B) INEPT spectra from POPE sample with the headgroup and glycerol backbone peaks labeled.
-    (C) 2D R-PDLF spectra
-    (D) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
-    (E) Experimetal S-DROSS curves giving signs of the order parameters.
-  }
-  \todo{A, B etc. labels to be put in the figure.}
-\end{figure}
-
-\begin{figure}[]
-%  \centering
-  \includegraphics[width=0.8\textwidth]{../Figs/PGexpRPDLF.eps}
-  \caption{\label{POPGspectra}
-    (A) Chemical structure of POPG with the labeling of headgroup and glycerol backbone carbons.
-    (B) INEPT spectra from POPG sample with the headgroup and glycerol backbone peaks labeled.
-    (C) 2D R-PDLF spectra
-    (D) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
-    (E) Experimetal S-DROSS curves giving signs of the order parameters.
-  }
-\end{figure}
-
-\begin{figure}[]
-%  \centering
-  \includegraphics[width=0.8\textwidth]{../Figs/SIMPSON.pdf}
-  \caption{\label{POPGsimpson}
-    Simpson simulaton of S-DROSS curve of $\beta$-carbon of POPG.
-  }
-\end{figure}
-
-\clearpage
-
-\section{Changes of PG headroup order parameters upon addition of PC}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPC.eps}
-  \caption{\label{HGorderparametersPGvsPC}
-    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
-    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
-  }
-\end{figure}
-
-\clearpage
-
-\section{Sodium binding to POPC simulations}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/NAdensPC.eps}
-  \caption{\label{NAdensPC}
-    Sodium (solid line) and choride ion density profiles along membrane normal
-    from different simulations with PC lipids.
-  }
-  \todo{Discussion about differences to the NMRlipids II to be discussed once we have the details on ions models.}
-\end{figure}
-
-\clearpage
-
-\section{Calcium binding to POPC:POPG (4:1) mixtures}
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC4.eps}
-  \caption{\label{changesWITHCaClPG1PC4}
-    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (4:1)
-    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{macdonald87} and simulations.
-    The changes with respect to the systems without CaCl$_2$ are shown for other data than
-    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
-  }
-  \todo{Lipid17ecc data to be analyzed and added.}
-\end{figure*}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/CAdensPG1PC4.eps}
-  \caption{\label{CAdensPG1PC4}
-    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (4:1) mixtures with different force fields.
-  }
-\end{figure}
-
-\section{Changes in headgroup conformations upon addition of CaCl$_2$}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MSlipids.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in POPC Slipids dihedrals with increasing amount of CaCl$_2$.
-  }
-\end{figure}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MSlipids.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in POPG Slipids dihedrals with increasing amount of CaCl$_2$.
-  }
-\end{figure}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in POPC lipid17 dihedrals with increasing amount of CaCl$_2$.
-  }
-\end{figure}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in POPG lipid17 dihedrals with increasing amount of CaCl$_2$.
-  }
-\end{figure}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in POPC lipid17ecc dihedrals with increasing amount of CaCl$_2$.
-  }
-\end{figure}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in POPG lipid17ecc dihedrals with increasing amount of CaCl$_2$.
-  }
-\end{figure}
-
-\clearpage
 \bibliography{refs.bib}

### 2020-12-02 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 Decrease in headgroup order parameters of PC lipids upon addition of charges is
-usually captured qualitatitely despite inaccurate lipid conformational ensembles in the models \cite{catte16},
+usually qualitatitely captured in simulations despite the inaccuries in lipid conformational ensembles \cite{catte16},
 but situation with PS lipids is more complicated \cite{antila19,meclr20}.
-Here we observe in figure \ref{changesWITHCaClPG} that
-Lipid17 and Slipids force fields correctly capture the PG $\beta$-carbon order parameter response to CaCl$_2$
-even thought the calcium binding affinity was overestimated.
+Here, Lipid17 and Slipids force fields correctly capture the PG $\beta$-carbon order parameter response to CaCl$_2$
+in figure \ref{changesWITHCaClPG} even thought the calcium binding affinity was overestimated.
 %based on the comparison of PC headgroup order parameter changes with experiments.
@@
 %but decrease of the order parameter is associated with decreased P-N vector angle in all force fields.
-Analysis of the changes in PG conformational ensemble from these models suggest that
-calcium is affecting also to the headgroup glycerol conformations
+Heavy atom dihedral angle distributions calculated from these simulations
+%the changes in PG conformational ensemble from these models
+suggest that also the headgroup glycerol conformations in PG are affected by calcium
 (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}),
@@
 %Even thought the change in headgroup orientation upon addition of calcium is smaller
-in PG than in PC headgroup, possible due to the compenating effects from the changes in
+in PG than in PC headgroup, possible due to the compensating effects from the changes in
 phosphate and glycerol regions. 
 However, none of the simulations captures and calcium binding affinity and conformational ensemble of PG lipids
-simulationeously, and differences between force fields are observed for the response of PG $\alpha$ carbon to
-the added calcium for which experimental data is not available.
+simultaneously, and experimental data to evaluate 
+%differences between force fields are observed for
+the response of $\alpha$-carbon order parameters to
+the added calcium in PG is not available.
 Therefore, more accurate force fields are required for the solid analysis of PG conformations enembles in different ionic conditions.
@@
 
-\subsection{Protein bound lipid conformations }
-\begin{figure*}[!h]
+Despite the limited cabability of simulations to interpret some experimental data due to
+the inaccuries in force fields, we can conclude that the experimentally observed changes
+in headgroup order parameters upon addition of charges into bilayer arise from relatively
+small changes in conformational ensembles. These changes can be characterised by
+mild changes in dihedral angle distributions, rather than restriction of lipids into
+fixed conformations. Therefore, lipid headgroups remain in disordered state sampling large
+space of different conformations, thereby being able to interact with different molecules
+in various ways, also when charged molecules are bound to membranes.
+
+\subsection{Protein bound lipid conformations}
+
+While our results in previous sections suggest that lipid headgroups
+sample large conformational space in liquid lamellar phase, lipids
+can bound to proteins in fixed conformations which can be then detected
+using crystallography or cryo-EM. Such protein bound lipid conformations are
+available within protein structures deposited in The Protein Data Bank (PDB)
+available at \url{http://www.rcsb.org/} \cite{berman00}. With our criterias described
+in the methods, we found
+?? conformations of PC lipids,
+?? conformations of PE lipids,
+?? conformations of PG lipids, and
+?? conformations of PS lipids from PDB.
+
+\begin{figure}[]
   \centering
-  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPCfromPDB.eps}
-  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPEfromPDB.eps}
-  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPGfromPDB.eps}
-  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPSfromPDB.eps}
-  \caption{\label{structures}
+  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSALLfromPDB.eps}
+%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPEfromPDB.eps}
+%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPGfromPDB.eps}
+%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPSfromPDB.eps}
+  \caption{\label{dihedralsFROMpdb}
     Dihedral distributions from simulations and lipid structures in PDB.
   }
-\end{figure*}
+\end{figure}
 
-\clearpage
+To analyze the relation of lipid headgroup conformational ensembles between
+liquid lamellar and protein bound states, we calculated the heavy atom dihedral angle
+distributions also from the protein conformations found from the PDB.
+The results in figure \ref{dihedralsFROMpdb} do not reveal major differences
+between different lipid headgroups when bound to proteins.
+Some preference for trans state in C$_\alpha$-C$_\beta$ bond of PC lipids, and
+for positive angles in O$_\alpha$-C$_\alpha$ and P-O$_\alpha$ of PS lipids with respect
+to other lipids may be present in the data, but more statistics is required for solid
+conclusions. Therefore, the differences in conformational ensembles between different
+lipids observed in liquid lamellar state are not clearly visible in the protein bound states.
+
+Almost all dihedral angles are found from the protein bounds states for all dihedrals
+except for P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P, which seem to avoid
+the cis conformations. The wide range of available dihedrals is similar to our results
+in liquid lamellar state, except for O$_\alpha$-C$_\alpha$-C$_\beta$-/C$_\gamma$ and
+g$_1$-g$_2$-g$_3$-O$_{g_3}$ dihedrals which seem to be more restricted in the liquid
+lamellar phase (Fig. \ref{structures}). This suggests that lipids compromize their
+preferred conformations when bound to proteins to minimize the intermolecular interactions
+with proteins.
+
 \section{Conclusions}
+
+We have measured the C-H bond order parameters with the signs for headgroup and glycerol
+backbone regions of the most most abundant biological phopholipids, PC, PE, PG and PS.
+Combining this experimental data with the large amount of simulation data collected 
+using the NMRlipids open collaboration, we were able to resolve the differences between
+conformational ensembles of lipid headgroup leading to the differences observed in NMR experiments.
+
+Our results indicate that lipid headgroups are flexible and sample large conformational space
+in biologically relevant liquid lamellar state, also in membranes containing charged molecules,
+which have the largest affect on lipid headgroups in NMR experiments.
+Therefore, lipids can bind to proteins and other biomolecules in multiple different conformations,
+as indeed observed in the analysis of protein bound lipid conformations available from PDB.
+The weak correlation between lipid structures in liquid lamellar phase and protein bound state
+suggestst that the selective lipid binding to proteins is dominated by the intermolecular lipid-protein
+interactions, and the differences in conformational ensembles between different lipid types
+play only a minor role.
+
+Our results pave the way to understand how lipids regulate membrane protein function.
+For example, the resolved lipid conformational ensembles in liquid lamellar phase and
+realistic MD simulations enables accurate analysis of lipid-protein interactions energetics.
+Furthermore, our results demonstrate how NMR data can be combined with the large amount of
+indexed MD simulation data to find the most realistic conformational ensembles of biomolecules
+in membrane environment. This paves the way toward standardized methods to resolve the quality
+evaluated conformational ensembles of disordered biomolecules in membrane environment.
+In the NMRlipids project we aim to build a general databank of quality evaluated
+conformational ensembles of lipids and other disordered biomolecules based primarily on
+MD simulations and NMR data.
 

### 2020-12-02 Manuscript/manuscriptPGPEsuppl.tex
Update manuscript

@@
 \section{Comparison of headgroup order parameters from different force fields to experiments}
-The poor performance of headgroup order parameters in Berger model can be probably explained by
-ring like structures seen in Fig. 6 in Ref. \citenum{mukhopadhyay04}, which is a typical feature
-for Berger based lipid force fields containing explicit hydrogen atoms in the head group \cite{zhao08,henin09,dahlberg10}.
-The poor performance of glycerol backbone of Slipids simulations is systemically observed also
-for other lipids in previous studies \cite{botan15,antila19}.
-\todo{Should we comment more the relative quality of different force fields and/or make the subjective force field ranking figures?
-https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/8}
+
+The quality of PE and PG headgroup conformational ensembles in different simulations against NMR experiments
+is evaluated in figures \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}
+using C-H bond order parameters as in our previous studies for PC and PS lipids \cite{botan15,antila19}.
+Conclusions are the same for all lipids: None of the force fields correctly captures the lipid headgroup
+conformational ensembles, but CHARMM36 gives results closest to experiments.
+
+%The poor performance of headgroup order parameters in Berger model can be probably explained by
+%ring like structures seen in Fig. 6 in Ref. \citenum{mukhopadhyay04}, which is a typical feature
+%for Berger based lipid force fields containing explicit hydrogen atoms in the head group \cite{zhao08,henin09,dahlberg10}.
+%The poor performance of glycerol backbone of Slipids simulations is systemically observed also
+%for other lipids in previous studies \cite{botan15,antila19}.
+%\todo{Should we comment more the relative quality of different force fields and/or make the subjective force field ranking figures?
+%https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/8}
+
 \begin{figure}[]
@@
 
-\pagebreak
+\clearpage
 
-\subsection{PC headgroup interactions with PE and PG}
+\section{Lipid headgroups in mixtures with PC and PE or PG lipids}
 \begin{figure*}[]
@@
 
-In experiments, the PC headgroup order parameters increase
-with the addition of negatively charged PG or PS lipids, but
-are not affected by the addition of zwitterionic PE and SM lipids
-or cholesterol (Fig.~\ref{HGorderparametersPCvsPEPG}).
-This can be explained by the electrometer concept, which suggests that the 
-headgroup dipole tilts more parallel to the membrane plane upon addition of
-negative charge to the membrane~\cite{seelig87, scherer87,antila18}.
-The response of PC headgroup order parameters to PE by the tested CHARMM36 and
-Berger-OPLS force fields, although CHARMM36 slightly overestimates the changes (Fig.~\ref{HGorderparametersPCvsPEPG}).
+Headgroup order parameters of PC lipids are unchanged upon addition
+of zwitterionic lipids or cholesterol in experiments, but increase
+upon addition of negatively charged PG or PS lipids because
+headgroup dipole tilts more parallel to the membrane plane after incorporation
+of negative charges into the membrane~\cite{seelig87, scherer87,antila18}.
+The response of PC headgroup order parameters to the addition of PE or PG lipids
+from different simulations is compared with experiments in figure~\ref{HGorderparametersPCvsPEPG}.
+by the tested CHARMM36 and
+Berger-OPLS force fields, although CHARMM36 slightly overestimates the changes 
 The good performance of Berger-OPLS simulations is notable because the response of headgroup order parameters
 to cholesterol was significantly overestimated by the Berger/Höltje force field in our previous work \cite{botan15}.
-\todo{This is text by P. Fuchs, copied from the blog. \\
-Area results in nm$^2$, the error is $<=$ 0.003 nm$^2$ \\
-- pure POPC \\
-CHARMM36: 0.624 \\
-Berger : 0.649 \\
-- POPC/POPE 50:50 \\
-CHARMM36 : POPC 0.609, POPE 0.557 \\
-Berger-hacked: POPC 0.637, POPE 0.632 \\
------ \\
-One can see that CHARMM 36 predicts a drop in the area on going from pure POPC to POPC/POPE 50:50. This means that POPC pack tightly to POPE. 
-In contrast, the values for Berger are not that changed. 
-The POPE value predicted by CHARMM 36 (in the mixture POPC/POPE 50:50) is much smaller than that predicted by Berger.\\
--------\\
-The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
-which supports the more condensed PE bilayer.
-This is interesting, but not exactly the core message of the manuscript.
-Maybe we should mention this very briefly? For example, we could just report the areas per lipid (without distinguishing PC and PE)
-and mention the difference between CHARMM36 and Berger. I have opened an issue for this: https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/7
-}
+%\todo{This is text by P. Fuchs, copied from the blog. \\
+%Area results in nm$^2$, the error is $<=$ 0.003 nm$^2$ \\
+%- pure POPC \\
+%CHARMM36: 0.624 \\
+%Berger : 0.649 \\
+%- POPC/POPE 50:50 \\
+%CHARMM36 : POPC 0.609, POPE 0.557 \\
+%Berger-hacked: POPC 0.637, POPE 0.632 \\
+%----- \\
+%One can see that CHARMM 36 predicts a drop in the area on going from pure POPC to POPC/POPE 50:50. This means that POPC pack tightly to POPE. 
+%In contrast, the values for Berger are not that changed. 
+%The POPE value predicted by CHARMM 36 (in the mixture POPC/POPE 50:50) is much smaller than that predicted by Berger.\\
+%-------\\
+%The experimental acyl chain order parameters for POPE \cite{pare98} seem larger than reported for POPC \cite{ferreira13},
+%which supports the more condensed PE bilayer.
+%This is interesting, but not exactly the core message of the manuscript.
+%Maybe we should mention this very briefly? For example, we could just report the areas per lipid (without distinguishing PC and PE)
+%and mention the difference between CHARMM36 and Berger. I have opened an issue for this: https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/7
+%}
 

### 2020-12-03 scripts/DataBankINFO/POPCPOPE11berger300K.py
Add P-N vector angle results

@@ -1,24 +1,24 @@
-DOI="10.5281/zenodo.3266240"
+DOI="10.5281/zenodo.1402449"
 
 user_information = """
-POPC-POPG 7:3 310K gromos-ckp
+POPC:POPE 1:1 300K berger
 #NMRLIPIDS BEGIN
 
 @SIM
-@MAPPING=POPC,mappingPOPCgromos-ckp.txt,POPG,mappingPOPGgromos-ckp.txt
-@SYSTEM=POPCPOPG7:3_T310K
+@MAPPING=POPC,mappingPOPCberger.txt,POPE,mappingPOPEberger.txt
+@SYSTEM=POPE_T300K
 @SOFTWARE=gromacs
-@FF=gromos-ckp
+@FF=berger
 @FF_SOURCE=??
 @FF_DATE=?/?/????
-@TRJ=total_4_500.xtc
-@TPR=md_0.tpr
-@PREEQTIME=400
-@TIMELEFTOUT=0
-@UNITEDATOM=POPC,GROMOS_CKP_POPC,POPG,GROMOS_CKP_POPG
+@TRJ=md_dt100_OK_centered.xtc
+@TPR=md.tpr
+@PREEQTIME=0
+@TIMELEFTOUT=100
+@UNITEDATOM=POPC,Berger_POPC,POPE,Berger_POPE
 
 @POPC=POPC
-@POPG=LPOG
+@POPG=POPG
 @POPS=POPS
 @POPE=POPE
 

### 2020-12-03 scripts/DataBankINFO/POPCPOPG73gromos_ckp310Ktmp.py

@@ -5,7 +5,7 @@ POPC-POPG 7:3 310K gromos-ckp
 #NMRLIPIDS BEGIN
 
 @SIM
-@MAPPING=POPC,mappingPOPCgromos-ckp.txt,POPG,mappingPOPGgromos-ckp.txt
+@MAPPING=POPG,mappingPOPGgromos-ckp.txt
 @SYSTEM=POPCPOPG7:3_T310K
 @SOFTWARE=gromacs
 @FF=gromos-ckp
@@ -13,8 +13,8 @@ POPC-POPG 7:3 310K gromos-ckp
 @FF_DATE=?/?/????
 @TRJ=total_4_500.xtc
 @TPR=md_0.tpr
-@PREEQTIME=400
-@TIMELEFTOUT=0
+@PREEQTIME=0
+@TIMELEFTOUT=400
 @UNITEDATOM=POPC,GROMOS_CKP_POPC,POPG,GROMOS_CKP_POPG
 
 @POPC=POPC

### 2020-12-03 scripts/DataBankINFO/POPCberger300K.py

@@ -1,24 +1,24 @@
-DOI="10.5281/zenodo.3266240"
+DOI="10.5281/zenodo.1402417"
 
 user_information = """
-POPC-POPG 7:3 310K gromos-ckp
+POPC 300K berger
 #NMRLIPIDS BEGIN
 
 @SIM
-@MAPPING=POPC,mappingPOPCgromos-ckp.txt,POPG,mappingPOPGgromos-ckp.txt
-@SYSTEM=POPCPOPG7:3_T310K
+@MAPPING=POPC,mappingPOPCberger.txt
+@SYSTEM=POPC_T300K
 @SOFTWARE=gromacs
-@FF=gromos-ckp
+@FF=berger
 @FF_SOURCE=??
 @FF_DATE=?/?/????
-@TRJ=total_4_500.xtc
-@TPR=md_0.tpr
-@PREEQTIME=400
-@TIMELEFTOUT=0
-@UNITEDATOM=POPC,GROMOS_CKP_POPC,POPG,GROMOS_CKP_POPG
+@TRJ=md_dt100_OK_centered.xtc
+@TPR=md.tpr
+@PREEQTIME=0
+@TIMELEFTOUT=100
+@UNITEDATOM=POPC,Berger_POPC
 
 @POPC=POPC
-@POPG=LPOG
+@POPG=POPG
 @POPS=POPS
 @POPE=POPE
 

### 2020-12-03 scripts/DataBankINFO/POPCgromos_ckp310K.py

@@ -1,12 +1,12 @@
-DOI="10.5281/zenodo.3266240"
+DOI="10.5281/zenodo.3247435"
 
 user_information = """
-POPC-POPG 7:3 310K gromos-ckp
+POPC 310K gromos-ckp
 #NMRLIPIDS BEGIN
 
 @SIM
-@MAPPING=POPC,mappingPOPCgromos-ckp.txt,POPG,mappingPOPGgromos-ckp.txt
-@SYSTEM=POPCPOPG7:3_T310K
+@MAPPING=POPC,mappingPOPCgromos-ckp.txt
+@SYSTEM=POPC_T310K
 @SOFTWARE=gromacs
 @FF=gromos-ckp
 @FF_SOURCE=??
@@ -15,7 +15,7 @@ POPC-POPG 7:3 310K gromos-ckp
 @TPR=md_0.tpr
 @PREEQTIME=400
 @TIMELEFTOUT=0
-@UNITEDATOM=POPC,GROMOS_CKP_POPC,POPG,GROMOS_CKP_POPG
+@UNITEDATOM=POPC,GROMOS_CKP_POPC
 
 @POPC=POPC
 @POPG=LPOG

### 2020-12-03 scripts/calcPNvectors.py

@@ -121,7 +121,7 @@ for subdir, dirs, files in os.walk(r'../Data/Simulations/'):
                         mapping_file = './mapping_files/'+readme['MAPPING_DICT'][molname] # readme.get('MAPPING')[0][0]
                                                 
                         try:
-                            atoms = read_mapping_filePAIR(mapping_file, atom1 , atom2, molname)
+                            atoms = read_mapping_filePAIR(mapping_file, atom1 , atom2, readme[molname])
                         except:
                             print(atom1 + " and " + atom2 + " not found in the mapping file.")
                             continue

### 2020-12-03 scripts/DataBankINFO/AddData.py
file paths to def file fixed

@@ -29,7 +29,7 @@ import os
 
 #For quering webs
 import urllib.request
-from urllib.error import URLError,HTTPError
+from urllib.error import URLError,HTTPError #,ContentTooShortError
 
 # From time monitoring
 from tqdm import tqdm
@@ -618,8 +618,6 @@ for sim in sims_working_links:
            # print("Downloading of the file " + str(file_name) + " failed.")
             #break
             continue
-
-
 # ## Calculate hash downloaded files
 
 # In[15]:
@@ -986,7 +984,7 @@ for sim in sims_working_links:
     if unitedAtom:
         for key in sim['UADICTIONARY']:
         #construct order parameter definition file for CH bonds from mapping file
-            def_file = open(str(dir_wrk) + '/tmp/' + str(ID) + '/' + key + '.def', 'w')
+            def_file = open(strdir_tmp) + '/' + str(ID) + '/' + key + '.def', 'w')
 
             mapping_file = sim['MAPPING_DICT'][key]
             previous_line = ""
@@ -1018,9 +1016,9 @@ for sim in sims_working_links:
             def_file.close()
 
         #Add hydrogens to trajectory and calculate order parameters with buildH
-            ordPfile = str(dir_wrk) + '/tmp/' + str(ID) + '/' + key + '_order_parameter.dat'
+            ordPfile = open(strdir_tmp) + '/' + str(ID) + '/' + key + '_order_parameter.dat' #ONKS NÄÄ VÄÄRIN
             topfile = sim.get('GRO')
-            deffile = str(dir_wrk) + '/tmp/' + str(ID) + '/' + key + '.def'
+            deffile = open(strdir_tmp) + '/' + str(ID) + '/' + key + '.def' #ONKS TÄÄ VÄÄRIN
             lipidname = sim['UADICTIONARY'][key]
         #    print(lipidname)
             buildH_calcOP_test.main(topfile,lipidname,deffile,xtcwhole,ordPfile)

### 2020-12-03 scripts/DataBankINFO/AddData.py
file paths to def file fixed

@@ -984,7 +984,7 @@ for sim in sims_working_links:
     if unitedAtom:
         for key in sim['UADICTIONARY']:
         #construct order parameter definition file for CH bonds from mapping file
-            def_file = open(strdir_tmp) + '/' + str(ID) + '/' + key + '.def', 'w')
+            def_file = str(strdir_tmp) + '/' + str(ID) + '/' + key + '.def', 'w')
 
             mapping_file = sim['MAPPING_DICT'][key]
             previous_line = ""
@@ -1016,9 +1016,9 @@ for sim in sims_working_links:
             def_file.close()
 
         #Add hydrogens to trajectory and calculate order parameters with buildH
-            ordPfile = open(strdir_tmp) + '/' + str(ID) + '/' + key + '_order_parameter.dat' #ONKS NÄÄ VÄÄRIN
+            ordPfile = str(strdir_tmp) + '/' + str(ID) + '/' + key + '_order_parameter.dat' 
             topfile = sim.get('GRO')
-            deffile = open(strdir_tmp) + '/' + str(ID) + '/' + key + '.def' #ONKS TÄÄ VÄÄRIN
+            deffile = str(strdir_tmp) + '/' + str(ID) + '/' + key + '.def' 
             lipidname = sim['UADICTIONARY'][key]
         #    print(lipidname)
             buildH_calcOP_test.main(topfile,lipidname,deffile,xtcwhole,ordPfile)

### 2020-12-03 scripts/DataBankINFO/POPGgromos_ckp310K.py
Add P-N vector results

@@ -1,0 +1,52 @@
+DOI="10.5281/zenodo.3484548"
+
+user_information = """
+POPG 310K gromos-ckp
+#NMRLIPIDS BEGIN
+
+@SIM
+@MAPPING=POPG,mappingPOPGgromos-ckp.txt
+@SYSTEM=POPG_T310K
+@SOFTWARE=gromacs
+@FF=gromos-ckp
+@FF_SOURCE=??
+@FF_DATE=?/?/????
+@TRJ=total_4_500.xtc
+@TPR=md_1.tpr
+@PREEQTIME=400
+@TIMELEFTOUT=0
+@UNITEDATOM=POPG,GROMOS_CKP_POPG
+
+@POPC=POPC
+@POPG=LPOG
+@POPS=POPS
+@POPE=POPE
+
+@POT=K
+@SOD=NA
+@CLA=CL
+@CAL=CA
+@SOL=SOL
+
+@NPOPC=[0,0]
+@NPOPG=[0,0]
+@NPOPS=[0,0]
+@NPOPE=[0,0]
+
+@NPOT=0
+@NSOD=0
+@NCLA=0
+@NCAL=0
+@NSOL=0
+
+@TEMPERATURE=0
+@TRJLENGTH=0
+
+
+#NMRLIPIDS END
+
+"""
+
+
+# Working directory
+dir_wrk  = "/media/osollila/Data/tmp/DATABANK/"

### 2020-12-03 scripts/DataBankINFO/AddData.py
fuxed def file paths

@@ -984,7 +984,7 @@ for sim in sims_working_links:
     if unitedAtom:
         for key in sim['UADICTIONARY']:
         #construct order parameter definition file for CH bonds from mapping file
-            def_file = str(strdir_tmp) + '/' + str(ID) + '/' + key + '.def', 'w')
+            def_file = open(str(dir_tmp) + '/' + str(ID) + '/' + key + '.def', 'w')
 
             mapping_file = sim['MAPPING_DICT'][key]
             previous_line = ""
@@ -1016,9 +1016,9 @@ for sim in sims_working_links:
             def_file.close()
 
         #Add hydrogens to trajectory and calculate order parameters with buildH
-            ordPfile = str(strdir_tmp) + '/' + str(ID) + '/' + key + '_order_parameter.dat' 
+            ordPfile = str(dir_tmp) + '/' + str(ID) + '/' + key + '_order_parameter.dat' 
             topfile = sim.get('GRO')
-            deffile = str(strdir_tmp) + '/' + str(ID) + '/' + key + '.def' 
+            deffile = str(dir_tmp) + '/' + str(ID) + '/' + key + '.def' 
             lipidname = sim['UADICTIONARY'][key]
         #    print(lipidname)
             buildH_calcOP_test.main(topfile,lipidname,deffile,xtcwhole,ordPfile)

### 2020-12-04 scripts/DataBankINFO/POPCPOPG7:3slipidsINFOpeon.py
Add P-N orientation data

@@ -20,11 +20,11 @@ POPC:POPG 7:3 slipids from A. Peon
 @POPS=POPS
 @POPE=POPE
 
-@POT=M_K_M
-@SOD=M_NA_M
-@CLA=M_CL_M
-@CAL=M_CA_M
-@SOL=M_SOL_M
+@POT=K
+@SOD=NA
+@CLA=CL
+@CAL=CA
+@SOL=SOL
 
 @NPOPC=[0,0]
 @NPOPG=[0,0]

### 2020-12-04 scripts/DataBankINFO/POPGslipidsT310KINFO.py

@@ -22,11 +22,11 @@ POPG Slipids T=310K
 @POPS=POPS
 @POPE=POPE
 
-@POT=M_K_M
-@SOD=M_NA_M
-@CLA=M_CL_M
-@CAL=M_CA_M
-@SOL=M_SOL_M
+@POT=K
+@SOD=NA
+@CLA=CL
+@CAL=CA
+@SOL=SOL
 
 @NPOPC=[0,0]
 @NPOPG=[0,0]

### 2020-12-07 Manuscript/manuscriptPGPEsuppl.tex
Update manuscript

@@
 
-\section{Lipid headgroups in mixtures with PC and PE or PG lipids}
+\section{PC headgroup in mixtures with PE or PG lipids}
 \begin{figure*}[]
@@
   }
-  \todo{P-N vector angles from Berger-OPLS and GROMOS-CKP simulations are yet to be analyzed.}
 \end{figure*}
@@
 from different simulations is compared with experiments in figure~\ref{HGorderparametersPCvsPEPG}.
-by the tested CHARMM36 and
-Berger-OPLS force fields, although CHARMM36 slightly overestimates the changes 
-The good performance of Berger-OPLS simulations is notable because the response of headgroup order parameters
-to cholesterol was significantly overestimated by the Berger/Höltje force field in our previous work \cite{botan15}.
+%
+%
 %\todo{This is text by P. Fuchs, copied from the blog. \\
@@
 %}
+%
+%
+None of the simulations reproduce neither the experimentally observed increase in PC headgroup order parameters
+with increasing amount of PG nor the related tilting of the headgroup more parallel with the membrane.
+Similar observations in our previous work for PS lipids were explained by the overestimated counterion
+binding affinity that neturalizes the effect of added negative charge \cite{antila19}.
+%In all force fields except Slipids, the order parameters of different
+%hydrogens attached to the $\alpha$-carbon are responsing differently when mixed with PE or PG lipids
+%\todo{Maybe we should figure out what is the reason for this? \\
+%  Maybe we should analyze the P-N vector angle from different simulations? \\
+%  https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/10}.
+All simulations except Berger-OPLS predict tilting of P-N headgroup outwards from the membrane and
+decrease of PC headgroup order parameters upon addition of PE lipids.
+These results are not in line with experiments where the PC headgroup order parameters are not affected by zwitterionic lipids \cite{scherer87}.
+The good performance of Berger-OPLS simulations in here is surprising because
+headgroup conformational enemble is not very close to experiments in this model and
+the response of headgroup order parameters
+to cholesterol was significantly overestimated by the Berger/H{\"o}ltje force field in our previous work \cite{botan15}.
 
-None of the force fields fully reproduces the PC headgroup order parameter response to the increasing amount of
-PG, which may be related to the counterion binding affinity (see also the next section) \cite{antila19}.
-In all force fields except Slipids, the order parameters of different
-hydrogens attached to the $\alpha$-carbon are responsing differently when mixed with PE or PG lipids
-\todo{Maybe we should figure out what is the reason for this? \\
-  Maybe we should analyze the P-N vector angle from different simulations? \\
-  https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/10}.
-
-For $\beta$-carbon order parameter in PG headgroup, experiments report
-mild increase \cite{macdonald87} or no change~\cite{borle85} upon addition 
-of PC lipids (Fig. \ref{HGorderparametersPGvsPCchange}). 
-Simulations with all the tested force fields give only very small changes also for
-the $\alpha$-carbon order parameter (Figs. \ref{HGorderparametersPGvsPC} and \ref{HGorderparametersPGvsPCchange}). 
-Therefore, the simulations are generally in line with experiments, suggesting that the
-interactions with PC do not essentially effect the PG headgroup structure.
-This suggests that the interactions between PG and PC headgroups are captured better in
-simulations than for PS headgroup, where
-all the force fieds significantly overestimated the structural response of PS headgroup
-to the interactions with PC lipids \cite{antila18}.
+\clearpage
+\section{PG headgroup in mixtures with PC lipids}
+Changes in other than PC lipid headgroup with changing membrane composition are less
+extensively characterized in the literature.
+The $\beta$-carbon order parameter in PG headgroup increases
+mildly \cite{macdonald87} or is unchanged~\cite{borle85} upon increasing amount
+of PC lipids (Fig. \ref{HGorderparametersPGvsPCchange}), but experimental data from
+$\alpha$-carbon is not available. Also the tested force fields predict very small changes
+for the $\beta$-carbon order parameter, while the P-N vector tilt and its response to the
+increased amount of PC varies significantly between force fields in figure \ref{HGorderparametersPGvsPCchange}. 
+Therefore, more experimental data and more accurate force fields are still required
+to resolve the PG conformational ensembles in mixtures with other lipids.
+%the simulations are generally in line with experiments, suggesting that the
+%interactions with PC do not essentially effect the PG headgroup structure.
+%This suggests that the interactions between PG and PC headgroups are captured better in
+%simulations than for PS headgroup, where
+%all the force fieds significantly overestimated the structural response of PS headgroup
+%to the interactions with PC lipids \cite{antila18}.
 \begin{figure}[]
@@
   }
-  \todo{P-N angles from Slipids and GROMOS-CKP yet to be calculated.}
 \end{figure}
 
+%
+%\clearpage
+%\subsection{Sodium binding to PE and PG lipid bilayers}
+%Sodium binding affinity to PE lipids has not been measured experimentally,
+%but large differences to PC would be surprising. 
+%In simulations, the sodium binding affinity to POPE depends on the used force field (Fig. \ref{NAdensPE}),
+%but lesser extend than reported previously for PC~\cite{catte16}.
+%\todo{This will be finished once we have all the simulation details and Lipid17 simulations with correct dihedrals from issue https://github.com/NM%RLipids/NMRlipidsIVPEandPG/issues/12},
+%Because some simulation and ion parameters are not identical with the previous work~\cite{catte16},
+%we compare POPE results to the POPC simulations ran with identical parameters (Fig. \ref{NAdensPC}).
+%In Lipid17 with the strongest sodium binding affinity to POPE,
+%the binding affinity is approximately similar to POPC.
+%Slipids and CHARMM36 exhibit slightly, and
+%GROMOS-CKP subtantially weaker binding to POPE than to POPC.
+%%The Lipid17 exhibits stronger sodium binding affinity
+%%to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
+%%Experimental data for sodium binding to PE lipids is not available, while
+%%simulations in agreement with electrometer data from NMR suggests that
+%%sodium binding to POPC is weaker than in any of the simulations
+%%here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}.
+%Assuming that the binding
+%to POPE would be similar than to POPC, the sodium binding affinity to POPE
+%is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
+%but substantially overestimated in Lipid17 simulation.
+%%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
+%%previous Amber lipid version, Lipid14.
+%%\todo{The potential differences in the used ion models should be checked}.
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
+%  \caption{\label{NAdensPE}
+%    Sodium (solid line) and choride ion density profiles along membrane normal
+%    from different simulations with PE lipids.
+%  }
+%\end{figure}
+
+%Simulations with PG lipids give similar dependence on force field as observed in POPE simulations:
+%Lipid 17 simulations with Dang ion parameters exhibits stronger counter-ion binding affinity to pure POPG bilayer
+%than CHARMM36, Slipids, and GROMOS-CKP simulations, which are roughly similar (Fig. \ref{CIdensPG}).
+%Lipid17 also exhibits less increase in POPC headgroup
+%%$\beta$ carbon
+%order parameters upon addition of POPG than other simulations (Fig. \ref{HGorderparametersPCvsPEPG}),
+%and lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$).
+%In our previous study about PS lipids \cite{antila18}, such behaviour was related to 
+%%the results suggest that
+%the overestimated counterions binding and shielding the electrostatic repulsion between PG headgroups
+%in bilayers.% simulated with Lipid17 parameters.
+%%the underestimated increase of PC
+%%headgroup order parameters upon addition of negaticely charged lipids
+%%was related
+%%to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
+%%Similar correlation is observed here in %other than GROMOS-CKP simulation:
+%%%Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
+%%
+%Even though the area per lipid in CHARMM36, Slipids, and GROMOS-CKP simulations is in good agreement
+%with experiments (Fig. \ref{CIdensPG}), the experimental increase in 
+%%In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of
+%POPC headgroup order parameters upon addition of POPG are not fully reproduced (Fig. \ref{HGorderparametersPCvsPEPG}).
+%%more than in CHARMM36 and Slipids simulations even though the binding affinities in these
+%%simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
+%%are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
+%%and other decreases \todo{Can we figure out the reason for this?}.
+%Therefore we conclude that the counter-ion binding affinity is overestimated in Lipid17 simulations,
+%% with the strongest affinity among the tested models,
+%while the other simulations are more realistic, but slight overbinding cannot be excluded.
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
+%  \caption{\label{CIdensPG}
+%    Counterion densities and area per lipids from simulations with PG lipids.
+%    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
+%  }
+%\end{figure}
+
+%\clearpage
+%
+%\section{Changes of PG headroup order parameters upon addition of PC}
+
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPC.eps}
+%  \caption{\label{HGorderparametersPGvsPC}
+%    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
+%    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
+%  }
+%\end{figure}
+
+%\clearpage
+
+%\section{Sodium binding to POPC simulations}
+
+%\begin{figure}[]
+%  \centering
+%  \includegraphics[width=8.0cm]{../Figs/NAdensPC.eps}
+%  \caption{\label{NAdensPC}
+%    Sodium (solid line) and choride ion density profiles along membrane normal
+%    from different simulations with PC lipids.
+%  }
+%  \todo{Discussion about differences to the NMRlipids II to be discussed once we have the details on ions models.}
+%\end{figure}
+
 \clearpage
-\subsection{Sodium binding to PE and PG lipid bilayers}
-Sodium binding affinity to PE lipids has not been measured experimentally,
-but large differences to PC would be surprising. 
-In simulations, the sodium binding affinity to POPE depends on the used force field (Fig. \ref{NAdensPE}),
-but lesser extend than reported previously for PC~\cite{catte16}.
-\todo{This will be finished once we have all the simulation details and Lipid17 simulations with correct dihedrals from issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12},
-Because some simulation and ion parameters are not identical with the previous work~\cite{catte16},
-we compare POPE results to the POPC simulations ran with identical parameters (Fig. \ref{NAdensPC}).
-In Lipid17 with the strongest sodium binding affinity to POPE,
-the binding affinity is approximately similar to POPC.
-Slipids and CHARMM36 exhibit slightly, and
-GROMOS-CKP subtantially weaker binding to POPE than to POPC.
-%The Lipid17 exhibits stronger sodium binding affinity
-%to POPE bilayers than Slipids, CHARMM36 and GROMOS-CKP (Fig. \ref{NAdensPE}).
-%Experimental data for sodium binding to PE lipids is not available, while
-%simulations in agreement with electrometer data from NMR suggests that
-%sodium binding to POPC is weaker than in any of the simulations
-%here (Fig. \ref{NAdensPC})~\cite{catte16,melcr18}.
-Assuming that the binding
-to POPE would be similar than to POPC, the sodium binding affinity to POPE
-is potentially realistic in CHARMM36, Slipids, and GROMOS-CKP simulations here,
-but substantially overestimated in Lipid17 simulation.
-%On the other hand, the sodium binding affinity is stronger to POPE in Lipid17 simulations than to POPC in simulations with
-%previous Amber lipid version, Lipid14.
-%\todo{The potential differences in the used ion models should be checked}.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/NAdensPE.eps}
-  \caption{\label{NAdensPE}
-    Sodium (solid line) and choride ion density profiles along membrane normal
-    from different simulations with PE lipids.
-  }
-\end{figure}
-
-Simulations with PG lipids give similar dependence on force field as observed in POPE simulations:
-Lipid 17 simulations with Dang ion parameters exhibits stronger counter-ion binding affinity to pure POPG bilayer
-than CHARMM36, Slipids, and GROMOS-CKP simulations, which are roughly similar (Fig. \ref{CIdensPG}).
-Lipid17 also exhibits less increase in POPC headgroup
-%$\beta$ carbon
-order parameters upon addition of POPG than other simulations (Fig. \ref{HGorderparametersPCvsPEPG}),
-and lower area per molecule (59.5 {\AA}$^2$) than in experiments (66.1 {\AA}$^2$).
-In our previous study about PS lipids \cite{antila18}, such behaviour was related to 
-%the results suggest that
-the overestimated counterions binding and shielding the electrostatic repulsion between PG headgroups
-in bilayers.% simulated with Lipid17 parameters.
-%the underestimated increase of PC
-%headgroup order parameters upon addition of negaticely charged lipids
-%was related
-%to the overestimated counterion binding affinity, which would overcompensates the effect of negative charge.
-%Similar correlation is observed here in %other than GROMOS-CKP simulation:
-%%Lipid 17 simulations, which exhibits stronger counter-ion binding affinity to pure POPG bilayer (Fig. \ref{CIdensPG})
-%
-Even though the area per lipid in CHARMM36, Slipids, and GROMOS-CKP simulations is in good agreement
-with experiments (Fig. \ref{CIdensPG}), the experimental increase in 
-%In GROMOS-CKP simulations, however, the $\beta$-carbon order parameter of
-POPC headgroup order parameters upon addition of POPG are not fully reproduced (Fig. \ref{HGorderparametersPCvsPEPG}).
-%more than in CHARMM36 and Slipids simulations even though the binding affinities in these
-%simulations are similar. On the other hand, the responses of different hydrogens attached in $\alpha$-carbons
-%are different in all simulations except in Slipids, most sever in Lipid17 and GROMOS-CKP where other in increases
-%and other decreases \todo{Can we figure out the reason for this?}.
-Therefore we conclude that the counter-ion binding affinity is overestimated in Lipid17 simulations,
-% with the strongest affinity among the tested models,
-while the other simulations are more realistic, but slight overbinding cannot be excluded.
-\begin{figure}[]
-  \centering
-  \includegraphics[width=9.0cm]{../Figs/CIdensPG.eps}
-  \caption{\label{CIdensPG}
-    Counterion densities and area per lipids from simulations with PG lipids.
-    Experimental area for POPG at 303 K is 66.1 {\AA}$^2$ and 67 {\AA}$^2$ for DPPC at 323 K \cite{pan12b}.
-  }
-\end{figure}
-
-\section{Calcium density distributions}
+%\section{Calcium density distributions}
+\section{Calcium binding to POPC:POPG mixtures}
 
@@
     Calcium ion density profiles along membrane normal from simulations of POPC:POPG (1:1) mixtures with different force fields.
+    The changes in the order parameters upon addition of CaCl$_2$ are compared with experiments in figure \ref{changesWITHCaClPG} in the main text.
   }
@@
 
-\clearpage
-
-\section{Changes of PG headroup order parameters upon addition of PC}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/HGorderparametersPGvsPC.eps}
-  \caption{\label{HGorderparametersPGvsPC}
-    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
-    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
-  }
-\end{figure}
-
-\clearpage
-
-\section{Sodium binding to POPC simulations}
-
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{../Figs/NAdensPC.eps}
-  \caption{\label{NAdensPC}
-    Sodium (solid line) and choride ion density profiles along membrane normal
-    from different simulations with PC lipids.
-  }
-  \todo{Discussion about differences to the NMRlipids II to be discussed once we have the details on ions models.}
-\end{figure}
-
-\clearpage
-
-\section{Calcium binding to POPC:POPG (4:1) mixtures}
 \begin{figure*}[]
@@
   }
-  \todo{Lipid17ecc data to be analyzed and added.}
 \end{figure*}

### 2020-12-08 Manuscript/manuscriptPGPEsuppl.tex
Update SI

@@
 
+\clearpage
 \section{Changes in headgroup conformations upon addition of CaCl$_2$}
@@
 
+\clearpage
 \section{Simulated systems}
@@
   \end{minipage}
+  \todo{Simulations with added NaCl are not currently used here, maybe should be removed from the table?} \\
   \todo{Citation for CHARMM36 PE?} \\
@@
   \end{minipage}
+    \todo{Simulations with added NaCl are not currently used here, maybe should be removed from the table?} \\
     \todo{Citations and ion model for CHARMM36?} \\
-  \todo{Lipid17 simulation with correct dihedral potentials still coming.} \\
+  \todo{Lipid17 simulation with ions with correct dihedral potentials still coming?} \\
   \todo{Citation and ion model for GROMOS-CKP?}
@@
     \todo{Citation and ion model for GROMOS-CKP?} \\
-  \todo{Citation and description for ''Berger'' model?} \\
-  \todo{Lipid17 POPC and POPC:POPG mixtures (https://doi.org/10.5281/zenodo.3241242 and https://doi.org/10.5281/zenodo.3237656)
-    should be added after simulated with corrected dihedrals.}\\
   \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
@@
       \hline
+      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG8020} \\
+      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG8020100mMCaCl} \\
+      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 1.0& 350:88 & 24840 & 475 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG80201000mMCaCl} \\
       POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \cite{Lipid17eccPOPCPOPG5050} \\
@@
   \end{minipage}
-    \todo{Citation and ion model for GROMOS-CKP?} \\
   \todo{Citation and description for ''Berger'' model?} \\
-  \todo{Lipid17 POPC and POPC:POPG mixtures (https://doi.org/10.5281/zenodo.3241242 and https://doi.org/10.5281/zenodo.3237656)
-    should be added after simulated with corrected dihedrals.}\\
   \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}

### 2020-12-08 Manuscript/manuscriptPGPEsuppl.tex
Update SI

@@
 
-\noindent {\it POPC:POPG mixture with additional calcium} \todo{Simulation details by J. Madsen.}
+\noindent {\it POPC:POPG mixture with additional calcium} \todo{Simulation details by A. Kiirikki.}
 
-\noindent {\it POPC:POPG mixture with additional NaCl} \todo{Simulation details by A. Peon.}
+\noindent {\it POPC and POPC:POPG (7:3) mixture} \todo{Simulation details by A. Peon.}
 
@@
   I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
+
+\noindent {\it POPC:POPG mixture with additional CaCl} \todo{Simulation details by M. Javanainen.}
 
@@
 
+\noindent {\it POPG} \todo{Simulation details by A. Peon.}
+
+\noindent {\it POPC:POPG mixture} \todo{Simulation details by A. Peon.}
+
+\subsection{OPLS-MacRog}
+
+\noindent {\it POPE} \todo{Simulation details by M. Javanainen and P. Fuchs.}
+
+\noindent {\it POPC:POPE mixtures} \todo{Simulation details by P. Fuchs.}
+
 \subsection{Lipid17}
 
-\subsection{ECC-LIPID POPG}
+\noindent {\it POPE} \todo{Simulation details by A. Peon.}
+
+\noindent {\it POPG} \todo{Simulation details by A. Peon.}
+
+\noindent {\it POPC:POPG} \todo{Simulation details by S. Virtanen or O. H. S. Ollila.}
+
+\subsection{Lipid17ecc}
+
+\todo{This is to be finished and POPC:POPG mixtures to be described}
 In ECC-lipid models, electronic continuum correction (ECC) is applied to 

### 2020-12-08 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 
+Lipid composition of biomembranes vary greatly between different
+organelles and organism, and different lipid types interact with
+proteins in many different ways regulating their function \cite{??}.
+Lipids may directly interact with proteins by specific binding sites
+or indirectly affect protein functions by altering membrane properties
+or charge \cite{??}. Specific interactions with certain lipid headgroups
+are known to be essential for the function of several proteins \cite{??}.
+However, it is not clear if the specific binding is driven by
+the differences in acessible conformational states between lipid types or
+by specific intermolecular interactions between lipids and proteins.
+
+Differences between preferred conformations of different lipid types
+are observed in lipid crystal structures \cite{buldt81,pascher92},
+but lipids are not in solid state in biomembranes. 
+On the other hand, NMR experiments enable detection of lipid headgroups
+in physiological liquid states also in cells \cite{gally81,scherer87,seelig90}.
+These studies suggest that the glycerol backbone structures are largely similar irrespectively of the headroup \cite{gally81}, 
+glycerol backbone and headgroup structure and behaviour are similar in model membranes and in bacteria \cite{gally81,scherer87,seelig90},
+and the headgroup structures are similar in PC, PE and PG lipids, while headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. 
+%Extensive discussion about structural details of PE, PG or PS headgroups do not exists (as far as I know), 
+%In contrast to PC lipids (see \cite{botan15} and references therein).
+%Some attempts to resolve conformational ensembles from NMR for PC and PE lipids have been made,
+%but lesser extend for PG or PS lipids \cite{seelig77c,davis83,Semchyschyn04}.
+Attempts to resolve the conformational ensembles of lipid headgroup rely mainly on
+absolute values of C-H bond order parameters measured using $^2$H NMR~\cite{seelig77c,davis83,Semchyschyn04}.
+However, appropriate models to interpret the experimental data have not been available and
+also the signs of the order parameters are required to fully capture the details in conformational ensembles  \cite{botan15,ollila16}.
+%determine the structural 
+%Classical molecular dynamics simulations could potentially give such ensembles and therefore enable
+%the detailed studies of lipid headgroup behaviour in complex biomolecular systems, but current
+%force fields are not accurate enough to reproduce the correct conformational ensembles for PC and PS headgroups \cite{botan15,antila19}.
+
 PE and PG lipids are most common lipids in bacteria \cite{sohlenkamp16}.
@@
 
-Structural details of lipid headgroups are mainly studied using NMR experiments, which
-suggest that the glycerol backbone structures are largely similar irrespectively of the headroup \cite{gally81}, 
-glycerol backbone and headgroup structure and behaviour are similar in model membranes and in bacteria \cite{gally81,scherer87,seelig90},
-and the headgroup structures are similar in PC, PE and PG lipids, while headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. 
-%Extensive discussion about structural details of PE, PG or PS headgroups do not exists (as far as I know), 
-%In contrast to PC lipids (see \cite{botan15} and references therein).
-Some attempts to resolve conformational ensembles from NMR for PC and PE lipids have been made,
-but lesser extend for PG or PS lipids \cite{seelig77c,davis83,Semchyschyn04}.
-Classical molecular dynamics simulations could potentially give such ensembles and therefore enable
-the detailed studies of lipid headgroup behaviour in complex biomolecular systems, but current
-force fields are not accurate enough to reproduce the correct conformational ensembles for PC and PS headgroups \cite{botan15,antila19}.
-Several MD simulations of PE and PG lipids have been published especially in the context of modeling
-inner membrane of Gram-negative
-bacteria \cite{devries04,murzyn05,pedersen06,zhao07,gurtovenko08,zhao08,henin09,kukol09,tsai12,dickson12,venable13,dickson14,berglund15}
-\todo{There may be some relevant publication missing from here},
-but evaluation of glycerol backbone and headgroup structures against experiments is rare \cite{henin09}.
+%Several MD simulations of PE and PG lipids have been published especially in the context of modeling
+%inner membrane of Gram-negative
+%bacteria \cite{devries04,murzyn05,pedersen06,zhao07,gurtovenko08,zhao08,henin09,kukol09,tsai12,dickson12,venable13,dickson14,berglund15}
+%\todo{There may be some relevant publication missing from here},
+%but evaluation of glycerol backbone and headgroup structures against experiments is rare \cite{henin09}.
 

### 2020-12-10 Manuscript/manuscriptPGPE.tex
Update intro

@@
 % insert abstract here
-Abstract
+The force field giving the best description for glycerol backbone and headgroup structures
+of PC, PS, PG and PE headgroups (CHARMM36) reproduces the essential differences in order parameters
+between these headgroups, and therefore enables the analysis of structural differences between the
+headgroups.
+
 \end{abstract}
@@
 
-Lipid composition of biomembranes vary greatly between different
-organelles and organism, and different lipid types interact with
-proteins in many different ways regulating their function \cite{??}.
-Lipids may directly interact with proteins by specific binding sites
-or indirectly affect protein functions by altering membrane properties
-or charge \cite{??}. Specific interactions with certain lipid headgroups
-are known to be essential for the function of several proteins \cite{??}.
-However, it is not clear if the specific binding is driven by
-the differences in acessible conformational states between lipid types or
-by specific intermolecular interactions between lipids and proteins.
+Chemical compositions of hydrophilic lipid headgroups vary between different
+organelles and organisms, and different lipid types
+regulate protein functions in many different ways \cite{lee03,vanmeer08}.
+Lipids can directly bind to proteins or indirectly affect protein
+functions by altering membrane properties or charge \cite{lee03,lemmon08}.
+While the specific interactions with certain lipid headgroups
+are known to be essential for the function of several proteins \cite{lee11,lemmon08},
+it is not clear if the specificity is driven by the differences in accessible
+conformational states between lipid types or
+by specific intermolecular lipid-protein interactions.
 
-Differences between preferred conformations of different lipid types
-are observed in lipid crystal structures \cite{buldt81,pascher92},
-but lipids are not in solid state in biomembranes. 
-On the other hand, NMR experiments enable detection of lipid headgroups
-in physiological liquid states also in cells \cite{gally81,scherer87,seelig90}.
-These studies suggest that the glycerol backbone structures are largely similar irrespectively of the headroup \cite{gally81}, 
-glycerol backbone and headgroup structure and behaviour are similar in model membranes and in bacteria \cite{gally81,scherer87,seelig90},
+Lipid conformational ensembles in physiologically relevant lamellar liquid phase
+are typically derived from NMR experiments, particularly from 
+%Attempts to resolve the conformational ensembles of lipid headgroup rely mainly on
+C-H bond order parameters measured using $^2$H NMR~\cite{seelig77c,davis83,Semchyschyn04}
+which can be detected even from cells \cite{gally81,scherer87,seelig90}.
+%but lipids are not in solid state in biomembranes. 
+%On the other hand, NMR experiments enable detection of lipid headgroups
+%in physiological liquid states also in cells \cite{gally81,scherer87,seelig90}.
+These studies suggest that the glycerol backbone structures are largely similar irrespectively of the headgroup \cite{gally81}, 
 and the headgroup structures are similar in PC, PE and PG lipids, while headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. 
@@
 %but lesser extend for PG or PS lipids \cite{seelig77c,davis83,Semchyschyn04}.
-Attempts to resolve the conformational ensembles of lipid headgroup rely mainly on
-absolute values of C-H bond order parameters measured using $^2$H NMR~\cite{seelig77c,davis83,Semchyschyn04}.
-However, appropriate models to interpret the experimental data have not been available and
-also the signs of the order parameters are required to fully capture the details in conformational ensembles  \cite{botan15,ollila16}.
+However, these results are based only on the absolute values while 
+%we have recently demonstrated that also
+the necessity of order parameters signs in capturing the lipid conformational ensembles has been resently demonstrated \cite{botan15,ollila16,ferreira16}.
+Furthermore, the detailed understanding of lipid conformational ensembles
+is limited by the lack of universal models to interpret the experiments \cite{pezeshkian18,akutsu20}.
+%do not exist and
+%the proposed structures are often based on lipid crystal structures \cite{pezeshkian18,akutsu20}.
+%which were solved many decades ago \cite{buldt81,pascher92}.
 %determine the structural 
@@
 
-PE and PG lipids are most common lipids in bacteria \cite{sohlenkamp16}.
-Zwitterionic PE is the second most abundant glycerophospholipid in eukaryotic cells
-and has been related to the diseases \cite{vance15,calzada16,patel17}.
-Anionic PG lipids are less abundant, but is also proposed to be fundamental for terrestrial life \cite{furse17}.
-PE and PG affect membrane protein functionality \cite{hariharan18} and bind to various proteins \cite{yeagle14}.
-PE headgroup is also prone for negative membrane curvature and causes membrane fusion \cite{Chernomordik08,calzada16}.
-Therefore, the PE and PG headgroup structures play probably essential roles in 
-many biological processes.
+Stuctures of different lipid types in protein bound states can be extracted from the protein data bank (PDB) \cite{berman00},
+but their relation to the conformational ensembles in liquid lamellar state remains unclear \cite{marsh13b}.
+In addition to the changes in lipid conformational ensembles upon binding to proteins,
+%PC headgroup conformations do not change upon addition of zwitterionic lipids or cholesterol \cite{scherer87,ferreira13} into a membrane,
+%but are sensitive to the presence of charged molecules in membranes \cite{seelig87}.
+also the experimentally measured response of lipid headgroup to membrane bound charges remains poorly understood 
+%The response to charges is explained by the change in the orientation of lipid headgroup dipole vector due to electrostatic interactions,
+%but the detailed changes are not yet understood
+due to the lack of suitable models to interpret the lipid conformational ensembles in liquid lamellar state \cite{Semchyschyn04}.
+
+Here, we use natural abundance $^{13}$C NMR experiments and MD simulations from the NMRlipids open collaboration
+to resolve the differences in conformational ensembles between PC, PE, PG and PS lipid headgroups.
+%open collaboration and order parameters of glycerol backbone and headgroup
+%to evaluate the accuracy of PE and PG heagroup structures, and the cation binding affinity
+Zwitterionic PC and PE are the most common lipids in eukaryotes and bacteria, respectively \cite{vanmeer08,sohlenkamp16}.
+PE is also the second most abundant glycerophospholipid in eukaryotic cells
+and has been related to various diseases \cite{vance15,calzada16,patel17}.
+PS and PG are the most common negatively charged lipids in eukaryotes and bacteria, respectively,
+and affect membrane protein functionality and signaling \cite{lemmon08,leventis10,sohlenkamp16,hariharan18}.
+%PE and PG lipids are most common lipids in bacteria \cite{sohlenkamp16}.
+%Anionic PG lipids are less abundant, but is also proposed to be fundamental for terrestrial life \cite{furse17}.
+%PE and PG affect membrane protein functionality \cite{hariharan18} and
+All the studied lipids specifically bind to various proteins \cite{yeagle14}.
+%PE headgroup is also prone for negative membrane curvature and causes membrane fusion \cite{Chernomordik08,calzada16}.
+%Therefore, the PE and PG headgroup structures play probably essential roles in 
+%many biological processes.
+Furthermore, we use our results to elucidate lipid-protein interactions and the effect of charges on lipid conformations.
+
+The lipid conformational ensembles in liquid lamellar state paves the way toward understanding the
+specific binding of different lipid types to membrane proteins and how they regulate the protein function.
+Because glycerol backbone and headgroup structures of PC lipids are similar in model membranes and in bacteria \cite{gally81,scherer87,seelig90},
+the results from model systems could be used to understand the biological role of lipid headgroup conformational ensembles in different lipid types.
 
@@
 
-Derivation of different lipid headgroup conformational ensembles in liquid state has been
-inconclusive due to lack of suitable experimental data and tools to interpret the conformational ensembles.
+%Derivation of different lipid headgroup conformational ensembles in liquid state has been
+%inconclusive due to lack of suitable experimental data and tools to interpret the conformational ensembles.
 
-Besides the structure, also ion binding may regulate biophysical activity of especially negatively
-charged lipid headgroups \cite{seelig90}. Monovalent cation (except Lithium) binding to zwitterionic PC and anionic PS headgroups is
-very weak, while multivalent ion binding is stronger but still weak \cite{cevc90,tocanne90,roux90,catte16,antila19}. The ion binding affinity
-data for PE is more scarce \cite{marra85}, but large differences to PC would be surprising.
-Negatively charged lipids are suggested to bear same cation binding constants than zwitterionic lipids, but
-the amount of bound ions to negatively charged membranes would still be larger because
-the concentration of cations in the vicinity of membranes would be higher  \cite{seelig90}.
+%Besides the structure, also ion binding may regulate biophysical activity of especially negatively
+%charged lipid headgroups \cite{seelig90}. Monovalent cation (except Lithium) binding to zwitterionic PC and anionic PS headgroups is
+%very weak, while multivalent ion binding is stronger but still weak \cite{cevc90,tocanne90,roux90,catte16,antila19}. The ion binding affinity
+%data for PE is more scarce \cite{marra85}, but large differences to PC would be surprising.
+%Negatively charged lipids are suggested to bear same cation binding constants than zwitterionic lipids, but
+%the amount of bound ions to negatively charged membranes would still be larger because
+%the concentration of cations in the vicinity of membranes would be higher  \cite{seelig90}.
+%
 %is stronger
@@
 %\end{displayquote}
-On the other hand, anionic PS lipids are proposed chelate with calcium ions \cite{hauser85,feigenson86,roux91}.
-In simulations, the cation binding affinity to PC and PS membranes is typically overestimated \cite{catte16,antila19},
-which can be improved by applying the ECC to the partial charges of the force fields \cite{melcr18,melcr19}.
-
-Here, we use open collaboration and order parameters of glycerol backbone and headgroup
-to evaluate the accuracy of PE and PG heagroup structures, and the cation binding affinity
-
-to anionic membranes containing PG lipids in the current MD simulation force fields.
-The force field giving the best description for glycerol backbone and headgroup structures
-of PC, PS, PG and PE headgroups (CHARMM36) reproduces the essential differences in order parameters
-between these headgroups, and therefore enables the analysis of structural differences between the
-headgroups.
+%
+%On the other hand, anionic PS lipids are proposed chelate with calcium ions \cite{hauser85,feigenson86,roux91}.
+%In simulations, the cation binding affinity to PC and PS membranes is typically overestimated \cite{catte16,antila19},
+%which can be improved by applying the ECC to the partial charges of the force fields \cite{melcr18,melcr19}.
 

### 2020-12-10 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 \subsection{Experimental C--H bond order parameters}
-The headgroup and glycerol backbone C--H bond order parameter magnitudes and signs of POPE and POPG
-were determined by measuring the chemical-shift resolved dipolar splittings
-with a R-type Proton Detected Local Field (R-PDLF) experiment~\cite{dvinskikh04} and
-%The corresponding order parameter signs were measured with a
-S-DROSS experiments~\cite{gross97} using natural abundance $^{13}$C solid state NMR spectroscopy
+The headgroup and glycerol backbone C--H bond order parameters of POPE and POPG
+were measured using natural abundance $^{13}$C solid state NMR spectroscopy
 as described previously \cite{ferreira13,ferreira16}.
+The magnitudes of order parameters were determined from the chemical-shift resolved dipolar splittings
+using a R-type Proton Detected Local Field (R-PDLF) experiment~\cite{dvinskikh04} and
+the signs from S-DROSS experiments~\cite{gross97} combined with SIMPSON simulations \cite{bak00}.
 %The experiments were done in a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz.
@@
 %The following experimental setups were used.
-POPE and POPG powder were purchased from Avanti polar lipids.
-The NMR experiments were identical to our previous work~\cite{antila19}.
-\todo{Is this enough and correct, or should we repeat some methods from the NMRlipidsIVps paper?}
+The NMR experiments were identical as in our previous work~\cite{antila19}.
+POPE and POPG powder were purchased from Avanti polar lipids \todo{Which isomer of POPG we have in the experiments?}.
+%\todo{Is this enough and correct, or should we repeat some methods from the NMRlipidsIVps paper?}
 The POPE experiments were recorded at 310~K and POPG experiments at 298~K, where the bilayers are in the liquid disordered phase \cite{marsh13}.
 
-Absolute values of the headgroup and glycerol backbone order parameters from PE and PG lipids are measured
-previously using $^2$H NMR~\cite{seelig76,gally81,wohlgemuth80,borle85}. Because also the order parameter
-signs bear essential information about the lipid structures \cite{botan15,ollila16}, we measured the
-magnitudes and signs of POPE and POPG C--H bond headgroup and glycerol backbone order parameter in liquid phase
-using the 2D-RPDLF and S-DROSS experiments, as described previously \cite{ferreira13,ferreira16,antila19}.
-For POPE, the glycerol backbone and $\alpha$-carbon peaks in INEPT spectra were assigned based on
-previously measured POPC spectra~\cite{ferreira13} and
-the $\beta$-carbon peak was assigned based on $^{13}$C chemical shift table for amines available
-at \url{https://www.chem.wisc.edu/areas/reich/nmr/c13-data/cdata.htm} (Fig. \ref{POPEspectra}).
-For POPG, the glycerol backbone peaks in INEPT spectra were assigned based on
-previously measured POPC spectra~\cite{ferreira13}, while $\alpha$ and  $\gamma$-carbon peaks
-\todo{How were these assigned?} (Fig. \ref{POPGspectra}). The numerical value of the $\beta$-carbon
-order parameter could not be determined, because its peak overlapped with the g$_2$ peak from glycerol backbone in POPG.
-However, the order parameter of $\beta$-carbon is expected to be clearly smaller than for g$_2$
-based on previous $^2$H NMR measurements \cite{wohlgemuth80,gally81,borle85}.
-Therefore, the beginning of the S-DROSS curve gives the sign for g$_2$ order parameter and end for $\beta$ (Fig. \ref{POPGspectra} (E)).
-This is confirmed with SIMPSON calculations using negative value for g$_2$ and positive value for $\beta$ order parameter (Fig. \ref{POPGsimpson}).
+%Absolute values of the headgroup and glycerol backbone order parameters from PE and PG lipids are measured
+%previously using $^2$H NMR~\cite{seelig76,gally81,wohlgemuth80,borle85}. Because also the order parameter
+%signs bear essential information about the lipid structures \cite{botan15,ollila16}, we measured the
+%magnitudes and signs of POPE and POPG C--H bond headgroup and glycerol backbone order parameter in liquid phase
+%using the 2D-RPDLF and S-DROSS experiments, as described previously \cite{ferreira13,ferreira16,antila19}.
+
+Glycerol backbone peaks from both lipids, and $\alpha$-carbon peak from POPE in the INEPT spectra 
+were assigned based on previously measured POPC spectra~\cite{ferreira13}.
+%For POPE, the glycerol backbone and $\alpha$-carbon peaks in INEPT spectra were assigned based on
+%previously measured POPC spectra~\cite{ferreira13} and
+The $\beta$-carbon peak from POPE was assigned based on $^{13}$C chemical shift table for amines available
+at \url{https://www.chem.wisc.edu/areas/reich/nmr/c13-data/cdata.htm}.
+%For POPG, the glycerol backbone peaks in INEPT spectra were assigned based on
+%previously measured POPC spectra~\cite{ferreira13},
+\todo{How were $\alpha$ and  $\gamma$-carbon peaks assigned in POPG?}.
+The $\beta$-carbon peak from POPG overlapped with the g$_2$ peak from glycerol backbone
+because their chemical environments are similar in POPG.
 \todo{Details to be checked by Tiago}.
@@
 Molecular dynamics simulation data were collected using
-the Open Collaboration method \cite{botan15}, with
-the NMR\-lipids Project blog (\url{nmrlipids.blogspot.fi}) and
-GitHub repository (\url{github.com/NMRlipids/NMRlipidsIVotherHGs})
-as the communication platforms.
+the NMRlipids Open Collaboration \cite{botan15}, which is running at
+\url{nmrlipids.blogspot.fi} and \url{github.com/NMRlipids/NMRlipidsIVotherHGs}.
 The simulated systems of pure PE and PG bilayers without additional ions
@@
 and in the NMRlipids/MATCH repository (\url{github.com/NMRlipids/MATCH}).
+
+\subsection{Analysis of molecular dynamics simulation data}
+The big data set of MD simulations was analysed in the NMRlipids databank manner.
+Unique naming convention for lipid atoms in each force field was defined using the mapping files
+and analysis for all simulations indexed in NMRlipids databank manner were performed
+using python codes.
 
@@
 
-\subsection{Analysis of molecular dynamics simulation data}
-The big data set of MD simulations was analysed in the NMRlipids databank manner.
-Unique naming convention for lipid atoms in each force field was defined using the mapping files
-and analysis for all simulations indexed in NMRlipids databank manner were performed
-using python codes.
-
 \subsection{Analysis of lipid conformations bound to proteins}
@@
      Chemical structure, refocused-INEPT spectrum, 2D R-PDLF spectra, and S-DROSS data (from top to bottom) of {\bf A)} POPE  and {\bf B)} POPG.
+      Full NMR spectra are shown Figs. \ref{POPEspectra} and \ref{POPGspectra}.
      {\bf C)} Chemical structure of POPC and POPS.
@@
    }
-   \todo{This is a sketch, needs a lot of polishing.} \\
+   \todo{This is a sketch, Tiago Ferreira will make a new figure.} \\
   \todo{D) could be clarified as Fig. 2 in the NMRlipids IVps paper.}
@@
 
-To experimentally characterize the differences in lipid headgroup conformational ensembles
-in liquid bilayer phase, we measured the C-H bond order parameters magnitudes and signs
-from POPG and POPE bilayers.
-Order parameters for POPE are in good agreement with previous $^2$H NMR experiments \cite{seelig76}
-and similar to our previous results for POPC \cite{ferreira16}.
-In POPG, the $\beta$ and g$_2$ carbons have similar chemical environment and their peaks
-overlap in the NMR spectra (Fig. \ref{HGorderParameters} B).
-The signs of these order parameters were solved using SIMPSON simulations to interpre the
-S-DROSS experiments in similar fashion as we did previously for PS lipids \cite{antila19}.
-The resulting order parameters are compared with our previously published results
-for POPC \cite{ferreira13,ferreira16} and POPS \cite{antila19} in figure \ref{HGorderParameters} D.
+To experimentally characterize the headgroup conformational ensembles of lipids
+that are not bound to proteins in electrostatically neutral cell membrane,
+we measured the C-H bond order parameters and their signs from of POPG and POPE
+in liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
+Determination of headgroup and glycerol backbone order parameters and their signs
+was straightforward from the
+%R-PDLF and S-DROSS
+data in Figs. \ref{HGorderParameters}, \ref{POPEspectra} and \ref{POPGspectra}
+for all the C-H bonds, except the $\beta$ and g$_2$ carbons in POPG.
+These carbons have overlapping peaks in the INEPT spectra due to their similar chemical environments,
+and only the magnitude of the larger order parameter could be determined from the R-PDLF spectra (Fig. \ref{HGorderParameters} B).
+Nevertheless, based on previous $^2$H NMR measurements \cite{wohlgemuth80,gally81,borle85},
+we assigned the larger order parameter to the g$_2$ carbon
+%order parameter of $\beta$-carbon is expected to be clearly smaller than for 
+and used the literature value for the $\beta$-carbon in SIMPSON simulations to determine the signs.
+The decrease in the beginning of the S-DROSS curve suggests that the sign of larger g$_2$ order parameter
+is negative and later increase suggests that sign of smaller $\beta$ order parameter is positive (Fig. \ref{HGorderParameters} (B)).
+This interpretation is confirmed by SIMPSON calculations in Fig \ref{POPGsimpson}.
+%using negative value for g$_2$ and positive value for $\beta$ order parameter (Fig. \ref{POPGsimpson}).
+%The signs of these order parameters were solved using SIMPSON simulations to interpre the
+%S-DROSS experiments in similar fashion as we did previously for PS lipids \cite{antila19}.
 
-%Our experimental order parameters, including the signs, from different lipid headgroups
-%combined with the literature data are shown in Fig. \ref{HGorderParameters}.
 %Also the headgroup order parameters of PC lipids are close PE, althought the latter gives systematically
@@
 
-The most distinct order parameters are observed for PS and PG headgroups.
-In PS, the $\alpha$-carbon order parameter exhibits significant forking
+Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and
+headgroups from this and previous studies are collected in Fig. \ref{HGorderParameters} D),
+where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
+%The resulting order parameters are compared with our previously published results
+%for POPC \cite{ferreira13,ferreira16} and POPS \cite{antila19} in figure \ref{HGorderParameters} D.
+%Order parameters for POPE are in good agreement with previous $^2$H NMR experiments \cite{seelig76}
+%and similar to our previous results for POPC \cite{ferreira16}.
+Overall agreement of order parameters determined by different authors and different techniques
+for same lipid headgroup is very good in here and previous studies \cite{botan15,ollila16,antila19},
+suggesting that the observed differences between lipid types arise from differences in
+headgroup chemistry rather than inaccuracies in experiments. 
+The most distinct order parameters are observed for PS headgroups,
+for which the $\alpha$-carbon order parameter exhibits significant forking
 and the $\beta$-carbon has more negative value than in other lipids.
-In PG headgroup,
+On the other hand, the $\beta$-carbon order parameter of PG headgroup
 %the $\alpha$-carbon order parameter is similar to PE and PG, while
-the $\beta$-carbon order parameter has positive sign in contrast to all the other lipids.
+has positive sign in contrast to all the other lipids.
 Notably, this has not been observed in traditional $^2$H NMR experiments,
@@
 positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG.
+Essential differences between PC and PE headgroups are not observed.
 
@@
 
-To characterize the differences in lipid headgroup conformational ensembles
-leading to the distinct order parameters in PS and PG lipids, we calculate the
-the distributions of heavy atom dihedral angles from CHARMM36 simulations.
-Among available simulation models, this force field has least problems in
-reproducing the experimental lipid headgroup and glycerol backbone order parameters in different
-lipids \cite{botan15,antila19} (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
-Importantly, it captures the experimentally observed distinct order parameters for PS and PG headgroups (Fig. \ref{??}).
-The dihedral distributions in figure \ref{structures} show significant similarity between different lipids for all the other
-bonds, except for the last two bonds in the headgroup end, O$_\alpha$-C$_\alpha$ and  C$_\alpha$-C$_\beta$.
-These differences explaing the disctinct order parameters of
-PS and PG headgroups with respenct to the other lipids.
-The difference between PC and PE lipids in the O$_\alpha$-C$_\alpha$ bond dihedral
-may be an artefact because the $\beta$-carbon order parameter in PC is poorly reproduced by the CHARMM36 force field \cite{botan15}.
-The differences between lipids are reflected also to the 
-angle between headgroup dipole and membrane normal,
-which decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig. \ref{structures}).
-
-%
-%As in previous NMRlipids project results for PC and PS lipids \cite{botan15,antila19},
-%none of the MD simulation force fields correctly captures all the
-%headgroup and glycerol backbone order parameters of PE and PG lipids
-%(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG})
-%that would enable a straightforward interpretation of conformational ensembles.
-%
-%show wide variation between different force fields
-%and none of the force fields reproduce all values within experimental error bars.
-%(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}),
-%as observed previously also for PC and PS lipids \cite{botan15,antila19}.
-%The Slipid simulations were able to capture the essential differences between PC and PS lipid
-%headgroups \cite{antila19}, but this is not the case for PE and PG lipids.
-%Without further discussion about poorly performing force fields, we focus on more detailed analysis
-%of
-%
-%Nevertheless, CHARMM36 force field, which gives the results closest to the experiments for all lipids,
-%captures the essential differences between
-%PC, PS, PG and PE headroup order parameters  (Fig.~\ref{structures}) with the exception of $\beta$-carbon order parameter of PC
-%
-%give the best overall agreement with experiments for
-%the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
-%studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
-%not within the experimental error bars, it 
-%
-%which is too negative when compared with other lipids or experiments \cite{botan15}.
-%
-% As already pointed out in our previous study \cite{antila19},
-
-The analysis of dihedral angle distributions reveal the free rotation of dihedral around
-phosphorus-oxygen bonds in all lipids as all possible angles are observed in the distributions.
-Results in figure \ref{structure} support the recently proposed models where this free rotation 
-decouples the dynamics and conformational ensembles between acyl chains and headgroup in all lipids containing
-the phosphorus group \cite{antila21}. 
-However, some dihedral angles of phosphorus-oxygen bonds are less likely for PS, which
-%suggests that the decoupling may be weaker for PS lipids
-possibly explains the more rigid headgroup structures proposed for PS lipids~\cite{browning80,buldt81}.
+To resolve the conformational ensembles of lipid headgroups,
+we compared the headgroup and glycerol backbone C-H bond order parameters
+from different MD simulation force fields to the experimental data. 
+None of the force fields reproduced the order parameters
+within experimental accuracy for any of the lipids analysed in this work
+(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}) or in our previous studies \cite{botan15,antila19}.
+%has least problems in
+%reproducing the experimental lipid headgroup and glycerol backbone order parameters in different
+%lipids \cite{botan15,antila19} (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
+Nevertheless, the CHARMM36 force field performs best for all the headgroups and
+%Importantly, it captures the experimentally observed distinct order parameters for PS and PG headgroups (Fig. \ref{??}).
+%To characterize the differences in lipid headgroup conformational ensembles
+reproduces the distinct order parameters of PS and PG lipids (Fig. \ref{structures} B).
+To understand the structural origin of these distinct order parameters,
+we calculated the distributions of heavy atom dihedral angles from CHARMM36 simulations in figure \ref{structures} C).
+%The dihedral distributions in figure \ref{structures} show significant similarity between different lipids for all the other
+%bonds, except
+Major differences between headgroups are observed only for the last two dihedrals
+in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$,
+which prefer {\it gauche$^-$} conformations for PG and {\it gauche$^+$} for PS.
+Rest of the dihedrals are similar between different lipids, with the exception of
+%However, some dihedral angles of phosphorus-oxygen bonds are less likely for PS, which
+small differences for PS lipids. Therefore we suggest that the main differences between
+lipid headgroups leading to distinct order parameter occur in the choline part, while
+also changes in phosphate region may contribute in PS lipids, which could also
+explain the more rigid headgroup structures~\cite{browning80,buldt81}.
+%These differences explaing the disctinct order parameters of
+%PS and PG headgroups with respenct to the other lipids.
+%The differences between lipids are reflected also to
+Furthermore, the angle between headgroup dipole and membrane normal
+decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig. \ref{structures} B)).
+However, the differences between PC and PE in P-O$_\alpha$-C$_\alpha$-C$_\beta$ dihedral
+and P-N vector dipole may be an artefacts as the $\beta$-carbon order parameter in PC
+is too negative the CHARMM36 force field, thereby not being equal to the order parameter
+in PE as observed in experiments \cite{botan15}.
 
@@
 %althought this rotation is slower in PS lipids where crossing certain barriers are slower.
-Despite the differences in dihedral distributions of two the last bonds at the headgroup end,
-leading to the distinct order parameters for PS and PG lipids,
-lipids with different headgroup sample dihedral angles within approximately same ranges.
-Altogether, our results suggest that lipid headgroups and very flexible thereby being
-able to adopt multiple conformations when interacting with proteins, ions or other biomolecules.
-Furthermore, approximately same conformations are present liquid lamellar phase for different
-headgroups studied here, althought the probability weight for the conformations are slightly different.
+Despite the differences in distributions
+%of two the last bonds at the headgroup end,
+%leading to the distinct order parameters for PS and PG lipids,
+%lipids with
+between headgroups, the sampled dihedral angles are within approximately same ranges for all headgroup types.
+Therefore, we suggest that all lipid headgroups are very flexible thereby being
+able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
+%Furthermore, approximately same conformations are present liquid lamellar phase for different
+%headgroups studied here, althought the probability weight for the conformations are slightly different.
 %the order parameter experiments suggest that the glycerol backbone conformations in all lipids
@@
 %of PS lipids are previously related to the more rigid structure of the headgroup~\cite{browning80,buldt81,antila19}.
+%The analysis of dihedral angle distributions reveal the free rotation of dihedral around
+%phosphorus-oxygen bonds in all lipids as all possible angles are observed in the distributions.
+Our results support the models that propose free rotation around phosphate group for PC lipids,
+which decouples the dynamics and conformatios between acyl chains and headgroups \cite{klauda08c,antila21b},
+and suggests that these models could be applied for all lipids containing the phosphorus group.
+The wide range of observed conformations suggest that the structures in lipid crystals \cite{buldt81,pascher92}
+play only a minor role, and that the models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20}
+are not sufficient to capture the large conformational space of lipids in liquid lamellar state.
 
 \subsection{Lipid conformational ensembles in lipid bilayers with bound ions}
-%Phosphatidylcholine headgroup conformational ensemble and orientation are unchanged
-%upon addition of neutral molecules such as cholesterol or sphingomyelin into a bilayer,
-%while
-Order parameters of $\alpha$ and $\beta$ carbons in PC lipid headgroup are known to depend
-linearly on the incorporation of charged molecules into a lipid bilayer,
-which is explained by tilting of the headgroup dipole orientation \cite{??}.
-%are kn affect the headgroup dipole vector tilt which can
-%be detected by measuring C-H bond order parameters of $\alpha$ and $\beta$ carbons.
-Such changes have been observed upon addition of charged lipids, proteins, surfactants,
-drugs, or ions \cite{??}, but how charges affect the lipid conformational ensemble remains unknown.
-%Headgroup order parameters of PC lipids decrease or increase upon addition of positive
-%or negative charge into membrane, respectively, while they remain 
-%\cite{??}.
-%This has been explained by the tilting of the headgroup dipole due to electrostatic
-%interactions with bound charges, and has been used to measure the binding affinity
-%of ions to membranes \cite{??}.
-%The headgroup order parameters of PC lipids can be used to measure the ion binding
-%affinity to lipid bilayers because their magnitude is linearly proportional to the amount of bound charge in bilayer \cite{seelig87,catte16}.
-%This molecular electrometer concept can be used also for bilayers containing PC lipids mixed with charged lipids \cite{borle85,macdonald87,roux90,antila19}.
-%The headgroup order parameters can be used to evaluate MD simulations against experimental data,
-%because they can be directly calculated from MD simulations \cite{catte16}.
 
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPCsurf_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
-%  \caption{\label{DIHSwithSURF}
-%
-%  }
-%\end{figure}
+Charged lipids, proteins, surfactants, drugs, and ions incorporated in membranes
+reorient the headgroup dipole in PC lipids, thereby affecting the order parameters of
+lipid headgroups \cite{seelig87}. However, detailed understanding on
+lipid conformational ensembles in membranes bearing electric charge is still lacking \cite{Semchyschyn04}.
+%Order parameters of $\alpha$ and $\beta$ carbons in PC lipid headgroup are known to depend
+%linearly on the incorporation of charged molecules into a lipid bilayer,
+%which is explained by tilting of the headgroup dipole orientation \cite{??}.
+%%are kn affect the headgroup dipole vector tilt which can
+%%be detected by measuring C-H bond order parameters of $\alpha$ and $\beta$ carbons.
+%Such changes have been observed upon addition of charged lipids, proteins, surfactants,
+%drugs, or ions \cite{??}, but how charges affect the lipid conformational ensemble remains unknown.
 
-The experimentally measured decrease of PC headgroup order parameters upon addition of cationic surfactants into a bilayer
-are well captured by CHARMM36 simulations in figure \ref{changesWITHsurf} A,
-suggesting that these simulations can be used to interpret how lipid headgroup conformational ensembles
-respond to the binding of positively charged into a membrane.
-Charactization of conformational ensembles using heavy atom dihedral angle distributions in figure \ref{changesWITHsurf} B
-reveal that choline region is essentially unchanged upon addition of carge.
-The major changes upon addition of charge are observed in dihedrals related to the phosphate oxygens,
-while small change is observed also in the g$_2$-g$_3$ bond in the glycerol backbone.
-This result is in line with the recently proposed model suggesting that the flexible phosphate enables the headgroup
-orient according to the charge accumulated into a membrane \cite{antila21}.
+To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge,
+%suggesting that these simulations can be used to interpret the charge induced changes in lipid conformational ensembles.
+we first analyzed CHARMM36 simulations which correctly capture the experimentally measured decrease in PC headgroup order parameters
+upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
+Heavy atom dihedral angle distributions in figure \ref{changesWITHsurf} B
+reveal that the decrease of trans state probabilities in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
+dihedrals are the major effects of positive charge on lipid headgroup conformations.
+Choline region remains essentially unchanged and only minor changes are observed in other dihedrals
+even thought almost half of the molecules in membranes are cationic surfactants.
+%The major changes upon addition of charge are observed in dihedrals related to the phosphate oxygens,
+%while small change is observed also in the g$_2$-g$_3$ bond in the glycerol backbone.
+%This result is in line with the recently proposed model suggesting that the flexible phosphate enables the headgroup
+%orient according to the charge accumulated into a membrane \cite{antila21}.
+We also tried to resolve PC lipid headgroup conformational ensembles in mixed membrane with
+negatively charged lipids \cite{scherer87},
+%Headgroup order parameters of PC lipids respond similarly also to the addition of charged lipids in experiments \cite{??}.
+%To resolve the conformational ensembles of lipid headgroups also in biologically relevant
+%mixtures, % of zwitterionic and charged lipids,
+%we compared the response of headgroup order parameters in
+%PC lipids to the addition of PE or PG (Fig. \ref{??}),
+%as well as in PG lipids to the addition of PC (Fig. \ref{??}),
+%between simulations and experiments. However, the
+but the accuracy of currently available force fields
+was not sufficient for this (Fig. \ref{HGorderparametersPCvsPEPG}).
+%to capture the headgroup conformational ensembles
+%in mixed membranes: 
+%The best performing force field for single componen membranes, CHARMM36, overestimates
+%the effect of PE to PC conformations and underestimates the response PC to the addition
+%of anionic PG headgroup. Previously we explained similar result for PC/PS mixtures  
+%%as we previously observed also for mixtures with PS lipids \cite{antila19}.
+%%The latter may arise from
+%by overbinding of the sodium counterions \cite{antila19}.
 
@@
 
-Headgroup order parameters of PC lipids respond similarly also to the addition of charged lipids in experiments \cite{??}.
-To resolve the conformational ensembles of lipid headgroups also in biologically relevant
-mixtures, % of zwitterionic and charged lipids,
-we compared the response of headgroup order parameters in
-PC lipids to the addition of PE or PG (Fig. \ref{??}),
-as well as in PG lipids to the addition of PC (Fig. \ref{??}),
-between simulations and experiments. However, the accuracy of currently available force fields
-is not sufficient to capture the headgroup conformational ensembles
-in mixed membranes: 
-The best performing force field for single componen membranes, CHARMM36, overestimates
-the effect of PE to PC conformations and underestimates the response PC to the addition
-of anionic PG headgroup. Previously we explained similar result for PC/PS mixtures  
-%as we previously observed also for mixtures with PS lipids \cite{antila19}.
-%The latter may arise from
-by overbinding of the sodium counterions \cite{antila19}.
-
-In order to analyze how conformational ensemble of PC and PG lipids response to the
-bound ions, we compared the changes of headgroup order parameters
-in POPC:POPG (1:1) and (4:1) mixtures upon addition of CaCl$_2$ from different simulations
-to the experimental data available in the literature \cite{borle85,macdonald87}
-in figures \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4}.
+Also binding of ions may affect the zwitterionic and charged lipid headgroup conformational
+ensembles in physiological conditions. The resolve this effect,
+%In order to analyze how conformational ensemble of PC and PG lipids response to the
+%bound ions,
+we compared the changes of headgroup order parameters
+in POPC:POPG mixtures 
+upon addition of CaCl$_2$ between different simulations and experiments \cite{borle85,macdonald87}
+in figures \ref{changesWITHCaClPG} (molar ratio 1:1) and \ref{changesWITHCaClPG1PC4} (molar ratio 4:1).
 As in our previous studies \cite{catte16,melcr18,antila19,melcr20},  

### 2020-12-11 user-instructions/user-instructions.tex
updated user instructions

--- 2020-11-30 user_instructions/user_instructions.tex
@@
 
-\section*{User input}
+\subsection*{Compulsory user input}
+The following parameters are compulsory and have a strict format.
 
 \subsubsection*{DOI}
-DOI is the DOI access number of the simulation. Use the version DOI.
-
-\subsubsection*{SYSTEM}
-SYSTEM is name of the system.
+DOI is the DOI access number of the simulation. Use the version DOI on Zenodo.
 
@@
 
+\subsubsection*{TRJ}
+TRJ stands for the name of the trajectory file.
+\subsubsection*{TPR}
+TPR stands for the name of the topology file.
+
+\subsubsection*{PREEQTIME}
+PREEQTIME means the time used for pre-equilibrating the system. This should be in nano seconds. 
+\subsubsection*{TIMELEFTOUT}
+TIMELEFTOUT stands for the length of the simulation in nano seconds to be left out at the beginning of the simulation. This value is used in the script to leave out the given length from the analysis!
+
+\subsubsection*{UNITEDATOM}
+The handling of united atom simulations is enabled by a separate script called buildH\_calcOP.py.
+In case of a united atom simulation, the user has to give the names of the lipids and the corresponding names of those lipids as they are in the dic\_lipids.py dictionary used by buildH\_calcOP.py script.
+In case of an all atom simulation this can be given as an empty string.
+
+\subsubsection*{Molecule names}
+The databank has default names for lipids, ions and solvent.
+The user must also provide the names of the molecules, ions and solvent that are used in the simulation to match the names used by the databank.
+If the name of the molecule is not in the databank it needs to be added.
+The names provided by the user must be the same as in the tpr file.
+If a lipid consists of more than one residue the name of the head group residue is given.
+
+\subsubsection*{dir\_wrk}
+The user has to give the path of the working directory.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%free form parameters%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subsection*{Free form user input}
+The form of how the values of these parameters are written is not essential for the script to work properly.
+\subsubsection*{SYSTEM}
+SYSTEM is name of the system.
 \subsubsection*{FF}
@@
 
-\subsubsection*{TRJ}
-TRJ stands for the name of the trajectory file.
-\subsubsection*{TPR}
-TPR stands for the name of the topology file.
-\subsubsection*{PREEQTIME}
-PREEQTIME means the time used for pre-equilibrating the system. This should be in nano seconds.
-\subsubsection*{TIMELEFTOUT}
-TIMELEFTOUT stands for the length of the simulation in nano seconds to be left out at the beginning of the simulation.
-\subsubsection*{UNITEDATOM}
-The handling of united atom simulations is enabled by a separate script called buildH\_calcOP.py.
-In case of a united atom simulation, the user has to give the names of the lipids and the corresponding names of those lipids as they are in the dic\_lipids.py dictionary used by buildH\_calcOP.py script.
-In case of an all atom simulation this can be given as an empty string.
+\subsection*{Automatically analyzed parameters}
+The following parameters are read automatically from the trajectory and topology files.
 
-\subsubsection*{Molecule names}
-The user must also provide the names of the molecules, ions and solvent that are used.
-These are obtained from the tpr file.
-If a lipid consists of more than one residue the name of the head group residue is given.
-\subsubsection*{dir\_wrk}
-Finally the user has to give the path of the working directory.
+\subsubsection*{Molecule numbers}
+Numbers of lipid molecules (NPOPC, NPOPG, etc.) per membrane leaflet are calculated by determining on which side of the center of mass of the membrane the center of mass of the head group of each lipid molecule is located.
+\newline \\
+\noindent Numbers of other molecules such as solvent and ions (NSOL, NPOT, NSOD, etc.) are read from the topology file.
+
+\subsubsection*{Temperature}
+Temperature of the simulation is read from the topology file.
+
+\subsubsection*{Trajectory length}
+The length of a trajectory is read from the trajectory file.
 

### 2020-12-14 scripts/DataBankINFO/AddData.py
molecule_ff_dict dictionary added for specifying force fields for different molecules

@@ -29,7 +29,7 @@ import os
 
 #For quering webs
 import urllib.request
-from urllib.error import URLError,HTTPError #,ContentTooShortError
+from urllib.error import URLError,HTTPError
 
 # From time monitoring
 from tqdm import tqdm
@@ -202,7 +202,40 @@ molecule_numbers_dict = {
                         },
     
                 }
-               
+
+molecule_ff_dict = {
+                'FFPOPC' : {"REQUIRED": False,
+                                "TYPE": "string",
+                           },
+                'FFPOPG' : {"REQUIRED": False,
+                                "TYPE": "string",
+                           },
+                'FFPOPS' : {"REQUIRED": False,
+                                "TYPE": "string",
+                           },
+                'FFPOPE' : {"REQUIRED": False,
+                                "TYPE": "string",
+                           },
+                'FFCHOL' : {"REQUIRED": False,
+                                "TYPE": "string",
+                           },
+		'FFPOT' : {"REQUIRED": False,
+                            "TYPE" : "string",
+                        },
+                'FFSOD' : {"REQUIRED": False,
+                            "TYPE" : "string",
+                        },
+                'FFCLA' : {"REQUIRED": False,
+                            "TYPE" : "string",
+                        },
+                'FFCAL' : {"REQUIRED": False,
+                            "TYPE" : "string",
+                        },
+                'FFSOL' : {"REQUIRED": False,
+                            "TYPE" : "string",
+                        },
+               }
+
                 
                 
 
@@ -268,7 +301,7 @@ gromacs_dict = {
                              "TYPE" : "string",
  #                        "EXTENSION": ("txt"),
                              },
-            'UNITEDATOM' : {"REQUIRED": False,
+            'UNITEDATOM' : {"REQUIRED": True,
                             "TYPE" : "string",
                          },
                }
@@ -427,8 +460,8 @@ for sim in sims_valid_software:
         if key_sim.upper() in ("ID", "SOFTWARE"):
             #print("NOT REQUIRED")
             continue
-    #Anne: check if key is in molecules_dict or molecule_numbers_dict too
[…]

### 2020-12-15 Manuscript/manuscriptPGPE.tex
Update manuscript

@@
 % insert abstract here
-The force field giving the best description for glycerol backbone and headgroup structures
-of PC, PS, PG and PE headgroups (CHARMM36) reproduces the essential differences in order parameters
-between these headgroups, and therefore enables the analysis of structural differences between the
-headgroups.
+  Chemistry of lipid headgroups, the water facing components of cell membranes, varies between organisms and
+  organelles and regulate cell functions via lipid-protein interactions.
+  Because lipid membranes are in liquid state in physiological conditions,
+  the conformational ensembles of lipids and their dependence of headgroup chemistry are difficult to resolve experimentally.
+  Here, we combine solid state NMR experiments and MD simulations from NMRlipids open collaboration to
+  resolve conformational ensembles of PC, PS, PG and PE lipid headgroups in liquid lamellar phase in various biologically relevant conditions.
+  Interpretation of NMR experiments using the best simulations from the NMRlipids project suggests that all lipid headgroups sample wide range of
+  conformations in neutral and charged cell membranes, and that the differences in headgroup chemistry
+  appears in the probability weights of different conformations. Together with the analysis of protein-bound-lipids
+  from the protein data bank (PDB), this suggests that lipids can bind to proteins in wide range of conformations
+  independently on the headgroup chemistry. Therefore, the specific lipid-protein interactions are likely driven by
+  protein-lipid interactions, rather than conformational restrictions of lipids. Our results paves the way to
+  comprehensive understanding of lipid-protein interaction energetics, and understanding complex biomolecular assemblies
+  such as membrane proteins.
+
+  %The force field giving the best description for glycerol backbone and headgroup structures
+  %of PC, PS, PG and PE headgroups (CHARMM36) reproduces the essential differences in order parameters
+  %between these headgroups, and therefore enables the analysis of structural differences between the
+  %headgroups.
 
@@
 
-\subsection{Analysis of lipid conformations bound to proteins}
-Dihedral angles of all available conformations in the PDB databank were calculated using the
-API access to the databank.
-
 %\clearpage
@@
 \end{figure*}
+
+\subsection{Analysis of lipid conformations bound to proteins}
+Lipid structures from PDB (\url{http://www.rcsb.org/})
+were searched using PDBe REST API (\url{www.ebi.ac.uk/pdbe/pdbe-rest-api}).
+First, all PDB entries containing PC, PE, PG, or PS lipid headgroups were searched.
+The ligand names to identify the lipids were:
+PLC, PX4, 6PL, LIO, HGX, PC7, PC8, P1O, 6O8, XP5, EGY, PLD, SBM, HXG, and PCW for PC;
+8PE, PTY, 3PE, PEH, PEF, 6OE, 6O9, 9PE, PEV, 46E, SBJ, L9Q, PEK, EPH, ZPE, 9TL, 9Y0, 6OU, LOP, and PEE for PE;
+PGT, PGK, LHG, 44G, PGV, OZ2, D3D, PGW, DR9, P6L, PG8, H3T, and GOT for PG; and
+PSF, PS6, Q3G, P5S, D39, PS2, 17F, and 8SP for PS.
+All PDB structures containing these ligands were downloaded and the dihedral angles of
+the first lipid in the latest version of the structures within each PDB code were calculated  
+using MDTraj python library \cite{mcgibbon15}.
+The used Jupyter notebook is available from the project's GitHub repository ({\it scripts/pdbSEARCH.ipynb}).
 
@@
 
+\begin{figure*}[bt]
+  \centering
+  \includegraphics[width=18.0cm]{../Figs/figure3.eps}
+  \caption{\label{changesWITHsurf}
+    {\bf A)} Modulation of PC headgroup order parameters and P-N vector angle upon addition of cationic surfactant
+    from CHARMM36 simulations compared with experimental data \cite{??}.
+    {\bf B)} Changes in PC headgroup conformational ensembles upon increasing amount of positive charge in bilayer,
+    characterized by the heavy atom dihedral distributions, from CHARMM36 simulations.
+  }
+\end{figure*}
+
 To resolve the conformational ensembles of lipid headgroups,
@@
 
-\begin{figure*}[bt]
-  \centering
-  \includegraphics[width=18.0cm]{../Figs/figure3.eps}
-  \caption{\label{changesWITHsurf}
-    {\bf A)} Modulation of PC headgroup order parameters and P-N vector angle upon addition of cationic surfactant
-    from CHARMM36 simulations compared with experimental data \cite{??}.
-    {\bf B)} Changes in PC headgroup conformational ensembles upon increasing amount of positive charge in bilayer,
-    characterized by the heavy atom dihedral distributions, from CHARMM36 simulations.
-  }
-\end{figure*}
-
 In conclusion, all lipid headgroups sample very broad conformational ensembles in liquid lamellar phase.
@@
 
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC1.eps}
+  \caption{\label{changesWITHCaClPG}
+    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
+    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
+    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
+    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
+    The changes with respect to the systems without CaCl$_2$ are shown for other data than
+    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
+    Calsium density distributions are shown in figure \ref{CAdensPG}.
+    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
+    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
+  }
+\end{figure*}
+
 To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge,
 %suggesting that these simulations can be used to interpret the charge induced changes in lipid conformational ensembles.
-we first analyzed CHARMM36 simulations which correctly capture the experimentally measured decrease in PC headgroup order parameters
+we first analyzed CHARMM36 simulations which correctly captures the experimentally measured decrease in PC headgroup order parameters
 upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
@@
 
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC1.eps}
-  \caption{\label{changesWITHCaClPG}
-    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
-    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
-    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
-    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
-    The changes with respect to the systems without CaCl$_2$ are shown for other data than
-    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
-    Calsium density distributions are shown in figure \ref{CAdensPG}.
-    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
-    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
-  }
-\end{figure*}
-
-Also binding of ions may affect the zwitterionic and charged lipid headgroup conformational
-ensembles in physiological conditions. The resolve this effect,
-%In order to analyze how conformational ensemble of PC and PG lipids response to the
-%bound ions,
-we compared the changes of headgroup order parameters
+Also binding of ions may affect the lipid headgroup conformational ensembles in physiological conditions.
+The resolve this effect, we compared the changes of headgroup order parameters
 in POPC:POPG mixtures 
@@
 in figures \ref{changesWITHCaClPG} (molar ratio 1:1) and \ref{changesWITHCaClPG1PC4} (molar ratio 4:1).
-As in our previous studies \cite{catte16,melcr18,antila19,melcr20},  
-%The decrease of POPC headgroup order parameters in mixtures with POPG lipids with
-%increasing CaCl$_2$ concentration is overestimated in Slipids and Lipid17
-%order parameters decrease in the response to bound cations in qualitative agreement with
-%experiments, but
-the calcium binding affinity to membranes is typically overestimated in simulations,
-except by CHARMM36 with the NBfix correction which underestimates the binding affinity,
-and the implicit inclusion of electronic polarizability improves the results.
-Lipid17ecc model with the implicit inclusion of electronic polarizability gives
-the most realistic response of PC lipid headgroup order parameters to the binding of calcium.
-In this model, the main effect of calcium to the lipid conformational ensemble is
-%the decrease is headgroup order parameters observed in this model is
-the slight change of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral distribution from trans state
-to eclipsed conformations (Fig. \ref{DIHSwithCAlipid17eccPOPC}).
-This is in line with the changes observed in CHARMM36 simulations upon addition of charged surgactants (Fig. \ref{changesWITHsurf}),
-despite the major differences in lipid headgroup conformational ensembles between
-these models without ions (Fig. \ref{structures} vs. Fig. \ref{DIHSwithCAlipid17eccPOPC}).
+The results are in line with our previous studies:
+most force fields overestimate the calcium binding \cite{catte16,antila19}, but CHARMM36 with the NBfix correction
+underestimates the binding affinity \cite{antila19}, and the implicit inclusion of electronic polarizability
+using the electronic continuum correction (ECC) improves the results \cite{melcr18,melcr20}.
+The most realistic response of PC lipid headgroup order parameters to the calcium binding
+is observed in the Lipid17ecc simulations with the ECC included, where
+the decreased probability for trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
+is the main structural consequence of bound ions (Fig. \ref{DIHSwithCAlipid17eccPOPC}).
+Similar but smaller effect is seen in CHARMM36 simulations with less bound calcium (Fig. \ref{DIHSwithCAcharmm36POPC}).
+
+%distribution from
+%to eclipsed conformations 
+%In this model, the main effect of calcium to the lipid conformational ensemble is
+%the slight change of 
+%This is in line with the changes observed in CHARMM36 simulations upon addition of charged surgactants (Fig. \ref{changesWITHsurf}),
+%despite the major differences in lipid headgroup conformational ensembles between
+%these models without ions (Fig. \ref{structures} vs. Fig. \ref{DIHSwithCAlipid17eccPOPC}).
+
 %of lipid headgroup is not 
@@
 
-Decrease in headgroup order parameters of PC lipids upon addition of charges is
-usually qualitatitely captured in simulations despite the inaccuries in lipid conformational ensembles \cite{catte16},
-but situation with PS lipids is more complicated \cite{antila19,meclr20}.
-Here, Lipid17 and Slipids force fields correctly capture the PG $\beta$-carbon order parameter response to CaCl$_2$
-in figure \ref{changesWITHCaClPG} even thought the calcium binding affinity was overestimated.
-%based on the comparison of PC headgroup order parameter changes with experiments.
-%While applying ECC to Lipid17 improved the PC headgroup order parameter response and binding affinity,
-%The response of PG $\beta$-carbon order parameter to calcium is too weak in CHARMM36 and ECC-lipid17 force fields,
-%but decrease of the order parameter is associated with decreased P-N vector angle in all force fields.
-Heavy atom dihedral angle distributions calculated from these simulations
-%the changes in PG conformational ensemble from these models
-suggest that also the headgroup glycerol conformations in PG are affected by calcium
-(Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}),
-in contrast to PC lipids where only conformations near phosphate were affected.
-%dihedral angle distribution around O$_\alpha$-C$_\alpha$ bond is affected by calcium binding (Fig. ??).
-On the other hand, the upward tilting of the headgroup dipole is weaker
-%Even thought the change in headgroup orientation upon addition of calcium is smaller
-in PG than in PC headgroup, possible due to the compensating effects from the changes in
-phosphate and glycerol regions. 
-However, none of the simulations captures and calcium binding affinity and conformational ensemble of PG lipids
-simultaneously, and experimental data to evaluate 
-%differences between force fields are observed for
-the response of $\alpha$-carbon order parameters to
-the added calcium in PG is not available.
-Therefore, more accurate force fields are required for the solid analysis of PG conformations enembles in different ionic conditions.
+%Decrease in headgroup order parameters of PC lipids upon addition of charges is
+%usually qualitatitely captured in simulations despite the inaccuries in lipid conformational ensembles \cite{catte16},
+%but situation with PS lipids is more complicated \cite{antila19,melcr20}.
+Also the headgroup order parameters of charged lipids are affected by the binding of ions,
+but their response is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
+In our simulations, the Lipid17 and Slipids force fields give the most realistic response of PG $\beta$-carbon to CaCl$_2$
+in figure \ref{changesWITHCaClPG} even thought the calcium binding affinity was overestimated in these models.
+%Even thought the
+%change in headgroup orientation upon addition of calcium is smaller
+%On the other hand, the
+These models suggest that the upward tilting of the headgroup dipole is weaker in PG than in PC headgroup (\ref{changesWITHCaClPG}),
+%Heavy atom dihedral angle distributions calculated from these simulations
+%suggest that also the headgroup glycerol conformations in PG are affected by calcium
+but dihedral distributions are more sensitive to the bound ions in PG 
+(Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
+%in contrast to PC lipids where only conformations near phosphate were affected.
+%, possible due to the compensating effects from the changes in
+%phosphate and glycerol regions. 
+However, none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids
+simultaneously, experimental data to evaluate the response of $\alpha$-carbon order parameters to
+the added calcium in PG is not available, and changes in dihedrals upon addition of cacium differ between
+the best models (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
+Therefore, more accurate force fields are required for the detailed analysis of the interactions between ions
+and charged lipids, as concluded previously also for PS lipids \cite{antila19,melcr20}.
 
@@
 
-Despite the limited cabability of simulations to interpret some experimental data due to
-the inaccuries in force fields, we can conclude that the experimentally observed changes
+Despite the limited capability of simulations to interpret some of the experimental data due to
+the inaccuries in force fields, we conclude that the experimentally observed changes
 in headgroup order parameters upon addition of charges into bilayer arise from relatively
-small changes in conformational ensembles. These changes can be characterised by
-mild changes in dihedral angle distributions, rather than restriction of lipids into
+small changes in conformational ensembles. These changes eventuate from
+mild changes in dihedral angle distributions, rather than from restriction of lipids into
 fixed conformations. Therefore, lipid headgroups remain in disordered state sampling large
 space of different conformations, thereby being able to interact with different molecules
-in various ways, also when charged molecules are bound to membranes.
+in various ways, also in membranes bearing electric charge.
 
@@
 
-While our results in previous sections suggest that lipid headgroups
-sample large conformational space in liquid lamellar phase, lipids
-can bound to proteins in fixed conformations which can be then detected
-using crystallography or cryo-EM. Such protein bound lipid conformations are
-available within protein structures deposited in The Protein Data Bank (PDB)
-available at \url{http://www.rcsb.org/} \cite{berman00}. With our criterias described
-in the methods, we found
-?? conformations of PC lipids,
-?? conformations of PE lipids,
-?? conformations of PG lipids, and
-?? conformations of PS lipids from PDB.
+To analyze lipid conformations in protein bound states,
+we calculated the heavy atom dihedral distributions from 
+all lipid conformations available within protein structures
+deposited in the PDB \cite{berman00} in figure \ref{dihedralsFROMpdb}.
+We found 176 PC, 198 PE, 70 PG, and 41 PS conformations, which
+present lipids that are tightly bound to proteins in fixed conformations
+that can be determined as a part of protein structure using crystallography, NMR, or cryo-EM.
 
@@
 
-To analyze the relation of lipid headgroup conformational ensembles between
-liquid lamellar and protein bound states, we calculated the heavy atom dihedral angle
-distributions also from the protein conformations found from the PDB.
-The results in figure \ref{dihedralsFROMpdb} do not reveal major differences
-between different lipid headgroups when bound to proteins.
-Some preference for trans state in C$_\alpha$-C$_\beta$ bond of PC lipids, and
-for positive angles in O$_\alpha$-C$_\alpha$ and P-O$_\alpha$ of PS lipids with respect
-to other lipids may be present in the data, but more statistics is required for solid
-conclusions. Therefore, the differences in conformational ensembles between different
-lipids observed in liquid lamellar state are not clearly visible in the protein bound states.
+Similarly to the liquid lamellar state, lipid dihedrals have very wide range of angles
+when bound to proteins. Only %restrictions seem to be in 
+%Almost all dihedral angles are found from the protein bounds states for all dihedrals
+%except for
+P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals
+seem to avoid cis conformations as observed also in liquid lamellar state.
+The %range of available angles for
+O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and
+g$_1$-g$_2$-g$_3$-O$_{g_3}$ dihedrals are even less restricted in protein bound state
+than in liquid lamellar phase where cis and anti ecplipsed state were not present (Fig. \ref{structures}). 
+Major differences between different lipid headgroups are not observed when bound to proteins.
+Slight preference for the trans state in O$_\alpha$-C$_\alpha$-C$_\beta$-/C$_\gamma$ dihedral of PC,
+and for positive angles in P-O$_\alpha$-C$_\alpha$-C$_\gamma$ dihedral of PS lipids with respect
+to other lipids could be present in the data, but the statistics is not sufficient for conclusions.
+Therefore, the differences in conformational ensembles between different
+lipids in liquid lamellar state in Fig. \ref{structures} are not seen in the protein bound states.
 
-Almost all dihedral angles are found from the protein bounds states for all dihedrals
-except for P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P, which seem to avoid
-the cis conformations. The wide range of available dihedrals is similar to our results
-in liquid lamellar state, except for O$_\alpha$-C$_\alpha$-C$_\beta$-/C$_\gamma$ and
-g$_1$-g$_2$-g$_3$-O$_{g_3}$ dihedrals which seem to be more restricted in the liquid
-lamellar phase (Fig. \ref{structures}). This suggests that lipids compromize their
-preferred conformations when bound to proteins to minimize the intermolecular interactions
-with proteins.
+This suggests that lipids can bind to proteins in wide range of conformations independently
+of headgroup chemistry. Thereby, lipids compromize their preferred conformations when binding
+to proteins and the binding of lipids to specific locations would be driven by 
+the intermolecular interactions between proteins and lipids.
+However, it is important to note that lipids are often not the main target in
+structures deposited in PDB and therefore their conformations may be less reliable than proteins.
+Stereochemical violations and structures deviating from lipid crystals have been
+previously proposed to indicate inaccuracies in lipid structures in PDB \cite{marsh13b,pezeshkian18}. 
+However, we see large deviations from lipid crystals structures also in conformational ensembles
+that reproduce the NMR data in liquid lamellar phase, thereby proposing that such deviations
+are realistic also in protein bound states.
 
@@
 
-We have measured the C-H bond order parameters with the signs for headgroup and glycerol
-backbone regions of the most most abundant biological phopholipids, PC, PE, PG and PS.
-Combining this experimental data with the large amount of simulation data collected 
-using the NMRlipids open collaboration, we were able to resolve the differences between
-conformational ensembles of lipid headgroup leading to the differences observed in NMR experiments.
+%We have measured the C-H bond order parameters with the signs for headgroup and glycerol
+%backbone regions of the
+Conformational ensembles resolved using solid state NMR experiments and MD simulations
+from NMRlipids open collaboration revealed that headgroups of the most abundant biological
+phopholipids, PC, PE, PG and PS, sample wide range of different conformations in lamellar liquid state.
+%Combining this experimental data with the large amount of simulation data collected 
+%using the NMRlipids open collaboration, we were able to resolve the differences between
+%conformational ensembles of lipid headgroup leading to the differences observed in NMR experiments.
+%Our results indicate that lipid headgroups are flexible and sample large conformational space
+%in biologically relevant liquid lamellar state,
+Differences in NMR order parameters between different headgroups can be explained
+by the changes in dihedral angle distributions, suggesting that
+similar conformations are accessed by all headgroups, but with different probabilities.
+Also the changes in order parameters upon addition of charged molecules, such as cationic lipids, surfactants, or drugs,
+in membranes can be explained by the changes in dihedral angle probabilities, particularly close to the phosphate region.
 
-Our results indicate that lipid headgroups are flexible and sample large conformational space
-in biologically relevant liquid lamellar state, also in membranes containing charged molecules,
-which have the largest affect on lipid headgroups in NMR experiments.
-Therefore, lipids can bind to proteins and other biomolecules in multiple different conformations,
-as indeed observed in the analysis of protein bound lipid conformations available from PDB.
-The weak correlation between lipid structures in liquid lamellar phase and protein bound state
-suggestst that the selective lipid binding to proteins is dominated by the intermolecular lipid-protein
-interactions, and the differences in conformational ensembles between different lipid types
-play only a minor role.
+Wide range of conformations are observed also in lipids that are tightly bound to proteins in PDB,
+%Therefore, lipids can bind to proteins and other biomolecules in multiple different conformations,
+%as indeed observed in the analysis of protein bound lipid conformations available from PDB.
+%The weak correlation between lipid structures in liquid lamellar phase and protein bound state
+suggesting that the specific binding of lipids to proteins is dominated by the intermolecular lipid-protein
+interactions, while the differences in conformational ensembles between different lipid types
+play a minor role. Therefore, lipids can conform themself to multiple binding positions locating
+in various proteins.
 
 Our results pave the way to understand how lipids regulate membrane protein function.
-For example, the resolved lipid conformational ensembles in liquid lamellar phase and
-realistic MD simulations enables accurate analysis of lipid-protein interactions energetics.
-Furthermore, our results demonstrate how NMR data can be combined with the large amount of
-indexed MD simulation data to find the most realistic conformational ensembles of biomolecules
-in membrane environment. This paves the way toward standardized methods to resolve the quality
-evaluated conformational ensembles of disordered biomolecules in membrane environment.
-In the NMRlipids project we aim to build a general databank of quality evaluated
-conformational ensembles of lipids and other disordered biomolecules based primarily on
-MD simulations and NMR data.
+Lipid conformational ensembles in liquid lamellar phase are necessary
+for the detailed analysis of lipid-protein interaction energetics,
+%the
+%realistic MD simulations enables accurate
+%analysis of
+particularly for the entropic part. 
+Furthermore, our results demonstrate how the NMRlipids databank, containing MD simulations and NMR data,
+can be used to resolve
+%NMR data can be combined with the large amount of
+%indexed MD simulation data to find the most realistic
+conformational ensembles of disordered biomolecules in membrane environment,
+and how this information can be coupled with structural data in the PDB.
+We believe that this combination brings us closer to the comprehensive understanding of
+complex biomolecular assemblies containg both disordered and structural regions, such as membrane
+protein complexes.
+%This paves the way toward standardized methods to resolve the quality
+%evaluated conformational ensembles of disordered biomolecules in membrane environment.
+%In the NMRlipids project we aim to build a general databank of quality evaluated
+%conformational ensembles of lipids and other disordered biomolecules based primarily on
+%MD simulations and NMR data.
 

### 2020-12-15 Manuscript/manuscriptPGPEsuppl.tex
Update manuscript

@@
     DPPG with 100mM NaCl from Ref.~\citenum{wohlgemuth80},% contains 10mM PIPES and,
-    and E.Coli PG results from Ref.~\citenum{gally81}).
+    and E.Coli PG results from Ref.~\citenum{gally81})
     and simulations with different force fields.
@@
     Modulation of POPC headgroup order parameters with increasing amount of POPE (left) and POPG (right) in bilayer
-    from experiments \cite{scherer87,macdonald87} and simulations with different force fields.
-    Signs are determined as discussed in \cite{botan15,ollila16}.
+    from experiments at 298 K \cite{scherer87,macdonald87} and simulations with different force fields
+    (temperatures listed in tables \ref{systemsMIX} and \ref{systemsMIX2} are between 298-310 K).
+    Signs are determined as discussed in Refs. \citenum{botan15,ollila16}.
   }
@@
 binding affinity that neturalizes the effect of added negative charge \cite{antila19}.
+\todo{Difference between Åqvist and Dang in Slipids results probably supports this,
+but this has to checked from density profiles.}
 %In all force fields except Slipids, the order parameters of different
@@
     Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
-    from experiments \cite{borle85,macdonald87} and simulations with different force fields.
+    from experiments at 298 K \cite{borle85,macdonald87} and simulations with different force fields at 310 K.
   }
@@
   \centering
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
-  \caption{\label{DIHSwithCA}
-    Changes in POPC CHARMM36 dihedrals with increasing amount of CaCl$_2$.
-  }
-\end{figure}
-
-\begin{figure}[]
-  \centering
   \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17ecc.png}
@@
     Changes in POPC lipid17ecc dihedrals with increasing amount of CaCl$_2$.
+  }
+\end{figure}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+  \caption{\label{DIHSwithCAcharmm36POPC}
+    Changes in POPC CHARMM36 dihedrals with increasing amount of CaCl$_2$.
   }
@@
     % The lipid force fields named as in our previous work~\cite{botan15}.
-  }\label{systemsMIX}
+  }\label{systemsMIX2}
   \begin{minipage}[t]{\textwidth}

### 2020-12-16 scripts/DataBankINFO/AddData.py
AddData.py prints error and quits running if downloads fail

@@ -29,10 +29,12 @@ import os
 
 #For quering webs
 import urllib.request
-from urllib.error import URLError,HTTPError
+from urllib.error import URLError,HTTPError,ContentTooShortError
 
 # From time monitoring
 from tqdm import tqdm
+
+import socket
 
 # Python program to find SHA256 hash string of a file
 import hashlib
@@ -622,7 +624,9 @@ for sim in sims_required_entries:
 
 # In[14]:
 
+socket.setdefaulttimeout(30)
 
+download_failed = False
 
 # Create temporary directory where to download files and analyze them
 dir_tmp = os.path.join(dir_wrk, "tmp_6-" + str(randint(100000, 999999)))
@@ -631,6 +635,7 @@ if (not os.path.isdir(dir_tmp)):
     os.mkdir(dir_tmp)
 
 for sim in sims_working_links:
+    file_sizes={}
     print("ID {0}".format(sim["ID"]), flush=True)
     software_sim = software_dict[sim['SOFTWARE'].upper()]
     dir_sim = os.path.join(dir_tmp, str(sim["ID"]))
@@ -647,11 +652,41 @@ for sim in sims_working_links:
                 for file_provided in tqdm(value_sim, desc = key_sim):
                     file_url = download_link(DOI, file_provided[0])
                     file_name = os.path.join(dir_sim, file_provided[0])
+                    #get the size of the file to be downloaded
+                    url_size = urllib.request.urlopen(file_url).length
                     if (not os.path.isfile(file_name)):
-                        response = urllib.request.urlretrieve(file_url, file_name)
-                          
+                        print("downloading")
+                        try:
+                            response = urllib.request.urlretrieve(file_url, file_name)
+                        except ContentTooShortError:
+                            download_failed = True
+                            print("Content too short error.")
+                        except HTTPError as e:
+                            download_failed = True
+                            print(e)
+                        except URLError as ue:
+                            download_failed = True
+                            print("failed to download")
+                        except socket.timeout as se:
+                            download_failed = True
+                            print("socket time out")
+                        except Exception as ee:
+                            download_failed = True
+                            print(ee)
+                #check if the file is fully downloaded
+                    size = os.path.getsize(file_name)
+                    print("size of the file " + file_provided[0] + " to be downloaded: " + str(url_size))
+                    print("size of the file " + file_provided[0] + " after download: " + str(size) )
+                    if url_size != size:
[…]

### 2020-12-16 user-instructions/user-instructions.tex
updated user instructions

@@
 To add a simulation to the database the user has to give some essential information about the simulation.
-This is done by writing a info file (*INFO.py) which is passed to AddData.py.
+This is done by writing an info file (*INFO.py) which is passed to AddData.py.
 \newline \\
@@
 The user must also provide the names of the molecules, ions and solvent that are used in the simulation to match the names used by the databank.
-If the name of the molecule is not in the databank it needs to be added.
 The names provided by the user must be the same as in the tpr file.
 If a lipid consists of more than one residue the name of the head group residue is given.
+If the name of the molecule is not in the databank it needs to be added.
+The names of the molecules are listed in dictionaries called lipids\_dict and molecules\_dict which are in the script.
+If a new molecule is added it needs to be added to molecule\_numbers\_dict and molecule\_ff\_dict too.
 
@@
 FF\_DATE is the date when the forcefield parameters were created. The format is day/month/year.
-
+\subsubsection*{Individual force field names for molecules}
+The user can also give forcefield names to different molecules individually. These parameters are named as FFPOPC, FFPOT, FFSOL etc.
 \subsection*{Automatically analyzed parameters}
 The following parameters are read automatically from the trajectory and topology files.
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%automatically analyzed parameters%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \subsubsection*{Molecule numbers}

### 2020-12-22 Manuscript/manuscriptPGPE.tex
Update manuscript
Update authors

@@
 
+\author{Am{\'e}lie Bacle}
+\affiliation{Paris, France}
+
 \author{Pavel Buslaev}
 \affiliation{University of Jyv{\"a}skyl{\"a}}
+
+\author{Rebeca Garc{\'i}a Fandi{\~n}o}
+\affiliation{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela, E-15782 Santiago de Compostela, Spain}
+\affiliation{CIQUP, Centro de Investigação em Química, Departamento de Química e Bioquímica, Faculdade de Ciências, Universidade do Porto, Porto, Portugal}
 
@@
 
+\author{Paula Milan Rodriguez}
+\affiliation{Paris, France}
+
 \author{Markus S. Miettinen}
@@
 \affiliation{Chemistry, University of Southampton, Highfield, Southampton SO17 1BJ, United Kingdom}
+
+\author{{\'A}ngel Pi{\~n}eiro}
+\affiliation{Departamento de F{\'i}sica Aplicada, Facultade de F{\'i}sica, Universidade de Santiago de Compostela, E-15782 Santiago de Compostela, Spain}
 
@@
    \todo{This is a sketch, Tiago Ferreira will make a new figure.} \\
-  \todo{D) could be clarified as Fig. 2 in the NMRlipids IVps paper.}
+%  \todo{D) could be clarified as Fig. 2 in the NMRlipids IVps paper.}
 \end{figure*}
@@
   \centering
-  \includegraphics[width=18.0cm]{../Figs/figure3.eps}
+  \includegraphics[width=18.0cm]{../Figs/HGorderparametersPCvsSURFchangeDIHEDRALS.eps}
   \caption{\label{changesWITHsurf}

### 2020-12-22 Manuscript/manuscriptPGPEsuppl.tex
Update authors

@@
 \affiliation{University of Jyv{\"a}skyl{\"a}}
+
+\author{Rebeca Garc{\'i}a Fandi{\~n}o}
+\affiliation{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela, E-15782 Santiago de Compostela, Spain}
+\affiliation{CIQUP, Centro de Investigação em Química, Departamento de Química e Bioquímica, Faculdade de Ciências, Universidade do Porto, Porto, Portugal}
 
@@
 \affiliation[University of Southampton]{Chemistry, University of Southampton, Highfield, Southampton SO17 1BJ, United Kingdom}
+
+\author{{\'A}ngel Pi{\~n}eiro}
+\affiliation{Departamento de F{\'i}sica Aplicada, Facultade de F{\'i}sica, Universidade de Santiago de Compostela, E-15782 Santiago de Compostela, Spain}
 

### 2021-04-23 Data/PDBdistributions/pdbtable.tex

\documentclass{article}

%\usepackage{longtable}

\title{Tables of PDB structures used to analyze protein-bound lipids}

\begin{document}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & molecule name \\
\hline
1ein & Lipase \\ 
3uu5 & Proton-gated ion channel \\ 
3p4w & Proton-gated ion channel \\ 
3p50 & Proton-gated ion channel \\ 
3uu6 & Proton-gated ion channel \\ 
3uu8 & Proton-gated ion channel \\ 
3uub & Proton-gated ion channel \\ 
4dos & Nuclear receptor subfamily 5 group A member 2 \\ 
3h1j & MITOCHONDRIAL UBIQUINOL-CYTOCHROME-C \\
     &REDUCTASE COMPLEX CORE PROTEIN I \\ 
1lpa & Colipase \\ 
4hfc & Proton-gated ion channel \\ 
5mvm & Proton-gated ion channel \\ 
5ejz & Cellulose synthase catalytic subunit [UDP-forming] \\ 
5j0z & Proton-gated ion channel \\ 
5mzr & Proton-gated ion channel \\ 
5mvn & Proton-gated ion channel \\ 
5mur & Proton-gated ion channel \\ 
5mzt & Proton-gated ion channel \\ 
4hfe & Proton-gated ion channel \\ 
4hfi & Proton-gated ion channel \\ 
4ila & Proton-gated ion channel \\ 
4ilb & Proton-gated ion channel \\ 
4hfh & Proton-gated ion channel \\ 
4hfb & Proton-gated ion channel \\ 
4ilc & Proton-gated ion channel \\ 
4qh5 & Proton-gated ion channel \\ 
4p02 & Cellulose synthase catalytic subunit [UDP-forming] \\ 
4p00 & Cellulose synthase catalytic subunit [UDP-forming] \\ 
4hfd & Proton-gated ion channel \\ 
5eiy & Cellulose synthase catalytic subunit [UDP-forming] \\ 
4zzc & Proton-gated ion channel \\ 
4il9 & Proton-gated ion channel \\ 
6yj4 & NADH-ubiquinone oxidoreductase chain 3 \\ 
3rw0 & Ion\_trans domain-containing protein \\ 
4zxh & Carrier domain-containing protein \\ 
5klv & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
5okd & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
    \end{tabular}
    \caption{PDBIDs and molecules names used in the analysis of PC}
    \label{tab:my_label}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & molecule name \\
\hline
5vb2 & Ion\_trans domain-containing protein \\ 
4ms2 & Ion\_trans domain-containing protein \\ 
5vb8 & Ion\_trans domain-containing protein \\ 
6mwa & Ion\_trans domain-containing protein \\ 
6mwg & Ion\_trans domain-containing protein \\ 
6mwd & Ion\_trans domain-containing protein \\ 
6mvv & Ion\_trans domain-containing protein \\ 
6haw & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
6mwb & Ion\_trans domain-containing protein \\ 
6mvw & Ion\_trans domain-containing protein \\ 
6c1k & Ion\_trans domain-containing protein \\ 
5yua & Ion\_trans domain-containing protein \\ 
6c1p & Ion\_trans domain-containing protein \\ 
6c1e & Ion\_trans domain-containing protein \\ 
6c1m & Ion\_trans domain-containing protein \\ 
6p6x & Ion\_trans domain-containing protein \\ 
6p6y & Ion\_trans domain-containing protein \\ 
6ubs & Glycine receptor subunit alphaZ1 \\ 
6mvy & Ion\_trans domain-containing protein \\ 
6xvf & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
2h26 & T-cell surface glycoprotein CD1b \\ 
3b7q & CRAL-TRIO domain-containing protein YKL091C \\ 
3b7z & CRAL-TRIO domain-containing protein YKL091C \\ 
6z5r & Light-harvesting complex 1 alpha chain \\ 
6z5s & Light harvesting complex 1 Protein W \\ 
6c15 & T-cell surface glycoprotein CD1c \\ 
1dmh & Catechol 1,2-dioxygenase \\ 
1dlm & Catechol 1,2-dioxygenase \\ 
1dlt & Catechol 1,2-dioxygenase \\ 
1dlq & Catechol 1,2-dioxygenase \\ 
3n9t & INTRADIOL\_DIOXYGENAS domain-containing protein \\ 
5kym & 1-acyl-sn-glycerol-3-phosphate acyltransferase \\ 
2hh1 & Reaction center protein L chain \\ 
2hg9 & Reaction center protein L chain \\ 
5mzq & Proton-gated ion channel \\ 
6qpc & Anoctamin-6 \\ 
6qp6 & Anoctamin-6 \\ 
5irz & Transient receptor potential cation channel subfamily V member 1 \\ 
5irx & Transient receptor potential cation channel subfamily V member 1 \\ 
    \end{tabular}
    \caption{PDBIDs and molecules names used in the analysis of PC}
    \label{tab:my_label}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & molecule name \\
\hline
4i9j & 1-phosphatidylinositol phosphodiesterase \\ 
5ej1 & Cellulose synthase catalytic subunit [UDP-forming] \\ 
6s7o & Dolichyl-diphosphooligosaccharide--protein glycosyltransferase subunit STT3A \\ 
6c26 & Dolichyl-diphosphooligosaccharide--protein glycosyltransferase subunit STT3 \\ 
6s7t & Dolichyl-diphosphooligosaccharide--protein glycosyltransferase subunit STT3B \\ 
1lsh & Vitellogenin \\ 
6uz8 & Short transient receptor potential channel 6 \\ 
4tso & DELTA-actitoxin-Afr1a \\ 
4tsp & DELTA-actitoxin-Afr1a \\ 
4tsq & DELTA-actitoxin-Afr1a \\ 
6iej & Cytosolic phospholipase A2 \\ 
2a1l & Phosphatidylinositol transfer protein beta isoform \\ 
1t27 & Phosphatidylinositol transfer protein alpha isoform \\ 
4f2a & Cholesteryl ester transfer protein \\ 
4nab & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
4uu1 & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
5d3i & Toll-like receptor 2 \\ 
5uph & Lysosome membrane protein 2 \\ 
5ylu & Potassium-transporting ATPase alpha chain 1 \\ 
2obd & Cholesteryl ester transfer protein \\ 
5xa7 & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
5xab & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
5xaa & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
6yu4 & Sodium-dependent transporter \\ 
6vyk & Mechanosensitive channel MscS \\ 
6rld & Small-conductance mechanosensitive channel \\ 
6jxi & Potassium-transporting ATPase alpha chain 1 \\ 
6jxh & Potassium-transporting ATPase alpha chain 1 \\ 
6jxj & Potassium-transporting ATPase alpha chain 1 \\ 
5xa8 & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
5xa9 & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
5ncq & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
5ylv & Potassium-transporting ATPase alpha chain 1 \\ 
7d91 & Sodium/potassium-transporting ATPase subunit alpha-1 \\ 
7k4b & Transient receptor potential cation channel subfamily V member 6 \\ 
7ddl & Sodium/potassium-transporting ATPase subunit alpha-1 \\ 
7bt2 & Sarcoplasmic/endoplasmic reticulum calcium ATPase 2 \\ 
7ddh & Sodium/potassium-transporting ATPase subunit alpha-1 \\ 
7ddj & Sodium/potassium-transporting ATPase subunit alpha-1 \\ 
7ddk & Sodium/potassium-transporting ATPase subunit alpha-1 \\ 
7k4a & Transient receptor potential cation channel subfamily V member 6 \\ 
7ado & ER membrane protein complex subunit 1 \\ 
7d94 & Sodium/potassium-transporting ATPase subunit alpha-1 \\ 
6pw5 & Ion\_trans domain-containing protein \\ 
    \end{tabular}
    \caption{PDBIDs and molecules names used in the analysis of PC}
    \label{tab:my_label}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & molecule name \\
\hline
3cxh & Cytochrome b-c1 complex subunit 1, mitochondrial \\
3cx5 & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
5klv & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
4rpe & Linoleate 9/13-lipoxygenase \\ 
6ymx & Cytochrome c oxidase subunit 1 \\ 
6giq & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
6nhg & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
3ar5 & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
3ar4 & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
2dqs & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
2agv & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
3jtc & Endothelial protein C receptor \\ 
3nan & Calcium-transporting ATPase \\ 
3w5c & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
3nam & Calcium-transporting ATPase \\ 
3w5b & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
3zuy & Transporter \\ 
3zux & Transporter \\ 
3w5a & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
2eau & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
3ar3 & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
3ar7 & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
3ar6 & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
1l8j & Endothelial protein C receptor \\ 
1lqv & Endothelial protein C receptor \\ 
4cz8 & Na\_H\_Exchanger domain-containing protein \\ 
4ycn & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
5jmn & Multidrug efflux pump subunit AcrB \\ 
4j2t & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
3w5d & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
4v3e & IT4VAR07 CIDRA \\ 
4v3d & HB3VAR03 CIDRA DOMAIN \\ 
5a1s & Citrate-sodium symporter \\ 
4ycm & Sarcoplasmic/endoplasmic reticulum calcium ATPase 1 \\ 
4n31 & Signal peptidase I \\ 
2vwa & Early transcribed membrane protein 13 \\ 
3pg7 & Neurofibromin \\ 
5td3 & Catechol 1,2-dioxygenase \\ 
5vxt & Catechol 1,2-dioxygenase \\ 
6wlw & V-type proton ATPase 21 kDa proteolipid subunit \\ 
6wbn & Pannexin-1 \\ 
6wbg & Pannexin-1 \\ 
6wbf & Pannexin-1 \\ 
6wbm & Pannexin-1 \\ 
    \end{tabular}
    \caption{PDBIDs and molecules names used in the analysis of PC}
    \label{tab:my_label}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & name \\
\hline
6z16 & Multisubunit Na+/H+ antiporter, A subunit \\ 
6ymy & Cytochrome c oxidase subunit 1 \\ 
6wm2 & V-type proton ATPase subunit E 1 \\ 
1p84 & Ubiquinol-cytochrome C reductase complex core protein I \\ 
3szd & Probable porin \\ 
3m9i & Lens fiber major intrinsic protein \\ 
5ejz & Cellulose synthase catalytic subunit [UDP-forming] \\ 
4p02 & Cellulose synthase catalytic subunit [UDP-forming] \\ 
4pd4 & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
4p00 & Cellulose synthase catalytic subunit [UDP-forming] \\ 
4qn9 & N-acyl-phosphatidylethanolamine-hydrolyzing phospholipase D \\ 
5ej1 & Cellulose synthase catalytic subunit [UDP-forming] \\ 
6zk9 & NADH dehydrogenase [ubiquinone] flavoprotein 1, mitochondrial \\ 
6zkb & Mitochondrial complex I, 49 kDa subunit \\ 
1m57 & Cytochrome c oxidase subunit 1 \\ 
1n69 & Saposin-B \\ 
1m56 & Cytochrome c oxidase subunit 1 \\ 
5weh & Cytochrome c oxidase subunit 1 \\ 
6zqv & Genome polyprotein \\ 
2x72 & Rhodopsin \\ 
2qgu & Probable signal peptide protein \\ 
3wmm & Photosynthetic reaction center cytochrome c subunit \\ 
1yp0 & Steroidogenic factor 1 \\ 
1zdt & Steroidogenic factor 1 \\ 
1kb9 & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
1eys & PHOTOSYNTHETIC REACTION CENTER \\ 
5y5s & Photosynthetic reaction center cytochrome c subunit \\ 
5w7l & Bac\_transf domain-containing protein \\ 
6t15 & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
5z39 & PH domain-containing protein \\ 
6hu9 & Cytochrome b-c1 complex subunit 1, mitochondrial \\ 
5irz & Transient receptor potential cation channel subfamily V member 1 \\ 
5irx & Transient receptor potential cation channel subfamily V member 1 \\ 
6cud & Short transient receptor potential channel 3 \\ 
6o72 & Transient receptor potential cation channel subfamily M member 8 \\ 
6o6r & Transient receptor potential cation channel subfamily M member 8 \\ 
2e2x & Neurofibromin \\ 
3peg & Neurofibromin \\ 
3p7z & Neurofibromin \\ 
    \end{tabular}
    \caption{PDBIDs and molecules names used in the analysis of PC}
    \label{tab:my_label}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & name \\
\hline
6rd5 & ASA-10: Polytomella F-ATP synthase associated subunit 10 \\ 
6xiw & Sodium leak channel non-selective protein \\ 
6lum & Succinate dehydrogenase subunit C \\ 
6tyi & Biopolymer transport protein ExbB \\ 
5jwy & Phosphatidylglycerophosphatase B \\ 
6uza & Short transient receptor potential channel 6 \\ 
3l9r & Ig-like domain-containing protein \\ 
6or2 & Membrane protein, MmpL family protein \\ 
6w2y & Gamma-aminobutyric acid type B receptor subunit 1 \\ 
2ein & Cytochrome c oxidase subunit 1 \\ 
2eil & Cytochrome c oxidase subunit 1 \\ 
2eij & Cytochrome c oxidase subunit 1 \\ 
3ag3 & Cytochrome c oxidase subunit 1 \\ 
3abm & Cytochrome c oxidase subunit 1 \\ 
2y69 & Cytochrome c oxidase subunit 1 \\ 
2eik & Cytochrome c oxidase subunit 1 \\ 
2dyr & Cytochrome c oxidase subunit 1 \\ 
2r40 & Ecdysone receptor \\ 
1r20 & NR LBD domain-containing protein \\ 
3ae5 & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3ixp & NR LBD domain-containing protein \\ 
3vr8 & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3vr9 & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
1zoy & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
1g2n & NR LBD domain-containing protein \\ 
1nek & Succinate dehydrogenase flavoprotein subunit \\ 
1r1k & Ecdysone receptor \\ 
3ae8 & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3ae9 & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3aeb & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3ae2 & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3ae4 & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3ae3 & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3abv & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3ae6 & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3aea & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3ae1 & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3aef & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
1nen & Succinate dehydrogenase flavoprotein subunit \\ 
4ytm & Succinate dehydrogenase flavoprotein \\ 
3vra & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
3vrb & Succinate dehydrogenase [ubiquinone] flavoprotein subunit, mitochondrial \\ 
    \end{tabular}
    \caption{PDBIDs and molecules names used in the analysis of PC}
    \label{tab:my_label}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & name \\
\hline
4g33 & Linoleate 9/13-lipoxygenase \\ 
5ir4 & Lipoxygenase LoxA \\ 
5lc8 & Linoleate 9/13-lipoxygenase \\ 
5ir5 & Lipoxygenase LoxA \\ 
4g32 & Linoleate 9/13-lipoxygenase \\ 
6a9j & Autophagy-related protein 2 \\ 
6adq & Cytochrome aa3 subunit 2 \\ 
7bh1 & Potassium-transporting ATPase potassium-binding subunit \\ 
7bgy & Potassium-transporting ATPase potassium-binding subunit \\ 
7lc3 & Potassium-transporting ATPase potassium-binding subunit \\ 
7bh2 & Potassium-transporting ATPase potassium-binding subunit \\ 
5jxd & Tumor necrosis factor alpha-induced protein 8 \\ 
5h5a & Mitochondrial distribution and morphology protein 12 \\ 
6pqo & Transient receptor potential cation channel subfamily A member 1 \\ 
6pqp & Transient receptor potential cation channel subfamily A member 1 \\ 
6x16 & Glutamate transporter homologue GltPh \\ 
6vyi & Diacylglycerol O-acyltransferase 1 \\ 
6w98 & F5/8 type C domain-containing protein \\ 
6uz3 & Sodium channel protein type 5 subunit alpha,Green fluorescent protein \\ 
6uw4 & Transient receptor potential cation channel subfamily V member 3 \\ 
6x15 & Glutamate transporter homologue GltPh \\ 
6agf & Sodium channel protein type 4 subunit alpha \\ 
6br8 & Protein A6 homolog \\ 
6br9 & Protein A6 homolog \\ 
6vz1 & Diacylglycerol O-acyltransferase 1 \\ 
6uwf & Glutamate transporter homolog \\ 
6uz0 & Sodium channel protein type 5 subunit alpha,Green fluorescent protein \\ 
6pqq & Transient receptor potential cation channel subfamily A member 1 \\ 
7k18 & Green fluorescent protein \\ 
7kv8 & Envelope protein E \\ 
7cu3 & Sodium leak channel non-selective protein \\ 
2qjk & Cytochrome b \\ 
2qjy & Cytochrome b \\ 
2fyn & Cytochrome b \\ 
3ayg & COX1 domain-containing protein \\ 
2qjp & Cytochrome b \\ 
3ayf & COX1 domain-containing protein \\ 
5kkz & Cytochrome b \\ 
5kli & Cytochrome b \\ 
6qq6 & Nitric oxide reductase subunit B \\ 
1pp9 & Ubiquinol-cytochrome C reductase complex core protein I, mitochondrial \\ 
    \end{tabular}
    \caption{PDBIDs and molecules names used in the analysis of PC}
    \label{tab:my_label}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & name \\
\hline
3h1l & Mitochondrial ubiquinol-cytochrome-c reductase complex core protein i \\ 
3tgu & Mitochondrial ubiquinol-cytochrome-c reductase complex core protein i \\ 
3h1k & Mitochondrial ubiquinol-cytochrome-c reductase complex core protein i \\ 
3l70 & Mitochondrial ubiquinol-cytochrome-c reductase complex core protein i \\ 
3l75 & Mitochondrial ubiquinol-cytochrome-c reductase complex core protein i \\ 
3l71 & Mitochondrial ubiquinol-cytochrome-c reductase complex core protein i \\ 
2bcc & Cytochrome b-c1 complex subunit 1, mitochondrial \\  
    \end{tabular}
    \caption{PDBIDs and molecules names used in the analysis of PC}
    \label{tab:my_label}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & molecule name \\
\hline
3mtx & Lymphocyte antigen 86 \\ 
4c7r & Glycine betaine transporter BetP \\ 
4doj & Glycine betaine transporter BetP \\ 
5azb & Phosphatidylglycerol--prolipoprotein diacylglyceryl transferase \\ 
4opc & Digeranylgeranylglycerophospholipid reductase \\ 
4phz & unknown peptide \\ 
5azc & Phosphatidylglycerol--prolipoprotein diacylglyceryl transferase \\ 
6z5r & Light-harvesting complex 1 alpha chain \\ 
6z5s & Light harvesting complex 1 Protein W \\ 
6tyi & Biopolymer transport protein ExbB \\ 
6tjv & NAD(P)H-quinone oxidoreductase subunit 1 \\ 
2hhk & Reaction center protein L chain \\ 
2cyd & V-type sodium ATPase subunit K \\ 
2bhw & Chlorophyll a-b binding protein AB80, chloroplastic \\ 
1rwt & Chlorophyll a-b binding protein, chloroplastic \\ 
2axt & Photosystem II protein D1 1 \\ 
2bl2 & V-type sodium ATPase subunit K \\ 
3jcu & Photosystem II protein D1 \\ 
5t85 & Antibody 10E8 FAB HEAVY CHAIN \\ 
2ein & Cytochrome c oxidase subunit 1 \\ 
2eil & Cytochrome c oxidase subunit 1 \\ 
2eij & Cytochrome c oxidase subunit 1 \\ 
2irv & Rhomboid protease GlpG \\ 
3ag3 & Cytochrome c oxidase subunit 1 \\ 
3abm & Cytochrome c oxidase subunit 1 \\ 
2y69 & Cytochrome c oxidase subunit 1 \\ 
2eik & Cytochrome c oxidase subunit 1 \\ 
2dyr & Cytochrome c oxidase subunit 1 \\ 
3oz2 & Digeranylgeranylglycerophospholipid reductase \\ 
4i7z & Cytochrome b6 \\ 
5wki & T-cell surface glycoprotein CD1b \\ 
5wl1 & T-cell surface glycoprotein CD1b \\ 
2r9r & Voltage-gated potassium channel subunit beta-2 \\ 
3wmm & Photosynthetic reaction center cytochrome c subunit \\ 
4jta & Voltage-gated potassium channel subunit beta-2 \\ 
4jtc & Voltage-gated potassium channel subunit beta-2 \\ 
4jtd & Voltage-gated potassium channel subunit beta-2 \\ 
3lnm & Voltage-gated potassium channel subunit beta-2 \\ 
\end{tabular}
    \caption{PDBIDs and molecule names of PG lipids}
    \label{tab:my_label}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & name \\
\hline
6tos & Orexin receptor type 1 \\ 
6to7 & Orexin receptor type 1 \\ 
6tot & Orexin receptor type 1 \\ 
6tp6 & Orexin receptor type 1 \\ 
6tod & Orexin receptor type 1 \\ 
6cju & Cyclic nucleotide-binding domain-containing protein \\ 
6cjt & Cyclic nucleotide-binding domain-containing protein \\ 
6bw5 & UDP-N-acetylglucosamine--dolichyl-phosphate N-acetylglucosaminephosphotransferase \\ 
6bw6 & UDP-N-acetylglucosamine--dolichyl-phosphate N-acetylglucosaminephosphotransferase \\ 
6tp4 & Orexin receptor type 1 \\ 
6tq6 & Orexin receptor type 1 \\ 
6tq7 & Orexin receptor type 1 \\ 
6tq4 & Orexin receptor type 1 \\ 
6v35 & Calcium-activated potassium channel subunit alpha-1 \\ 
6tp3 & Orexin receptor type 1 \\ 
6tq9 & Orexin receptor type 1 \\ 
7cge & Lipid asymmetry maintenance ABC transporter permease subunit MlaE \\ 
1ymt & Steroidogenic factor 1 \\ 
7kc4 & Protein Wnt-8a \\ 
1yok & Nuclear receptor subfamily 5 group A member 2 \\ 
4nh2 & Ammonium transporter \\ 
4oni & Nuclear receptor subfamily 5 group A member 2 \\ 
5eg1 & Microcin-J25 export ATP-binding/permease protein McjD \\ 
3tx7 & Catenin beta-1 \\ 
6rck & Outer membrane protein C \\ 
6uxp & Bcl-2 homologous antagonist/killer \\ 
6hsy & Toluene tolerance protein Ttg2D \\ 
\end{tabular}
    \caption{PDBIDs and molecule names of PG lipids}
    \label{tab:my_label}
\end{table}

\begin{table}[]
    \centering
    \begin{tabular}{c|c}
PDBID & name \\
\hline
3kaa & Hepatitis A virus cellular receptor 2 homolog \\ 
3bib & T-cell immunoglobulin and mucin domain-containing protein 4 \\ 
1dsy & Protein kinase C alpha type \\ 
6wlw & V-type proton ATPase 21 kDa proteolipid subunit \\ 
6wm2 & V-type proton ATPase subunit E 1 \\ 
6b8o & Triggering receptor expressed on myeloid cells 2 \\ 
6snd & LN01 light chain \\ 
6roh & Probable phospholipid-transporting ATPase DRS2 \\ 
3d9s & Aquaporin-5 \\ 
5dye & Aquaporin-5 \\ 
5c5x & Aquaporin-5 \\ 
6u9w & P2X purinoceptor 7 \\ 
6u9v & P2X purinoceptor 7 \\ 
4b2z & Oxysterol-binding protein homolog 6 \\ 
6lcp & Phospholipid-transporting ATPase \\ 
6lcr & Phospholipid-transporting ATPase \\ 
6i3y & PRELI domain-containing protein 1, mitochondrial \\ 
6vp0 & Diacylglycerol O-acyltransferase 1 \\ 
6tt7 & ATP synthase subunit alpha \\ 
6sp2 & Membrane protein TMS1d \\ 
7bsu & ATP11C \\ 
7bsv & ATP11C \\ 
6k7m & Phospholipid-transporting ATPase \\ 
2hj6 & Reaction center protein L chain \\ 
4hyt & Sodium/potassium-transporting ATPase subunit alpha-1 \\ 
4res & Sodium/potassium-transporting ATPase subunit alpha-1 \\ 
6uxn & Bcl-2 homologous antagonist/killer \\ 
6nhh & Cytochrome b \\ 
    \end{tabular}
    \caption{PDBIDs and molecule names of protein structures used in the analysis of PS headgroup structures.}
    \label{tab:my_label}
\end{table}

\end{document}
